### 2021-10-22 tex/paper.wecs/wecs_tgrs.tex
estrutura+ (22out21)

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
 \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath
%\usepackage{amsaddr}		
%\usepackage{amsthm}		
%\usepackage{amsaddr}		
\usepackage{amsfonts}

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

\def\bSig\mathbf{\Sigma}
\newcommand{\VS}{V\&S}
\usepackage{url}
\usepackage{soul} % allows to pass a line in words
\newcommand{\Z}{\mathbb{Z}} % Set of integers
\newcommand{\re}{\mathbb{R}} % Set of real numbers
\newcommand{\sgn}{\mathrm{sign}} % Signal operador
\newcommand{\ci}{\mathrm{i}}
\def\sT{\mbox{\tiny$T$}}
\def\E{\mbox{{\rm E\,}}}
\def\cov{\mbox{{\rm cov\,}}}
\def\corr{\mbox{{\rm corr\,}}}
\def\var{\mbox{{\rm var\,}}}
\newcommand{\vgamma}{\pmb{\gamma}}
\newcommand{\vrho}{\pmb{\rho}}
\newcommand{\vbeta}{\pmb{\beta}}
\newcommand{\vxi}{\pmb{\xi}}
\newcommand{\vepsilon}{\pmb{\epsilon}}
\def\shalf{\mbox{{\tiny$\frac{1}{2}$}}}
\def\half{\mbox{{\small$\frac{1}{2}$}}}
%\newtheorem{theorem}{Theorem}{\bf}{\rm}  
\newtheorem{lemma}{Lemma}

\newcommand{\vA}{{\textbf A}}
\newcommand{\vD}{{\textbf D}}
\newcommand{\vd}{{\textbf d}}
\newcommand{\vW}{{\textbf W}}
\newcommand{\vR}{{\textbf R}}
\newcommand{\vU}{{\textbf U}}
\newcommand{\vI}{{\textbf I}}
\newcommand{\vX}{{\textbf X}}
\newcommand{\vf}{{\textbf f}}
\newcommand{\vu}{{\textbf u}}
\newcommand{\vT}{{\textbf T}}
\newcommand{\vt}{{\textbf t}}

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Wavelet Spatio-Temporal Change Detection on multi-temporal POLSAR images}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Rodney~Fonseca,
        Alu\'{i}sio~Pinheiro,
        and~Abdourrahmane~Atto,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
\thanks{This work was supported by FAPESP grants 2016/24469-6 and 2018/04654-9 and CNPq grants 309230/2017-9 and 310991/2020-0.}
\thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
\thanks{A.~Atto is with the LISTIC - Polytech Annecy-Chamb\'{e}ry, Universit\'{e} de Savoie, 74944 Annecy le Vieux Cedex, France (e-mail: Abdourrahmane.Atto@univ-savoie.fr).}}
% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~X, No.~X, Month~20XX}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised sparse procedure to detect spatio-temporal change points on multi-temporal Polarimetric SAR (POLSAR) images or even on sequences of very high resolution images. The procedure is based on multiscale approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients. We present two complimentary multiscale measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images. We show WECS performance on synthetic multi-temporal image data. We also apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana. The images were captured from November 08, 2015 to December 09 2017.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Change detection, multi-temporal images, satellite images, wavelets.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{W}{e} propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR/POLSAR images. WECS is based upon correlation screening for energy apportionment on wavelet approximations. The spatial character of the change detection is attained on pixel level. The method is fast, scalable, linearly updatable, and the resulting measures are sparse.

A review for change detection in multi-temporal remote sensing is given by \cite{ban2016change}. 
Different proposals for this purpose may be found  in the literature. They vary in their motivations as well as in their applicability.  Change detection in multi-temporal hyperspectral images is discussed in \cite{bovolo2015time},  \cite{liu2019review}, and \cite{matsunaga2017current}.  \cite{jia2018novel} pursue change detection techniques via non-local means and principal component analysis. Compressed projection and image fusion are employed by \cite{hou2014unsupervised}. Deep learning by slow feature analysis for change detection is the subject of \cite{du2019unsupervised}. \cite{chen2020change} proposes a change detection method driven by adaptive parameter estimation.

Besides different methodological paradigms, several areas of application receive special attention.  For instance, urban change detection applications via polarimetric SAR (POLSAR) images are discussed in \cite{ansari2020urban}. \cite{song2018multi} discusses land cover change detection in mountainous terrain via multi-temporal and multi-sensor remote sensing images. \cite{ru2021multi} studies multi-temporal scene classification and scene change detection. Deforestation change detection is discussed by \cite{barreto2016deforestation}.

Wavelet methods present many advantages for a plethora of applications \cite{vidakovic1999statistical} thanks to wavelet capabilities in capturing multiscale/multiresolution information. Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical} use different wavelet methods for change detection in satellite images.

The motivation for our proposed method is multi-fold. We aim for a fast and accurate method. We would also like this method to be easily updatable when a new observation is captured. Finally, scalability was a concern as well. We propose a wavelet-based procedure for change detection in multi-temporal remore sensing images (WECS). It is unsupervised and built on ultra-high dimensional correlation screening \cite{fan2020statistical} for the wavelet coefficients.  We present two complimentary wavelet measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images. The procedure presents some advantages. It is unsupervised, fast and updatable, thus allowing for real-time change detection. Moreover, it is sparse and scalable.

The rest of the text goes as follows. Section \ref{section_method} introduces the problem and presents the proposed method. We show WECS performance on synthetic multi-temporal image data in Section \ref{section_validation}. In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.  Section \ref{section_discussion} concludes the paper with a discussion.

\section{Methodology}\label{section_method}

Let $\mathcal{I}(1),\ldots,\mathcal{I}(m)$ be a set of matrices representing the log-images of some region of interest. These images may be relative to one POLSAR channel or a combination of channels; this will be specified when appropriate. Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}(m)$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time. We shall address these tasks by analyzing the bidimensional discrete wavelet decomposition of $\mathcal{I}(m)$ at the $J$-th approximation level, i.e., 
such that
\begin{equation} 
\mathcal{I}(m)=\mathcal{I}_J(m)+\vepsilon_J(m),
\end{equation}
%\label{eq_modellogimage}
where $\mathcal{I}_J(m)$ is the wavelet approximation of log-image $\mathcal{I}(m)$ at level $J$. We denote $\mathcal{I}_J(m)$'s approximation coefficients matrix by $\vX(m)$ \cite{morettin2017wavelets,vidakovic1999statistical}. The advantages of doing so is that we linearize possible speckle (multiplicative) noise in the original images by taking logarithms to obtain $\mathcal{I}(m)$ and later perform a smoothing to obtain $\mathcal{I}_J(m)$, whose approximation coefficients matrix is $\vX(m)$. Such approach is also described in the literature as multiplicative wavelet transform \cite{atto2016wavelet}.

We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\mathcal I(m)\right\}$ as
\begin{equation} 
\sum_{m=1}^n\|\mathcal I_J(m)\|^2_2=n\|\bar{\mathcal I}\|^2_2+\sum_{m=1}^n\|\mathcal I_J(m)-\bar{\mathcal I}\|^2_2,\label{eq_ANOVAlogimage}
\end{equation}
where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}_J(m)$. The same can be done on the wavelet domain with $\vX(m)$, which is the way we shall proceed in the following steps.

We discuss here two procedures which are similar, but may yield different results depending on the nature of relevant changes. The first procedure makes use of (\ref{eq_ANOVAlogimage}), so that we establish an {\it average wavelet-approximated image} and detect time points for which images differ from the {\it characteristic} image. Changes are detected by absolute correlations between individual wavelet approximation coefficients time series and the time series of overall wavelet energy. The second procedure also uses absolute correlations but instead of an average wavelet image energy, we compute  the overall energy differences for subsequent wavelet-smoothed images and their approximation coefficients. The former should help us detect image changes from an overall behavior over time whilst the latter should also detect changes on non-stationary set-ups.  In practive, both procedures shall be performed in the wavelet domain by using $\{\vX(m)\}$ instead of $\{\mathcal{I}_J(m)\}$.

The average image $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}_J(m)$ has approximation coefficients matrix given by $\bar{\vX} = n^{-1}\sum_{m=1}^n \vX(m)$. Let $X_{k,l}(m)$ and $\bar{X}_{k,l}$ be the entry $(k,l)$ of the matrices $\vX(m)$ and $\bar{\vX}$, respectively. We take the matrix $\vD(m) = [D_{k.l}(m)]$, where $D_{k,l}(m)=(X_{k,l}(m)-\bar{X}_{k,l})^2$. We then 
analyze the time series given by
\begin{equation} 
\vd(m)= \sum_{k,l}D_{k,l}(m) = \sum_{k,l}(X_{k,l}(m) - \bar{X}_{k,l})^2,
\label{eq_defdm}
\end{equation}
$m=1,\ldots,n$, which displays the temporal variation with respect to $\bar{\vX}$ of spatial energies.

The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy. Define the  $n\times p$ matrix
\begin{equation}
 \vD=\left(
 \begin{array}{c}
 vec(\vD(1))^T\\
 \vdots\\
 vec(\vD(n))^T\\
 \end{array}
 \right),
\label{eq_defmatrixD}
\end{equation}
where $vec(\vD(m))$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX(m)$, $m=1,\ldots,n$. Sparsity \cite{johnstone2009statistical} on the wavelet coefficients plays a special role here.  We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $e_d$, is such that $e_d<<p$. This can be represented as the following linear model
\begin{equation} 
\vd=\vD\vbeta^{(d)}+\vxi^{(d)}
\label{sparsemodel_d}
\end{equation}
where $\vbeta^{(d)}$ is sparse, i.e., it has $p-e_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
%\citep{jonhstone2009}

 In order to identify spatio-temporal changes we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows. For each squared mean-corrected approximation coefficient time series, given by $\vD(m)$, consider its Pearson correlation with the mean-corrected total approximation energy, given by $\vd$:
\begin{equation}
R_{k,l}^{(d)}= \corr\left( \vD_{k,l}, \vd\right),
\label{def_Rkld}
\end{equation}
where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.

We have a matrix $\vR^{(d)}=[R_{k,l}^{(d)}]$ of correlations of ultra-high dimension. Define the set of 
{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
$\mathcal{M}^{*d}=\{(k,l): \textit{change in } \mathcal{I}(m)\textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are caused by changes in approximation coefficients of index} $(k,l)\}$.  This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta^{(d)}$ in (\ref{sparsemodel_d}). 
We build the empirical set of selected indices by
\begin{equation}
\mathcal{M}_{\tau}^{(d)}=\{(k,l):|R_{k,l}^{(d)}|>\tau_d\},
\label{def_Mtaud}
\end{equation}
where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}^{(d)}$ contains its target set satisfies
\[
P(\mathcal{M}_{\tau}^{(d)}\supset\mathcal{M}^{*d})\rightarrow 1,
\]
as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.

Analogously we take $\vT(m) = [T_{k,l}(m)]$, where $T_{k,l}(m)=(X_{k,l}(m+1)-X_{k,l}(m))^2$, for $m=1,\ldots,n-1$. We then analyze the time series given by
\begin{equation}
\vt(m)= \sum_{k,l}T_{k,l}(m) = \sum_{k,l}(X_{k,l}(m+1) - X_{k,l}(m))^2,
\label{eq_deftm}
\end{equation}
$m=1,\ldots,n-1$.

The highest values of $\{\vt(m)\}$ represent the images for which the most expressive changes take place between time points $m$ and $m+1$, where changes here are measured through $\mathbb{L}_2$ energy. Define the  $(n-1)\times p$ matrix
\begin{equation}
 \vT=\left(
 \begin{array}{c}
 vec(\vT(1))^T\\
 \vdots\\
 vec(\vT(n))^T\\
 \end{array}
 \right),
\label{eq_defmatrixT}
\end{equation}
where $vec(\vT(m))$ is the $p\times 1$ vector of wavelet coefficients for time $m=1,\ldots,n-1$. We again suppose a handful of coefficients drive the changes given by $\vt$, so that the effective dimension, say $e_t$, is such that $e_t<<p$. This can be represented as the following linear model
\begin{equation} 
\vt=\vT\vbeta^{(t)}+\vxi^{(t)}
\label{sparsemodel_t}
\end{equation}
where $\vbeta^{(t)}$ is sparse, i.e. $\vbeta^{(t)}$ has $p-e_t$ null elements, and $\vxi^{(t)}$ is some $(n-1)\times 1$ random vector of errors.

For each squared mean-corrected approximation coefficient time series, given by $\vT_{k,l} = (T_{k,l}(1),\ldots,T_{k,l}(n))^T$, consider its Pearson correlation with the mean-corrected total approximation energy, given by $\vt$:
\begin{equation}
R_{k,l}^{(t)}= \corr\left( \vT_{k,l}, \vt\right).
\label{def_Rklt}
\end{equation}

We have a matrix $\vR^{(t)}=[R_{k,l}^{(t)}]$ of correlations of ultra-high dimension. Define the set of 
{\it important} indices for changes in subsequent images as $\mathcal{M}^{*t}=\{(k,l):$ \textit{Change in }$\mathcal{I}(m+1)$ \textit{with respect to }$\mathcal{I}(m)$ \textit{ for some }$m=1,\ldots,n-1$\textit{ are caused by the approximation coefficients of index }$(k,l)\}$. We build a set of selected indices by
\[
\mathcal{M}_{\tau}^{(t)}=\{(k,l):|R_{k,l}^{(t)}|>\tau_t\},
\] 
where $\tau_t>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions,
\[
P(\mathcal{M}_{\tau}^{(t)}\supset\mathcal{M}^{*t})\rightarrow 1,
\]
as $n\rightarrow\infty$ \cite{fan2020statistical}.

Therefore, if we define
\begin{eqnarray}
\mathcal{M}_{\tau}&=&\mathcal{M}_{\tau}^{(t)}\cup\mathcal{M}_{\tau}^{(d)},\label{def_Mpop}\\
\mathcal{M}^{*}&=& \mathcal{M}^{*d}\cup\mathcal{M}^{*t}\label{def_Memp},
\end{eqnarray}
we have the following consistency property:
\[
P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1.
\]
as $n\rightarrow\infty$.

Thence, if we compute $\{\vd(m)\}$ and $\{\vt(m)\}$, and build $\mathcal{M}_{\tau}=\mathcal{M}_{\tau}^{(t)}\cup\mathcal{M}_{\tau}^{(d)}$, the consistency of the screening methods above guarantees asymptotic coverage of all approximation coefficients strongly associated with changes with respect to the average image as well as immediate previous image with a high probability, as long as the required regularity conditions hold.

Further geometrical motivation for our proposal is given as follows.  We argue the case of $\vd$ but the same may be written regarding $\vt$ as well. As defined by (\ref{eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values. This segregates the multi-temporal images, since the former time points identify the images in which significant changes occur, while the latter indices identify time points with no major changes. Consider $U>L>0$ such that the $s_d$ highest values of $\vd$ are larger than $U$, and the $n-s_d$ smallest values of $\vd$ are smaller than $L$. We also take $\delta=U-L$. The indices defined by  (\ref{def_Mtaud}) are such that   
\[
\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau_d,
\]
i.e., such that  $\sum_{m=1}^nD_{k,l}(m)\vd(m)>\tau_d\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
\[
\left|\sum_{m:\vd(m)>U} D_{k,l}(m)\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}(m)\right|>\Delta,
\]
for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$). Thence, when we employ correlation screening we select the two-dimensional wavelet indices which have the closest empirical directions to the vector of image temporal changes. Thus we are performing a truly spatio-temporal change detection in a single procedure.

\section{Validation on synthetic data}\label{section_validation}

In this section we apply the change detection methods above on synthetic data of multi-temporal images. 
The synthetic multi-temporal images ($n=4$) are shown in Figure \ref{F:EllipsoidChanges}. The first image, $I(1)$, presents three elongated ellipses. Changes consist of three different types of ellipses that are successively added to the original image $I(1)$. The second image, $I(2)$, has new large ellipses added. Smaller ellipses are then added to form $I(3)$ and $I(4)$.

\begin{figure}[htb!]
\centering
%\includegraphics[width=40pc]{EllipsoidChanges/ellipses_all}
\includegraphics[scale=.17]{ellipses_t1}
\includegraphics[scale=.17]{ellipses_t2}
\includegraphics[scale=.17]{ellipses_t3}
\includegraphics[scale=.17]{ellipses_t4}
\caption{Synthetic multi-temporal ($n=4$) images. Features and changes come as ellipses and dots.}
\label{F:EllipsoidChanges}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=.1]{total_changes}(a)
\includegraphics[scale=.1]{corr_changes_dm}(b)
\includegraphics[scale=.1]{corr_changes_tm}(c)
\includegraphics[scale=.1]{corr_changes_logratios}(d)
\includegraphics[scale=.1]{corr_changes_dm_nowavelets}(e)
\includegraphics[scale=.1]{corr_changes_tm_nowavelets}(f)
\caption{Synthetic images with changing ellipses. (a) Image composed by the total changes over time. 
(b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Proposed db2 WECS $\vt(m)$ with $J=2$.
(d) Standard approach. (e) $\vd(m)$ without wavelets. (f) $\vt(m)$  without wavelets. 
}
\label{F:Changes_methods_images}
\end{figure}

Figure \ref{F:Changes_methods_images} illustrates the simulated synthetic images, the proposed wavelet detection methods, and three classic detection methods, as well. Panel (a) presents the total changes with respect to image $I(1)$. Panels (b) and (c) show the results by the proposed methods using Daubechies wavelet with two null moments (db2) and $J=2$ by $\vd(m)$ and $\vt(m)$, respectively. Panel (d) presents the results by a standard approach (aggregated log-ratios). Finally, in Panels (e) and (f)  we can see the results if $\vd(m)$ and $\vt(m)$ are performed purely on the spatial domain, without multiscale approximations. The spatio-temporal advantages of the proposed wavelet  $\vd(m)$ and $\vt(m)$ are clear in Figure \ref{F:Changes_methods_images}. A slight advantage for the detection of small ellipses is attained by $\vd(m)$ over $\vt(m)$.

We compute ROC curves to compare the detection performance of different methods. For this we simulate noisy versions of the synthetic images illustrated by Figure \ref{F:EllipsoidChanges}. The change detection methods are then employed. Each method generates a correlation matrix between the real image of total changes and the estimated one.

Each ROC curve presents how close the magnitude variation of change measures is to the variation of the image of total changes in the following way:
\begin{enumerate}
\item Let $R$ be the matrix of change measures. Compute the range $[r_{\min},r_{\max}]$ of the values in $R$;
\item Let $(r_{(1)},\ldots,r_{(100)})$ be equally space values between $r_{\min}$ and $r_{\max}$;
\item For each $k=1,\ldots,n$, check how many pixels are such that $R_{i,j}>r_{(k)}$ coincide with the pixels $(i,j)$ where a change really occurs on the image of total changes. Dividing this number by the total number of changes gives the true positive rate.
\item For each $k=1,\ldots,n$, check how many pixels are such that $R_{i,j}>r_{(k)}$  do not coincide with the pixels $(i,j)$ where a change really occurs. Dividing this number by the total number of pixels where changes do not occur gives the false positive rate.
\end{enumerate}

\begin{figure}[htp!]
\includegraphics[scale=.13]{methods_comparison}\hspace{-.5cm}(a) 
\includegraphics[scale=.13]{levels_comparison_WithReference}\hspace{-.5cm}(b)\\
\includegraphics[scale=.13]{families_comparison_WithReference}\hspace{-.5cm}(c)
\includegraphics[scale=.13]{dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
\includegraphics[scale=.13]{levels_comparison_NoReference}\hspace{-.5cm}(e)
\includegraphics[scale=.13]{families_comparison_NoReference}\hspace{-.5cm}(f)
\caption{ROC curves for detection of changing ellipses in synthetic images and different methods. (a) The proposed methods in black (db2 WECS $\vd(m)$) and green (db2 WECS $\vt(m)$) vs three non-wavelet methods: standard (red stars); $\vd(m)$ (blue); and $\vt(m)$ (red circles). 
(b) db2 $\vd(m)$ with different levels.
(c) $\vd(m)$ with different wavelet bases and J=2;
(d) The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line).  
(e) db2 $\vt(m)$ with different levels. (f) $\vt(m)$ with different wavelet bases and J=2. }
\label{F:EllipsoidChanges_details}
\end{figure}

Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the synthetic data as follows. The effects of wavelet bases, level of decomposition, deep-learning feature extraction and $\vd(m)$/$\vt(m)$ usage are shown on the ROC curves. We employ the following wavelet bases: Haar;  Daubechies db2; Daubechies db4; Coiflets coif4; Symlets sym2 ; and Symlets sym4. Panels (c) and (f) present the ROC curves for the proposed methods under the aforementioned bases for $\vd(m)$ and $\vt(m)$, respectively. On both instances $J=2$ is employed. The results for $\vd(m)$ are much more robust to basis variation than the ones for $\vt(m)$. Combining the ROC curves's comparison from both panels, Daubechies db2 is the best choice.  Panels (b) and (e) present the ROC curves for different levels of decomposition under the aforementioned bases for $\vd(m)$ and $\vt(m)$, respectively. On both instances db2 is employed, and five levels are considered: $J=1,2,3,4,5$. Levels $J=2,3$ have a clear better performance for $\vd(m)$ (with a slight advantage to $J=2$), whilst $J=1$ is competitive for $\vt(m)$. The overall performance  of $J=2$ warrants its use for the rest of the comparisons. Panel (d) shows how the proposed method performs with or without images' deep learning feature extraction from a residual learning network \cite{zhang2017beyond}. The change detection method is the proposed db2 $\vd(m)$ with $J=2$. We can see that the ROC curves for images treated with deep-learning methods or treated with wavelet based methods are almost identical. The WECS runs in 3.35s, while the deep-learning based $\vd(m)$ runs in 559.12s on a notebook. The configuration of the notebook is: OS - Ubuntu 18.04.5 LTS; RAM 7.7 GB; Intel\textregistered Core\texttrademark $ $i7-7500U CPU @ 2.70GHz x 4; graphics - Intel\textregistered HD Graphics 620 (KBL GT2); GNOME - 3.28.2; OS type - 64-bit. We finally have in Panel (a) the proposed db2 $\vd(m)$ with $J=2$ and db2 $\vt(m)$ with $J=2$  compared to three other non-wavelet methods. These are $\vd(m)$ and $\vt(m)$ where wavelet decomposition is not performed, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated log-ratios of $\{\mathcal{I}(m)\}$. The ROC curves in Panel (a) clearly show that  the proposed db2 $\vd(m)$ with $J=2$ outperforms the rest.

We may summarize these results as: the proposed wavelet $\vd(m)$ method presents a superior performance. It is also equipped with the following nice properties: (i) it is scalable; (ii) it is sparse; (iii) it is parsimonious; (iv) it can be easily adapted to be linearly updated when a new image is acquired; and (v) it is fast. Hence, the proposed wavelet change-detection procedure can be used as a real-time change detection tool for long time series of large images.

\section{Real Data Results}\label{section_realdata}

We employed the proposed change detection method on a series of 85 multi-date satellite images. The images were taken on a forest region at the border of Brazil and the French Guiana from November 08, 2015 to December 09, 2017. Each image has two channels and 1200 by 1000 pixels. We perform three change detection wavelet analyses: VV Polarization Channel;  VH Polarization Channel; and the Combined Image by Euclidean norm.

A multi-resolution analysis (MRA) based on a Symlet basis with filter of length 16 (symlet 8) is built.  The log-images are approximated at levels $J=1,2,3,4$. Table \ref{T:approxenergy} shows the 85 images' average energy for each approximation level. We notice that roughly 99\% of the energy is recovered with $J=1$, and more than  90\% with $J=2$. The VV channel shows better overall energy recovery than the VH channel. For $J=4$ and $J=3$, the Euclidean combination of the polarization channels increases the energy representation percentage.  For $J=2$, VV channel and combined channels are equivalent. VH has 4\% less energy than VV for $J=3$, and $10\%$ less  for $J=4$.

\begin{table*}[h!]
\caption{Wavelet Approximation Mean Energy Percentage for log-images. Forest region at the border of Brazil and the French Guiana from November 08, 2015 to December 09, 2017. $n=85$ multi-date satellite images. Each image has two channels and 1200 by 1000 pixels. VV Polarization Channel;  VH Polarization Channel; and the Combined Image by Euclidean norm. Approximation $J=1,2,3,4$.}
\centering
\begin{tabular}{cccc|cccc|cccc}
\hline
\multicolumn{12}{c}{\sc Mean Approximated Energy Percentage}\\
\hline
\multicolumn{4}{c|}{VV Channel}&\multicolumn{4}{c|}{VH Channel}&\multicolumn{4}{c}{Combined Channels}\\
\hline
$J=4$&$J=3$&$J=2$&$J=1$&$J=4$&$J=3$&$J=2$&$J=1$&$J=4$&$J=3$&$J=2$&$J=1$\\
\hline
0.803&0.847&0.924&0.990&0.763&0.814&0.908&0.988&0.816&0.858&0.931&0.991\\
\hline
\end{tabular}\label{T:approxenergy}
\end{table*}

Figures \ref{F:squared_J1-4_VV}-\ref{F:squared_J1-4_euclid} show the series of squared deviations $\vd(m)$ and $\vt(m)$, for  the VV channel, VH channel, and Euclidean combination, respectively.  An overall feature on this data is that the VV polarization presents much higher energy than the VH. The amount of energy related to changes  is ten times higher on the former compared to the latter's.

Regarding change time points, in each figure, we can notice a pattern of peaks which are common to all approximation levels. They are time points:
\begin{list}{}{}
\item (a) 14, 43, 54, and 58 by the coefficients' squared deviations on the average VV image;
\item (b) 14, 43, 54, and 58 by the coefficients' squared deviations on the consecutive VV images;
\item (c) 14, 38, 41, 43, 54, 56, and 58 by the coefficients' squared deviations on the average VH image;
\item (d) 14, 38, 41, 43, 54-58 by the coefficients' squared deviations on the consecutive VH images;
\item (e) 14, 43, 54, and 58 by the coefficients' squared deviations on the average combined channels image; and
\item (e) 14, 43, 54, and 58 by the coefficients' squared deviations on the consecutive combined channels images.
%%%%%%%% VV 12 14 15 16 (42) 43 54 56 58 d(m)
%%%%%%%% VV 12 14 15 16 (42) 43 54 56 58 t(m)
%%%%%%%% VH  12 14 15 38 41 43 54 56 58
%%%%%%%% euclid 12 14 15 16 43 54 56 58
\end{list}

\begin{figure}[htp!]
\includegraphics[scale=.12]{J1_VV_squared_meandev}(a)
\includegraphics[scale=.12]{consecdif_J1_VV_squared_meandev}(b)\\
\includegraphics[scale=.12]{J2_VV_squared_meandev}(c)
\includegraphics[scale=.12]{consecdif_J2_VV_squared_meandev}(d)\\
\includegraphics[scale=.12]{J3_VV_squared_meandev}(e)
\includegraphics[scale=.12]{consecdif_J3_VV_squared_meandev}(f)\\
\includegraphics[scale=.12]{J4_VV_squared_meandev}(g)
\includegraphics[scale=.12]{consecdif_J4_VV_squared_meandev}(h)
\caption{{\sc VV Polarization Channel} Series of squared deviations $\vd(m)$ and $\vt(m)$. The red horizontal line represents the median value and the yellow horizontal line represents their median plus two times their absolute median deviation.  $\vd(m)$ - Approximation Levels:  (a) $J=1$; (c) $J=2$; (e) $J=3$; (g) $J=4$. $\vt(m)$ - Approximation Levels:  (b) $J=1$; (d) $J=2$; (f) $J=3$; (h) $J=4$. 
}
\label{F:squared_J1-4_VV}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=.12]{J1_VH_squared_meandev}(a)
\includegraphics[scale=.12]{consecdif_J1_VH_squared_meandev}(b)\\
\includegraphics[scale=.12]{J2_VH_squared_meandev}(c)
\includegraphics[scale=.12]{consecdif_J2_VH_squared_meandev}(d)\\
\includegraphics[scale=.12]{J3_VH_squared_meandev}(e)
\includegraphics[scale=.12]{consecdif_J3_VH_squared_meandev}(f)\\
\includegraphics[scale=.12]{J4_VH_squared_meandev}(g)
\includegraphics[scale=.12]{consecdif_J4_VH_squared_meandev}(h)
\caption{{\sc VH Polarization  Channel} Series of squared deviations $\vd(m)$ and $\vt(m)$. The red horizontal line represents the median value and the yellow horizontal line represents their median plus two times their absolute median deviation.  $\vd(m)$ - Approximation Levels:  (a) $J=1$; (c) $J=2$; (e) $J=3$; (g) $J=4$. $\vt(m)$ - Approximation Levels:  (b) $J=1$; (d) $J=2$; (f) $J=3$; (h) $J=4$. 
}
\label{F:squared_J1-4_VH}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=.12]{J1_euclid_squared_meandev}(a)
\includegraphics[scale=.12]{consecdif_J1_euclid_squared_meandev}(b)\\
\includegraphics[scale=.12]{J2_euclid_squared_meandev}(c)
\includegraphics[scale=.12]{consecdif_J2_euclid_squared_meandev}(d)\\
\includegraphics[scale=.12]{J3_euclid_squared_meandev}(e)
\includegraphics[scale=.12]{consecdif_J3_euclid_squared_meandev}(f)\\
\includegraphics[scale=.12]{J4_euclid_squared_meandev}(g)
\includegraphics[scale=.12]{consecdif_J4_euclid_squared_meandev}(h)
\caption{{\sc Combined Channels} Series of squared deviations $\vd(m)$ and $\vt(m)$. The red horizontal line represents the median value  and the yellow horizontal line represents their median plus two times their absolute median deviation.  $\vd(m)$ - Approximation Levels:  (a) $J=1$; (c) $J=2$; (e) $J=3$; (g) $J=4$. $\vt(m)$ - Approximation Levels:  (b) $J=1$; (d) $J=2$; (f) $J=3$; (h) $J=4$. 
}
\label{F:squared_J1-4_euclid}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=.12]{J3_VV_squared_meandev}(a)
\includegraphics[scale=.12]{consecdif_J3_VV_squared_meandev}(b)\\
\includegraphics[scale=.12]{J3_VV_sqrdif_change_locations}(c)
\includegraphics[scale=.12]{consecdif_J3_VV_sqrdif_change_locations}(d)\\
\includegraphics[scale=.12]{J3_VV_sqrdif_lowest_cor_locations}(e)
\includegraphics[scale=.12]{consecdif_J3_VV_sqrdif_lowest_cor_locations}(f)\\
\includegraphics[scale=.12]{perc01_J3_VV_sqrdif_change_locations}(g)
\includegraphics[scale=.12]{perc01_consecdif_J3_VV_sqrdif_change_locations}(h)\\
\includegraphics[scale=.12]{perc01_J3_VV_sqrdif_lowest_cor_locations}(i)
\includegraphics[scale=.12]{perc01_consecdif_J3_VV_sqrdif_lowest_cor_locations}(j)
\caption{{\sc VV Polarization Channel} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$. Red horizontal line represents the median value and yellow line represents two absolute median deviations beyond the median. Squared approximation coefficient deviations: 0.1\% highest absolute correlations - (c) $\vd(m)$;  (d) $\vt(m)$; 0.1\% smallest absolute correlations - (e) $\vd(m)$; (f) $\vt(m)$; 0.01\% highest absolute correlations - (g) $\vd(m)$;  (h) $\vt(m)$;  0.01\% smallest absolute correlations - (i) $\vd(m)$; (j)  $\vt(m)$. 
}  
\label{F:squared_meandev_J3_VV}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=.12]{J3_VH_squared_meandev}(a)
\includegraphics[scale=.12]{consecdif_J3_VH_squared_meandev}(b)

\includegraphics[scale=.12]{J3_VH_sqrdif_change_locations}(c)
\includegraphics[scale=.12]{consecdif_J3_VH_sqrdif_change_locations}(d)

\includegraphics[scale=.12]{J3_VH_sqrdif_lowest_cor_locations}(e)
\includegraphics[scale=.12]{consecdif_J3_VH_sqrdif_lowest_cor_locations}(f)

\includegraphics[scale=.12]{perc01_J3_VH_sqrdif_change_locations}(g)
\includegraphics[scale=.12]{perc01_consecdif_J3_VH_sqrdif_change_locations}(h)

\includegraphics[scale=.12]{perc01_J3_VH_sqrdif_lowest_cor_locations}(i)
\includegraphics[scale=.12]{perc01_consecdif_J3_VH_sqrdif_lowest_cor_locations}(j)

\caption{{\sc VH Polarization Channel} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$. Red horizontal line represents the median value and yellow line represents two absolute median deviations beyond the median. Squared approximation coefficient deviations: 0.1\% highest absolute correlations - (c) $\vd(m)$;  (d) $\vt(m)$; 0.1\% smallest absolute correlations - (e) $\vd(m)$; (f) $\vt(m)$; 0.01\% highest absolute correlations - (g) $\vd(m)$;  (h) $\vt(m)$;  0.01\% smallest absolute correlations - (i) $\vd(m)$; (j)  $\vt(m)$. 
}  
\label{F:squared_meandev_J3_VH}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=.12]{J3_euclid_squared_meandev}(a)
\includegraphics[scale=.12]{consecdif_J3_euclid_squared_meandev}(b)\\
\includegraphics[scale=.12]{J3_euclid_sqrdif_change_locations}(c)
\includegraphics[scale=.12]{consecdif_J3_euclid_sqrdif_change_locations}(d)\\
\includegraphics[scale=.12]{J3_euclid_sqrdif_lowest_cor_locations}(e)
\includegraphics[scale=.12]{consecdif_J3_euclid_sqrdif_lowest_cor_locations}(f)\\
\includegraphics[scale=.12]{perc01_J3_euclid_sqrdif_change_locations}(g)
\includegraphics[scale=.12]{perc01_consecdif_J3_euclid_sqrdif_change_locations}(h)\\
\includegraphics[scale=.12]{perc01_J3_euclid_sqrdif_lowest_cor_locations}(i)
\includegraphics[scale=.12]{perc01_consecdif_J3_euclid_sqrdif_lowest_cor_locations}(j)
\caption{{\sc Combined Channels} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$. Red horizontal line represents the median value and yellow line represents two absolute median deviations beyond the median. Squared approximation coefficient deviations: 0.1\% highest absolute correlations - (c) $\vd(m)$;  (d) $\vt(m)$; 0.1\% smallest absolute correlations - (e) $\vd(m)$; (f) $\vt(m)$; 0.01\% highest absolute correlations - (g) $\vd(m)$;  (h) $\vt(m)$;  0.01\% smallest absolute correlations - (i) $\vd(m)$; (j)  $\vt(m)$. 
} 
\label{F:squared_meandev_J3_euclid}
\end{figure}

Figures \ref{F:squared_meandev_J3_VV}-\ref{F:squared_meandev_J3_euclid} present a timewise comparison between overall energy variations, $\vd(m)$ and $\vt(m)$ and their respective individual coefficients. Each figure has ten panels. Panels on the left and right deal with the first and second change detection methods respectively. A full description is give in each figure's caption. The overall conclusion from these figures is that as expected there are temporal variations between the methods regarding change detection (Panels (a)-(b)). The correlation screening detects the most relevant indices, as well as the least relevant ones  (Panels (c)-(f)). 
These conclusions hold for analyses based upon VV, VH or combined polarizations, but the VV polarization signal is much stronger than VH's. A slight advantage is perceived for the $\vd(m)$ method as opposed to $\vt(m)$'s. This makes sense, since we are dealing here with data from a forest region over a long time, and seasonal changes will be perceived more easily on the first proposed method.

\begin{table*}[ht!]
\caption{Absolute correlation thresholds and number of selected coefficients for $n=85$ multi-temporal images of $1200\times 1000$ . Correlation was computed between approximation coefficients and approximation total energy at level $J=2$ for each image. }\label{tabela_quantis}
\centering
\begin{tabular}{c|cccr||c|cccr}
\hline
Qtile &\multicolumn{4}{c||}{\sc Correlation Thresh} &Qtile & \multicolumn{4}{c}{\sc Correlation Thresh}  \\
Level &VV&VH&Comb& Coeffs &Level &VV&VH&Comb& Coeffs \\
\hline
0.50&0.281&0.199&0.292&600000&0.99&0.647&0.582&0.668&12000\\
0.55&0.300&0.218&0.311&540000&0.991&0.651&0.587&0.673&10800\\
0.60&0.319&0.239&0.331&480000& 0.992&0.656&0.594&0.678&9600\\
0.65&0.339&0.260&0.352&420000& 0.993&0.662&0.601&0.684&8400\\
0.70&0.361&0.282&0.375&360000& 0.994&0.669&0.608&0.690&7200\\
0.75&0.385&0.307&0.399&300000& 0.995&0.676&0.617&0.697&6000\\
0.80&0.412&0.335&0.428&240000& 0.996&0.685&0.626&0.705&4800\\
0.85&0.445&0.368&0.462&180000& 0.997&0.696&0.639&0.716&3600\\
0.90&0.487&0.410&0.506&120000& 0.998&0.710&0.654&0.729&2400\\
0.95&0.549&0.472&0.570&60000& 0.999&0.731&0.676&0.748&1200\\
\hline
\end{tabular}
\end{table*}

\begin{figure}[htp!]
\includegraphics[scale=.12]{J2_VV_image_cor}(a)
\includegraphics[scale=.12]{J2_VH_image_cor}(b)\\
\includegraphics[scale=.12]{J2_euclid_image_cor}(c)
\includegraphics[scale=.12]{J3_VV_image_cor}(d)\\ 
\includegraphics[scale=.12]{J3_VH_image_cor}(e)
\includegraphics[scale=.12]{J3_euclid_image_cor}(f)

\caption{Absolute correlation matrices $\vR^{(d)}$ between mean-corrected squared coefficients at approximation levels and overall mean-corrected total energy $\{\vd(m)\}$.  $J=2$ (a)-(c);  $J=3$ (d)-(f). $n=85$ multi-temporal images of $1200\times 1000$ ($J=10$). The color bar on the right gives the magnitude of correlation at all positions. (a) {\sc VV  Polarization - Channel} $J=2$. (b) {\sc VH  Polarization - Channel} $J=2$. (c) {\sc Combined Channels} $J=2$.  (d)  {\sc VV  Polarization - Channel} $J=3$. (e) {\sc VH  Polarization - Channel} $J=3$. (f) {\sc Combined Channels} $J=3$} 
\label{F:image_corr_J2J3}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=.12]{consecdif_J2_VV_image_cor}(a)
\includegraphics[scale=.12]{consecdif_J2_VH_image_cor}(b)\\
\includegraphics[scale=.12]{consecdif_J2_euclid_image_cor}(c)
\includegraphics[scale=.12]{consecdif_J3_VV_image_cor}(d)\\ 
\includegraphics[scale=.12]{consecdif_J3_VH_image_cor}(e)
\includegraphics[scale=.12]{consecdif_J3_euclid_image_cor}(f)

\caption{Absolute correlation matrices $\vR^{(t)}$ between squared differences of consecutive log-images' coefficients at approximation levels and their overall total energy $\{\vt(m)\}$. Levels $J=2$ (a)-(c);  $J=3$ (d)-(f). $n=85$ multi-temporal images of $1200\times 1000$. The color bar on the right gives the magnitude of correlation at all positions. (a) {\sc VV  Polarization - Channel} $J=2$. (b) {\sc VH  Polarization - Channel} $J=2$. (c) {\sc Combined Channels} $J=2$. (d)  {\sc VV  Polarization - Channel} $J=3$. (e) {\sc VH  Polarization - Channel} $J=3$. (f) {\sc Combined Channels} $J=3$.} 
\label{F:image_corr_J2J3_consec}
\end{figure}

Figures \ref{F:image_corr_J2J3}-\ref{F:image_corr_J2J3_consec} shows the absolute correlation images for the VV, VH and combined channels for levels $J=2$ and $J=3$. We can notice that high correlation coefficients for $J=2$ are high correlation coefficients for $J=3$ as well. Moreover, there is a clear spatial connection between polarizations and between $\vd(m)$- and $\vt(m)$-based  analyses. On the other hand, correlations are more efficiently segregated when we move: from $J=2$ to $J=3$; from VH to VV polarization; or from  $\vt(m)$ to $\vd(m)$.

\section{Discussion}\label{section_discussion}

We present a novel way of detecting changes in multi-temporal satellite images, WECS. The procedure is based on wavelet energies from both the estimated individual coefficients as well as the whole image approximation. It makes use of correlation screening for ultra-high dimensional data. The proposed method's performance is shown using both synthetic and real data. The proposed method yields spatio-temporal  change points. Its performance is similar to a deep-learning feature extraction method, but the computational cost of the proposed method is 180 times smaller than the cost for the latter. Therefore, we may say that WECS may be used on images with equivalent performance for a fraction of the computational cost. Because of its reliance on wavelet representation and correlation screening, it is sparse, very fast and scalable. Finally, it is easily adapted to be updatable, so that real-time change detection is feasible even with a portable computer.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

%
%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

% use section* for acknowledgment
%\section*{Acknowledgment}

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{bibfile}
%
%% <OR> manually copy in the resultant .bbl file
%% set second argument of \begin to the number of references
%% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}
%[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{../Photo/rodney}}]{Rodney Fonseca}
%was born in Brazil, in 1993. He received the Master’s
%degrees in Statistics from the Federal University of Pernambuco,
%Recife, Brazil, in 2017, and the Ph.D. degree in Statistics, from Campinas State University, Campinas, Brazil, in 2021.
%His research interests include nonparametric statistics, time series, graph signal processing and regression models.
%\end{IEEEbiography}
%\vfill
%
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{../Photo/aluisio}}]{Alu\'{i}sio~Pinheiro}
%was born in Brazil, in 1967. He received the Master’s degree in Statistics from Campinas State University, Campinas, Brazil, in 1992, and the Ph.D. degree in Statistics from University of North Carolina, Chapel Hill, US, in 1997.
%He is currently Professor at Campinas State University. His research interests concern U-statistics, wavelets, nonparametric statistics and time series.
%Dr. Pinheiro was the 2012 Pranab Kumar Sen Distinguished Visiting Professor of Biostatistics, UNC-Chapel Hill. He has been the Brazilian Statistics Association General Secretary (2010-2012) and Treasurer (2018-2020).
%\end{IEEEbiography}
%
%
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{../Photo/abdou}}]{Abdourrahmane Mahamane Atto}
%was born in Niger, in 1974. He received the Master’s
%degrees in applied mathematics and pure mathematics from the University of Abomey-Calavi,
%Cotonou, Benin, 2001 and 2002, respectively, the
%Master’s degree of advanced studies in electrical engineering from the École Polytechnique d’Abomey-Calavi, Cotonou, in 2002, and the Ph.D. degree
%in applied mathematics, codelivered by TELECOM
%Bretagne, Brest, France, and the University of
%Rennes I, Rennes, France, in 2008.
%He worked with École Polytechnique d’Abomey-Calavi, from 2002 to 2003,
%Ecole des Mines, de l’Industrie et de la Géologie, Niamey, Niger, from 2003
%to 2004, TELECOM Bretagne, from 2008 to 2009, and Institut Polytechnique
%de Bordeaux, Talence, France, from 2009 to 2011. Since September 2011, he
%has been an Associate Professor at the University of Savoie, Polytech Annecy-Chambéry, Annecy, France. His research interests concern mathematical methods and models for signal, image, and information processing.
%\end{IEEEbiography}
%\vfill

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2021-10-22 tex/paper.wecs/wecs_tgrs.tex
update figures in the tex

@@
 %\includegraphics[width=40pc]{EllipsoidChanges/ellipses_all}
-\includegraphics[scale=.17]{ellipses_t1}
-\includegraphics[scale=.17]{ellipses_t2}
-\includegraphics[scale=.17]{ellipses_t3}
-\includegraphics[scale=.17]{ellipses_t4}
+\includegraphics[scale=.17]{../../figs/ellipses_t1}
+\includegraphics[scale=.17]{../../figs/ellipses_t2}
+\includegraphics[scale=.17]{../../figs/ellipses_t3}
+\includegraphics[scale=.17]{../../figs/ellipses_t4}
 \caption{Synthetic multi-temporal ($n=4$) images. Features and changes come as ellipses and dots.}
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.1]{total_changes}(a)
-\includegraphics[scale=.1]{corr_changes_dm}(b)
-\includegraphics[scale=.1]{corr_changes_tm}(c)
-\includegraphics[scale=.1]{corr_changes_logratios}(d)
-\includegraphics[scale=.1]{corr_changes_dm_nowavelets}(e)
-\includegraphics[scale=.1]{corr_changes_tm_nowavelets}(f)
+\includegraphics[scale=.1]{../../figs/total_changes}(a)
+\includegraphics[scale=.1]{../../figs/corr_changes_dm}(b)
+\includegraphics[scale=.1]{../../figs/corr_changes_tm}(c)
+\includegraphics[scale=.1]{../../figs/corr_changes_logratios}(d)
+\includegraphics[scale=.1]{../../figs/corr_changes_dm_nowavelets}(e)
+\includegraphics[scale=.1]{../../figs/corr_changes_tm_nowavelets}(f)
 \caption{Synthetic images with changing ellipses. (a) Image composed by the total changes over time. 
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.13]{methods_comparison}\hspace{-.5cm}(a) 
-\includegraphics[scale=.13]{levels_comparison_WithReference}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{families_comparison_WithReference}\hspace{-.5cm}(c)
-\includegraphics[scale=.13]{dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
-\includegraphics[scale=.13]{levels_comparison_NoReference}\hspace{-.5cm}(e)
-\includegraphics[scale=.13]{families_comparison_NoReference}\hspace{-.5cm}(f)
+\includegraphics[scale=.13]{../../figs/methods_comparison}\hspace{-.5cm}(a) 
+\includegraphics[scale=.13]{../../figs/levels_comparison_WithReference}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.13]{../../figs/families_comparison_WithReference}\hspace{-.5cm}(c)
+\includegraphics[scale=.13]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
+\includegraphics[scale=.13]{../../figs/levels_comparison_NoReference}\hspace{-.5cm}(e)
+\includegraphics[scale=.13]{../../figs/families_comparison_NoReference}\hspace{-.5cm}(f)
 \caption{ROC curves for detection of changing ellipses in synthetic images and different methods.
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{J1_VV_squared_meandev}(a)
-\includegraphics[scale=.12]{consecdif_J1_VV_squared_meandev}(b)\\
-\includegraphics[scale=.12]{J2_VV_squared_meandev}(c)
-\includegraphics[scale=.12]{consecdif_J2_VV_squared_meandev}(d)\\
-\includegraphics[scale=.12]{J3_VV_squared_meandev}(e)
-\includegraphics[scale=.12]{consecdif_J3_VV_squared_meandev}(f)\\
-\includegraphics[scale=.12]{J4_VV_squared_meandev}(g)
-\includegraphics[scale=.12]{consecdif_J4_VV_squared_meandev}(h)
+\includegraphics[scale=.12]{../../figs/J1_VV_squared_meandev}(a)
+\includegraphics[scale=.12]{../../figs/consecdif_J1_VV_squared_meandev}(b)\\
+\includegraphics[scale=.12]{../../figs/J2_VV_squared_meandev}(c)
+\includegraphics[scale=.12]{../../figs/consecdif_J2_VV_squared_meandev}(d)\\
+\includegraphics[scale=.12]{../../figs/J3_VV_squared_meandev}(e)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_squared_meandev}(f)\\
+\includegraphics[scale=.12]{../../figs/J4_VV_squared_meandev}(g)
+\includegraphics[scale=.12]{../../figs/consecdif_J4_VV_squared_meandev}(h)
 \caption{{\sc VV Polarization Channel} Series of squared deviations $\vd(m)$ and $\vt(m)$.
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{J1_VH_squared_meandev}(a)
-\includegraphics[scale=.12]{consecdif_J1_VH_squared_meandev}(b)\\
-\includegraphics[scale=.12]{J2_VH_squared_meandev}(c)
-\includegraphics[scale=.12]{consecdif_J2_VH_squared_meandev}(d)\\
-\includegraphics[scale=.12]{J3_VH_squared_meandev}(e)
-\includegraphics[scale=.12]{consecdif_J3_VH_squared_meandev}(f)\\
-\includegraphics[scale=.12]{J4_VH_squared_meandev}(g)
-\includegraphics[scale=.12]{consecdif_J4_VH_squared_meandev}(h)
+\includegraphics[scale=.12]{../../figs/J1_VH_squared_meandev}(a)
+\includegraphics[scale=.12]{../../figs/consecdif_J1_VH_squared_meandev}(b)\\
+\includegraphics[scale=.12]{../../figs/J2_VH_squared_meandev}(c)
+\includegraphics[scale=.12]{../../figs/consecdif_J2_VH_squared_meandev}(d)\\
+\includegraphics[scale=.12]{../../figs/J3_VH_squared_meandev}(e)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_squared_meandev}(f)\\
+\includegraphics[scale=.12]{../../figs/J4_VH_squared_meandev}(g)
+\includegraphics[scale=.12]{../../figs/consecdif_J4_VH_squared_meandev}(h)
 \caption{{\sc VH Polarization  Channel} Series of squared deviations $\vd(m)$ and $\vt(m)$.
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{J1_euclid_squared_meandev}(a)
-\includegraphics[scale=.12]{consecdif_J1_euclid_squared_meandev}(b)\\
-\includegraphics[scale=.12]{J2_euclid_squared_meandev}(c)
-\includegraphics[scale=.12]{consecdif_J2_euclid_squared_meandev}(d)\\
-\includegraphics[scale=.12]{J3_euclid_squared_meandev}(e)
-\includegraphics[scale=.12]{consecdif_J3_euclid_squared_meandev}(f)\\
-\includegraphics[scale=.12]{J4_euclid_squared_meandev}(g)
-\includegraphics[scale=.12]{consecdif_J4_euclid_squared_meandev}(h)
+\includegraphics[scale=.12]{../../figs/J1_euclid_squared_meandev}(a)
+\includegraphics[scale=.12]{../../figs/consecdif_J1_euclid_squared_meandev}(b)\\
+\includegraphics[scale=.12]{../../figs/J2_euclid_squared_meandev}(c)
+\includegraphics[scale=.12]{../../figs/consecdif_J2_euclid_squared_meandev}(d)\\
+\includegraphics[scale=.12]{../../figs/J3_euclid_squared_meandev}(e)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_squared_meandev}(f)\\
+\includegraphics[scale=.12]{../../figs/J4_euclid_squared_meandev}(g)
+\includegraphics[scale=.12]{../../figs/consecdif_J4_euclid_squared_meandev}(h)
 \caption{{\sc Combined Channels} Series of squared deviations $\vd(m)$ and $\vt(m)$.
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{J3_VV_squared_meandev}(a)
-\includegraphics[scale=.12]{consecdif_J3_VV_squared_meandev}(b)\\
-\includegraphics[scale=.12]{J3_VV_sqrdif_change_locations}(c)
-\includegraphics[scale=.12]{consecdif_J3_VV_sqrdif_change_locations}(d)\\
-\includegraphics[scale=.12]{J3_VV_sqrdif_lowest_cor_locations}(e)
-\includegraphics[scale=.12]{consecdif_J3_VV_sqrdif_lowest_cor_locations}(f)\\
-\includegraphics[scale=.12]{perc01_J3_VV_sqrdif_change_locations}(g)
-\includegraphics[scale=.12]{perc01_consecdif_J3_VV_sqrdif_change_locations}(h)\\
-\includegraphics[scale=.12]{perc01_J3_VV_sqrdif_lowest_cor_locations}(i)
-\includegraphics[scale=.12]{perc01_consecdif_J3_VV_sqrdif_lowest_cor_locations}(j)
+\includegraphics[scale=.12]{../../figs/J3_VV_squared_meandev}(a)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_squared_meandev}(b)\\
+\includegraphics[scale=.12]{../../figs/J3_VV_sqrdif_change_locations}(c)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_sqrdif_change_locations}(d)\\
+\includegraphics[scale=.12]{../../figs/J3_VV_sqrdif_lowest_cor_locations}(e)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_sqrdif_lowest_cor_locations}(f)\\
+\includegraphics[scale=.12]{../../figs/perc01_J3_VV_sqrdif_change_locations}(g)
+\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VV_sqrdif_change_locations}(h)\\
+\includegraphics[scale=.12]{../../figs/perc01_J3_VV_sqrdif_lowest_cor_locations}(i)
+\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VV_sqrdif_lowest_cor_locations}(j)
 \caption{{\sc VV Polarization Channel} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$.
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{J3_VH_squared_meandev}(a)
-\includegraphics[scale=.12]{consecdif_J3_VH_squared_meandev}(b)
-
-\includegraphics[scale=.12]{J3_VH_sqrdif_change_locations}(c)
-\includegraphics[scale=.12]{consecdif_J3_VH_sqrdif_change_locations}(d)
-
-\includegraphics[scale=.12]{J3_VH_sqrdif_lowest_cor_locations}(e)
-\includegraphics[scale=.12]{consecdif_J3_VH_sqrdif_lowest_cor_locations}(f)
-
-\includegraphics[scale=.12]{perc01_J3_VH_sqrdif_change_locations}(g)
-\includegraphics[scale=.12]{perc01_consecdif_J3_VH_sqrdif_change_locations}(h)
-
-\includegraphics[scale=.12]{perc01_J3_VH_sqrdif_lowest_cor_locations}(i)
-\includegraphics[scale=.12]{perc01_consecdif_J3_VH_sqrdif_lowest_cor_locations}(j)
+\includegraphics[scale=.12]{../../figs/J3_VH_squared_meandev}(a)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_squared_meandev}(b)
+
+\includegraphics[scale=.12]{../../figs/J3_VH_sqrdif_change_locations}(c)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_sqrdif_change_locations}(d)
+
+\includegraphics[scale=.12]{../../figs/J3_VH_sqrdif_lowest_cor_locations}(e)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_sqrdif_lowest_cor_locations}(f)
+
+\includegraphics[scale=.12]{../../figs/perc01_J3_VH_sqrdif_change_locations}(g)
+\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VH_sqrdif_change_locations}(h)
+
+\includegraphics[scale=.12]{../../figs/perc01_J3_VH_sqrdif_lowest_cor_locations}(i)
+\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VH_sqrdif_lowest_cor_locations}(j)
 
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{J3_euclid_squared_meandev}(a)
-\includegraphics[scale=.12]{consecdif_J3_euclid_squared_meandev}(b)\\
-\includegraphics[scale=.12]{J3_euclid_sqrdif_change_locations}(c)
-\includegraphics[scale=.12]{consecdif_J3_euclid_sqrdif_change_locations}(d)\\
-\includegraphics[scale=.12]{J3_euclid_sqrdif_lowest_cor_locations}(e)
-\includegraphics[scale=.12]{consecdif_J3_euclid_sqrdif_lowest_cor_locations}(f)\\
-\includegraphics[scale=.12]{perc01_J3_euclid_sqrdif_change_locations}(g)
-\includegraphics[scale=.12]{perc01_consecdif_J3_euclid_sqrdif_change_locations}(h)\\
-\includegraphics[scale=.12]{perc01_J3_euclid_sqrdif_lowest_cor_locations}(i)
-\includegraphics[scale=.12]{perc01_consecdif_J3_euclid_sqrdif_lowest_cor_locations}(j)
+\includegraphics[scale=.12]{../../figs/J3_euclid_squared_meandev}(a)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_squared_meandev}(b)\\
+\includegraphics[scale=.12]{../../figs/J3_euclid_sqrdif_change_locations}(c)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_sqrdif_change_locations}(d)\\
+\includegraphics[scale=.12]{../../figs/J3_euclid_sqrdif_lowest_cor_locations}(e)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_sqrdif_lowest_cor_locations}(f)\\
+\includegraphics[scale=.12]{../../figs/perc01_J3_euclid_sqrdif_change_locations}(g)
+\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_euclid_sqrdif_change_locations}(h)\\
+\includegraphics[scale=.12]{../../figs/perc01_J3_euclid_sqrdif_lowest_cor_locations}(i)
+\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_euclid_sqrdif_lowest_cor_locations}(j)
 \caption{{\sc Combined Channels} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$.
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{J2_VV_image_cor}(a)
-\includegraphics[scale=.12]{J2_VH_image_cor}(b)\\
-\includegraphics[scale=.12]{J2_euclid_image_cor}(c)
-\includegraphics[scale=.12]{J3_VV_image_cor}(d)\\ 
-\includegraphics[scale=.12]{J3_VH_image_cor}(e)
-\includegraphics[scale=.12]{J3_euclid_image_cor}(f)
+\includegraphics[scale=.12]{../../figs/J2_VV_image_cor}(a)
+\includegraphics[scale=.12]{../../figs/J2_VH_image_cor}(b)\\
+\includegraphics[scale=.12]{../../figs/J2_euclid_image_cor}(c)
+\includegraphics[scale=.12]{../../figs/J3_VV_image_cor}(d)\\ 
+\includegraphics[scale=.12]{../../figs/J3_VH_image_cor}(e)
+\includegraphics[scale=.12]{../../figs/J3_euclid_image_cor}(f)
 
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{consecdif_J2_VV_image_cor}(a)
-\includegraphics[scale=.12]{consecdif_J2_VH_image_cor}(b)\\
-\includegraphics[scale=.12]{consecdif_J2_euclid_image_cor}(c)
-\includegraphics[scale=.12]{consecdif_J3_VV_image_cor}(d)\\ 
-\includegraphics[scale=.12]{consecdif_J3_VH_image_cor}(e)
-\includegraphics[scale=.12]{consecdif_J3_euclid_image_cor}(f)
+\includegraphics[scale=.12]{../../figs/consecdif_J2_VV_image_cor}(a)
+\includegraphics[scale=.12]{../../figs/consecdif_J2_VH_image_cor}(b)\\
+\includegraphics[scale=.12]{../../figs/consecdif_J2_euclid_image_cor}(c)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_image_cor}(d)\\ 
+\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_image_cor}(e)
+\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_image_cor}(f)
 

### 2021-10-25 codes/EllispseChangeAnalysis_Deep.m
Codes to analyze synthetic images

@@ -1,0 +1,154 @@
+
+%%
+close all
+clear;
+%%
+im1 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');  
+im2 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
+im3 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
+im4 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
+%
+figure
+imshow(im1)
+figure
+imshow(im2)
+figure
+imshow(im3)
+figure
+imshow(im4)
+%
+% Images are too big, consider subsampling
+subsamplingfactor = 2;
+im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
+totalchanges = imread('GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
+totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
+%
+figure
+imshow(totalchanges)
+title('Total changes')
+
+%%
+eps = .00001;
+im1log = log(eps+double(im1));
+im2log = log(eps+double(im2));
+im3log = log(eps+double(im3));
+im4log = log(eps+double(im4));
+n=4;
+imRef = im1log + im2log + im3log + im4log;
+    imRef = imRef/n;
+%% comparison of wavelet denoising with mean image
+
+wname = 'db2';
+J = 2;
+
+tic
+[X1,~,~,~] = swt2(im1log,J,wname);
+    X1 = X1(:,:,J); % focus only on level J approximations ?
+[X2,~,~,~] = swt2(im2log,J,wname);
+    X2 = X2(:,:,J); % focus only on level J approximations ?
+[X3,~,~,~] = swt2(im3log,J,wname);
+    X3 = X3(:,:,J); % focus only on level J approximations ?
+[X4,~,~,~] = swt2(im4log,J,wname);
+    X4 = X4(:,:,J); % focus only on level J approximations ?
+toc
+
[…]

### 2021-10-25 codes/EllispseChangeAnalysis_families.m

@@ -1,0 +1,326 @@
+
+% Evaluation of wavelet bases when mean approximation coefficients are used
+
+%%
+close all
+clear;
+%%
+im1 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');  
+im2 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
+im3 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
+im4 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
+%
+figure
+imshow(im1)
+figure
+imshow(im2)
+figure
+imshow(im3)
+figure
+imshow(im4)
+%
+% Images are too big, consider subsampling
+subsamplingfactor = 2;
+im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
+totalchanges = imread('GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
+totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
+%
+figure
+imshow(totalchanges)
+title('Total changes')
+
+%%
+eps = .00001;
+im1log = log(eps+double(im1));
+im2log = log(eps+double(im2));
+im3log = log(eps+double(im3));
+im4log = log(eps+double(im4));
+n=4;
+imRef = im1log + im2log + im3log + im4log;
+    imRef = imRef/n;
+
+% decomposition level fixed
+J = 2;
+
+%% Case 1 : reference image 
+
+% Wavelet basis: Haar
+wname = 'haar';
+
+[X1,~,~,~] = swt2(im1log,J,wname);
+    X1 = X1(:,:,J); % focus only on level J approximations ?
+[X2,~,~,~] = swt2(im2log,J,wname);
+    X2 = X2(:,:,J); % focus only on level J approximations ?
+[X3,~,~,~] = swt2(im3log,J,wname);
[…]

### 2021-10-25 codes/EllispseChangeAnalysis_levels.m

@@ -1,0 +1,283 @@
+
+% Evaluation of different levels when mean approximation coefficients are used
+
+%%
+close all
+clear;
+%%
+im1 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');  
+im2 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
+im3 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
+im4 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
+%
+figure
+imshow(im1)
+figure
+imshow(im2)
+figure
+imshow(im3)
+figure
+imshow(im4)
+%
+% Images are too big, consider subsampling
+subsamplingfactor = 1;
+im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
+totalchanges = imread('GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
+totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
+%
+figure
+imshow(totalchanges)
+title('Total changes')
+
+%%
+eps = .00001;
+im1log = log(eps+double(im1));
+im2log = log(eps+double(im2));
+im3log = log(eps+double(im3));
+im4log = log(eps+double(im4));
+n=4;
+imRef = im1log + im2log + im3log + im4log;
+    imRef = imRef/n;
+
+% decomposition level fixed
+wname = 'db2';
+
+%% Case 1 : reference image 
+
+% Wavelet basis: Haar
+J = 1;
+
+[X1,~,~,~] = swt2(im1log,J,wname);
+    X1 = X1(:,:,J); % focus only on level J approximations ?
+[X2,~,~,~] = swt2(im2log,J,wname);
+    X2 = X2(:,:,J); % focus only on level J approximations ?
+[X3,~,~,~] = swt2(im3log,J,wname);
[…]

### 2021-10-25 codes/EllispseChangeAnalysis_methods.m

@@ -1,0 +1,165 @@
+
+% Evaluation of different methods
+
+%%
+close all
+clear;
+%%
+im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
+im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
+im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
+im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
+%
+figure
+imshow(im1)
+figure
+imshow(im2)
+figure
+imshow(im3)
+figure
+imshow(im4)
+
+figure
+subplot(1,4,1); imshow(im1); title('I(1)')
+subplot(1,4,2); imshow(im2); title('I(2)')
+subplot(1,4,3); imshow(im3); title('I(3)')
+subplot(1,4,4); imshow(im4); title('I(4)')
+
+%
+% Images are too big, consider subsampling
+subsamplingfactor = 2;
+im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
+im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
+totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
+totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
+%
+mImage = figure;
+imshow(totalchanges)
+title('Total changes', 'FontSize', 17)
+%saveas(mImage,sprintf('total_changes.jpg'))
+
+%%
+eps = .00001;
+im1log = log(eps+double(im1));
+im2log = log(eps+double(im2));
+im3log = log(eps+double(im3));
+im4log = log(eps+double(im4));
+n=4;
+imRef = im1log + im2log + im3log + im4log;
+imRef = imRef/n;
+
+%% comparison with mean image
+
+wname = 'db2';
+J = 2;
+
+[X1,~,~,~] = swt2(im1log,J,wname);
+    X1 = X1(:,:,J); % focus only on level J approximations ?
+[X2,~,~,~] = swt2(im2log,J,wname);
+    X2 = X2(:,:,J); % focus only on level J approximations ?
[…]

### 2021-10-25 codes/F1Scorecurve.m

@@ -1,0 +1,46 @@
+function [pD,pFA]=F1Scorecurve(img,im0)
+
+lxy=numel(im0);
+% Nombre reel de pixel ayant change
+CP=find(im0~=0);
+nCP=numel(CP);
+%%%%%%%%%%%%%%%%
+u=0;
+v=1;
+pas=(v-u)/100;
+t=u:pas:v;
+%t=u+2*pas:pas:v-2*pas;
+lt=numel(t);
+pD=zeros(1,lt);
+pFA=pD;
+for k=1:lt
+    %imtemp=255.*im2bw(img,k/T);
+    imtemp=255.*double(img>t(k));
+    % Detection
+    D=find(imtemp==255);
+    nD=numel(D);
+    % True positive
+    imgdiff=2*imtemp-im0;
+    TP=find(imgdiff==255);
+    nTP=numel(TP);
+    % Precision
+    pD(k)=nTP/nD;
+    % Recall
+    FN=find(imgdiff==-255);
+    nFN=numel(FN);
+    pD(k)=nTP/(nBD + );
+    % Probabilite de fausse alarm
+    pFA(k)=(nD-nTP)/(lxy-nCP);
+end
+
+%y=reshape(imtemp,1,lxy);
+%z=reshape(im0,1,lxy);
+
+%%% Plot ROC curves
+figure
+plot(pFA,pD);
+title('ROC Curve')
+xlabel('False positive rate')
+ylabel('True positive rate')
+axis([0 1 0 1]);
+axis square

### 2021-10-25 codes/ROCcurveNew.m

@@ -1,0 +1,43 @@
+function [pD,pFA]=ROCcurveNew(img,im0)
+
+lxy=numel(im0);
+% Nombre reel de pixel ayant change
+CP=find(im0~=0);
+nCP=numel(CP);
+%%%%%%%%%%%%%%%%
+u=min(min(img));
+%u=0;
+v=max(max(img));
+pas=(v-u)/100;
+t=u:pas:v;
+%t=u+2*pas:pas:v-2*pas;
+lt=numel(t);
+pD=zeros(1,lt);
+pFA=pD;
+for k=1:lt
+    %imtemp=255.*im2bw(img,k/T);
+    imtemp=255.*double(img>t(k));
+    % Detection
+    D=find(imtemp==255);
+    nD=numel(D);
+    % Bonne detection
+    imgdiff=2*imtemp-im0;
+    BD=find(imgdiff==255);
+    nBD=numel(BD);
+    % Probabilte de detection
+    pD(k)=nBD/nCP;
+    % Probabilite de fausse alarm
+    pFA(k)=(nD-nBD)/(lxy-nCP);
+end
+
+%y=reshape(imtemp,1,lxy);
+%z=reshape(im0,1,lxy);
+
+%%% Plot ROC curves
+figure
+plot(pFA,pD);
+title('ROC Curve')
+xlabel('False positive rate')
+ylabel('True positive rate')
+axis([0 1 0 1]);
+axis square

### 2021-10-26 tex/paper.wecs/wecs_tgrs.tex
Updates on latex
updates on latex

@@
 % Do not put math or special symbols in the title.
-\title{Wavelet Spatio-Temporal Change Detection on multi-temporal POLSAR images}
+\title{Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images}
 %
@@
 
-\author{Rodney~Fonseca,
+\author{Rodney~V.~Fonseca, Rog\'{e}rio~G.~Negri,
         Alu\'{i}sio~Pinheiro,
@@
 \begin{abstract}
-We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised sparse procedure to detect spatio-temporal change points on multi-temporal Polarimetric SAR (POLSAR) images or even on sequences of very high resolution images.
+We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised sparse procedure to detect spatio-temporal change points on multi-temporal SAR images or even on sequences of very high resolution images.
 The procedure is based on multiscale approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients.
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{W}{e} propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR/POLSAR images.
+\IEEEPARstart{W}{e} propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images.
 WECS is based upon correlation screening for energy apportionment on wavelet approximations.
@@
 
-Let $\mathcal{I}(1),\ldots,\mathcal{I}(m)$ be a set of matrices representing the log-images of some region of interest.
-These images may be relative to one POLSAR channel or a combination of channels; this will be specified when appropriate.
+Let $\mathcal{I}(1),\ldots,\mathcal{I}(m)$ be a set of matrices representing the images of some region of interest.
+These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate.
 Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}(m)$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
-We shall address these tasks by analyzing the bidimensional discrete wavelet decomposition of $\mathcal{I}(m)$ at the $J$-th approximation level, i.e.,
-such that
+We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}(m)$.
+Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}.
+After application of this wavelet transform to $\mathcal{I}(m)$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX(m)$, a smooth version of $\mathcal{I}(m)$ with the same dimension.
+The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX(m)$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}(m)$.
+Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX(m)$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
+
+We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX(m)\right\}$ as
 \begin{equation} 
-\mathcal{I}(m)=\mathcal{I}_J(m)+\vepsilon_J(m),
+\sum_{m=1}^n\|\vX(m)\|^2=n\|\bar{\vX}\|^2+\sum_{m=1}^n\|\mathcal \vX(m)-\bar{\vX}\|^2,\label{E:eq_ANOVAlogimage}
 \end{equation}
-%\label{eq_modellogimage}
-where $\mathcal{I}_J(m)$ is the wavelet approximation of log-image $\mathcal{I}(m)$ at level $J$.
-We denote $\mathcal{I}_J(m)$'s approximation coefficients matrix by $\vX(m)$ \cite{morettin2017wavelets,vidakovic1999statistical}.
-The advantages of doing so is that we linearize possible speckle (multiplicative) noise in the original images by taking logarithms to obtain $\mathcal{I}(m)$ and later perform a smoothing to obtain $\mathcal{I}_J(m)$, whose approximation coefficients matrix is $\vX(m)$.
-Such approach is also described in the literature as multiplicative wavelet transform \cite{atto2016wavelet}.
-
-We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\mathcal I(m)\right\}$ as
-\begin{equation} 
-\sum_{m=1}^n\|\mathcal I_J(m)\|^2_2=n\|\bar{\mathcal I}\|^2_2+\sum_{m=1}^n\|\mathcal I_J(m)-\bar{\mathcal I}\|^2_2,\label{eq_ANOVAlogimage}
-\end{equation}
-where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}_J(m)$. The same can be done on the wavelet domain with $\vX(m)$, which is the way we shall proceed in the following steps.
-
-We discuss here two procedures which are similar, but may yield different results depending on the nature of relevant changes.
-The first procedure makes use of (\ref{eq_ANOVAlogimage}), so that we establish an {\it average wavelet-approximated image} and detect time points for which images differ from the {\it characteristic} image.
-Changes are detected by absolute correlations between individual wavelet approximation coefficients time series and the time series of overall wavelet energy.
-The second procedure also uses absolute correlations but instead of an average wavelet image energy, we compute  the overall energy differences for subsequent wavelet-smoothed images and their approximation coefficients.
-The former should help us detect image changes from an overall behavior over time whilst the latter should also detect changes on non-stationary set-ups.
-In practive, both procedures shall be performed in the wavelet domain by using $\{\vX(m)\}$ instead of $\{\mathcal{I}_J(m)\}$.
-
-The average image $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}_J(m)$ has approximation coefficients matrix given by $\bar{\vX} = n^{-1}\sum_{m=1}^n \vX(m)$.
+where $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX(m)$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
+The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX(m)$ compared to an {\it average wavelet-approximated image} $\bar{\vX}$, what motivates us to such deviations to detect change points in time.
+Since each element (pixel) in $\vX(m)$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space.
+Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further.
+Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX(m)-\vX(m-1)\|^2$, $d(\vX(m),\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
+
 Let $X_{k,l}(m)$ and $\bar{X}_{k,l}$ be the entry $(k,l)$ of the matrices $\vX(m)$ and $\bar{\vX}$, respectively.
-We take the matrix $\vD(m) = [D_{k.l}(m)]$, where $D_{k,l}(m)=(X_{k,l}(m)-\bar{X}_{k,l})^2$.
+Computing $D_{k,l}(m)=(X_{k,l}(m)-\bar{X}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD(m) = [D_{k.l}(m)]$.
 We then
@@
 \vd(m)= \sum_{k,l}D_{k,l}(m) = \sum_{k,l}(X_{k,l}(m) - \bar{X}_{k,l})^2,
-\label{eq_defdm}
+\label{E:eq_defdm}
 \end{equation}
-$m=1,\ldots,n$, which displays the temporal variation with respect to $\bar{\vX}$ of spatial energies.
+$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\vX}$.
 
@@
 Define the  $n\times p$ matrix
-\begin{equation}
+\begin{equation*}
  \vD=\left(
@@
 \label{eq_defmatrixD}
-\end{equation}
+\end{equation*}
 where $vec(\vD(m))$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX(m)$, $m=1,\ldots,n$.
-Sparsity \cite{johnstone2009statistical} on the wavelet coefficients plays a special role here.
-We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $e_d$, is such that $e_d<<p$.
+Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.
+We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d<<p$.
 This can be represented as the following linear model
 \begin{equation} 
-\vd=\vD\vbeta^{(d)}+\vxi^{(d)}
-\label{sparsemodel_d}
+\vd=\vD\vbeta+\vxi
+\label{E:sparsemodel_d}
 \end{equation}
-where $\vbeta^{(d)}$ is sparse, i.e., it has $p-e_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
+where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
 %\citep{jonhstone2009}
 
- In order to identify spatio-temporal changes we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
-For each squared mean-corrected approximation coefficient time series, given by $\vD(m)$, consider its Pearson correlation with the mean-corrected total approximation energy, given by $\vd$:
-\begin{equation}
-R_{k,l}^{(d)}= \corr\left( \vD_{k,l}, \vd\right),
+In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
+For each local squared mean deviation time series, given by individual elements of $\vD(m)$ across $m=1,\ldots,n$, consider its Pearson correlation with the overall squared mean deviations, given by $\vd$:
+\begin{equation*}
+R_{k,l}= \corr\left( \vD_{k,l}, \vd\right),
 \label{def_Rkld}
-\end{equation}
+\end{equation*}
 where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.
 
-We have a matrix $\vR^{(d)}=[R_{k,l}^{(d)}]$ of correlations of ultra-high dimension. Define the set of 
+We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
 {\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
-$\mathcal{M}^{*d}=\{(k,l): \textit{change in } \mathcal{I}(m)\textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are caused by changes in approximation coefficients of index} $(k,l)\}$.
-This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta^{(d)}$ in (\ref{sparsemodel_d}).
+$\mathcal{M}^{*d}=\{(k,l): \textit{change in } \mathcal{I}(m)\textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{is affected by changes in approximation coefficients of index} $(k,l)\}$.
+This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
 We build the empirical set of selected indices by
 \begin{equation}
-\mathcal{M}_{\tau}^{(d)}=\{(k,l):|R_{k,l}^{(d)}|>\tau_d\},
-\label{def_Mtaud}
+\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau_d\},
+\label{E:def_Mtaud}
 \end{equation}
-where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}^{(d)}$ contains its target set satisfies
+where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
 \[
-P(\mathcal{M}_{\tau}^{(d)}\supset\mathcal{M}^{*d})\rightarrow 1,
+P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
 \]
@@
 
-Analogously we take $\vT(m) = [T_{k,l}(m)]$, where $T_{k,l}(m)=(X_{k,l}(m+1)-X_{k,l}(m))^2$, for $m=1,\ldots,n-1$. We then analyze the time series given by
-\begin{equation}
-\vt(m)= \sum_{k,l}T_{k,l}(m) = \sum_{k,l}(X_{k,l}(m+1) - X_{k,l}(m))^2,
-\label{eq_deftm}
-\end{equation}
-$m=1,\ldots,n-1$.
-
-The highest values of $\{\vt(m)\}$ represent the images for which the most expressive changes take place between time points $m$ and $m+1$, where changes here are measured through $\mathbb{L}_2$ energy.
-Define the  $(n-1)\times p$ matrix
-\begin{equation}
- \vT=\left(
- \begin{array}{c}
- vec(\vT(1))^T\\
- \vdots\\
- vec(\vT(n))^T\\
- \end{array}
- \right),
-\label{eq_defmatrixT}
-\end{equation}
-where $vec(\vT(m))$ is the $p\times 1$ vector of wavelet coefficients for time $m=1,\ldots,n-1$.
-We again suppose a handful of coefficients drive the changes given by $\vt$, so that the effective dimension, say $e_t$, is such that $e_t<<p$.
-This can be represented as the following linear model
-\begin{equation} 
-\vt=\vT\vbeta^{(t)}+\vxi^{(t)}
-\label{sparsemodel_t}
-\end{equation}
-where $\vbeta^{(t)}$ is sparse, i.e. $\vbeta^{(t)}$ has $p-e_t$ null elements, and $\vxi^{(t)}$ is some $(n-1)\times 1$ random vector of errors.
-
-For each squared mean-corrected approximation coefficient time series, given by $\vT_{k,l} = (T_{k,l}(1),\ldots,T_{k,l}(n))^T$, consider its Pearson correlation with the mean-corrected total approximation energy, given by $\vt$:
-\begin{equation}
-R_{k,l}^{(t)}= \corr\left( \vT_{k,l}, \vt\right).
-\label{def_Rklt}
-\end{equation}
-
-We have a matrix $\vR^{(t)}=[R_{k,l}^{(t)}]$ of correlations of ultra-high dimension. Define the set of 
-{\it important} indices for changes in subsequent images as $\mathcal{M}^{*t}=\{(k,l):$ \textit{Change in }$\mathcal{I}(m+1)$ \textit{with respect to }$\mathcal{I}(m)$ \textit{ for some }$m=1,\ldots,n-1$\textit{ are caused by the approximation coefficients of index }$(k,l)\}$.
-We build a set of selected indices by
-\[
-\mathcal{M}_{\tau}^{(t)}=\{(k,l):|R_{k,l}^{(t)}|>\tau_t\},
-\] 
-where $\tau_t>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions,
-\[
-P(\mathcal{M}_{\tau}^{(t)}\supset\mathcal{M}^{*t})\rightarrow 1,
-\]
-as $n\rightarrow\infty$ \cite{fan2020statistical}.
-
-Therefore, if we define
-\begin{eqnarray}
-\mathcal{M}_{\tau}&=&\mathcal{M}_{\tau}^{(t)}\cup\mathcal{M}_{\tau}^{(d)},\label{def_Mpop}\\
-\mathcal{M}^{*}&=& \mathcal{M}^{*d}\cup\mathcal{M}^{*t}\label{def_Memp},
-\end{eqnarray}
-we have the following consistency property:
-\[
-P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1.
-\]
-as $n\rightarrow\infty$.
-
-Thence, if we compute $\{\vd(m)\}$ and $\{\vt(m)\}$, and build $\mathcal{M}_{\tau}=\mathcal{M}_{\tau}^{(t)}\cup\mathcal{M}_{\tau}^{(d)}$, the consistency of the screening methods above guarantees asymptotic coverage of all approximation coefficients strongly associated with changes with respect to the average image as well as immediate previous image with a high probability, as long as the required regularity conditions hold.
-
 Further geometrical motivation for our proposal is given as follows.
-We argue the case of $\vd$ but the same may be written regarding $\vt$ as well.
-As defined by (\ref{eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values.
+As defined by (\ref{E:eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values.
 This segregates the multi-temporal images, since the former time points identify the images in which significant changes occur, while the latter indices identify time points with no major changes.
@@
 We also take $\delta=U-L$.
-The indices defined by  (\ref{def_Mtaud}) are such that
+The indices defined by  (\ref{E:def_Mtaud}) are such that
 \[

### 2021-10-27 codes/EllispseChangeAnalysis_Deep.m
updates on codes

@@ -3,27 +3,19 @@
 close all
 clear;
 %%
-im1 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');  
-im2 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
-im3 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
-im4 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
+im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
+im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
+im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
+im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
 %
-figure
-imshow(im1)
-figure
-imshow(im2)
-figure
-imshow(im3)
-figure
-imshow(im4)
-%
+
 % Images are too big, consider subsampling
-subsamplingfactor = 2;
+subsamplingfactor = 16;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
-totalchanges = imread('GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
+totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
 totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
 %
 figure
@@ -76,6 +68,7 @@ end
 R = R./max(R(:));
 R = reshape(R,NbRows, NbCols);
 [pD2,pFA2]=ROCcurveNew(R,255*totalchanges); close
+[vp,vF1_db2,~,~] = F1Scorecurve(R,255*totalchanges); close
 %
 figure
 imshow(R)
@@ -83,7 +76,7 @@ title('d(m): wavelet db2, J=2')
 
 
 %% deep learning denoiser instead of using a wavelet transform
-cd deeplearning/
+
[…]

### 2021-10-27 codes/EllispseChangeAnalysis_families.m

@@ -5,27 +5,19 @@
 close all
 clear;
 %%
-im1 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');  
-im2 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
-im3 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
-im4 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
-%
-figure
-imshow(im1)
-figure
-imshow(im2)
-figure
-imshow(im3)
-figure
-imshow(im4)
+im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
+im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
+im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
+im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
+
 %
 % Images are too big, consider subsampling
-subsamplingfactor = 2;
+subsamplingfactor = 8;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
-totalchanges = imread('GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
+totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
 totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
 %
 figure
@@ -60,8 +52,8 @@ wname = 'haar';
     X4 = X4(:,:,J); % focus only on level J approximations ?
 
 % % mean approximation coefficients
-imRef_wave = X1 + X2 + X3 + X4;
-    imRef_wave = imRef_wave/n;
+% imRef_wave = X1 + X2 + X3 + X4;
[…]

### 2021-10-27 codes/EllispseChangeAnalysis_levels.m

@@ -5,27 +5,19 @@
 close all
 clear;
 %%
-im1 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');  
-im2 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
-im3 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
-im4 = imread('ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
+im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
+im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
+im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
+im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
 %
-figure
-imshow(im1)
-figure
-imshow(im2)
-figure
-imshow(im3)
-figure
-imshow(im4)
-%
+
 % Images are too big, consider subsampling
-subsamplingfactor = 1;
+subsamplingfactor = 8;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
-totalchanges = imread('GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
+totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
 totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
 %
 figure
@@ -83,6 +75,7 @@ end
 R = R./max(R(:));
 R = reshape(R,NbRows, NbCols);
 [pD1,pFA1]=ROCcurveNew(R,255*totalchanges); close
+[vp,vF1_haar,~,~] = F1Scorecurve(R,255*totalchanges); close
 %
 figure
 imshow(R)
@@ -125,6 +118,7 @@ end
 R = R./max(R(:));
[…]

### 2021-10-27 codes/EllispseChangeAnalysis_methods.m

@@ -10,24 +10,10 @@ im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');
 im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
 im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
 %
-figure
-imshow(im1)
-figure
-imshow(im2)
-figure
-imshow(im3)
-figure
-imshow(im4)
-
-figure
-subplot(1,4,1); imshow(im1); title('I(1)')
-subplot(1,4,2); imshow(im2); title('I(2)')
-subplot(1,4,3); imshow(im3); title('I(3)')
-subplot(1,4,4); imshow(im4); title('I(4)')
 
 %
 % Images are too big, consider subsampling
-subsamplingfactor = 2;
+subsamplingfactor = 8;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
@@ -38,7 +24,7 @@ totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
 mImage = figure;
 imshow(totalchanges)
 title('Total changes', 'FontSize', 17)
-%saveas(mImage,sprintf('total_changes.jpg'))
+%saveas(mImage,sprintf('../figs/total_changes.jpg'))
 
 %%
 eps = .00001;
@@ -88,12 +74,14 @@ end
 R = R./max(R(:));
 R = reshape(R,NbRows, NbCols);
 [pD,pFA]=ROCcurveNew(R,255*totalchanges); close
+[vp,vF1_wave,~,~] = F1Scorecurve(R,255*totalchanges); close
 %
 mImage = figure;
 imshow(R)
 title('db2 WECS d(m), J=2', 'FontSize', 17)
-%saveas(mImage,sprintf('corr_changes_dm.jpg'))
-    
+%saveas(mImage,sprintf('../figs/corr_changes_dm.jpg'))
+
+
 %% No wavelets for comparison with mean image
 X1 = im1log;
 X2 = im2log;
@@ -121,11 +109,12 @@ end
 R = R./max(R(:));
 R = reshape(R,NbRows, NbCols);
[…]

### 2021-10-27 codes/F1Scorecurve.m

@@ -1,21 +1,23 @@
-function [pD,pFA]=F1Scorecurve(img,im0)
+function [t,F1,pD,Re]=F1Scorecurve(img,im0)
 
-lxy=numel(im0);
 % Nombre reel de pixel ayant change
 CP=find(im0~=0);
 nCP=numel(CP);
 %%%%%%%%%%%%%%%%
-u=0;
-v=1;
+u=0.01;
+v=.99;
 pas=(v-u)/100;
 t=u:pas:v;
+vQuant = quantile(img(:),1-t); % quantiles in decreasing order
 %t=u+2*pas:pas:v-2*pas;
 lt=numel(t);
 pD=zeros(1,lt);
-pFA=pD;
+Re=zeros(1,lt);
+F1=zeros(1,lt);
 for k=1:lt
-    %imtemp=255.*im2bw(img,k/T);
-    imtemp=255.*double(img>t(k));
+    % the p% largest correlations, are those larger that the (1-p)th
+    % quantile of R
+    imtemp=255.*double(img>vQuant(k));
     % Detection
     D=find(imtemp==255);
     nD=numel(D);
@@ -24,13 +26,13 @@ for k=1:lt
     TP=find(imgdiff==255);
     nTP=numel(TP);
     % Precision
-    pD(k)=nTP/nD;
+    pD(k)= nTP/nD;
     % Recall
     FN=find(imgdiff==-255);
     nFN=numel(FN);
-    pD(k)=nTP/(nBD + );
-    % Probabilite de fausse alarm
-    pFA(k)=(nD-nTP)/(lxy-nCP);
+    Re(k)= nTP/(nTP + nFN);
+    % F1-score
+    F1(k)=2*pD(k)*Re(k)/(pD(k) + Re(k));
 end
 
 %y=reshape(imtemp,1,lxy);
@@ -38,9 +40,9 @@ end
 
 %%% Plot ROC curves
 figure
-plot(pFA,pD);
-title('ROC Curve')
-xlabel('False positive rate')
-ylabel('True positive rate')
+plot(t,F1);
+%title('F1-score')
+xlabel('p')
+ylabel('F1-score')
 axis([0 1 0 1]);
 axis square

### 2021-10-27 codes/denoiseImageCPU.m

@@ -1,0 +1,106 @@
+function I = denoiseImageCPU(A, net)
+%denoiseImage Denoise an image using deep neural network
+%
+%   B = denoiseImage(A, net) estimates denoised single image or batch of
+%       images B from a noisy image or batch of noisy images A using a
+%       pre-trained denoising deep neural network specified as net.
+%
+%   NOTE: This function requires the Deep Learning Toolbox.
+%   
+%   Class Support
+%   -------------
+%
+%    In case of a single image, A can be a H-by-W or H-by-W-by-C matrix of
+%    class uint8, uint16, single, double. In case of a batch of images, A
+%    should be a H-by-W-by-C-by-N matrix of class uint8, uint16, single,
+%    double where C denotes the number of image channels and N the batch
+%    size. net is a network. B is a matrix of same class and shape as A.
+%
+%   Notes
+%   -----
+%
+%   1. net can be obtained by using the function denoisingNetwork. It can
+%      also be a user defined custom pre-trained denoising network.
+%
+%   2. It is expected that net would be trained to handle images with same
+%      channel format (RGB or grayscale) as A.
+%
+%   3. The learning framework is assumed to be residual and the final
+%      denoised image would be the difference of the input image and the estimated
+%      residual output from the network layers.
+%
+%   4. The net is assumed to be a denoising network with a convolution layer
+%      before the last regression layer and output of this function
+%      corresponds to the response of this final convolution layer.
+%
+%   Example 
+%   -------
+%
+%   net = denoisingNetwork('dncnn');
+%   I = imread('cameraman.tif');
+%
+%   noisyI = imnoise(I, 'gaussian', 0, 0.01);
+%   denoisedI = denoiseImage(noisyI, net);
+%
+%   figure
+%   imshowpair(noisyI,denoisedI,'montage');
+%
+%   See also dnCNNLayers, denoisingImageDatastore, denoisingNetwork 
+
+
+%   Copyright 2017-2019 The MathWorks, Inc.
+
+matlab.images.internal.errorIfgpuArray(A, net);
+images.internal.requiresNeuralNetworkToolbox(mfilename);
+
+narginchk(2,2);
+
+validateInputImage(A);
+validateInputNetwork(net);
+
[…]

### 2021-10-27 codes/EllispseChangeAnalysis_Deep.m
updates on codes and .tex

@@ -10,7 +10,7 @@ im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');
 %
 
 % Images are too big, consider subsampling
-subsamplingfactor = 16;
+subsamplingfactor = 2;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
@@ -140,7 +140,7 @@ legend('db2 WECS $\textbf{d}(m)$, $J=2$', '\textbf{d}(m): deep learning denoiser
     'interpreter','latex','Location','southeast', 'FontSize', 12)
 legend('boxoff')
 hold off
-%saveas(mImage,sprintf('../figs/dm_comparison_wavelet_deepL.jpg'))
+saveas(mImage,sprintf('../figs/dm_comparison_wavelet_deepL.jpg'))
 
 mImage = figure;
 hold on
@@ -155,7 +155,7 @@ legend('db2 WECS $\textbf{d}(m)$, $J=2$', '\textbf{d}(m): deep learning denoiser
     'interpreter','latex','Location','southeast', 'FontSize', 12)
 legend('boxoff')
 hold off
-%saveas(mImage,sprintf('../figs/dm_comparison_wavelet_deepL_F1score.jpg'))
+saveas(mImage,sprintf('../figs/dm_comparison_wavelet_deepL_F1score.jpg'))
 
 
 return

### 2021-10-27 codes/EllispseChangeAnalysis_families.m

@@ -12,7 +12,7 @@ im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');
 
 %
 % Images are too big, consider subsampling
-subsamplingfactor = 8;
+subsamplingfactor = 2;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
@@ -315,7 +315,7 @@ legend('haar', 'db2', 'db4',...
     'coif4', 'sym4', 'sym2','Location','southeast', 'FontSize', 12)
 legend('boxoff')
 hold off
-%saveas(mImage,sprintf('../figs/families_comparison.jpg'))
+saveas(mImage,sprintf('../figs/families_comparison.jpg'))
 
 mImage = figure;
 hold on
@@ -334,7 +334,7 @@ legend('haar', 'db2', 'db4',...
     'coif4', 'sym4', 'sym2','Location','northeast', 'FontSize', 12)
 legend('boxoff')
 hold off
-%saveas(mImage,sprintf('../figs/families_comparison_F1score.jpg'))
+saveas(mImage,sprintf('../figs/families_comparison_F1score.jpg'))
 
 
 

### 2021-10-27 codes/EllispseChangeAnalysis_levels.m

@@ -12,7 +12,7 @@ im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');
 %
 
 % Images are too big, consider subsampling
-subsamplingfactor = 8;
+subsamplingfactor = 2;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
@@ -39,7 +39,6 @@ wname = 'db2';
 
 %% Case 1 : reference image 
 
-% Wavelet basis: Haar
 J = 1;
 
 [X1,~,~,~] = swt2(im1log,J,wname);
@@ -75,7 +74,7 @@ end
 R = R./max(R(:));
 R = reshape(R,NbRows, NbCols);
 [pD1,pFA1]=ROCcurveNew(R,255*totalchanges); close
-[vp,vF1_haar,~,~] = F1Scorecurve(R,255*totalchanges); close
+[vp,vF1_J1,~,~] = F1Scorecurve(R,255*totalchanges); close
 %
 figure
 imshow(R)
@@ -271,7 +270,7 @@ legend('$J=1$', '$J=2$', '$J=3$','$J=4$', '$J=5$',...
     'interpreter','latex','Location','southeast', 'FontSize', 12)
 legend('boxoff')
 hold off
-%saveas(mImage,sprintf('../figs/levels_comparison.jpg'))
+saveas(mImage,sprintf('../figs/levels_comparison.jpg'))
 
 mImage = figure;
 hold on
@@ -280,7 +279,7 @@ xlabel('$p$','interpreter','latex', 'FontSize', 13)
 ylabel('$F_1$-score','interpreter','latex', 'FontSize', 13)
 axis([0 1 0 0.7]);
 axis square
-plot(vp,vF1_haar,':+')
+plot(vp,vF1_J1,':+')
 plot(vp,vF1_J2,'-.or')
 plot(vp,vF1_J3,':*')
 plot(vp,vF1_J4,':^')
@@ -289,7 +288,7 @@ legend('$J=1$', '$J=2$', '$J=3$','$J=4$', '$J=5$',...
     'interpreter','latex','Location','northeast', 'FontSize', 12)
 legend('boxoff')
 hold off
[…]

### 2021-10-27 codes/EllispseChangeAnalysis_methods.m

@@ -13,7 +13,7 @@ im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');
 
 %
 % Images are too big, consider subsampling
-subsamplingfactor = 8;
+subsamplingfactor = 2;
 im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
 im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
@@ -24,7 +24,7 @@ totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
 mImage = figure;
 imshow(totalchanges)
 title('Total changes', 'FontSize', 17)
-%saveas(mImage,sprintf('../figs/total_changes.jpg'))
+saveas(mImage,sprintf('../figs/total_changes.jpg'))
 
 %%
 eps = .00001;
@@ -79,7 +79,7 @@ R = reshape(R,NbRows, NbCols);
 mImage = figure;
 imshow(R)
 title('db2 WECS d(m), J=2', 'FontSize', 17)
-%saveas(mImage,sprintf('../figs/corr_changes_dm.jpg'))
+saveas(mImage,sprintf('../figs/corr_changes_dm.jpg'))
 
 
 %% No wavelets for comparison with mean image
@@ -114,13 +114,13 @@ R = reshape(R,NbRows, NbCols);
 mImage = figure;
 imshow(R)
 title('d(m) without wavelets', 'FontSize', 17)
-%saveas(mImage,sprintf('../figs/corr_changes_dm_nowavelets.jpg'))
+saveas(mImage,sprintf('../figs/corr_changes_dm_nowavelets.jpg'))
 
 
 %% Standard change detection
 
 S = abs(im4log - im3log) + abs(im3log - im2log) + abs(im2log - im1log);
-S = S./max(R(:));
+S = S./max(S(:));
 %
 [pD1,pFA1]=ROCcurveNew(S,255*totalchanges); close
 [vp,vF1_std,~,~] = F1Scorecurve(R,255*totalchanges); close
@@ -128,7 +128,7 @@ S = S./max(R(:));
 mImage = figure;
 imshow(S)
[…]

### 2021-10-29 tex/paper.wecs/wecs_tgrs.tex
updates on latex

@@
 \newcommand{\vR}{{\textbf R}}
+\newcommand{\vS}{{\textbf S}}
 \newcommand{\vU}{{\textbf U}}
@@
 \begin{equation} 
-\sum_{m=1}^n\|\vX(m)\|^2=n\|\bar{\vX}\|^2+\sum_{m=1}^n\|\mathcal \vX(m)-\bar{\vX}\|^2,\label{E:eq_ANOVAlogimage}
+\sum_{m=1}^n\|\vX(m)\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX(m)-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
 \end{equation}
-where $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX(m)$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
-The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX(m)$ compared to an {\it average wavelet-approximated image} $\bar{\vX}$, what motivates us to such deviations to detect change points in time.
+where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}(m)$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX(m)$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
+The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX(m)$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
 Since each element (pixel) in $\vX(m)$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space.
@@
 
-Let $X_{k,l}(m)$ and $\bar{X}_{k,l}$ be the entry $(k,l)$ of the matrices $\vX(m)$ and $\bar{\vX}$, respectively.
-Computing $D_{k,l}(m)=(X_{k,l}(m)-\bar{X}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD(m) = [D_{k.l}(m)]$.
+Let $X_{k,l}(m)$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX(m)$ and $\bar{\mathcal{I}}$, respectively.
+Computing $D_{k,l}(m)=(X_{k,l}(m)-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD(m) = [D_{k.l}(m)]$.
 We then
@@
 \begin{equation} 
-\vd(m)= \sum_{k,l}D_{k,l}(m) = \sum_{k,l}(X_{k,l}(m) - \bar{X}_{k,l})^2,
+\vd(m)= \sum_{k,l}D_{k,l}(m) = \sum_{k,l}(X_{k,l}(m) - \bar{\mathcal{I}}_{k,l})^2,
 \label{E:eq_defdm}
 \end{equation}
-$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\vX}$.
+$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.
 
@@
 {\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
-$\mathcal{M}^{*d}=\{(k,l): \textit{change in } \mathcal{I}(m)\textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{is affected by changes in approximation coefficients of index} $(k,l)\}$.
+$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}(m)\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
 This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
-We build the empirical set of selected indices by
+We build the empirical set of indices flagged as corresponding to change locations by
 \begin{equation}
-\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau_d\},
+\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
 \label{E:def_Mtaud}
@@
 as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.
+In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{*}$ when the number of observations $n$ is large.
 
@@
 \[
-\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau_d,
+\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau,
 \]
-i.e., such that  $\sum_{m=1}^nD_{k,l}(m)\vd(m)>\tau_d\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
+i.e., such that  $\sum_{m=1}^nD_{k,l}(m)\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
 \[
@@
 The synthetic multi-temporal images ($n=4$) are shown in Figure \ref{F:EllipsoidChanges}.
-The first image, $I(1)$, presents three elongated ellipses.
+The first image, $\mathcal{I}(1)$, presents three elongated ellipses.
 Changes consist of three different types of ellipses that are successively added to the original image $I(1)$.
-The second image, $I(2)$, has new large ellipses added.
-Smaller ellipses are then added to form $I(3)$ and $I(4)$.
+The second image, $\mathcal{I}(2)$, has new large ellipses added.
+Smaller ellipses are then added to form $\mathcal{I}(3)$ and $\mathcal{I}(4)$.
+All the changes made to image $\mathcal{I}(1)$ can be seen in Figure \ref{F:Changes_methods_images}(a), which displays a matrix of zeros and ones that correspond, respectively, to locations without and with changes along time.
+Applying WECS to these images, we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation.
+An example of $\vR$ is presented in Figure \ref{F:Changes_methods_images}(b).
+For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a matrix of zeros and ones that can be compared with the total true changes displayed in Figure \ref{F:Changes_methods_images}(a).
 
@@
 \includegraphics[scale=.1]{../../figs/corr_changes_dm}(b)
-\includegraphics[scale=.1]{../../figs/corr_changes_tm}(c)
-\includegraphics[scale=.1]{../../figs/corr_changes_logratios}(d)
-\includegraphics[scale=.1]{../../figs/corr_changes_dm_nowavelets}(e)
-\includegraphics[scale=.1]{../../figs/corr_changes_tm_nowavelets}(f)
-\caption{Synthetic images with changing ellipses. (a) Image composed by the total changes over time. 
-(b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Proposed db2 WECS $\vt(m)$ with $J=2$.
-(d) Standard approach. (e) $\vd(m)$ without wavelets. (f) $\vt(m)$  without wavelets. 
-}
+\includegraphics[scale=.1]{../../figs/corr_changes_logratios}(c)
+\includegraphics[scale=.1]{../../figs/corr_changes_dm_nowavelets}(d)
+\caption{Synthetic images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
 \label{F:Changes_methods_images}
@@
 
-Figure \ref{F:Changes_methods_images} illustrates the simulated synthetic images, the proposed wavelet detection methods, and three classic detection methods, as well.
-Panel (a) presents the total changes with respect to image $I(1)$.
-Panels (b) and (c) show the results by the proposed methods using Daubechies wavelet with two null moments (db2) and $J=2$ by $\vd(m)$ and $\vt(m)$, respectively.
-Panel (d) presents the results by a standard approach (aggregated log-ratios).
-Finally, in Panels (e) and (f)  we can see the results if $\vd(m)$ and $\vt(m)$ are performed purely on the spatial domain, without multiscale approximations.
-The spatio-temporal advantages of the proposed wavelet  $\vd(m)$ and $\vt(m)$ are clear in Figure \ref{F:Changes_methods_images}.
-A slight advantage for the detection of small ellipses is attained by $\vd(m)$ over $\vt(m)$.
-
-We compute ROC curves to compare the detection performance of different methods.
-For this we simulate noisy versions of the synthetic images illustrated by Figure \ref{F:EllipsoidChanges}.
-The change detection methods are then employed.
-Each method generates a correlation matrix between the real image of total changes and the estimated one.
-
-Each ROC curve presents how close the magnitude variation of change measures is to the variation of the image of total changes in the following way:
+Figure \ref{F:Changes_methods_images} illustrates a comparison of different approaches to detect accumulated changes.
+Panel (b) shows the result of WECS using Daubechies wavelet with two null moments (db2) and $J=2$.
+Panel (c) presents the result of using aggregated log-ratios, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{il} = \sum_{m=1}^3	|\log\left(\mathcal{I}_{k,l}(m+1)/\mathcal{I}_{k,l}(m)\right)|$.
+Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}(m)$ instead of $\vX(m)$ in the WECS formulation.
+The spatio-temporal advantages of the proposed wavelet  $\vd(m)$ is clear in Figure \ref{F:Changes_methods_images}(b), which shows that a wavelet smoothing has an expressive capacity of measuring the the changes through correlations of coefficients' deviations.
+
+We compute ROC curves to compare the detection performance of different methods and to verify the influence of some tuning parameters of wavelet smoothing: the resolution level $J$ and the choice of wavelet basis.
+Each detection method generates a matrix of change detection measures (correlations in the case of WECS).
+The ROC curves present the performance of change detection by applying a threshold on these measures, in the following way:
 \begin{enumerate}
@@
 \item Let $(r_{(1)},\ldots,r_{(100)})$ be equally space values between $r_{\min}$ and $r_{\max}$;
-\item For each $k=1,\ldots,n$, check how many pixels are such that $R_{i,j}>r_{(k)}$ coincide with the pixels $(i,j)$ where a change really occurs on the image of total changes.
+\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$ coincide with the element $(i,j)$ where a change really occurs on the image of total changes.
 Dividing this number by the total number of changes gives the true positive rate.
-\item For each $k=1,\ldots,n$, check how many pixels are such that $R_{i,j}>r_{(k)}$  do not coincide with the pixels $(i,j)$ where a change really occurs.
-Dividing this number by the total number of pixels where changes do not occur gives the false positive rate.
+\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$  do not coincide with the element $(i,j)$ where a change really occurs.
+Dividing this number by the total number of entries where changes do not occur gives the false positive rate.
+\item The ROC curve is the plot of true and false positive rates corresponding to each $k$.
 \end{enumerate}
@@
 \includegraphics[scale=.13]{../../figs/methods_comparison}\hspace{-.5cm}(a) 
-\includegraphics[scale=.13]{../../figs/levels_comparison_WithReference}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{../../figs/families_comparison_WithReference}\hspace{-.5cm}(c)
+\includegraphics[scale=.13]{../../figs/levels_comparison}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.13]{../../figs/families_comparison}\hspace{-.5cm}(c)
 \includegraphics[scale=.13]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
-\includegraphics[scale=.13]{../../figs/levels_comparison_NoReference}\hspace{-.5cm}(e)
-\includegraphics[scale=.13]{../../figs/families_comparison_NoReference}\hspace{-.5cm}(f)
 \caption{ROC curves for detection of changing ellipses in synthetic images and different methods.
-(a) The proposed methods in black (db2 WECS $\vd(m)$) and green (db2 WECS $\vt(m)$) vs three non-wavelet methods: standard (red stars); $\vd(m)$ (blue); and $\vt(m)$ (red circles).
+(a) The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue).
 (b) db2 $\vd(m)$ with different levels.
 (c) $\vd(m)$ with different wavelet bases and J=2;
-(d) The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line).  
-(e) db2 $\vt(m)$ with different levels. (f) $\vt(m)$ with different wavelet bases and J=2. }
+(d) The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line).}
 \label{F:EllipsoidChanges_details}
@@
 Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the synthetic data as follows.
-The effects of wavelet bases, level of decomposition, deep-learning feature extraction and $\vd(m)$/$\vt(m)$ usage are shown on the ROC curves.
+The effects of wavelet bases, level of decomposition, deep-learning feature extraction are shown on the ROC curves.
 We employ the following wavelet bases: Haar;  Daubechies db2; Daubechies db4; Coiflets coif4; Symlets sym2 ; and Symlets sym4.
-Panels (c) and (f) present the ROC curves for the proposed methods under the aforementioned bases for $\vd(m)$ and $\vt(m)$, respectively.
-On both instances $J=2$ is employed.
-The results for $\vd(m)$ are much more robust to basis variation than the ones for $\vt(m)$.
-Combining the ROC curves's comparison from both panels, Daubechies db2 is the best choice.
-Panels (b) and (e) present the ROC curves for different levels of decomposition under the aforementioned bases for $\vd(m)$ and $\vt(m)$, respectively.
-On both instances db2 is employed, and five levels are considered: $J=1,2,3,4,5$.
-Levels $J=2,3$ have a clear better performance for $\vd(m)$ (with a slight advantage to $J=2$), whilst $J=1$ is competitive for $\vt(m)$.
+Panel (c) presents the ROC curves for the proposed method under the aforementioned bases.
+On all instances $J=2$ is employed.
+Comparing the ROC curves of all options, we can notice that Daubechies db2 and Symlets sym2 are the best choices.
+Panel (b) presents the ROC curves for five different levels of decomposition $J=\{1,2,3,4,5\}$ under the Daubechies db2 basis.
+Levels $J=2,3$ have a clear better performance, with a slight advantage to $J=2$.
 The overall performance  of $J=2$ warrants its use for the rest of the comparisons.
-Panel (d) shows how the proposed method performs with or without images' deep learning feature extraction from a residual learning network \cite{zhang2017beyond}.
-The change detection method is the proposed db2 $\vd(m)$ with $J=2$.
+Panel (d) shows how the proposed method performs in comparison to a deep learning feature extraction from a residual learning network \cite{zhang2017beyond}.
+WECS is applied with db2 wavelets and $J=2$.
 We can see that the ROC curves for images treated with deep-learning methods or treated with wavelet based methods are almost identical.
-The WECS runs in 3.35s, while the deep-learning based $\vd(m)$ runs in 559.12s on a notebook.
-The configuration of the notebook is: OS - Ubuntu 18.04.5 LTS; RAM 7.7 GB; Intel\textregistered Core\texttrademark $ $i7-7500U CPU @ 2.70GHz x 4; graphics - Intel\textregistered HD Graphics 620 (KBL GT2); GNOME - 3.28.2; OS type - 64-bit.
-We finally have in Panel (a) the proposed db2 $\vd(m)$ with $J=2$ and db2 $\vt(m)$ with $J=2$  compared to three other non-wavelet methods.
-These are $\vd(m)$ and $\vt(m)$ where wavelet decomposition is not performed, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated log-ratios of $\{\mathcal{I}(m)\}$.
-The ROC curves in Panel (a) clearly show that  the proposed db2 $\vd(m)$ with $J=2$ outperforms the rest.
-
-We may summarize these results as: the proposed wavelet $\vd(m)$ method presents a superior performance.
+The WECS runs in 0.42s, while the deep-learning based $\vd(m)$ runs in 64.68s on a notebook.
+The configuration of the notebook is: OS - Ubuntu 18.04.5 LTS; RAM 7.7 GB; Intel\textregistered Core\texttrademark $ $i7-8565U CPU @ 1.80GHz x 8; graphics - Mesa Intel\texttrademark UHD Graphics 620 ; GNOME - 3.36.8; OS type - 64-bit.
+We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
+The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated log-ratios of $\{\mathcal{I}(m)\}$.
+The ROC curves in Panel (a) clearly show that  the proposed WECS outperforms the rest.
+
+%We may summarize these results as: the proposed wavelet $\vd(m)$ method presents a superior performance.
 It is also equipped with the following nice properties: (i) it is scalable; (ii) it is sparse; (iii) it is parsimonious; (iv) it can be easily adapted to be linearly updated when a new image is acquired; and (v) it is fast.

### 2021-10-29 codes/EllispseChangeAnalysis_Deep.m
included code to make synthetic images

@@ -24,10 +24,10 @@ title('Total changes')
 
 %%
 eps = .00001;
-im1log = log(eps+double(im1));
-im2log = log(eps+double(im2));
-im3log = log(eps+double(im3));
-im4log = log(eps+double(im4));
+im1log = eps+double(im1);
+im2log = eps+double(im2);
+im3log = eps+double(im3);
+im4log = eps+double(im4);
 n=4;
 imRef = im1log + im2log + im3log + im4log;
     imRef = imRef/n;
@@ -85,10 +85,10 @@ net = denoisingNetwork('DnCNN'); % load a deep neural network that has learned t
 % X4 = denoiseImageCPU(im4log,net);
 
 tic
-X1 = denoiseImageCPU(im1,net);
-X2 = denoiseImageCPU(im2,net);
-X3 = denoiseImageCPU(im3,net);
-X4 = denoiseImageCPU(im4,net);
+X1 = denoiseImageCPU(im1log,net);
+X2 = denoiseImageCPU(im2log,net);
+X3 = denoiseImageCPU(im3log,net);
+X4 = denoiseImageCPU(im4log,net);
 toc
 
 %%

### 2021-10-29 codes/EllispseChangeAnalysis_families.m

@@ -26,10 +26,10 @@ title('Total changes')
 
 %%
 eps = .00001;
-im1log = log(eps+double(im1));
-im2log = log(eps+double(im2));
-im3log = log(eps+double(im3));
-im4log = log(eps+double(im4));
+im1log = eps+double(im1);
+im2log = eps+double(im2);
+im3log = eps+double(im3);
+im4log = eps+double(im4);
 n=4;
 imRef = im1log + im2log + im3log + im4log;
     imRef = imRef/n;

### 2021-10-29 codes/EllispseChangeAnalysis_methods.m

@@ -28,10 +28,10 @@ saveas(mImage,sprintf('../figs/total_changes.jpg'))
 
 %%
 eps = .00001;
-im1log = log(eps+double(im1));
-im2log = log(eps+double(im2));
-im3log = log(eps+double(im3));
-im4log = log(eps+double(im4));
+im1log = eps+double(im1);
+im2log = eps+double(im2);
+im3log = eps+double(im3);
+im4log = eps+double(im4);
 n=4;
 imRef = im1log + im2log + im3log + im4log;
 imRef = imRef/n;

### 2021-10-29 codes/generate_synthetic.m

@@ -1,0 +1,49 @@
+
+% Code to generate the synthetic images 
+
+%%
+clear all
+close all
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Loading images 
+im1=imread('../figs/GroundTruthEllipsoidChanges/ellipse1.tif');
+im1=double(~im1); % boolean ~ used to represent change point with 1
+im2=imread('../figs/GroundTruthEllipsoidChanges/ellipse2.tif');
+im2=double(~im2); 
+im3=imread('../figs/GroundTruthEllipsoidChanges/ellipse3.tif');
+im3=double(~im3); 
+im4=imread('../figs/GroundTruthEllipsoidChanges/ellipse4.tif');
+im4=double(~im4); 
+
+% noise level used
+sig = 1/10;
+
+% adding Gaussian noise to images
+im1_obs = 1+im1/2 + sig*randn(height(im1), width(im1));
+im2_obs = 1+(im2+im1)/2 + sig*randn(height(im2), width(im2));
+im3_obs = 1+(im1+im2+im3)/2 + sig*randn(height(im3), width(im3));
+im4_obs = 1+(im1+im2+im3+im4)/2 + sig*randn(height(im4), width(im4));
+
+colormap(gray(256)); imagesc(im2_obs)
+colormap(gray(256)); imagesc(im3_obs)
+colormap(gray(256)); imagesc(im4_obs)
+colormap(gray(256)); imagesc(im1_obs+im2_obs+im3_obs+im4_obs)
+
+% saving the synthetic images
+mImage = cast(100*(im1_obs),'uint8');
+imshow(mImage)
+imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png'))
+
+mImage = cast(100*(im2_obs),'uint8');
+imshow(mImage)
+imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png'))
+
+mImage = cast(100*(im3_obs),'uint8');
+imshow(mImage)
+imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png'))
+
+mImage = cast(100*(im4_obs),'uint8');
+imshow(mImage)
+imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png'))
+

### 2021-11-02 tex/paper.wecs/wecs_tgrs.tex
updates on code to analyze the real data

@@
 The synthetic multi-temporal images ($n=4$) are shown in Figure \ref{F:EllipsoidChanges}.
+These images are generated from the sum of two matrices: a signal matrix with 1/2 on entries where ellipses occur and zero elsewhere, and a noise matrix with random variables following a Gaussian distribution with zero mean and variance $1/100$.
 The first image, $\mathcal{I}(1)$, presents three elongated ellipses.
-Changes consist of three different types of ellipses that are successively added to the original image $I(1)$.
+Changes consist of three different types of ellipses that are successively added to the original image $\mathcal{I}(1)$.
 The second image, $\mathcal{I}(2)$, has new large ellipses added.
@@
 Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}(m)$ instead of $\vX(m)$ in the WECS formulation.
-The spatio-temporal advantages of the proposed wavelet  $\vd(m)$ is clear in Figure \ref{F:Changes_methods_images}(b), which shows that a wavelet smoothing has an expressive capacity of measuring the the changes through correlations of coefficients' deviations.
+%The spatio-temporal advantages of the proposed wavelet  $\vd(m)$ is clear in Figure \ref{F:Changes_methods_images}(b), which shows that a wavelet smoothing has an expressive capacity of measuring the the changes through correlations of coefficients' deviations, even for the changes corresponding to smaller ellipses.
 
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.13]{../../figs/methods_comparison}\hspace{-.5cm}(a) 
-\includegraphics[scale=.13]{../../figs/levels_comparison}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{../../figs/families_comparison}\hspace{-.5cm}(c)
-\includegraphics[scale=.13]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
+\includegraphics[scale=.14]{../../figs/methods_comparison}\hspace{-.6cm}(a) 
+\includegraphics[scale=.14]{../../figs/levels_comparison}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.14]{../../figs/families_comparison}\hspace{-.5cm}(c)
+\includegraphics[scale=.14]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
 \caption{ROC curves for detection of changing ellipses in synthetic images and different methods.
@@
 We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
-The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated log-ratios of $\{\mathcal{I}(m)\}$.
-The ROC curves in Panel (a) clearly show that  the proposed WECS outperforms the rest.
+The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}(m)\}$.
+The ROC curves in Panel (a) show that the proposed WECS outperforms its counterpart without wavelets and that it is competitive with the method using aggregation of differences.
 
-%We may summarize these results as: the proposed wavelet $\vd(m)$ method presents a superior performance.
-It is also equipped with the following nice properties: (i) it is scalable; (ii) it is sparse; (iii) it is parsimonious; (iv) it can be easily adapted to be linearly updated when a new image is acquired; and (v) it is fast.
-Hence, the proposed wavelet change-detection procedure can be used as a real-time change detection tool for long time series of large images.
+%\begin{figure}[htp!]
+%\includegraphics[scale=.14]{../../figs/methods_comparison_F1score}\hspace{-.5cm}(a) 
+%\includegraphics[scale=.14]{../../figs/levels_comparison_F1score}\hspace{-.5cm}(b)\\
+%\includegraphics[scale=.14]{../../figs/families_comparison_F1score}\hspace{-.5cm}(c)
+%\includegraphics[scale=.14]{../../figs/dm_comparison_wavelet_deepL_F1score}\hspace{-.5cm}(d) \\
+%\caption{$F_1$-scores for detection of changing ellipses in synthetic images and different methods.
+(a) The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue).
+%(b) db2 $\vd(m)$ with different levels.
+%(c) $\vd(m)$ with different wavelet bases and J=2;
+%(d) The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line).}
+%\label{F:EllipsoidChanges_details}
+%\end{figure}
 
@@
 We employed the proposed change detection method on a series of 85 multi-date satellite images.
-The images were taken on a forest region at the border of Brazil and the French Guiana from November 08, 2015 to December 09, 2017.
-Each image has two channels and 1200 by 1000 pixels.
-We perform three change detection wavelet analyses: VV Polarization Channel;  VH Polarization Channel; and the Combined Image by Euclidean norm.
+The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3.
+Each image has two channels (VV and VH) and 1538 by 1556 pixels.
+We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}(m) = (\mathrm{VV}_{k,l}(m)^2+\mathrm{VH}_{k,l}(m)^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.
 
-A multi-resolution analysis (MRA) based on a Symlet basis with filter of length 16 (symlet 8) is built.
-The log-images are approximated at levels $J=1,2,3,4$.
-Table \ref{T:approxenergy} shows the 85 images' average energy for each approximation level.
-We notice that roughly 99\% of the energy is recovered with $J=1$, and more than  90\% with $J=2$.
-The VV channel shows better overall energy recovery than the VH channel.
-For $J=4$ and $J=3$, the Euclidean combination of the polarization channels increases the energy representation percentage.
-For $J=2$, VV channel and combined channels are equivalent.
-VH has 4\% less energy than VV for $J=3$, and $10\%$ less  for $J=4$.
-
-\begin{table*}[h!]
-\caption{Wavelet Approximation Mean Energy Percentage for log-images.
-Forest region at the border of Brazil and the French Guiana from November 08, 2015 to December 09, 2017.
-$n=85$ multi-date satellite images.
-Each image has two channels and 1200 by 1000 pixels.
-VV Polarization Channel;  VH Polarization Channel; and the Combined Image by Euclidean norm.
-Approximation $J=1,2,3,4$.}
-\centering
-\begin{tabular}{cccc|cccc|cccc}
-\hline
-\multicolumn{12}{c}{\sc Mean Approximated Energy Percentage}\\
-\hline
-\multicolumn{4}{c|}{VV Channel}&\multicolumn{4}{c|}{VH Channel}&\multicolumn{4}{c}{Combined Channels}\\
-\hline
-$J=4$&$J=3$&$J=2$&$J=1$&$J=4$&$J=3$&$J=2$&$J=1$&$J=4$&$J=3$&$J=2$&$J=1$\\
-\hline
-0.803&0.847&0.924&0.990&0.763&0.814&0.908&0.988&0.816&0.858&0.931&0.991\\
-\hline
-\end{tabular}\label{T:approxenergy}
-\end{table*}
-
-Figures \ref{F:squared_J1-4_VV}-\ref{F:squared_J1-4_euclid} show the series of squared deviations $\vd(m)$ and $\vt(m)$, for  the VV channel, VH channel, and Euclidean combination, respectively.
-An overall feature on this data is that the VV polarization presents much higher energy than the VH.
-The amount of energy related to changes  is ten times higher on the former compared to the latter's.
-
-Regarding change time points, in each figure, we can notice a pattern of peaks which are common to all approximation levels. They are time points:
-\begin{list}{}{}
-\item (a) 14, 43, 54, and 58 by the coefficients' squared deviations on the average VV image;
-\item (b) 14, 43, 54, and 58 by the coefficients' squared deviations on the consecutive VV images;
-\item (c) 14, 38, 41, 43, 54, 56, and 58 by the coefficients' squared deviations on the average VH image;
-\item (d) 14, 38, 41, 43, 54-58 by the coefficients' squared deviations on the consecutive VH images;
-\item (e) 14, 43, 54, and 58 by the coefficients' squared deviations on the average combined channels image; and
-\item (e) 14, 43, 54, and 58 by the coefficients' squared deviations on the consecutive combined channels images.
-%%%%%%%% VV 12 14 15 16 (42) 43 54 56 58 d(m)
-%%%%%%%% VV 12 14 15 16 (42) 43 54 56 58 t(m)
-%%%%%%%% VH  12 14 15 38 41 43 54 56 58
-%%%%%%%% euclid 12 14 15 16 43 54 56 58
-\end{list}
+A multi-resolution analysis (MRA) based on a Symlet basis with filter of length 16 (symlet 8) is built.  The images are approximated at resolution level $J=2$.
 
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/J1_VV_squared_meandev}(a)
-\includegraphics[scale=.12]{../../figs/consecdif_J1_VV_squared_meandev}(b)\\
-\includegraphics[scale=.12]{../../figs/J2_VV_squared_meandev}(c)
-\includegraphics[scale=.12]{../../figs/consecdif_J2_VV_squared_meandev}(d)\\
-\includegraphics[scale=.12]{../../figs/J3_VV_squared_meandev}(e)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_squared_meandev}(f)\\
-\includegraphics[scale=.12]{../../figs/J4_VV_squared_meandev}(g)
-\includegraphics[scale=.12]{../../figs/consecdif_J4_VV_squared_meandev}(h)
-\caption{{\sc VV Polarization Channel} Series of squared deviations $\vd(m)$ and $\vt(m)$.
-The red horizontal line represents the median value and the yellow horizontal line represents their median plus two times their absolute median deviation.
-$\vd(m)$ - Approximation Levels:  (a) $J=1$; (c) $J=2$; (e) $J=3$; (g) $J=4$.
-$\vt(m)$ - Approximation Levels:  (b) $J=1$; (d) $J=2$; (f) $J=3$; (h) $J=4$.
-}
-\label{F:squared_J1-4_VV}
+\includegraphics[scale=.13]{../../figs/forest_vSumDifCoefSq}\hspace{-.5cm}(a) 
+\includegraphics[scale=.13]{../../figs/forest_wecs_abscorr}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.11]{../../figs/forest_wecs_change_space}\hspace{-.5cm}(c) 
+\includegraphics[scale=.13]{../../figs/forest_aggreg_ratios}\hspace{-.5cm}(d)\\ 
+\caption{Analysis of change points in time and space of the forest data.
+(a) Plot of $\vd(m)$, $m=1,\ldots,85$ with orange horizontal line indicating two times its median absolute deviation.
+(b) Matrix of absolute correlations obtained with WECS.
+(c) Change regions detected by highlighting the $[n/\log n]$ largest correlations.
+(d) Change measured with aggregation of absolute differences.}
+\label{F:forest_wecs}
 \end{figure}
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/J1_VH_squared_meandev}(a)
-\includegraphics[scale=.12]{../../figs/consecdif_J1_VH_squared_meandev}(b)\\
-\includegraphics[scale=.12]{../../figs/J2_VH_squared_meandev}(c)
-\includegraphics[scale=.12]{../../figs/consecdif_J2_VH_squared_meandev}(d)\\
-\includegraphics[scale=.12]{../../figs/J3_VH_squared_meandev}(e)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_squared_meandev}(f)\\
-\includegraphics[scale=.12]{../../figs/J4_VH_squared_meandev}(g)
-\includegraphics[scale=.12]{../../figs/consecdif_J4_VH_squared_meandev}(h)
-\caption{{\sc VH Polarization  Channel} Series of squared deviations $\vd(m)$ and $\vt(m)$.
-The red horizontal line represents the median value and the yellow horizontal line represents their median plus two times their absolute median deviation.
-$\vd(m)$ - Approximation Levels:  (a) $J=1$; (c) $J=2$; (e) $J=3$; (g) $J=4$.
-$\vt(m)$ - Approximation Levels:  (b) $J=1$; (d) $J=2$; (f) $J=3$; (h) $J=4$.
-}
-\label{F:squared_J1-4_VH}
+\includegraphics[scale=.3]{../../figs/forest_changes_time} 
+\caption{Images whose time $m$ correspond to points of $\vd(m)$ above two times its median absolute deviation in Figure \ref{F:forest_wecs}a.}
+\label{F:forest_change_times}
 \end{figure}
-
-\begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/J1_euclid_squared_meandev}(a)
-\includegraphics[scale=.12]{../../figs/consecdif_J1_euclid_squared_meandev}(b)\\
-\includegraphics[scale=.12]{../../figs/J2_euclid_squared_meandev}(c)
-\includegraphics[scale=.12]{../../figs/consecdif_J2_euclid_squared_meandev}(d)\\
-\includegraphics[scale=.12]{../../figs/J3_euclid_squared_meandev}(e)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_squared_meandev}(f)\\
-\includegraphics[scale=.12]{../../figs/J4_euclid_squared_meandev}(g)
-\includegraphics[scale=.12]{../../figs/consecdif_J4_euclid_squared_meandev}(h)
-\caption{{\sc Combined Channels} Series of squared deviations $\vd(m)$ and $\vt(m)$.
-The red horizontal line represents the median value  and the yellow horizontal line represents their median plus two times their absolute median deviation.
-$\vd(m)$ - Approximation Levels:  (a) $J=1$; (c) $J=2$; (e) $J=3$; (g) $J=4$.
-$\vt(m)$ - Approximation Levels:  (b) $J=1$; (d) $J=2$; (f) $J=3$; (h) $J=4$.
-}
-\label{F:squared_J1-4_euclid}
-\end{figure}
-
-\begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/J3_VV_squared_meandev}(a)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_squared_meandev}(b)\\
-\includegraphics[scale=.12]{../../figs/J3_VV_sqrdif_change_locations}(c)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_sqrdif_change_locations}(d)\\
-\includegraphics[scale=.12]{../../figs/J3_VV_sqrdif_lowest_cor_locations}(e)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_sqrdif_lowest_cor_locations}(f)\\
-\includegraphics[scale=.12]{../../figs/perc01_J3_VV_sqrdif_change_locations}(g)
-\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VV_sqrdif_change_locations}(h)\\
-\includegraphics[scale=.12]{../../figs/perc01_J3_VV_sqrdif_lowest_cor_locations}(i)
-\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VV_sqrdif_lowest_cor_locations}(j)
-\caption{{\sc VV Polarization Channel} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$.
-Red horizontal line represents the median value and yellow line represents two absolute median deviations beyond the median.
-Squared approximation coefficient deviations: 0.1\% highest absolute correlations - (c) $\vd(m)$;  (d) $\vt(m)$; 0.1\% smallest absolute correlations - (e) $\vd(m)$; (f) $\vt(m)$; 0.01\% highest absolute correlations - (g) $\vd(m)$;  (h) $\vt(m)$;  0.01\% smallest absolute correlations - (i) $\vd(m)$; (j)  $\vt(m)$.
-}  
-\label{F:squared_meandev_J3_VV}
-\end{figure}
-
-\begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/J3_VH_squared_meandev}(a)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_squared_meandev}(b)
-
-\includegraphics[scale=.12]{../../figs/J3_VH_sqrdif_change_locations}(c)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_sqrdif_change_locations}(d)
-
-\includegraphics[scale=.12]{../../figs/J3_VH_sqrdif_lowest_cor_locations}(e)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_sqrdif_lowest_cor_locations}(f)
-
-\includegraphics[scale=.12]{../../figs/perc01_J3_VH_sqrdif_change_locations}(g)
-\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VH_sqrdif_change_locations}(h)
-
-\includegraphics[scale=.12]{../../figs/perc01_J3_VH_sqrdif_lowest_cor_locations}(i)
-\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_VH_sqrdif_lowest_cor_locations}(j)
-
-\caption{{\sc VH Polarization Channel} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$.
-Red horizontal line represents the median value and yellow line represents two absolute median deviations beyond the median.
-Squared approximation coefficient deviations: 0.1\% highest absolute correlations - (c) $\vd(m)$;  (d) $\vt(m)$; 0.1\% smallest absolute correlations - (e) $\vd(m)$; (f) $\vt(m)$; 0.01\% highest absolute correlations - (g) $\vd(m)$;  (h) $\vt(m)$;  0.01\% smallest absolute correlations - (i) $\vd(m)$; (j)  $\vt(m)$.
-}  
-\label{F:squared_meandev_J3_VH}
-\end{figure}
-
-\begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/J3_euclid_squared_meandev}(a)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_squared_meandev}(b)\\
-\includegraphics[scale=.12]{../../figs/J3_euclid_sqrdif_change_locations}(c)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_sqrdif_change_locations}(d)\\
-\includegraphics[scale=.12]{../../figs/J3_euclid_sqrdif_lowest_cor_locations}(e)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_sqrdif_lowest_cor_locations}(f)\\
-\includegraphics[scale=.12]{../../figs/perc01_J3_euclid_sqrdif_change_locations}(g)
-\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_euclid_sqrdif_change_locations}(h)\\
-\includegraphics[scale=.12]{../../figs/perc01_J3_euclid_sqrdif_lowest_cor_locations}(i)
-\includegraphics[scale=.12]{../../figs/perc01_consecdif_J3_euclid_sqrdif_lowest_cor_locations}(j)
-\caption{{\sc Combined Channels} Series of squared mean deviations at level $J=3$: (a) $\vd(m)$; (b) $\vt(m)$.
-Red horizontal line represents the median value and yellow line represents two absolute median deviations beyond the median.
-Squared approximation coefficient deviations: 0.1\% highest absolute correlations - (c) $\vd(m)$;  (d) $\vt(m)$; 0.1\% smallest absolute correlations - (e) $\vd(m)$; (f) $\vt(m)$; 0.01\% highest absolute correlations - (g) $\vd(m)$;  (h) $\vt(m)$;  0.01\% smallest absolute correlations - (i) $\vd(m)$; (j)  $\vt(m)$.
-} 
-\label{F:squared_meandev_J3_euclid}
-\end{figure}
-
-Figures \ref{F:squared_meandev_J3_VV}-\ref{F:squared_meandev_J3_euclid} present a timewise comparison between overall energy variations, $\vd(m)$ and $\vt(m)$ and their respective individual coefficients.
-Each figure has ten panels.
-Panels on the left and right deal with the first and second change detection methods respectively.
-A full description is give in each figure's caption.
-The overall conclusion from these figures is that as expected there are temporal variations between the methods regarding change detection (Panels (a)-(b)).
-The correlation screening detects the most relevant indices, as well as the least relevant ones  (Panels (c)-(f)).
-These conclusions hold for analyses based upon VV, VH or combined polarizations, but the VV polarization signal is much stronger than VH's.
-A slight advantage is perceived for the $\vd(m)$ method as opposed to $\vt(m)$'s.
-This makes sense, since we are dealing here with data from a forest region over a long time, and seasonal changes will be perceived more easily on the first proposed method.
-
-\begin{table*}[ht!]
-\caption{Absolute correlation thresholds and number of selected coefficients for $n=85$ multi-temporal images of $1200\times 1000$ .
-Correlation was computed between approximation coefficients and approximation total energy at level $J=2$ for each image.
-}\label{tabela_quantis}
-\centering
-\begin{tabular}{c|cccr||c|cccr}
-\hline
-Qtile &\multicolumn{4}{c||}{\sc Correlation Thresh} &Qtile & \multicolumn{4}{c}{\sc Correlation Thresh}  \\
-Level &VV&VH&Comb& Coeffs &Level &VV&VH&Comb& Coeffs \\
-\hline
-0.50&0.281&0.199&0.292&600000&0.99&0.647&0.582&0.668&12000\\
-0.55&0.300&0.218&0.311&540000&0.991&0.651&0.587&0.673&10800\\
-0.60&0.319&0.239&0.331&480000& 0.992&0.656&0.594&0.678&9600\\
-0.65&0.339&0.260&0.352&420000& 0.993&0.662&0.601&0.684&8400\\
-0.70&0.361&0.282&0.375&360000& 0.994&0.669&0.608&0.690&7200\\
-0.75&0.385&0.307&0.399&300000& 0.995&0.676&0.617&0.697&6000\\
-0.80&0.412&0.335&0.428&240000& 0.996&0.685&0.626&0.705&4800\\
-0.85&0.445&0.368&0.462&180000& 0.997&0.696&0.639&0.716&3600\\
-0.90&0.487&0.410&0.506&120000& 0.998&0.710&0.654&0.729&2400\\
-0.95&0.549&0.472&0.570&60000& 0.999&0.731&0.676&0.748&1200\\
-\hline
-\end{tabular}
-\end{table*}
-
-\begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/J2_VV_image_cor}(a)
-\includegraphics[scale=.12]{../../figs/J2_VH_image_cor}(b)\\
-\includegraphics[scale=.12]{../../figs/J2_euclid_image_cor}(c)
-\includegraphics[scale=.12]{../../figs/J3_VV_image_cor}(d)\\ 
-\includegraphics[scale=.12]{../../figs/J3_VH_image_cor}(e)
-\includegraphics[scale=.12]{../../figs/J3_euclid_image_cor}(f)
-
-\caption{Absolute correlation matrices $\vR^{(d)}$ between mean-corrected squared coefficients at approximation levels and overall mean-corrected total energy $\{\vd(m)\}$.
-$J=2$ (a)-(c);  $J=3$ (d)-(f).
-$n=85$ multi-temporal images of $1200\times 1000$ ($J=10$).
-The color bar on the right gives the magnitude of correlation at all positions.
-(a) {\sc VV  Polarization - Channel} $J=2$.
-(b) {\sc VH  Polarization - Channel} $J=2$.
-(c) {\sc Combined Channels} $J=2$.
-(d)  {\sc VV  Polarization - Channel} $J=3$.
-(e) {\sc VH  Polarization - Channel} $J=3$.
-(f) {\sc Combined Channels} $J=3$}
-\label{F:image_corr_J2J3}
-\end{figure}
-
-\begin{figure}[htp!]
-\includegraphics[scale=.12]{../../figs/consecdif_J2_VV_image_cor}(a)
-\includegraphics[scale=.12]{../../figs/consecdif_J2_VH_image_cor}(b)\\
-\includegraphics[scale=.12]{../../figs/consecdif_J2_euclid_image_cor}(c)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VV_image_cor}(d)\\ 
-\includegraphics[scale=.12]{../../figs/consecdif_J3_VH_image_cor}(e)
-\includegraphics[scale=.12]{../../figs/consecdif_J3_euclid_image_cor}(f)
-
-\caption{Absolute correlation matrices $\vR^{(t)}$ between squared differences of consecutive log-images' coefficients at approximation levels and their overall total energy $\{\vt(m)\}$.
-Levels $J=2$ (a)-(c);  $J=3$ (d)-(f).
-$n=85$ multi-temporal images of $1200\times 1000$.
-The color bar on the right gives the magnitude of correlation at all positions.
-(a) {\sc VV  Polarization - Channel} $J=2$.
-(b) {\sc VH  Polarization - Channel} $J=2$.
-(c) {\sc Combined Channels} $J=2$.
-(d)  {\sc VV  Polarization - Channel} $J=3$.
-(e) {\sc VH  Polarization - Channel} $J=3$.
-(f) {\sc Combined Channels} $J=3$.}
-\label{F:image_corr_J2J3_consec}
-\end{figure}
-
-Figures \ref{F:image_corr_J2J3}-\ref{F:image_corr_J2J3_consec} shows the absolute correlation images for the VV, VH and combined channels for levels $J=2$ and $J=3$.
-We can notice that high correlation coefficients for $J=2$ are high correlation coefficients for $J=3$ as well.
-Moreover, there is a clear spatial connection between polarizations and between $\vd(m)$- and $\vt(m)$-based  analyses.
-On the other hand, correlations are more efficiently segregated when we move: from $J=2$ to $J=3$; from VH to VV polarization; or from  $\vt(m)$ to $\vd(m)$.
 
@@
 
-We present a novel way of detecting changes in multi-temporal satellite images, WECS.
+%We present a novel way of detecting changes in multi-temporal satellite images, WECS.
 The procedure is based on wavelet energies from both the estimated individual coefficients as well as the whole image approximation.

### 2021-11-02 codes/EllispseChangeAnalysis_methods.m

@@ -123,11 +123,11 @@ S = abs(im4log - im3log) + abs(im3log - im2log) + abs(im2log - im1log);
 S = S./max(S(:));
 %
 [pD1,pFA1]=ROCcurveNew(S,255*totalchanges); close
-[vp,vF1_std,~,~] = F1Scorecurve(R,255*totalchanges); close
+[vp,vF1_std,~,~] = F1Scorecurve(S,255*totalchanges); close
 %
 mImage = figure;
 imshow(S)
-title('Aggregation of log-ratios'                                                                                                                                                                                                                                                                                                                               , 'FontSize', 17)
+title('Aggregation of absolute differences'                                                                                                                                                                                                                                                                                                                               , 'FontSize', 17)
 saveas(mImage,sprintf('../figs/corr_changes_logratios.jpg'))
 
 %% Compare all results
@@ -143,7 +143,7 @@ plot(pFA,pD,'k')
 plot(pFA0,pD0,'g-d')
 plot(pFA1,pD1,':+')
 legend('db2 WECS $\textbf{d}(m)$, $J=2$', '$\textbf{d}(m)$: without wavelets', ...
-    'Aggregation of log-ratios','interpreter','latex', 'Location','southeast', 'FontSize', 12)
+    'Aggregation of absolute differences','interpreter','latex', 'Location','southeast', 'FontSize', 12)
 legend('boxoff')
 hold off
 saveas(mImage,sprintf('../figs/methods_comparison.jpg'))
@@ -159,7 +159,7 @@ plot(vp,vF1_wave,'k')
 plot(vp,vF1_nowave,'g-d')
 plot(vp,vF1_std,':+')
 legend('db2 WECS $\textbf{d}(m)$, $J=2$', '$\textbf{d}(m)$: without wavelets', ...
-    'Aggregation of log-ratios','interpreter','latex', 'Location','southeast', 'FontSize', 12)
+    'Aggregation of absolute differences','interpreter','latex', 'Location','southeast', 'FontSize', 12)
 legend('boxoff')
 hold off
 saveas(mImage,sprintf('../figs/methods_comparison_F1score.jpg'))

### 2021-11-02 codes/Forest_data_analysis.m

@@ -1,0 +1,161 @@
+
+clear 
+close all
+clc
+
+%% Compute the spatial coefficient of variation per polarization for every acqusition
+
+% dates corresponding to each image
+dates = readtable('../../Images/timeSeries/ascending/saida.txt','ReadVariableNames',false);
+[n,~] = size(dates);
+n = n - 2; % the last two elements in the table are not dates
+% Sorting the dates according to time
+tmp = dates;
+for m=1:n
+    tmp2 = char(tmp{m,:});
+    if tmp2(7)=='-'
+        tmp{m,:} = {insertAfter(tmp2,5,'0')};
+    end
+end
+[~, dates_idsort] = sortrows(tmp(1:n,1));
+dates = dates(dates_idsort,:);
+
+% reading the first image
+t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{1,:})),'r');
+Y = read(t);
+imagesc(sqrt(Y(:,:,1).^2 + Y(:,:,2).^2))
+
+%
+[Nx,Ny,Nc] = size(Y);
+% Nx * Ny : spatial dimension
+% Nc : number of channels, here: VV and VH
+
+%% Configurations for wavelet decompositions
+
+wname='sym8';
+%wname='db2';
+
+% saving the current boundary handling and then changing it to periodic.
+% The default mode was giving different norms of vC for different
+% decomposition levels J
+origMode = dwtmode('status','nodisplay');
+dwtmode('per');
+
+% % this value will be added to images when taking logs
+% eps = min(min(min(min(data(:,:,:,:)))));
+% eps = abs(eps)*(eps<0)  + .001;
+
+
+%% Analysis of squared mean deviations
+
+
+%********************************************%
+%******** analysis for combined chanel ************%
+
+% matrix of mean squared differences of approximation coefficients
+mDifCoefSq = zeros([Nx, Ny, n]);
+
+% mean image
+imRef = zeros(Nx,Ny);
+for m=1:n
+    t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{m,:})),'r');
+    Y = read(t);
+    imRef = imRef + double(Y(:,:,1).^2 + Y(:,:,2).^2);
+end
[…]

### 2021-11-03 tex/paper.wecs/wecs_tgrs.tex
Updates on application

@@
 
-We employed the proposed change detection method on a series of 85 multi-date satellite images.
+We employed the proposed change detection method on a series of 84 multi-date satellite images.
 The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3.
@@
 \includegraphics[scale=.13]{../../figs/forest_vSumDifCoefSq}\hspace{-.5cm}(a) 
-\includegraphics[scale=.13]{../../figs/forest_wecs_abscorr}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.11]{../../figs/forest_wecs_change_space}\hspace{-.5cm}(c) 
-\includegraphics[scale=.13]{../../figs/forest_aggreg_ratios}\hspace{-.5cm}(d)\\ 
+\includegraphics[scale=.14]{../../figs/forest_wecs_abscorr}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.10]{../../figs/forest_wecs_change_space}\hspace{-.5cm}(c) 
+\includegraphics[scale=.14]{../../figs/forest_aggreg_ratios}\hspace{-.5cm}(d)\\ 
 \caption{Analysis of change points in time and space of the forest data.

### 2021-11-03 codes/Forest_data_analysis.m

@@ -102,7 +102,7 @@ saveas(mImage,sprintf('../figs/forest_vSumDifCoefSq.jpg'))
 % checking the images corresponding to change points in time
 mImage = figure;
 cont = 1;
-for m = [4 8 41]
+for m = [25 27 30]
     t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{m,:})),'r');
     Y = read(t);
     data = double(Y(:,:,1).^2 + Y(:,:,2).^2);
@@ -157,5 +157,28 @@ imagesc(S)
 axis off; colorbar
 set(gca,'FontSize',13)
 saveas(mImage,sprintf('../figs/forest_aggreg_ratios.jpg'))
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% making a video with the image time series
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% create the video writer with 1 fps
+writer_im = VideoWriter('images-ts.avi');
+% number of frames shown in a second
+writer_im.FrameRate = 2;
+% open the video writer
+open(writer_im);
+% write the frames to the video
+for m = 1:n
+    t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{m,:})),'r');
+    Y = read(t);
+    data = double(Y(:,:,1).^2 + Y(:,:,2).^2);
+    imagesc(data)
+    title(sprintf('m=%2d',m)); axis off;
+    frame = getframe(gcf) ;
+    writeVideo(writer_im, frame);    
+end
+% close the writer object
+close(writer_im);
 
 

### 2021-11-03 tex/paper.wecs/wecs_tgrs.tex
updates on tex

@@
 In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
-For each local squared mean deviation time series, given by individual elements of $\vD(m)$ across $m=1,\ldots,n$, consider its Pearson correlation with the overall squared mean deviations, given by $\vd$:
+For each local squared mean deviation time series, given by individual elements of $\vD(m)$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
 \begin{equation*}
-R_{k,l}= \corr\left( \vD_{k,l}, \vd\right),
+R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
 \label{def_Rkld}
@@
 
-A multi-resolution analysis (MRA) based on a Symlet basis with filter of length 16 (symlet 8) is built.  The images are approximated at resolution level $J=2$.
+A multi-resolution analysis (MRA) based on a Symlet basis with filter of length 16 (symlet 8) is built.
+In order to have dimensions as power of 2, the matrices $\mathcal{I}_{k,l}$ were extended to a $2048\times2048$ matrix with $\mathcal{I}_{k,l}$ at the center and the remaining parts being completed with mirrored values at the borders.
+The wavelet transform at resolution level $J=2$ was applied to these matrices.
+Then, we are able to compute the squared mean differences vector $\vd$ and the matrix of absolute correlations $\vR$.
+
+The time change measures computed with $\vd$ are shown in Figure \ref{F:forest_wecs}a.
+The orange line in Figure \ref{F:forest_wecs}a represents the median absolute deviation of $\vd$, which allow us to notice times that differ expressively from the others.
+We can notice that times $m=25,27,30$ are highlighted as having expressive changes.
+The images  corresponding to these times can be seen in Figure \ref{F:forest_change_times}.
+
+The changes in space can be analyzed using the image obtained with $\vR$, which is displayed in Figure \ref{F:forest_wecs}b.
+Taking the $[n/\log n]$ ($n=1538\times1556$) largest absolute correlations as those corresponding to possible change points, we obtain a matrix of zeros and ones that is presented in Figure \ref{F:forest_wecs}c.
+The white regions in Figure \ref{F:forest_wecs}c (entries with value one) represent the change points, which seem to concentrate mainly on three regions: at the center, to the right of the river; at the top, on the left border of the river; and at the top left.
+Computing aggregated absolute differences to measure changes in space, we obtain Figure \ref{F:forest_wecs}d.
+The first two regions captured using WECS are also detected in Figure \ref{F:forest_wecs}d, but the former offers higher contrasts to these change regions.
 
@@
 
-%We present a novel way of detecting changes in multi-temporal satellite images, WECS.
-The procedure is based on wavelet energies from both the estimated individual coefficients as well as the whole image approximation.
+We present a novel way of detecting changes in multi-temporal satellite images, WECS.
+The procedure is based on wavelet energies from both the estimated individual coefficients as well as the whole mean image.
 It makes use of correlation screening for ultra-high dimensional data.
 The proposed method's performance is shown using both synthetic and real data.
-The proposed method yields spatio-temporal  change points.
-Its performance is similar to a deep-learning feature extraction method, but the computational cost of the proposed method is 180 times smaller than the cost for the latter.
-Therefore, we may say that WECS may be used on images with equivalent performance for a fraction of the computational cost.
-Because of its reliance on wavelet representation and correlation screening, it is sparse, very fast and scalable.
-Finally, it is easily adapted to be updatable, so that real-time change detection is feasible even with a portable computer.
+The proposed method is useful to detect spatio-temporal change points, which is illustrated on data analyses.
+The method is employed to analyze a time series of 84 images of a forest.
 

### 2021-11-03 tex/paper.wecs/wecs_tgrs.tex
included analysis of change/nonchange regions

@@
 The first two regions captured using WECS are also detected in Figure \ref{F:forest_wecs}d, but the former offers higher contrasts to these change regions.
+The comparison between the performance of these two methods can also be checked in Figure \ref{F:forest_roc_wecs_agg}, where a ROC curve is computed to check the correct detection of changing and not-changing regions.
+These change/nonchange regions were determined using [XXXXX].
+We can notice that WECS reaches high correct detection earlier than the method that uses aggregation of differences, but both methods do not have good performances concerning the nonchange regions.
 
@@
 \label{F:forest_change_times}
+\end{figure}
+
+\begin{figure}[htp!]
+\includegraphics[scale=.11]{../../figs/forest_change}\hspace{-.5cm}(a)
+\includegraphics[scale=.11]{../../figs/forest_nonchange}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.13]{../../figs/forest_roc_change}\hspace{-.5cm}(c) 
+\includegraphics[scale=.14]{../../figs/forest_roc_nonchange}\hspace{-.5cm}(d)
+\caption{Analysis of change points in time and space of the forest data.
+(a) Regions (in white) where changes should be detected.
+(a) Regions (in white) where changes are not expected to be detected.
+(c) ROC curve for detection of changing regions.
+(d) ROC curves for detection of nonchange regions.}
+\label{F:forest_roc_wecs_agg}
 \end{figure}

### 2021-11-03 codes/Forest_data_analysis.m

@@ -158,6 +158,57 @@ axis off; colorbar
 set(gca,'FontSize',13)
 saveas(mImage,sprintf('../figs/forest_aggreg_ratios.jpg'))
 
+
+% reading the reference image with change and nonchange locations
+t = Tiff('../figs/change_nonchange.tif','r');
+Y = read(t);
+mChange = double(Y(:,:,1));
+mImage = figure;
+imshow(mChange)
+saveas(mImage,sprintf('../figs/forest_change.jpg'))
+
+mNonchange = double(Y(:,:,2));
+mImage = figure;
+imshow(mNonchange)
+saveas(mImage,sprintf('../figs/forest_nonchange.jpg'))
+
+% detection of change regions
+[D_wecs,FA_wecs] = ROCcurveNew(mCorr/max(mCorr(:)),mChange); close
+[D_agg,FA_agg] = ROCcurveNew(S/max(S(:)),mChange); close
+
+mImage = figure;
+hold on
+xlabel('False positive rate', 'FontSize', 13)
+ylabel('True positive rate', 'FontSize', 13)
+axis([0 1 0 1]);
+axis square
+plot(FA_wecs,D_wecs,'k')
+plot(FA_agg,D_agg,'g-d')
+legend('WECS','Aggregation of absolute differences',...
+    'Location','southeast', 'FontSize', 12)
+legend('boxoff')
+hold off
+saveas(mImage,sprintf('../figs/forest_roc_change.jpg'))
+
+% detection of nonchange regions
+[D_wecs,FA_wecs] = ROCcurveNew(1 - mCorr/max(mCorr(:)),mNonchange); close
+[D_agg,FA_agg] = ROCcurveNew(1 - S/max(S(:)),mNonchange); close
+
+mImage = figure;
+hold on
+xlabel('False positive rate', 'FontSize', 13)
+ylabel('True positive rate', 'FontSize', 13)
+axis([0 1 0 1]);
+axis square
+plot(FA_wecs,D_wecs,'k')
+plot(FA_agg,D_agg,'g-d')
+legend('WECS','Aggregation of absolute differences',...
+    'Location','southeast', 'FontSize', 12)
+legend('boxoff')
+hold off
+saveas(mImage,sprintf('../figs/forest_roc_nonchange.jpg'))
+
+
[…]

### 2021-11-04 codes/saving_tif_image.m
file to save tif images

@@ -1,0 +1,16 @@
+% saving a matrix of doubles as tiff image
+filename = 'testtifffp.tiff';
+A = mCorr;
+t = Tiff(filename, 'w');
+tagstruct.ImageLength = size(A, 1);
+tagstruct.ImageWidth = size(A, 2);
+tagstruct.Compression = Tiff.Compression.None;
+tagstruct.SampleFormat = Tiff.SampleFormat.IEEEFP;
+tagstruct.Photometric = Tiff.Photometric.MinIsBlack;
+tagstruct.BitsPerSample = 64;
+tagstruct.SamplesPerPixel = size(A,3);
+tagstruct.PlanarConfiguration = Tiff.PlanarConfiguration.Chunky;
+t.setTag(tagstruct);
+t.write(A);
+t.close();
+

### 2021-11-06 codes/EllispseChangeAnalysis_families.m
code for a larger sequence of synthetic images

@@ -5,41 +5,42 @@
 close all
 clear;
 %%
-im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
-im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
-im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
-im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
 
-%
-% Images are too big, consider subsampling
-subsamplingfactor = 2;
-im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
-totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
-totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
-%
-figure
-imshow(totalchanges)
-title('Total changes')
+% code that generates a sequence of synthetic images
 
-%%
-eps = .00001;
-im1log = eps+double(im1);
-im2log = eps+double(im2);
-im3log = eps+double(im3);
-im4log = eps+double(im4);
-n=4;
-imRef = im1log + im2log + im3log + im4log;
-    imRef = imRef/n;
+% Subsampling of images
+subsamplingfactor = 8;
+% noise level used
+sig = 1;
+% size of the sequence (must be multiple of 4)
+n = 80;
+% running code
+generate_synthetic_sequence
+
+% output
+size(mI) % true images
+size(mY) % sequence of synthetic images
+% number of rows and columns
+n1 = size(mY,1);
+n2 = size(mY,2);
+
+% all changes that are expected to be detected
+totalchanges = (abs(mI(:,:,1)-mI(:,:,2)) + abs(mI(:,:,2)-mI(:,:,3)) + ...
+    abs(mI(:,:,3)-mI(:,:,4)) + abs(mI(:,:,4)-mI(:,:,1)))>0;
+mImage = figure;
+colormap(gray(256)); imagesc(totalchanges)
+title('Total changes', 'FontSize', 17)
[…]

### 2021-11-06 codes/EllispseChangeAnalysis_methods.m

@@ -24,7 +24,7 @@ totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
 mImage = figure;
 imshow(totalchanges)
 title('Total changes', 'FontSize', 17)
-saveas(mImage,sprintf('../figs/total_changes.jpg'))
+%saveas(mImage,sprintf('../figs/total_changes.jpg'))
 
 %%
 eps = .00001;

### 2021-11-06 codes/SeqEllispseChangeAnalysis.m

@@ -1,0 +1,269 @@
+
+% Evaluation of different levels and families on WECS and comparison with a
+% deep-learning denoiser
+
+%%
+close all
+clear;
+%%
+
+% code that generates a sequence of synthetic images
+
+% Subsampling of images
+subsamplingfactor = 2;
+% noise level used
+sig = 1;
+% size of the sequence (must be multiple of 4)
+n = 80;
+% running code
+generate_synthetic_sequence
+
+% output
+size(mI) % true images
+size(mY) % sequence of synthetic images
+% number of rows and columns
+n1 = size(mY,1);
+n2 = size(mY,2);
+
+% all changes that are expected to be detected
+totalchanges = (abs(mI(:,:,1)-mI(:,:,2)) + abs(mI(:,:,2)-mI(:,:,3)) + ...
+    abs(mI(:,:,3)-mI(:,:,4)) + abs(mI(:,:,4)-mI(:,:,1)))>0;
+mImage = figure;
+imshow(totalchanges)
+title('Total changes', 'FontSize', 17)
+saveas(mImage,sprintf('../figs/total_changes.jpg'))
+
+%% Analysis of decomposition level
+
+% wavelet basis used
+wname = 'db2'; 
+
+R_levels = zeros(n1,n2,5);
+
+for J=1:5
+    [tmp,~] = wecs(mY,wname,J);
+    R_levels(:,:,J) = tmp/max(tmp(:));
+    mImage = figure;
+    imshow(R_levels(:,:,J))
+    title(sprintf('WECS: %s, J=%d',wname,J))
+end
+
+
+% Compare all results
+
+[TP_J1,FP_J1] = ROCcurveNew(R_levels(:,:,1),255*totalchanges); close
+[TP_J2,FP_J2] = ROCcurveNew(R_levels(:,:,2),255*totalchanges); close
+[TP_J3,FP_J3] = ROCcurveNew(R_levels(:,:,3),255*totalchanges); close
+[TP_J4,FP_J4] = ROCcurveNew(R_levels(:,:,4),255*totalchanges); close
+[TP_J5,FP_J5] = ROCcurveNew(R_levels(:,:,5),255*totalchanges); close
+
+mImage = figure;
+hold on
+%title('ROC Curve', 'FontSize', 17)
[…]

### 2021-11-06 codes/SeqEllispseChangeAnalysis_methods.m

@@ -1,0 +1,169 @@
+
+% Evaluation of different methods
+
+%%
+close all
+clear;
+%%
+
+% code that generates a sequence of synthetic images
+
+% Subsampling of images
+subsamplingfactor = 2;
+% noise level used
+sig = 1;
+% size of the sequence (must be multiple of 4)
+n = 80;
+% running code
+generate_synthetic_sequence
+
+% output
+size(mI) % true images
+size(mY) % sequence of synthetic images
+% number of rows and columns
+n1 = size(mY,1);
+n2 = size(mY,2);
+
+% all changes that are expected to be detected
+totalchanges = (abs(mI(:,:,1)-mI(:,:,2)) + abs(mI(:,:,2)-mI(:,:,3)) + ...
+    abs(mI(:,:,3)-mI(:,:,4)) + abs(mI(:,:,4)-mI(:,:,1)))>0;
+mImage = figure;
+colormap(gray(256)); imagesc(totalchanges)
+title('Total changes', 'FontSize', 17)
+saveas(mImage,sprintf('../figs/total_changes.jpg'))
+
+% mean observed image
+imRef = mean(mY,3);
+
+%% WECS
+
+wname = 'db2'; % wavelet basis used
+J = 2; % resolution level of wavelet transform
+
+mX = zeros(n1,n2,n);
+for m=1:n
+    [tmp,~,~,~] = swt2(mY(:,:,m),J,wname);
+    mX(:,:,m) = tmp(:,:,J); % focus only on level J approximations
+end
+
+% matrix of squared mean differences
+mD = zeros(n1,n2,n);
+for m=1:n
+    mD(:,:,m) = (mX(:,:,m) - imRef).^2;
+end
+
+% vector of overall changes
+vd = reshape(sum(sum(mD,1),2),n,1);
+plot(1:n,vd)
+
+%
+R = zeros(n1,n2);
+for ii=1:n1
+   for jj=1:n2
+       tmp = abs(corrcoef(reshape(mD(ii,jj,:),n,1),vd));
+       R(ii,jj) = tmp(1,2);
+   end
+end
+R_wecs = R./max(R(:));
+
+%
+mImage = figure;
+imshow(R_wecs)
+title('db2 WECS d(m), J=2', 'FontSize', 17)
+saveas(mImage,sprintf('../figs/corr_changes_dm.jpg'))
+
+
+%% No wavelets for comparison with mean image
+
+% matrix of squared mean differences
+mD = zeros(n1,n2,n);
+for m=1:n
[…]

### 2021-11-06 codes/generate_synthetic.m

@@ -33,17 +33,17 @@ colormap(gray(256)); imagesc(im1_obs+im2_obs+im3_obs+im4_obs)
 % saving the synthetic images
 mImage = cast(100*(im1_obs),'uint8');
 imshow(mImage)
-imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png'))
+%imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png'))
 
 mImage = cast(100*(im2_obs),'uint8');
 imshow(mImage)
-imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png'))
+%imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png'))
 
 mImage = cast(100*(im3_obs),'uint8');
 imshow(mImage)
-imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png'))
+%imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png'))
 
 mImage = cast(100*(im4_obs),'uint8');
 imshow(mImage)
-imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png'))
+%imwrite(mImage,sprintf('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png'))
 

### 2021-11-06 codes/generate_synthetic_sequence.m

@@ -1,0 +1,55 @@
+
+% Code to generate a sequence synthetic images 
+
+%%
+% Parameters that must be specified
+
+% % Images are too big, consider subsampling
+% subsamplingfactor = 2;
+% % noise level used
+% sig = 1;
+% % size of the sequence (must be multiple of 4)
+% n = 80;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Loading images 
+im1=imread('../figs/GroundTruthEllipsoidChanges/ellipse1.tif');
+% boolean ~ used to represent change point with 1
+im1=double(~im1(1:subsamplingfactor:end,1:subsamplingfactor:end));
+
+im2=imread('../figs/GroundTruthEllipsoidChanges/ellipse2.tif');
+im2=double(~im2(1:subsamplingfactor:end,1:subsamplingfactor:end)); 
+im3=imread('../figs/GroundTruthEllipsoidChanges/ellipse3.tif');
+im3=double(~im3(1:subsamplingfactor:end,1:subsamplingfactor:end)); 
+im4=imread('../figs/GroundTruthEllipsoidChanges/ellipse4.tif');
+im4=double(~im4(1:subsamplingfactor:end,1:subsamplingfactor:end)); 
+% number of rows and columns
+n1 = size(im1,1);
+n2 = size(im1,2);
+
+
+sprintf('signal to noise ratios')
+sprintf('im1: %f',sqrt(sum(sum(im1.^2,1))/(n1*n2))/sig)
+sprintf('im2: %f',sqrt(sum(sum(im2.^2,1))/(n1*n2))/sig)
+sprintf('im3: %f',sqrt(sum(sum(im3.^2,1))/(n1*n2))/sig)
+sprintf('im4: %f',sqrt(sum(sum(im4.^2,1))/(n1*n2))/sig)
+
+% array of true images
+mI = zeros(n1,n2,4);
+mI(:,:,1) = im1; mI(:,:,2) = im2;
+mI(:,:,3) = im3; mI(:,:,4) = im4;
+clear im1 im2 im3 im4
+
+% array to store sequence of images
+mY = zeros(n1, n2, n);
+% sequence of images with noise
+cont = 1;
+while cont<=80
+    for m=1:4
+        mY(:,:,cont) = mI(:,:,m) + sig*randn(n1, n2);
+        cont = cont + 1;
+    end
+end
+
+

### 2021-11-06 codes/wecs.m

@@ -1,0 +1,45 @@
+function [R,vd] = wecs(mY,wname,J)
+% Function to apply the WECS to a time series of images
+% Input
+% mY: array of observed images
+% wname: wavelet basis
+% J: resolution level of wavelet decomposition
+
+% Output
+% R: matrix of correlations of local squared differences with vd
+% vd: overall squared mean differences
+
+% image dimension and sample size
+[n1,n2,n] = size(mY);
+
+% mean observed image
+imRef = mean(mY,3);
+
+mX = zeros(n1,n2,n);
+for m=1:n
+    [tmp,~,~,~] = swt2(mY(:,:,m),J,wname);
+    mX(:,:,m) = tmp(:,:,J); % focus only on level J approximations
+end
+
+% matrix of squared mean differences
+mD = zeros(n1,n2,n);
+for m=1:n
+    mD(:,:,m) = (mX(:,:,m) - imRef).^2;
+end
+
+% vector of overall changes
+vd = reshape(sum(sum(mD,1),2),n,1);
+%plot(1:n,vd)
+
+% matrix of correlations
+R = zeros(n1,n2);
+for ii=1:n1
+   for jj=1:n2
+       tmp = abs(corrcoef(reshape(mD(ii,jj,:),n,1),vd));
+       R(ii,jj) = tmp(1,2);
+   end
+end
+
+
+end
+

### 2021-11-06 tex/paper.wecs/wecs_tgrs.tex
updates on tex

@@
 In this section we apply the change detection methods above on synthetic data of multi-temporal images.
-The synthetic multi-temporal images ($n=4$) are shown in Figure \ref{F:EllipsoidChanges}.
-These images are generated from the sum of two matrices: a signal matrix with 1/2 on entries where ellipses occur and zero elsewhere, and a noise matrix with random variables following a Gaussian distribution with zero mean and variance $1/100$.
+The synthetic multi-temporal images ($n=80$) is obtained by repeating a sequence of four images with different types of changes plus a noise.
+Examples of the first four of these images are shown in Figure \ref{F:EllipsoidChanges}.
+These images are generated from the sum of two matrices: a signal matrix with ones on entries where ellipses occur and zero elsewhere, and a noise matrix with random variables following a standard Gaussian distribution.
 The first image, $\mathcal{I}(1)$, presents three elongated ellipses.
-Changes consist of three different types of ellipses that are successively added to the original image $\mathcal{I}(1)$.
-The second image, $\mathcal{I}(2)$, has new large ellipses added.
+The second image $\mathcal{I}(2)$ has shorter and larger ellipses added.
 Smaller ellipses are then added to form $\mathcal{I}(3)$ and $\mathcal{I}(4)$.
-All the changes made to image $\mathcal{I}(1)$ can be seen in Figure \ref{F:Changes_methods_images}(a), which displays a matrix of zeros and ones that correspond, respectively, to locations without and with changes along time.
+All the changes made that occur among subsequent images can be seen in Figure \ref{F:Changes_methods_images}(a), which displays a matrix of zeros and ones that correspond, respectively, to locations without and with changes along time.
 Applying WECS to these images, we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation.
@@
 %\includegraphics[width=40pc]{EllipsoidChanges/ellipses_all}
-\includegraphics[scale=.17]{../../figs/ellipses_t1}
-\includegraphics[scale=.17]{../../figs/ellipses_t2}
-\includegraphics[scale=.17]{../../figs/ellipses_t3}
-\includegraphics[scale=.17]{../../figs/ellipses_t4}
-\caption{Synthetic multi-temporal ($n=4$) images. Features and changes come as ellipses and dots.}
+\includegraphics[scale=.11]{../../figs/ellipses_t1}
+\includegraphics[scale=.11]{../../figs/ellipses_t2}
+\includegraphics[scale=.11]{../../figs/ellipses_t3}
+\includegraphics[scale=.11]{../../figs/ellipses_t4}
+\caption{Example of the first four synthetic multi-temporal images. Features and changes come as white ellipses and dots.}
 \label{F:EllipsoidChanges}
@@
 Panel (b) shows the result of WECS using Daubechies wavelet with two null moments (db2) and $J=2$.
-Panel (c) presents the result of using aggregated log-ratios, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{il} = \sum_{m=1}^3	|\log\left(\mathcal{I}_{k,l}(m+1)/\mathcal{I}_{k,l}(m)\right)|$.
+Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}(m) - \mathcal{I}_{k,l}(m-1)|$.
 Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}(m)$ instead of $\vX(m)$ in the WECS formulation.
@@
 \begin{figure}[htp!]
-\includegraphics[scale=.14]{../../figs/methods_comparison}\hspace{-.6cm}(a) 
-\includegraphics[scale=.14]{../../figs/levels_comparison}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.14]{../../figs/families_comparison}\hspace{-.5cm}(c)
-\includegraphics[scale=.14]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
+\includegraphics[scale=.13]{../../figs/methods_comparison}\hspace{-.6cm}(a) 
+\includegraphics[scale=.13]{../../figs/levels_comparison}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.13]{../../figs/families_comparison}\hspace{-.5cm}(c)
+\includegraphics[scale=.13]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
 \caption{ROC curves for detection of changing ellipses in synthetic images and different methods.
@@
 Panel (b) presents the ROC curves for five different levels of decomposition $J=\{1,2,3,4,5\}$ under the Daubechies db2 basis.
-Levels $J=2,3$ have a clear better performance, with a slight advantage to $J=2$.
+Levels $J=2,3$ have a clear better performance, with a slight advantage to $J=2$, since it is uses less levels on the decomposition.
 The overall performance  of $J=2$ warrants its use for the rest of the comparisons.
 Panel (d) shows how the proposed method performs in comparison to a deep learning feature extraction from a residual learning network \cite{zhang2017beyond}.
-WECS is applied with db2 wavelets and $J=2$.
-We can see that the ROC curves for images treated with deep-learning methods or treated with wavelet based methods are almost identical.
-The WECS runs in 0.42s, while the deep-learning based $\vd(m)$ runs in 64.68s on a notebook.
+WECS is applied with db2 wavelets and $J=2$, whereas the other method applies WECS's idea, the only difference being that $\vX(m)$ is replaced by another smooth version of $\mathcal{I}(m)$ that employs neural-networks trained to compete with wavelet smoothing.
+We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS.
+% The WECS runs in 0.42s, while the deep-learning based $\vd(m)$ runs in 64.68s on a notebook.
 The configuration of the notebook is: OS - Ubuntu 18.04.5 LTS; RAM 7.7 GB; Intel\textregistered Core\texttrademark $ $i7-8565U CPU @ 1.80GHz x 8; graphics - Mesa Intel\texttrademark UHD Graphics 620 ; GNOME - 3.36.8; OS type - 64-bit.
@@
 The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}(m)\}$.
-The ROC curves in Panel (a) show that the proposed WECS outperforms its counterpart without wavelets and that it is competitive with the method using aggregation of differences.
+The ROC curves in Panel (a) show that the proposed WECS outperforms both the other two methods.
 

### 2021-11-06 codes/SeqEllispseChangeAnalysis.m

@@ -33,6 +33,13 @@ imshow(totalchanges)
 title('Total changes', 'FontSize', 17)
 saveas(mImage,sprintf('../figs/total_changes.jpg'))
 
+% first four observed images
+for m=1:4
+    mImage = figure;
+    imshow(mY(:,:,m))
+    saveas(mImage,sprintf('../figs/ellipses_t%d.jpg',m))
+end
+
 %% Analysis of decomposition level
 
 % wavelet basis used

### 2021-11-09 tex/paper.wecs/wecs_tgrs.tex
updates on tex and application

@@
 We present two complimentary multiscale measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images.
-We show WECS performance on synthetic multi-temporal image data.
-We also apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana.
+We show WECS performance on simulated multi-temporal image data.
+We also apply the proposed method to a time series of 84 satellite images in the border region of Brazil and the French Guiana.
 The images were captured from November 08, 2015 to December 09 2017.
@@
 Section \ref{section_method} introduces the problem and presents the proposed method.
-We show WECS performance on synthetic multi-temporal image data in Section \ref{section_validation}.
+We show WECS performance on simulated multi-temporal image data in Section \ref{section_validation}.
 In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.
@@
 
-\section{Validation on synthetic data}\label{section_validation}
+\section{Experiments}
 
-In this section we apply the change detection methods above on synthetic data of multi-temporal images.
-The synthetic multi-temporal images ($n=80$) is obtained by repeating a sequence of four images with different types of changes plus a noise.
+\subsection{Simulated images}
+\label{section_validation}
+
+In this section we apply the change detection methods above on simulated data of multi-temporal images.
+The simulated multi-temporal images ($n=80$) is obtained by repeating a sequence of four images with different types of changes plus a noise.
 Examples of the first four of these images are shown in Figure \ref{F:EllipsoidChanges}.
@@
 \includegraphics[scale=.11]{../../figs/ellipses_t4}
-\caption{Example of the first four synthetic multi-temporal images. Features and changes come as white ellipses and dots.}
+\caption{Example of the first four simulated multi-temporal images. Features and changes come as white ellipses and dots.}
 \label{F:EllipsoidChanges}
@@
 \includegraphics[scale=.1]{../../figs/corr_changes_dm_nowavelets}(d)
-\caption{Synthetic images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
+\caption{Simulated images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
 \label{F:Changes_methods_images}
@@
 Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}(m)$ instead of $\vX(m)$ in the WECS formulation.
-%The spatio-temporal advantages of the proposed wavelet  $\vd(m)$ is clear in Figure \ref{F:Changes_methods_images}(b), which shows that a wavelet smoothing has an expressive capacity of measuring the the changes through correlations of coefficients' deviations, even for the changes corresponding to smaller ellipses.
 
@@
 \includegraphics[scale=.13]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
-\caption{ROC curves for detection of changing ellipses in synthetic images and different methods.
+\caption{ROC curves for detection of changing ellipses in simulated images and different methods.
 (a) The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue).
@@
 
-Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the synthetic data as follows.
+Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the simulated data as follows.
 The effects of wavelet bases, level of decomposition, deep-learning feature extraction are shown on the ROC curves.
@@
 We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS.
-% The WECS runs in 0.42s, while the deep-learning based $\vd(m)$ runs in 64.68s on a notebook.
-The configuration of the notebook is: OS - Ubuntu 18.04.5 LTS; RAM 7.7 GB; Intel\textregistered Core\texttrademark $ $i7-8565U CPU @ 1.80GHz x 8; graphics - Mesa Intel\texttrademark UHD Graphics 620 ; GNOME - 3.36.8; OS type - 64-bit.
 We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
@@
 
-%\begin{figure}[htp!]
-%\includegraphics[scale=.14]{../../figs/methods_comparison_F1score}\hspace{-.5cm}(a) 
-%\includegraphics[scale=.14]{../../figs/levels_comparison_F1score}\hspace{-.5cm}(b)\\
-%\includegraphics[scale=.14]{../../figs/families_comparison_F1score}\hspace{-.5cm}(c)
-%\includegraphics[scale=.14]{../../figs/dm_comparison_wavelet_deepL_F1score}\hspace{-.5cm}(d) \\
-%\caption{$F_1$-scores for detection of changing ellipses in synthetic images and different methods.
-(a) The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue).
-%(b) db2 $\vd(m)$ with different levels.
-%(c) $\vd(m)$ with different wavelet bases and J=2;
-%(d) The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line).}
-%\label{F:EllipsoidChanges_details}
-%\end{figure}
-
-\section{Real Data Results}\label{section_realdata}
+\subsection{Real Data}
+\label{section_realdata}
 
@@
 
-A multi-resolution analysis (MRA) based on a Symlet basis with filter of length 16 (symlet 8) is built.
-In order to have dimensions as power of 2, the matrices $\mathcal{I}_{k,l}$ were extended to a $2048\times2048$ matrix with $\mathcal{I}_{k,l}$ at the center and the remaining parts being completed with mirrored values at the borders.
-The wavelet transform at resolution level $J=2$ was applied to these matrices.
+A multi-resolution analysis based on a Symlet basis with filter of length 16 (symlet 8) is built.
+In order to have dimensions as power of 2, the matrices $\mathcal{I}_{k,l}$ are extended to a $2048\times2048$ matrix with $\mathcal{I}_{k,l}$ at the center and the remaining parts being completed with mirrored values at the borders.
+The wavelet transform at resolution level $J=2$ is applied to these matrices and only the portion corresponding to the $1538 \times 1556$ is kept for further processing.
 Then, we are able to compute the squared mean differences vector $\vd$ and the matrix of absolute correlations $\vR$.
 
-The time change measures computed with $\vd$ are shown in Figure \ref{F:forest_wecs}a.
+The time change measures computed with $\vd$ are shown in Figure \ref{F:forest_wecs}.
 The orange line in Figure \ref{F:forest_wecs}a represents the median absolute deviation of $\vd$, which allow us to notice times that differ expressively from the others.
-We can notice that times $m=25,27,30$ are highlighted as having expressive changes.
+We can notice that times $m=25,27,30$ are highlighted as having the most expressive changes.
 The images  corresponding to these times can be seen in Figure \ref{F:forest_change_times}.
 
-The changes in space can be analyzed using the image obtained with $\vR$, which is displayed in Figure \ref{F:forest_wecs}b.
-Taking the $[n/\log n]$ ($n=1538\times1556$) largest absolute correlations as those corresponding to possible change points, we obtain a matrix of zeros and ones that is presented in Figure \ref{F:forest_wecs}c.
-The white regions in Figure \ref{F:forest_wecs}c (entries with value one) represent the change points, which seem to concentrate mainly on three regions: at the center, to the right of the river; at the top, on the left border of the river; and at the top left.
-Computing aggregated absolute differences to measure changes in space, we obtain Figure \ref{F:forest_wecs}d.
-The first two regions captured using WECS are also detected in Figure \ref{F:forest_wecs}d, but the former offers higher contrasts to these change regions.
-The comparison between the performance of these two methods can also be checked in Figure \ref{F:forest_roc_wecs_agg}, where a ROC curve is computed to check the correct detection of changing and not-changing regions.
-These change/nonchange regions were determined using [XXXXX].
-We can notice that WECS reaches high correct detection earlier than the method that uses aggregation of differences, but both methods do not have good performances concerning the nonchange regions.
-
 \begin{figure}[htp!]
-\includegraphics[scale=.13]{../../figs/forest_vSumDifCoefSq}\hspace{-.5cm}(a) 
-\includegraphics[scale=.14]{../../figs/forest_wecs_abscorr}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.10]{../../figs/forest_wecs_change_space}\hspace{-.5cm}(c) 
-\includegraphics[scale=.14]{../../figs/forest_aggreg_ratios}\hspace{-.5cm}(d)\\ 
-\caption{Analysis of change points in time and space of the forest data.
-(a) Plot of $\vd(m)$, $m=1,\ldots,85$ with orange horizontal line indicating two times its median absolute deviation.
-(b) Matrix of absolute correlations obtained with WECS.
-(c) Change regions detected by highlighting the $[n/\log n]$ largest correlations.
-(d) Change measured with aggregation of absolute differences.}
+\centering
+\includegraphics[scale=.2]{../../figs/forest_vSumDifCoefSq}\hspace{-.5cm}
+\caption{Plot of $\vd(m)$, $m=1,\ldots,85$ with orange horizontal line indicating two times its median absolute deviation.}
 \label{F:forest_wecs}
@@
 \includegraphics[scale=.3]{../../figs/forest_changes_time} 
-\caption{Images whose time $m$ correspond to points of $\vd(m)$ above two times its median absolute deviation in Figure \ref{F:forest_wecs}a.}
+\caption{Images $\vX(m)$ from times $m$ correspond to top-3 highest values of $\vd(m)$ and the mean image $\bar{\mathcal{I}}$.}
 \label{F:forest_change_times}
@@
 
+The changes in space can be analyzed using the image obtained with $\vR$, which is displayed in Figure \ref{F:forest_wecs}c.
+Denoting the images' dimension as $N=1538\times1556$, if we take the $[N/\log N]$ largest absolute correlations as those corresponding to possible change points, we obtain a matrix of zeros and ones that is presented in Figure \ref{F:forest_wecs}e.
+The white regions in Figure \ref{F:forest_wecs}c (entries with value one) represent the change points, which seem to concentrate mainly on three regions: at the center, to the right of the river; at the top, on the left border of the river; and at the top left.
+Computing aggregated absolute differences to measure changes in space, we obtain Figure \ref{F:forest_wecs}d.
+An image of detected changes corresponding to measures in  Figure \ref{F:forest_wecs}d can be obtained applying a thresholding method for grayscale images.
+Figure \ref{F:forest_wecs}f shows the result of using Otsu's thresholding method \cite{otsu1979threshold} for aggregation of absolute differences.
+
+The comparison of performance for the two change detection measures considered here can be checked in Figure \ref{F:forest_roc_wecs_agg}, where a ROC curve is computed to check the correct detection of changes.
+The reference of correct change regions is shown in Figure \ref{F:forest_roc_wecs_agg}a, which are determined using [XXXXX].
+We can also compare the change detection methods using the $F_1$-score accuracy measure.
+Denoting $\mathrm{TP}$ as the number of true positives (change pixels correctly detected), $\mathrm{FP}$ as the number of false positives (nonchange pixels flagged as change point) and $\mathrm{FN}$ as the total of false negatives (change points flagged as nonchange point), the $F_1$-score is defined as
+\begin{equation*}
+F_1 = 2\frac{\mathrm{Pr}\cdot\mathrm{Re}}{\mathrm{Pr}+\mathrm{Re}}	,
+\end{equation*}
+where
+\begin{equation*}
+\mathrm{Pr}=\frac{\mathrm{TP}}{\mathrm{TP}+\mathrm{FP}}\quad 
+\text{and} \quad \mathrm{Re} = \frac{\mathrm{TP}}{\mathrm{TP}+\mathrm{FN}}.
+\end{equation*}
+Results of these three accuracy measures are presented in Table \ref{T:forest_accuracy}.
+Results for aggregation of absolutes differences correspond to two types of thresholds: Otsu's method and Kittler-Illingworth \cite{kittler1986minimum}.
+We can observe that WECS presents the highest value of $F_1$-score, which means it has a better performance that the competing method.
+
 \begin{figure}[htp!]
-\includegraphics[scale=.11]{../../figs/forest_change}\hspace{-.5cm}(a)
-\includegraphics[scale=.11]{../../figs/forest_nonchange}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{../../figs/forest_roc_change}\hspace{-.5cm}(c) 
-\includegraphics[scale=.14]{../../figs/forest_roc_nonchange}\hspace{-.5cm}(d)
-\caption{Analysis of change points in time and space of the forest data.
+\centering
+\includegraphics[scale=.10]{../../figs/forest_change}\hspace{-.5cm}(a)
+\includegraphics[scale=.15]{../../figs/forest_roc_change}\hspace{-.5cm}(b)\\
+\includegraphics[scale=.13]{../../figs/forest_wecs_abscorr}\hspace{-.5cm}(c)
+\includegraphics[scale=.13]{../../figs/forest_aggreg_ratios}\hspace{-.5cm}(d)\\
+\includegraphics[scale=.10]{../../figs/forest_wecs_change_space}\hspace{-.5cm}(e) 
+\includegraphics[scale=.10]{../../figs/forest_aggreg_change_space}\hspace{-.5cm}(f)\\ 
+\caption{Change points in space of the forest data.
 (a) Regions (in white) where changes should be detected.
-(a) Regions (in white) where changes are not expected to be detected.
-(c) ROC curve for detection of changing regions.
-(d) ROC curves for detection of nonchange regions.}
+(b) ROC curve for detection of changing regions.
+(c) Matrix of absolute correlations obtained with WECS.
+(d) Change measures from aggregation of absolute differences.
+(e) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension).
+(f) Change regions detected using aggregated absolute differences and Otsu's thresholding.}
 \label{F:forest_roc_wecs_agg}
 \end{figure}
+
+\begin{table}
+\caption{Accuracy measures $F_1$-score, Precision ($\mathrm{Pr}$) and Recall ($\mathrm{Re}$) computed from detection of changing regions in the forest data using WECS (taking the $[N/\log N]$ largest correlations as corresponding to change) and the aggregation of absolute differences with Otsu and Kittler-Illingworth's (KI) thresholds.}
+\centering
+\begin{tabular}{c|ccc}
+\hline
+  & Aggregated - Otsu & Aggregated - KI & WECS \\
+\hline
+$F_1$-score & 0.2231 & 0.2163 & 0.3253 \\
+$\mathrm{Pr}$ & 0.1369 & 0.1296 & 0.2390 \\
+$\mathrm{Re}$ & 0.6022 & 0.6526 & 0.5094 \\
+\hline
+\end{tabular}
+\label{T:forest_accuracy}
+\end{table}
 
@@
 It makes use of correlation screening for ultra-high dimensional data.
-The proposed method's performance is shown using both synthetic and real data.
+The proposed method's performance is shown using both simulated and real data.
 The proposed method is useful to detect spatio-temporal change points, which is illustrated on data analyses.

### 2021-11-09 codes/F1score.m

@@ -1,0 +1,30 @@
+function [F1,Pr,Re] = F1score(img,im0)
+% Input
+% img: matrix of computed using data with 0 (nonchanges) and 1 (change)
+% im0: matrix of true change regions, with values 0 (nonchange) and 255
+% (change)
+% Output
+% F1: F1 score
+% Pr: Precision
+% Re: Recall
+
+% total number of positives
+imtemp=255.*double(img);
+D=find(imtemp==255);
+nD=numel(D);
+% True positive
+imgdiff=2*imtemp-im0;
+TP=find(imgdiff==255);
+nTP=numel(TP);
+% Precision
+Pr= nTP/nD;
+% Recall
+FN=find(imgdiff==-255); % false negatives
+nFN=numel(FN);
+Re = nTP/(nTP + nFN);
+% F1-score
+F1 = 2*Pr*Re/(Pr + Re);
+
+
+end
+

### 2021-11-09 codes/Forest_data_analysis.m

@@ -63,6 +63,7 @@ for m=1:n
     imRef = imRef + double(Y(:,:,1).^2 + Y(:,:,2).^2);
 end
 imRef = imRef/n;
+% constant to make the mean matrix have norm one
 normconst = norm(imRef);
 imRef = imRef/normconst;
 
@@ -106,12 +107,20 @@ for m = [25 27 30]
     t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{m,:})),'r');
     Y = read(t);
     data = double(Y(:,:,1).^2 + Y(:,:,2).^2);
+    
+    % extending image to be able to use swt2
+    pwr2 = ceil(log2(min([Nx Ny])));
+    extens = ceil((2^pwr2 - min([Nx Ny]))/2);
+    tmp = wextend('2D','per',data,extens); % matrix whose dim is power of 2
+    % wavelet approximation for log of VV chanel
+    [tmp2,~,~,~] = swt2(tmp(1:2^pwr2,1:2^pwr2),J,wname);
+
     subplot(2,2,cont); cont = cont + 1;
-    imagesc(data)
+    colormap(gray(256)); imagesc(tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J))    
     title(sprintf('m=%2d',m)); axis off;
 end
 subplot(2,2,4)
-imagesc(imRef); title('Mean image'); axis off;
+colormap(gray(256)); imagesc(imRef); title('Mean image'); axis off;
 saveas(mImage,sprintf('../figs/forest_changes_time.jpg'))
 
 
@@ -130,12 +139,13 @@ axis off; colorbar
 set(gca,'FontSize',13)
 saveas(mImage,sprintf('../figs/forest_wecs_abscorr.jpg'))
 
-
 %histogram(mCorr(:))
 mImage = figure;
 cutoff = quantile(mCorr(:),1-1/log(Nx*Ny));
 imshow(mCorr>cutoff)
 saveas(mImage,sprintf('../figs/forest_wecs_change_space.jpg'))
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % Aggregation of differences
 S = zeros(Nx,Ny);
@@ -158,6 +168,14 @@ axis off; colorbar
 set(gca,'FontSize',13)
[…]

### 2021-11-09 codes/kittler.m

@@ -1,0 +1,36 @@
+function [T, V] = kittler(I, h, g)
+% T = KITTLER(I) computes a global threshold (T) of the image I
+% according to the method described by J. KITTLER and J. ILLINGWORTH.
+%
+% T = KITTLER([], H, G) computes a threshold for the image with
+% histogram H. The second argument (optional) G can contain graylevel 
+% values for the histogram bins (defaults to 0:(length(H)-1))
+%
+% Author: Bob Pepin (bobpepin@gmail.com)
+%
+% https://www.mathworks.com/matlabcentral/fileexchange/45685-kittler-illingworth-thresholding
+%
+if ~isempty(I)
+    %h = histc(I(:), 0:ceil(max(I(:))));
+    h = histc(I(:), 0:0.001:ceil(max(I(:))));
+end
+if nargin < 2
+    g = (0:(length(h)-1))';
+else
+    g = reshape(g, [], 1);
+end
+C = cumsum(h);
+M = cumsum(h .* g);
+S = cumsum(h .* g.^2);
+sigma_f = sqrt(S ./ C - (M./C).^2);
+Cb = C(end) - C;
+Mb = M(end) - M;
+Sb = S(end) - S;
+sigma_b = sqrt(Sb ./ Cb - (Mb./Cb).^2);
+P = C ./ C(end);
+V = P .* log(sigma_f) + (1-P).*log(sigma_b) - P.*log(P) - (1-P).*log(1-P);
+V(~isfinite(V)) = Inf;
+[~, idx] = min(V);
+%T = g(idx);
+T = g(idx)*0.001;
+end

### 2021-11-09 tex/paper.wecs/wecs_tgrs__trans08nov21.tex
sufoco-part1 (09nov21)

--- 2021-11-06 tex/paper.wecs/wecs_tgrs.tex
@@
-%% bare_jrnl.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% see http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% journal paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
+\documentclass[journal]{IEEEtran}
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-\documentclass[journal]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[journal]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
 \ifCLASSINFOpdf
@@
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
 
-% *** MATH PACKAGES ***
-%
 \usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
 %\usepackage{amsaddr}		
@@
 
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
-% Axel Sommerfeldt. This package may be useful when used in conjunction with 
-% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
-% submissions have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.4.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/pkg/endfloat
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
+%DEFS. RODNEY%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \def\bSig\mathbf{\Sigma}
@@
 
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
+\usepackage{hyperref}
+
+%DEFS. ROGERIO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\usepackage{graphicx}
+
+\usepackage{subfigure}
+\usepackage{multirow}
+\usepackage{color,colortbl}
+\usepackage{enumerate}
+\usepackage{amsmath,amssymb}
+\usepackage{booktabs}
+\usepackage{rotating}
+\usepackage{cite}
+\usepackage{balance}
+
+\usepackage{bm,bbm}
+\usepackage{datetime2}
+\usepackage[binary-units]{siunitx}
+
+\usepackage{mathrsfs}
+
+\usepackage{placeins}
+
+\usepackage[abs]{overpic}
+\usepackage{pict2e}
+%\usepackage{float}
+\usepackage{microtype}
+\usepackage{pdfpages}
+
+\makeatletter
+\newcommand{\thickhline}{\noalign {\ifnum 0=`}\fi \hrule height 1pt \futurelet \reserved@a \@xhline}
+\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
+\makeatother
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+%\hyphenation{op-tical net-works semi-conduc-tor}
 
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
 \title{Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
 
@@
         and~Abdourrahmane~Atto,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
-\thanks{This work was supported by FAPESP grants 2016/24469-6 and 2018/04654-9 and CNPq grants 309230/2017-9 and 310991/2020-0.}
+\thanks{This work was supported by FAPESP (Grants 2016/24469-6, 2018/04654-9 and 2021/01305-6) and CNPq (Grants 309230/2017-9 and 310991/2020-0).}
 \thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
-\thanks{A.~Atto is with the LISTIC - Polytech Annecy-Chamb\'{e}ry, Universit\'{e} de Savoie, 74944 Annecy le Vieux Cedex, France (e-mail: Abdourrahmane.Atto@univ-savoie.fr).}}
+\thanks{A.~Atto is with the LISTIC - Polytech Annecy-Chamb\'{e}ry, Universit\'{e} de Savoie, 74944 Annecy le Vieux Cedex, France (e-mail: Abdourrahmane.Atto@univ-savoie.fr).}
+\thanks{R.~G.~Negri are with the Department of Environmental Engineering, São Paulo State University, São José dos Campos, Brazil (e-mail: rogerio.negri@unesp.br).}
+}
 % <-this % stops a space
 %\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
-
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
 
@@
 {Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark.
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
 
@@
 The images were captured from November 08, 2015 to December 09 2017.
+\textcolor{magenta}{(lapidaremos ao fim)}
 \end{abstract}
@@
 \section{Introduction}
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps.
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by the IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-\IEEEPARstart{W}{e} propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images.
-WECS is based upon correlation screening for energy apportionment on wavelet approximations.
-The spatial character of the change detection is attained on pixel level.
-The method is fast, scalable, linearly updatable, and the resulting measures are sparse.
+
+%%%RGN: Iniciar com uma visão geral do problema (detecćão de mudanća em dados multitemporais).
+O que é (a ideia)?
+Porque é importante?
+Mencionar abordagens supervisionada e não supervisionadaA inclusão de um/dois parágrafos seria suficiente.
+laplace
+
+\textcolor{red}{\IEEEPARstart{C}{hange} detection... (completar)...}
 
@@
 
+%%%RGN: Eu acho muito interessante que esta pesquisa esteja destinada à dados SAR (lida bem com dados ruidosos, imagens de uma única banda/polarizaćão e contempla uma aplicaćão real com dados Sentinel), no entanto, sinto falta de uma justificativa que relacione o método à dados SAR.
+
+%%%RGN: Podemos omitir o "via polarimetric SAR (POLSAR)"?
 Besides different methodological paradigms, several areas of application receive special attention.
@@
 
+%%%RGN: o "wavelet methods" aqui não está claro se é destinado à detecćão de mudanćas.
 Wavelet methods present many advantages for a plethora of applications \cite{vidakovic1999statistical} thanks to wavelet capabilities in capturing multiscale/multiresolution information.
@@
 \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical} use different wavelet methods for change detection in satellite images.
+
+%%%RGN: acrescentar algo sobre Screening? (dai ficam amarrados os conceitos de wavelets e screening, abrindo caminho para o wecs)
+
+%%%RGN: acredito que esse paragrafo fica melhor aqui
+%%%RGN: o acrônimo WECS só foi definido no abstract (geralmente é necessário definir no abstract e depois no corpo do texto com maior detalhe)
+We propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images.
+WECS is based upon correlation screening for energy apportionment on wavelet approximations.
+The spatial character of the change detection is attained on pixel level.
+The method is fast, scalable, linearly updatable, and the resulting measures are sparse.
 
@@
 
-The rest of the text goes as follows.
+%%%RGN: Deixar para o final, pois a estrutura do paper pode ser alterada até a submissão.
+\textcolor{magenta}{The rest of the text goes as follows.
 Section \ref{section_method} introduces the problem and presents the proposed method.
@@
 In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.
-Section \ref{section_discussion} concludes the paper with a discussion.
+Section \ref{section_discussion} concludes the paper with a discussion.}
 
-\section{Methodology}\label{section_method}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Methodology}\label{section_method}
+%%%RGN: Temporário, depois alteramos.
+\section{Wavelet Energies Correlation Screening}\label{section_method}
+
+%%%%%%%%%%%%%%%%%%%%%
+\subsection{Overview}\label{methodOverview}
+
+Figure~\ref{figOverview} provides an overview of the proposed method for multitemporal image change detection.
+
+%%%RGN: Vamos fazer um diagrama geral explicando como funciona o método?
+(usar o draw.io) Com base no diagrama fazemos uma discussão em alto nível.
+A próxima sećão vem formalizando os elementos deste diagrama.
+Se fizer uma versão inicial deste diagrama, eu ajudo a polir.
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+\includegraphics[width=0.485\textwidth]{../../figs/coming-soon.png}
+\caption{Proposed method overview.}\label{figOverview}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%
+\subsection{Formalization}\label{methodFormalization}
 
@@
 We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}(m)$.
+
 Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}.
@@
 where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}(m)$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX(m)$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
+
 The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX(m)$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
@@
 where $vec(\vD(m))$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX(m)$, $m=1,\ldots,n$.
-Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.
-We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d<<p$.
+
+Sparsity on wavelet coefficients plays a special role here \cite{johnstone2009statistical}.
+We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d \lle p$.
 This can be represented as the following linear model
@@
 \end{equation}
-where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
+where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$.
+
+Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
 \[
@@
 This segregates the multi-temporal images, since the former time points identify the images in which significant changes occur, while the latter indices identify time points with no major changes.
-Consider $U>L>0$ such that the $s_d$ highest values of $\vd$ are larger than $U$, and the $n-s_d$ smallest values of $\vd$ are smaller than $L$.
+Consider $0<L<U$ such that the $s_d$ highest values of $\vd$ are larger than $U$, and the $n-s_d$ smallest values of $\vd$ are smaller than $L$.
 We also take $\delta=U-L$.
@@
 \]
-for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$).
+for some arbitrary $\Delta \ggt 0$ (which can be a function of $n$ and $J$).
 Thence, when we employ correlation screening we select the two-dimensional wavelet indices which have the closest empirical directions to the vector of image temporal changes.
@@
 
-\section{Validation on synthetic data}\label{section_validation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}\label{experiments}
+
+\textcolor{red}{
+Texto introdutório que explica o desenho do experimento.
+Aponta a primeira subseção como forma de avaliar a proposta com dados simulados, já a segunda subseção mostra uma aplicação prática envolvendo dados reais.
+Apenas uma explicação superficial.
+}
+
+%%%%%%%%%%%%%%%%%%%%
+%\subsection{Validation on synthetic data}\label{section_validation}
+\subsection{Simulated data analysis}\label{secExpSimulated}
 
@@
 
-\section{Real Data Results}\label{section_realdata}
+%\section{Real Data Results}\label{section_realdata}
+%%%%%%%%%%%%%%%%%%%%
+%\subsection{Validation on synthetic data}\label{section_validation}
+\subsection{Actual remote sensing application}\label{secExpActual}
 
 We employed the proposed change detection method on a series of 84 multi-date satellite images.
-The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3.
-Each image has two channels (VV and VH) and 1538 by 1556 pixels.
+The images were taken on a forest region at the border of Brazil and the French Guiana (Figure~\ref{figAE}) from 26th December 2015 to 3rd December 2017.
+Each image has two polarizations (VV and VH) and 1538 by 1556 pixels.
 We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}(m) = (\mathrm{VV}_{k,l}(m)^2+\mathrm{VH}_{k,l}(m)^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+\includegraphics[width=0.485\textwidth]{../../qgis/maps/StudyArea.pdf}
+\caption{Study area location. \textcolor{red}{(preliminar)}}\label{figAE}
+\end{figure}
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[26th December, 2015]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2015.pdf}\label{figImgSentinel2015}} 
+\subfigure[3rd December, 2017]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2017.pdf}\label{figImgSentinel2017}}
+}
+
+\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.23\textwidth]{../../figs/coming-soon.png}\label{figImgSentinelROIs}}
+
+\caption{\textcolor{red}{Inserir caption...}}\label{figImageRef}
+\end{figure}
 

### 2021-11-09 tex/paper.wecs/wecs_tgrs.tex
files to draw a diagram
Computing SNR for simulated and real data

@@
 where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
-%\citep{jonhstone2009}
 
@@
 where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[scale=.45]{../../drawio/diagram_wecs.drawio}
+\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
+\label{F:EllipsoidChanges}
+\end{figure}
 
@@
 The comparison of performance for the two change detection measures considered here can be checked in Figure \ref{F:forest_roc_wecs_agg}, where a ROC curve is computed to check the correct detection of changes.
+WECS attains high true positive rates before the aggregation method, but both methods design competitive results.
+This performance differs from the ROC curve results in the previous section because the real images have a mean signal to noise ratio (SNR) of 3.9327, whereas the simulated images were generated with SNR of 0.2309.
+Hence, WECS is expected to have performance closer to the aggregation method in less noisy applications.
 The reference of correct change regions is shown in Figure \ref{F:forest_roc_wecs_agg}a, which are determined using [XXXXX].
+
 We can also compare the change detection methods using the $F_1$-score accuracy measure.
@@
 It makes use of correlation screening for ultra-high dimensional data.
+Thereupon, WECS is expected to provide a sample of points in space in a way that such set, contains real change points with high probability.
 The proposed method's performance is shown using both simulated and real data.

### 2021-11-09 codes/Forest_data_analysis.m

@@ -69,6 +69,7 @@ imRef = imRef/normconst;
 
 % resolution level used on wavelet decompositions
 J = 2;
+meanSNR = 0;
 
 for m=1:n
     t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{m,:})),'r');
@@ -82,8 +83,14 @@ for m=1:n
     [tmp2,~,~,~] = swt2(tmp(1:2^pwr2,1:2^pwr2),J,wname);
     % sqared mean deviations of level J approx. coeff.
     mDifCoefSq(:,:,m) = (tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J) - imRef).^2;
+    % mean SNR of observed images
+    meanSNR = meanSNR + ...
+        norm(tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J),'fro')/(norm(data,'fro')*n);
 end
 clear Y t tmp tmp2
+
+% mean SNR for observed images
+meanSNR
 
 % vector with total sum of squared mean deviations for each image
 vSumDifCoefSq = reshape(sum(sum(mDifCoefSq,1),2),1,n);

### 2021-11-09 codes/SeqEllispseChangeAnalysis.m

@@ -7,7 +7,7 @@ close all
 clear;
 %%
 
-% code that generates a sequence of synthetic images
+% code that generates a sequence of simulated images
 
 % Subsampling of images
 subsamplingfactor = 2;
@@ -17,6 +17,13 @@ sig = 1;
 n = 80;
 % running code
 generate_synthetic_sequence
+
+% mean SNR of simulated images
+meanSNR = 0;
+for m=1:4
+    meanSNR = meanSNR + norm(mI(:,:,m),'fro')/(4*norm(mY(:,:,m),'fro'));
+end
+meanSNR
 
 % output
 size(mI) % true images

### 2021-11-11 tex/paper.wecs/wecs_tgrs.tex
notação no .tex

@@
 
-Let $\mathcal{I}(1),\ldots,\mathcal{I}(m)$ be a set of matrices representing the images of some region of interest.
+Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time.
 These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate.
-Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}(m)$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
-We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}(m)$.
+Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
+We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}^{(m)}$.
 Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}.
-After application of this wavelet transform to $\mathcal{I}(m)$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX(m)$, a smooth version of $\mathcal{I}(m)$ with the same dimension.
-The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX(m)$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}(m)$.
-Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX(m)$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
+After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smooth version of $\mathcal{I}^{(m)}$ with the same dimension.
+The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX^{(m)}$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}^{(m)}$.
+Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX^{(m)}$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
 
-We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX(m)\right\}$ as
+We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX^{(m)}\right\}$ as
 \begin{equation} 
-\sum_{m=1}^n\|\vX(m)\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX(m)-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
+\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
 \end{equation}
-where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}(m)$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX(m)$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
-The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX(m)$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
-Since each element (pixel) in $\vX(m)$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space.
+where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
+The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX^{(m)}$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
+Since each element (pixel) in $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space.
 Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further.
-Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX(m)-\vX(m-1)\|^2$, $d(\vX(m),\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
+Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX^{(m)}-\vX^{(m-1)}\|^2$, $d(\vX^{(m)},\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
 
-Let $X_{k,l}(m)$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX(m)$ and $\bar{\mathcal{I}}$, respectively.
-Computing $D_{k,l}(m)=(X_{k,l}(m)-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD(m) = [D_{k.l}(m)]$.
+Let $X_{k,l}^{(m)}$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\bar{\mathcal{I}}$, respectively.
+Computing $D_{k,l}^{(m)}=(X_{k,l}^{(m)}-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD^{(m)} = [D_{k,l}^{(m)}]$.
 We then
@@
 \begin{equation} 
-\vd(m)= \sum_{k,l}D_{k,l}(m) = \sum_{k,l}(X_{k,l}(m) - \bar{\mathcal{I}}_{k,l})^2,
+\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
 \label{E:eq_defdm}
@@
 \end{equation*}
-where $vec(\vD(m))$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX(m)$, $m=1,\ldots,n$.
+where $vec(\vD^{(m)})$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$.
 Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.
@@
 In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
-For each local squared mean deviation time series, given by individual elements of $\vD(m)$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
+For each local squared mean deviation time series, given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
 \begin{equation*}
@@
 {\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
-$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}(m)\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
+$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
 This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
@@
 \]
-i.e., such that  $\sum_{m=1}^nD_{k,l}(m)\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
+i.e., such that  $\sum_{m=1}^nD_{k,l}^{(m)}\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
 \[
-\left|\sum_{m:\vd(m)>U} D_{k,l}(m)\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}(m)\right|>\Delta,
+\left|\sum_{m:\vd(m)>U} D_{k,l}^{(m)}\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}^{(m)}\right|>\Delta,
 \]
@@
 These images are generated from the sum of two matrices: a signal matrix with ones on entries where ellipses occur and zero elsewhere, and a noise matrix with random variables following a standard Gaussian distribution.
-The first image, $\mathcal{I}(1)$, presents three elongated ellipses.
-The second image $\mathcal{I}(2)$ has shorter and larger ellipses added.
-Smaller ellipses are then added to form $\mathcal{I}(3)$ and $\mathcal{I}(4)$.
+The first image, $\mathcal{I}^{(1)}$, presents three elongated ellipses.
+The second image $\mathcal{I}^{(2)}$ has shorter and larger ellipses added.
+Smaller ellipses are then added to form $\mathcal{I}^{(3)}$ and $\mathcal{I}^{(4)}$.
 All the changes made that occur among subsequent images can be seen in Figure \ref{F:Changes_methods_images}(a), which displays a matrix of zeros and ones that correspond, respectively, to locations without and with changes along time.
@@
 Panel (b) shows the result of WECS using Daubechies wavelet with two null moments (db2) and $J=2$.
-Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}(m) - \mathcal{I}_{k,l}(m-1)|$.
-Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}(m)$ instead of $\vX(m)$ in the WECS formulation.
+Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}^{(m)} - \mathcal{I}_{k,l}^{(m-1)}|$.
+Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ in the WECS formulation.
 
@@
 Panel (d) shows how the proposed method performs in comparison to a deep learning feature extraction from a residual learning network \cite{zhang2017beyond}.
-WECS is applied with db2 wavelets and $J=2$, whereas the other method applies WECS's idea, the only difference being that $\vX(m)$ is replaced by another smooth version of $\mathcal{I}(m)$ that employs neural-networks trained to compete with wavelet smoothing.
+WECS is applied with db2 wavelets and $J=2$, whereas the other method applies WECS's idea, the only difference being that $\vX^{(m)}$ is replaced by another smooth version of $\mathcal{I}^{(m)}$ that employs neural-networks trained to compete with wavelet smoothing.
 We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS.
 We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
-The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}(m)\}$ instead of $\{\vX(m)\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}(m)\}$.
+The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}^{(m)}\}$ instead of $\{\vX^{(m)}\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}^{(m)}\}$.
 The ROC curves in Panel (a) show that the proposed WECS outperforms both the other two methods.
@@
 Each image has two channels (VV and VH) and 1538 by 1556 pixels.
-We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}(m) = (\mathrm{VV}_{k,l}(m)^2+\mathrm{VH}_{k,l}(m)^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.
+We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}^{(m)} = ((\mathrm{VV}_{k,l}^{(m)})^2+(\mathrm{VH}_{k,l}^{(m)})^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.
 
@@
 \includegraphics[scale=.3]{../../figs/forest_changes_time} 
-\caption{Images $\vX(m)$ from times $m$ correspond to top-3 highest values of $\vd(m)$ and the mean image $\bar{\mathcal{I}}$.}
+\caption{Images $\vX^{(m)}$ from times $m$ correspond to top-3 highest values of $\vd(m)$ and the mean image $\bar{\mathcal{I}}$.}
 \label{F:forest_change_times}

### 2021-11-12 tex/paper.wecs/wecs_tgrs.tex
updates on .tex

@@
 
+%DEFS. RODNEY%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \def\bSig\mathbf{\Sigma}
@@
 
+\usepackage{hyperref}
+
+%DEFS. ROGERIO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\usepackage{graphicx}
+
+\usepackage{subfigure}
+\usepackage{multirow}
+\usepackage{color,colortbl}
+\usepackage{enumerate}
+\usepackage{amsmath,amssymb}
+\usepackage{booktabs}
+\usepackage{rotating}
+\usepackage{cite}
+\usepackage{balance}
+
+\usepackage{bm,bbm}
+\usepackage{datetime2}
+\usepackage[binary-units]{siunitx}
+
+\usepackage{mathrsfs}
+
+\usepackage{placeins}
+
+\usepackage[abs]{overpic}
+\usepackage{pict2e}
+%\usepackage{float}
+\usepackage{microtype}
+\usepackage{pdfpages}
+
+\makeatletter
+\newcommand{\thickhline}{\noalign {\ifnum 0=`}\fi \hrule height 1pt \futurelet \reserved@a \@xhline}
+\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
+\makeatother
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % *** Do not adjust lengths that control margins, column widths, etc. ***
@@
 \thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
-\thanks{A.~Atto is with the LISTIC - Polytech Annecy-Chamb\'{e}ry, Universit\'{e} de Savoie, 74944 Annecy le Vieux Cedex, France (e-mail: Abdourrahmane.Atto@univ-savoie.fr).}}
+\thanks{A.~Atto is with the LISTIC - Polytech Annecy-Chamb\'{e}ry, Universit\'{e} de Savoie, 74944 Annecy le Vieux Cedex, France (e-mail: Abdourrahmane.Atto@univ-savoie.fr).}
+\thanks{R.~G.~Negri is with the Department of Environmental Engineering, São Paulo State University, São José dos Campos, Brazil (e-mail: rogerio.negri@unesp.br).}
+}
 % <-this % stops a space
@@
 Besides different methodological paradigms, several areas of application receive special attention.
-For instance, urban change detection applications via polarimetric SAR (POLSAR) images are discussed in \cite{ansari2020urban}.
 \cite{song2018multi} discusses land cover change detection in mountainous terrain via multi-temporal and multi-sensor remote sensing images.
@@
 
-\subsection{Simulated images}
-\label{section_validation}
+\subsection{Simulated data analysis}\label{secExpSimulated}
 
@@
 
-\subsection{Real Data}
-\label{section_realdata}
+\subsection{Actual remote sensing application}\label{secExpActual}
 
@@
 We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}^{(m)} = ((\mathrm{VV}_{k,l}^{(m)})^2+(\mathrm{VH}_{k,l}^{(m)})^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+\includegraphics[width=0.485\textwidth]{../../qgis/maps/StudyArea.pdf}
+\caption{Study area location. \textcolor{red}{(preliminar)}}\label{figAE}
+\end{figure}
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[26th December, 2015]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2015.pdf}\label{figImgSentinel2015}} 
+\subfigure[3rd December, 2017]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2017.pdf}\label{figImgSentinel2017}}
+}
+
+\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.23\textwidth]{../../figs/coming-soon.png}\label{figImgSentinelROIs}}
+
+\caption{\textcolor{red}{Inserir caption...}}\label{figImageRef}
+\end{figure}
 

### 2021-11-13 tex/paper.wecs/wecs_tgrs.tex
introduction on .tex

@@
 We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised sparse procedure to detect spatio-temporal change points on multi-temporal SAR images or even on sequences of very high resolution images.
-The procedure is based on multiscale approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients.
+The procedure is based on wavelet approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients.
 We present two complimentary multiscale measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images.
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{W}{e} propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images.
-WECS is based upon correlation screening for energy apportionment on wavelet approximations.
-The spatial character of the change detection is attained on pixel level.
-The method is fast, scalable, linearly updatable, and the resulting measures are sparse.
 
+\IEEEPARstart{C}{hange} detection is an important task performed in remote sensing image that allows researchers and engineers to identify and evaluate modifications on land surfaces captured by multi-temporal satellite images.
+Analyzing problems such as deforestation, rapid urbanization and glacier melting, for example, are of great importance to study the dynamics of bio-systems.
+Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods.
 A review for change detection in multi-temporal remote sensing is given by \cite{ban2016change}.
-Different proposals for this purpose may be found  in the literature.
-They vary in their motivations as well as in their applicability.
+
+Most methods used for change detection analysis can be classified either as supervised (training data is used to set up the method) or unsupervised (fully data-driven techniques).
+We shall focus in this work on unsupervised approaches, whose examples in the literature include the works of \cite{bruzzone2000automatic}, \cite{celik2010change}, \cite{quin2014mimosa} and \cite{saha2020change}, for example.
+Many other proposals of methods vary in their motivations as well as in their applicability.
 Change detection in multi-temporal hyperspectral images is discussed in \cite{bovolo2015time},  \cite{liu2019review}, and \cite{matsunaga2017current}.
@@
 \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical} use different wavelet methods for change detection in satellite images.
+
+We propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images.
+WECS is based upon correlation screening for energy apportionment on wavelet approximations.
+The spatial character of the change detection is attained on pixel level.
+The method is fast, scalable, linearly updatable, and the resulting measures are sparse.
 

### 2021-11-14 tex/paper.wecs/wecs_tgrs.tex
updates in the introduction

@@
 \begin{abstract}
-We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised sparse procedure to detect spatio-temporal change points on multi-temporal SAR images or even on sequences of very high resolution images.
+We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised procedure to detect spatio-temporal change points on multi-temporal SAR images.
 The procedure is based on wavelet approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients.
-We present two complimentary multiscale measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images.
 We show WECS performance on simulated multi-temporal image data.
-We also apply the proposed method to a time series of 84 satellite images in the border region of Brazil and the French Guiana.
-The images were captured from November 08, 2015 to December 09 2017.
+We also evaluate the proposed method on a time series of 84 satellite images in a forest region at the border of Brazil and the French Guiana.
+The proposed method displays good results in covering change regions, with the additional benefit of having simple and fast computation.
 \end{abstract}
@@
 \IEEEPARstart{C}{hange} detection is an important task performed in remote sensing image that allows researchers and engineers to identify and evaluate modifications on land surfaces captured by multi-temporal satellite images.
-Analyzing problems such as deforestation, rapid urbanization and glacier melting, for example, are of great importance to study the dynamics of bio-systems.
-Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods.
+Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping}, for example, are of great importance to study the dynamics of regions sensitive to climate changes and human activity.
+Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods to images available at larger and larger from time to time.
 A review for change detection in multi-temporal remote sensing is given by \cite{ban2016change}.
@@
 
-Besides different methodological paradigms, several areas of application receive special attention.
+%Besides different methodological paradigms, several areas of application receive special attention.
 \cite{song2018multi} discusses land cover change detection in mountainous terrain via multi-temporal and multi-sensor remote sensing images.
 \cite{ru2021multi} studies multi-temporal scene classification and scene change detection.
-Deforestation change detection is discussed by \cite{barreto2016deforestation}.
 
-Wavelet methods present many advantages for a plethora of applications \cite{vidakovic1999statistical} thanks to wavelet capabilities in capturing multiscale/multiresolution information.
+In this paper we analyze Synthetic Aperture Radar (SAR) images through an unsupervised wavelet method.
+Known to be not affected by weather, cloud and sunlight conditions, SAR images have been an important source of data for change detection methods in the literature \cite{bovolo2005detail}.
+Unsupervised analysis of such images consists in the steps of data denoising, pixel-level processing and change examination.
+Our idea is to explore the denoising property of wavelet transforms and apply a feature screening idea to identify change regions at pixels represented by wavelet coefficients.
+Wavelet methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multiscale/multiresolution information.
 Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}.
-\cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical} use different wavelet methods for change detection in satellite images.
+Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical}.
+However, the application of feature screening on wavelet coefficients is still a novel idea in change detection literature, and we show it has an interesting potential to provide good results even with simple algorithms.
 
-We propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images.
-WECS is based upon correlation screening for energy apportionment on wavelet approximations.
-The spatial character of the change detection is attained on pixel level.
-The method is fast, scalable, linearly updatable, and the resulting measures are sparse.
+We propose here a novel method for unsupervised spatio-temporal change detection in a time series of SAR images.
+Wavelet Energies Correlation Screening (WECS) is based upon correlation screening for energy apportionment on wavelet approximations.
+The spatial character of the change detection is attained on pixel level, whereas total energy measures for each image in the time series quantify changes happening across time.
+The idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}.
+Such method is usually employed in high-dimensional regression models to reduce the problem's dimension by subsetting the available covariates in such a way that true covariates are among the chosen ones with high probability \cite{fan2008sure}.
+We show that by applying the feature screening idea on multi-temporal images, we obtain a fast and accurate method to cover change regions with good detection rates.
 
-The motivation for our proposed method is multi-fold.
+%The motivation for our proposed method is multi-fold.
 We aim for a fast and accurate method.
@@
 \centering
-\includegraphics[scale=.45]{../../drawio/diagram_wecs.drawio}
+\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
 \caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
@@
 \section{Experiments}
+
+In this section we shall evaluate the performance of WECS on two types of data sets.
+Initially, we generate simulated data where the level of noise and true change regions are known.
+For these images, we analyze different aspects of concerning the application of WECS and compare its performance with a standard change detection method.
+In the following application, real satellite data is employed, where the performance of WECS is evaluated on predefined change regions.
 

### 2021-11-15 codes/EllispseChangeAnalysis_Deep.m
updates in tex

@@ -1,163 +1,0 @@
-
-%%
-close all
-clear;
-%%
-im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
-im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
-im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
-im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
-%
-
-% Images are too big, consider subsampling
-subsamplingfactor = 2;
-im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
-totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
-totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
-%
-figure
-imshow(totalchanges)
-title('Total changes')
-
-%%
-eps = .00001;
-im1log = eps+double(im1);
-im2log = eps+double(im2);
-im3log = eps+double(im3);
-im4log = eps+double(im4);
-n=4;
-imRef = im1log + im2log + im3log + im4log;
-    imRef = imRef/n;
-%% comparison of wavelet denoising with mean image
-
-wname = 'db2';
-J = 2;
-
-tic
-[X1,~,~,~] = swt2(im1log,J,wname);
-    X1 = X1(:,:,J); % focus only on level J approximations ?
-[X2,~,~,~] = swt2(im2log,J,wname);
-    X2 = X2(:,:,J); % focus only on level J approximations ?
-[X3,~,~,~] = swt2(im3log,J,wname);
-    X3 = X3(:,:,J); % focus only on level J approximations ?
-[X4,~,~,~] = swt2(im4log,J,wname);
-    X4 = X4(:,:,J); % focus only on level J approximations ?
-toc
-
-%
-D1 = (X1 - imRef).^2;
-D2 = (X2 - imRef).^2;
-D3 = (X3 - imRef).^2;
[…]

### 2021-11-15 codes/EllispseChangeAnalysis_families.m

@@ -1,344 +1,0 @@
-
-% Evaluation of wavelet bases when mean approximation coefficients are used
-
-%%
-close all
-clear;
-%%
-
-% code that generates a sequence of synthetic images
-
-% Subsampling of images
-subsamplingfactor = 8;
-% noise level used
-sig = 1;
-% size of the sequence (must be multiple of 4)
-n = 80;
-% running code
-generate_synthetic_sequence
-
-% output
-size(mI) % true images
-size(mY) % sequence of synthetic images
-% number of rows and columns
-n1 = size(mY,1);
-n2 = size(mY,2);
-
-% all changes that are expected to be detected
-totalchanges = (abs(mI(:,:,1)-mI(:,:,2)) + abs(mI(:,:,2)-mI(:,:,3)) + ...
-    abs(mI(:,:,3)-mI(:,:,4)) + abs(mI(:,:,4)-mI(:,:,1)))>0;
-mImage = figure;
-colormap(gray(256)); imagesc(totalchanges)
-title('Total changes', 'FontSize', 17)
-%saveas(mImage,sprintf('../figs/total_changes.jpg'))
-
-% mean observed image
-imRef = mean(mY,3);
-
-% decomposition level fixed
-J = 2;
-
-%%
-
-% Wavelet basis
-wname = 'haar';
-
-[X1,~,~,~] = swt2(im1log,J,wname);
-    X1 = X1(:,:,J); % focus only on level J approximations ?
-[X2,~,~,~] = swt2(im2log,J,wname);
-    X2 = X2(:,:,J); % focus only on level J approximations ?
-[X3,~,~,~] = swt2(im3log,J,wname);
-    X3 = X3(:,:,J); % focus only on level J approximations ?
-[X4,~,~,~] = swt2(im4log,J,wname);
-    X4 = X4(:,:,J); % focus only on level J approximations ?
-
-% % mean approximation coefficients
-% imRef_wave = X1 + X2 + X3 + X4;
-%     imRef_wave = imRef_wave/n;
-%
-D1 = (X1 - imRef).^2;
-D2 = (X2 - imRef).^2;
-D3 = (X3 - imRef).^2;
-D4 = (X4 - imRef).^2;
-d = [sum(D1(:)) sum(D2(:)) sum(D3(:)) sum(D4(:))];
[…]

### 2021-11-15 codes/EllispseChangeAnalysis_levels.m

@@ -1,296 +1,0 @@
-
-% Evaluation of different levels when mean approximation coefficients are used
-
-%%
-close all
-clear;
-%%
-im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
-im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
-im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
-im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
-%
-
-% Images are too big, consider subsampling
-subsamplingfactor = 2;
-im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
-totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
-totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
-%
-figure
-imshow(totalchanges)
-title('Total changes')
-
-%%
-eps = .00001;
-im1log = eps+double(im1);
-im2log = eps+double(im2);
-im3log = eps+double(im3);
-im4log = eps+double(im4);
-n=4;
-imRef = im1log + im2log + im3log + im4log;
-    imRef = imRef/n;
-
-% decomposition level fixed
-wname = 'db2';
-
-%% Case 1 : reference image 
-
-J = 1;
-
-[X1,~,~,~] = swt2(im1log,J,wname);
-    X1 = X1(:,:,J); % focus only on level J approximations ?
-[X2,~,~,~] = swt2(im2log,J,wname);
-    X2 = X2(:,:,J); % focus only on level J approximations ?
-[X3,~,~,~] = swt2(im3log,J,wname);
-    X3 = X3(:,:,J); % focus only on level J approximations ?
-[X4,~,~,~] = swt2(im4log,J,wname);
[…]

### 2021-11-15 codes/EllispseChangeAnalysis_methods.m

@@ -1,171 +1,0 @@
-
-% Evaluation of different methods
-
-%%
-close all
-clear;
-%%
-im1 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle1.png');
-im2 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle2.png');  
-im3 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle3.png');  
-im4 = imread('../figs/ImagesWithEllipsoidChanges/ImageEllipseSpeckle4.png');  
-%
-
-%
-% Images are too big, consider subsampling
-subsamplingfactor = 2;
-im1 = im1(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im2 = im2(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im3 = im3(1:subsamplingfactor:end,1:subsamplingfactor:end);
-im4 = im4(1:subsamplingfactor:end,1:subsamplingfactor:end);
-totalchanges = imread('../figs/GroundTruthEllipsoidChanges/TotalEllipseChanges.png');  
-totalchanges = totalchanges(1:subsamplingfactor:end,1:subsamplingfactor:end);
-%
-mImage = figure;
-imshow(totalchanges)
-title('Total changes', 'FontSize', 17)
-%saveas(mImage,sprintf('../figs/total_changes.jpg'))
-
-%%
-eps = .00001;
-im1log = eps+double(im1);
-im2log = eps+double(im2);
-im3log = eps+double(im3);
-im4log = eps+double(im4);
-n=4;
-imRef = im1log + im2log + im3log + im4log;
-imRef = imRef/n;
-
-%% comparison with mean image
-
-wname = 'db2';
-J = 2;
-
-[X1,~,~,~] = swt2(im1log,J,wname);
-    X1 = X1(:,:,J); % focus only on level J approximations ?
-[X2,~,~,~] = swt2(im2log,J,wname);
-    X2 = X2(:,:,J); % focus only on level J approximations ?
-[X3,~,~,~] = swt2(im3log,J,wname);
-    X3 = X3(:,:,J); % focus only on level J approximations ?
-[X4,~,~,~] = swt2(im4log,J,wname);
-    X4 = X4(:,:,J); % focus only on level J approximations ?
-
-% % mean approximation coefficients
-% imRef_wave = X1 + X2 + X3 + X4;
[…]

### 2021-11-15 codes/Forest_data_analysis.m

@@ -105,7 +105,7 @@ xlabel('$m$','interpreter','latex','FontSize',20); xlim([0 n])
 ylabel('$\textbf{d}(m)$','interpreter','latex','FontSize',20);
 set(gca,'FontSize',13)
 hold off
-saveas(mImage,sprintf('../figs/forest_vSumDifCoefSq.jpg'))
+%saveas(mImage,sprintf('../figs/forest_vSumDifCoefSq.jpg'))
 
 % checking the images corresponding to change points in time
 mImage = figure;
@@ -128,7 +128,7 @@ for m = [25 27 30]
 end
 subplot(2,2,4)
 colormap(gray(256)); imagesc(imRef); title('Mean image'); axis off;
-saveas(mImage,sprintf('../figs/forest_changes_time.jpg'))
+%saveas(mImage,sprintf('../figs/forest_changes_time.jpg'))
 
 
 % computing correlations
@@ -144,13 +144,13 @@ mImage = figure;
 imagesc(mCorr)
 axis off; colorbar
 set(gca,'FontSize',13)
-saveas(mImage,sprintf('../figs/forest_wecs_abscorr.jpg'))
+%saveas(mImage,sprintf('../figs/forest_wecs_abscorr.jpg'))
 
 %histogram(mCorr(:))
 mImage = figure;
 cutoff = quantile(mCorr(:),1-1/log(Nx*Ny));
 imshow(mCorr>cutoff)
-saveas(mImage,sprintf('../figs/forest_wecs_change_space.jpg'))
+%saveas(mImage,sprintf('../figs/forest_wecs_change_space.jpg'))
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-11-17 tex/paper.wecs/wecs_tgrs__trans16nov21.tex
ajustes (17nov21)

--- 2021-11-14 tex/paper.wecs/wecs_tgrs.tex
@@
 
+\usepackage{mathabx}
+
 \makeatletter
@@
         and~Abdourrahmane~Atto,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
-\thanks{This work was supported by FAPESP grants 2016/24469-6 and 2018/04654-9 and CNPq grants 309230/2017-9 and 310991/2020-0.}
+\thanks{This work was supported by FAPESP (Grants 2016/24469-6, 2018/04654-9, 2021/01305-6) and CNPq (Grants 309230/2017-9 and 310991/2020-0).}
 \thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
@@
 
-%Besides different methodological paradigms, several areas of application receive special attention.
-\cite{song2018multi} discusses land cover change detection in mountainous terrain via multi-temporal and multi-sensor remote sensing images.
-\cite{ru2021multi} studies multi-temporal scene classification and scene change detection.
-
 In this paper we analyze Synthetic Aperture Radar (SAR) images through an unsupervised wavelet method.
@@
 
-%The motivation for our proposed method is multi-fold.
-We aim for a fast and accurate method.
-We would also like this method to be easily updatable when a new observation is captured.
-Finally, scalability was a concern as well.
-We propose a wavelet-based procedure for change detection in multi-temporal remore sensing images (WECS).
-It is unsupervised and built on ultra-high dimensional correlation screening \cite{fan2020statistical} for the wavelet coefficients.
-We present two complimentary wavelet measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images.
-The procedure presents some advantages.
-It is unsupervised, fast and updatable, thus allowing for real-time change detection.
-Moreover, it is sparse and scalable.
-
 The rest of the text goes as follows.
@@
 
+\section{Theory Background}\label{section_theroy}
+
+\textcolor{red}{Não seria interessante inserir uma breve secao abordando a questão de destaquei em vermelho mais adiante (sobre wavelet) assim como a definicao de feature screening, ultr-dimension, etc...
+Fazendo isso estaria preparando o terreno para as discussões que seguem.}
+
 \section{Methodology}\label{section_method}
 
-Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time.
-These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate.
+%%%RGN: vou usar texto em azul para coisas a serem conversadas depois.
+
+%Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time.
+Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be an image time series defined on a support $\mathcal{S} = \left\lbrace 1,\ldots,u  \right\rbrace \times \left\lbrace 1,\ldots,v  \right\rbrace \subset \mathbbm{N}^2$, hence representing a region of interest over $n$ distinct instants.
+%
+\textcolor{blue}{These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate}.
 Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
@@
 Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}.
-After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smooth version of $\mathcal{I}^{(m)}$ with the same dimension.
-The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX^{(m)}$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}^{(m)}$.
-Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX^{(m)}$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
-
-We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX^{(m)}\right\}$ as
-\begin{equation} 
-\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
+
+\textcolor{red}{Existe alguma formalizacao minima que poderia ser inserida aqui a fim de: (i) explicar como o método funciona; (ii) como X(m) origina deste processo; (iii) qual o significado do J neste contexto.
+A falta da ligacao entre o ponto de partida e as proximas formalizacoes deixa um pouco de duvidas.
+Se quiser, podemos fazer um texto bem compacto e um diagrama para simplificar.}
+
+After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smoothed version of $\mathcal{I}^{(m)}$
+defined on the same support $\mathcal{S}$.
+%with the same dimension. 
+The higher $J\in\{1,\ldots,\log_2( \min{u,v} )\}$ is, the smoother $\vX^{(m)}$ gets. %, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}^{(m)}$.
+
+%Besides many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX^{(m)}$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
+Beyond many other aspects that can be involved in wavelet analysis of images, which may include different types of wavelet transform and basis as well the use of thresholding for detail coefficients, the current construction is focused on $\vX^{(m)}$ to provide a simple wavelet smoothing.
+Nevertheless, extensions based on distinct aspects are straightforward.
+
+We can then consider further apportioning the total $\mathbb{L}_2$ energy of the series $\left\{\vX^{(m)}\right\}_{m=1,\ldots,n}$ as:
+\begin{equation}\label{E:eq_ANOVAlogimage}
+\begin{split}
+\sum_{m=1}^n\|\vX^{(m)}\|_F^2
+= n\|\overline{\mathcal{I}}\|_F^2 + 2n\langle\overline{\vX}-\overline{\mathcal{I}},\overline{\mathcal{I}}\rangle + \\
++ \sum_{m=1}^n\|\mathcal \vX^{(m)} - \overline{\mathcal{I}}\|_F^2,
+\end{split}
 \end{equation}
-where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
-The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX^{(m)}$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
-Since each element (pixel) in $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space.
-Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further.
-Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX^{(m)}-\vX^{(m-1)}\|^2$, $d(\vX^{(m)},\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
-
-Let $X_{k,l}^{(m)}$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\bar{\mathcal{I}}$, respectively.
-Computing $D_{k,l}^{(m)}=(X_{k,l}^{(m)}-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD^{(m)} = [D_{k,l}^{(m)}]$.
-We then
-analyze the time series given by
-\begin{equation} 
-\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
-\label{E:eq_defdm}
+where $\overline{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$ and $\overline{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$; $\| \cdot \|_F$ and $\left\langle \cdot , \cdot \right\rangle_F$ represents the Forbenius norm and inner product, respectively.
+Naturally, both $\overline{\mathcal{I}}$ and $\overline{\vX}$ shares the support $\mathcal{S}$.
+
+% and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
+
+%\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,
+
+The most right-hand term in Equation~(\ref{E:eq_ANOVAlogimage}) measures the deviations 
+between the wavelet coefficients at distinct instants ($\vX^{(m)}; \ m=1,\ldots,n$) regarding the \textit{average image} (i.e., $\overline{\mathcal{I}}$).
+Such deviations may favor detecting relevant changes over the time.
+%us to such deviations to detect change points in time
+%in time of $\vX^{(m)}$ compared to an {\it average image} $\overline{\mathcal{I}}$, what motivates us to such deviations to detect change points in time. 
+%
+Since each element (i.e., a pixel/position $(k,l)\in \mathcal{S}$) of $\vX^{(m)}$ also has a corresponding sequence of deviations in time, 
+%the relation of each 
+the local deviation to the overall measure 
+%can also be quantified, what allows us to detect changes in space. 
+may also allow detecting the spatial changes. 
+%Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further.
+Among several approaches able to detect change events, the Pearson correlation coefficient rises as an convenient alternative.
+Furthermore, such measure \textcolor{blue}{shares connections with the idea of feature screening employed in high-dimensional regression, as explained further.}
+
+%%%RGN: Desnecessario? (e pode confundir...)
+%Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX^{(m)}-\vX^{(m-1)}\|^2$, $d(\vX^{(m)},\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
+
+%%%RGN: vou remover a "," nos indices de posica para despoluir/descarregar  a notacão, ok?
+
+Let $X_{kl}^{(m)}$ and $\overline{\mathcal{I}}_{kl}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\overline{\mathcal{I}}$, respectively. 
+%
+The matrix $\vD^{(m)}$ embraces the squared differences between $\vX^{(m)}$ and $\overline{\mathcal{I}}$, where $D_{kl}^{(m)} = \left( \vX_{kl}^{m} - \overline{\mathcal{I}}_{kl} \right)^2$.
+%
+Therefore, it is defined the temporal overall variation sequence $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$, whose elements are given by:
+\begin{equation}\label{E:eq_defdm}
+\vd(m)= \sum_{k=1}^{u} \sum_{l=1}^{v} D_{kl}^{(m)} %; \ m=1,\ldots,n
 \end{equation}
-$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.
-
-The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy.
-Define the  $n\times p$ matrix
-\begin{equation*}
+
+%$X_{kl}^{(m)}$ and $\overline{\mathcal{I}}_{kl}$ at each position $(k,l)$ as 
+%Computing $D_{kl}^{(m)}=(X_{k,l}^{(m)}-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD^{(m)} = [D_{k,l}^{(m)}]$. We then 
+%analyze the time series given by
+%\begin{equation} 
+%\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
+%\label{E:eq_defdm}
+%\end{equation}
+%$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.
+
+In this context, a instant $m$ in $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$ with high value stands for the image $\mathcal{I}^{(m)}$ where the most expressive changes take place.
+
+%The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy.
+
+%%%RGN: Seria esta a razão?
+In order to obtain an expression for the spatio-temporal fluctuations, let consider the following global matrix $\vD$ of order $n \times uv$:
+%Define the  $n\times p$ matrix
+\begin{equation}\label{eq_defmatrixD}
  \vD=\left(
  \begin{array}{c}
- vec(\vD(1))^T\\
+ vec(\vD^{(1)})^T\\
  \vdots\\
- vec(\vD(n))^T\\
+ vec(\vD^{(n)})^T\\
  \end{array}
  \right),
-\label{eq_defmatrixD}
+\end{equation}
+where $vec(\vD^{(m)})$ is a $uv \times 1$ vector representation, using lexicographic ordering, of the wavelet coefficients regarding the instant $m$.
+%%%RGN: esta parte fica redundante, pois já definimos os elementos envolvidos aqui...
+% and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$. 
+\textcolor{red}{(precisa mesmo estar transposto? E se definir como $1 \times uv$?)}
+
+%%%%RGN: Acho que a posicao desta citacao esta errada, nao?
+Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role in this formalization.
+% here.  
+We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s$, is such that $s \ll p$.
+This can be represented as the following linear model
+\begin{equation}\label{E:sparsemodel_d}
+\vd=\vD\vbeta+\vxi
+\end{equation}
+where $\vbeta \in \mathbbm{R}^{uv}$ is sparse, i.e., it has $(uv - s)$ null elements, and $\vxi^{(d)} \in \mathbbm{R}^{n}$ is a random vector.
+\textcolor{red}{(existe esse $(d)$ no $\xi$ mesmo?
+Como $\beta$ e $\xi$ são obtidos?)} % of errors
+%some $n\times 1$ random vector of errors.
+
+%%%RGN: vamos trocar $s_d$ por $s$? a adicao do d nao faz muita diferenca.
+
+In order to identify spatio-temporal changes, we employ the concept of ultra-high dimensional correlation screening \cite{fan2020statistical} discussed in Section~\ref{section_theroy}.
+%as follows. 
+For each local \textcolor{red}{squared mean deviation} time series, given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
+\begin{equation*}\label{def_Rkld}
+R_{kl}= |\corr \left( \vD_{kl}, \vd \right)|,
 \end{equation*}
-where $vec(\vD^{(m)})$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$.
-Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.
-We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d<<p$.
-This can be represented as the following linear model
-\begin{equation} 
-\vd=\vD\vbeta+\vxi
-\label{E:sparsemodel_d}
-\end{equation}
-where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
-
-In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
-For each local squared mean deviation time series, given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
-\begin{equation*}
-R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
-\label{def_Rkld}
-\end{equation*}
-where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.
+where $\vD_{kl} = \left\lbrace D_{kl}^{(1)},\ldots,D_{kl}^{(n)} \right\rbrace$ is the time series of \textcolor{red}{squared mean deviations} of wavelet coefficients at position $(k,l)$.
+% the two-dimensional index $\{k,l\}$.
+
+%%%RGN: este trecho acima (...where $\vD_{k,l}=(D_{k,l}(1)) pode ser omitido, pois ja definimos quem é D_kl antes da equacao 2. Ok?
+
+\textcolor{magenta}{RGN: Nos destaques em vermelho que fiz acima, é para chamar atenćão que nao se trata de uma squared mean deviation e sim uma squered deviation, certo?}
 
@@
 
-We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
-{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
-$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
+Consequently, rises the matrix $\vR$ from the elements $R_{kl}$ assigned to each $(k,l) \in \mathcal{S}$.% which contain the ultra-high dimension correlations.
+%
+%We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. 
+%Define the set of {\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as $\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
 This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
-We build the empirical set of indices flagged as corresponding to change locations by
+
+Defining a mapping of \textit{important} indices for changes over the image series with respect to $\overline{\mathcal I}$ as $\mathcal{M}^{\star} \subseteq \mathcal{S}$ where \textit{changes in $\{\mathcal{I}^{(m)}\}_{m=1,\ldots,n}$ with respect to $\overline{\mathcal{I}}$ are affected by changes in the images}.
+%
+Such mapping, when re-written using lexicographic order as adopted to express the columns of $\vR$ (Eq.~\ref{eq_defmatrixD}), coincides with the non-zero components of $\vbeta$ in Equation~\ref{E:sparsemodel_d}.
+
+%$\mathcal{M}^{\star}=\{(k,l) \in \mathcal{S}: \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
+%This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
+
+An empirical mapping of change locations may be stated by:
 \begin{equation}
-\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
+\mathcal{M}_{\tau} = \{(k,l) \in \mathcal{S} : |R_{kl}|>\tau\},
 \label{E:def_Mtaud}
 \end{equation}
-where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
-\[
-P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
-\]
-as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.
+where $\tau \in mathbbm{R}_{+}^{*}$ is a convenient threshold value \textcolor{blue}{defined as function of $n$ and $J$}. 
+Under \textcolor{red}{some regularity conditions}, the probability $P(\mathcal{M}_{\tau}\supset\mathcal{M}^{\star}) \rightarrow 1$ as $n\rightarrow\infty$ \cite{fan2020statistical}.
+In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ when the number of \textcolor{red}{observations} $n$ is large enough.
+%%%RNG: observations mesmo? ou instats?
+
+%We build the empirical set of indices flagged as corresponding to change locations by
+%\begin{equation}
+%\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
+%\label{E:def_Mtaud}
+%\end{equation}
+%where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
+%\[
+%P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
+%\]
+%as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.
 In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{*}$ when the number of observations $n$ is large.
+
+{\color{blue}
+RGN: Vamos omitir este trecho abaixo por um instante e depois retomamos? Por enquanto parece faltar uma ligacao clara entreos elementos
 
@@
 Thus we are performing a truly spatio-temporal change detection in a single procedure.
+}
 
@@
 
-\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.23\textwidth]{../../figs/coming-soon.png}\label{figImgSentinelROIs}}
+\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Samples.pdf}\label{figImgSentinelROIs}}
 

### 2021-11-18 codes/Forest_data_analysis.m
images of change time as tiff

@@ -108,9 +108,9 @@ hold off
 %saveas(mImage,sprintf('../figs/forest_vSumDifCoefSq.jpg'))
 
 % checking the images corresponding to change points in time
-mImage = figure;
-cont = 1;
 for m = [25 27 30]
+    mImage = figure;
+
     t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{m,:})),'r');
     Y = read(t);
     data = double(Y(:,:,1).^2 + Y(:,:,2).^2);
@@ -122,13 +122,15 @@ for m = [25 27 30]
     % wavelet approximation for log of VV chanel
     [tmp2,~,~,~] = swt2(tmp(1:2^pwr2,1:2^pwr2),J,wname);
 
-    subplot(2,2,cont); cont = cont + 1;
-    colormap(gray(256)); imagesc(tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J))    
-    title(sprintf('m=%2d',m)); axis off;
+    %subplot(2,2,cont); cont = cont + 1;
+    colormap(gray(256)); imagesc(tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J))
+    axis off
+    saveas(mImage,sprintf('../figs/forest_changes_time_m%2d.tiff',m))
 end
-subplot(2,2,4)
-colormap(gray(256)); imagesc(imRef); title('Mean image'); axis off;
-%saveas(mImage,sprintf('../figs/forest_changes_time.jpg'))
+mImage = figure;
+colormap(gray(256)); imagesc(imRef); axis off;
+axis off
+saveas(mImage,sprintf('../figs/forest_changes_time_mean.tiff',m))
 
 
 % computing correlations

### 2021-11-20 tex/paper.wecs/wecs_tgrs.tex
updates in tex
changes in tex - methodology

@@
 \newcommand{\vbeta}{\pmb{\beta}}
+\newcommand{\vepsilon}{\pmb{\epsilon}}
 \newcommand{\vxi}{\pmb{\xi}}
-\newcommand{\vepsilon}{\pmb{\epsilon}}
 \def\shalf{\mbox{{\tiny$\frac{1}{2}$}}}
@@
 \newcommand{\vt}{{\textbf t}}
+\newcommand{\vy}{{\textbf y}}
 
@@
 
+\section{Theory Background}\label{section_theroy}
+
+Wavelet methods have been widely applied to analyze images in signal processing literature, specially for tasks such as signal denoising and compression \cite{mallat1998wavelet}.
+The most common way of describing wavelet representations is as a multiresolution decomposition, where a signal is represented on approximation and detail coefficients, which provide coarse and finer details of the signal, respectively.
+In practice, the discrete wavelet transform of matrices (image) consists in applying low and high-pass convolution filters to its rows and columns \cite{mallat1989theory}.
+In case of smoothing, applying such low-pass filter $J$ times to rows and columns of a matrix $\mathcal{I}$, yields a smooth image $\vX$.
+The number $J$ is also called resolution level, a tuning parameter for wavelet smoothing.
+This wavelet smoothing is employed in the next section as a denoising step for image analysis.
+
+Another idea that shall be employed in this work is the feature screening procedure, a method originally designed for ultra-high dimensional regression models \cite{fan2008sure}.
+In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, where $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise.
+The problem setup is that $p$ is much larger than $n$, what makes standard regression methods unfeasible, but only a handful of the available covariates are relevant for the model, i.e., have a nonzero corresponding parameter $\beta_i$.
+The feature screening idea consists in computing the sample correlation $\mathrm{corr}(\vy,\vW_i)$ among response and explanatory variables, and then selecting those covariates whose correlation are among the highest values.
+Under suitable conditions, such method is known to select a set containing all true covariates with high probability.
+In the image change detection problem, our idea is that a similar approach could be used to detect change locations by taking an overall change measure as response variable and local (pixel) measures as potential covariates.
+
 \section{Methodology}\label{section_method}
 
-Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time.
+Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be an image time series defined on a support $\mathcal{S} = \left\lbrace 1,\ldots,u  \right\rbrace \times \left\lbrace 1,\ldots,v  \right\rbrace \subset \mathbbm{N}^2$, hence representing a region of interest over $n$ distinct instants.
+%
+\textcolor{blue}{These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate}.
+Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
+We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}^{(m)}$.
+Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}.
+
+After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smoothed version of $\mathcal{I}^{(m)}$
+defined on the same support $\mathcal{S}$. The higher $J\in\{1,\ldots,\log_2( \min{u,v} )\}$ is, the smoother $\vX^{(m)}$ gets.
+
+Beyond many other aspects that can be involved in wavelet analysis of images, which may include different types of wavelet transform and basis as well the use of thresholding for detail coefficients, the current construction is focused on $\vX^{(m)}$ to provide a simple wavelet smoothing.
+Nevertheless, extensions based on distinct aspects are straightforward.
+
+We can then consider further apportioning the total $\mathbb{L}_2$ energy of the series $\left\{\vX^{(m)}\right\}_{m=1,\ldots,n}$ as:
+\begin{equation}\label{E:eq_ANOVAlogimage}
+\begin{split}
+\sum_{m=1}^n\|\vX^{(m)}\|_F^2
+= n\|\overline{\mathcal{I}}\|_F^2 + 2n\langle\overline{\vX}-\overline{\mathcal{I}},\overline{\mathcal{I}}\rangle_F + \\
++ \sum_{m=1}^n\|\mathcal \vX^{(m)} - \overline{\mathcal{I}}\|_F^2,
+\end{split}
+\end{equation}
+where $\overline{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$ and $\overline{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$; $\| \cdot \|_F$ and $\left\langle \cdot , \cdot \right\rangle_F$ represents the Forbenius norm and inner product, respectively.
+Naturally, both $\overline{\mathcal{I}}$ and $\overline{\vX}$ shares the support $\mathcal{S}$.
+
+The most right-hand term in Equation~(\ref{E:eq_ANOVAlogimage}) measures the deviations 
+between the wavelet coefficients at distinct instants ($\vX^{(m)}; \ m=1,\ldots,n$) regarding the \textit{average image} (i.e., $\overline{\mathcal{I}}$).
+Such deviations may favor detecting relevant changes over the time.
+Since each element (i.e., a pixel/position $(k,l)\in \mathcal{S}$) of $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the local deviation to the overall measure
+may also allow detecting the spatial changes.
+Among several approaches able to detect change events, the Pearson correlation coefficient rises as an convenient alternative.
+Furthermore, such measure \textcolor{blue}{shares connections with the idea of feature screening employed in high-dimensional regression, as explained earlier.}
+
+Let $X_{kl}^{(m)}$ and $\overline{\mathcal{I}}_{kl}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\overline{\mathcal{I}}$, respectively. 
+%
+The matrix $\vD^{(m)}$ embraces the squared differences between $\vX^{(m)}$ and $\overline{\mathcal{I}}$, where $D_{kl}^{(m)} = \left( \vX_{kl}^{m} - \overline{\mathcal{I}}_{kl} \right)^2$.
+%
+Therefore, it is defined the temporal overall variation sequence $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$, whose elements are given by:
+\begin{equation}\label{E:eq_defdm}
+\vd(m)= \sum_{k=1}^{u} \sum_{l=1}^{v} D_{kl}^{(m)} %; \ m=1,\ldots,n
+\end{equation}
+
+In this context, a instant $m$ in $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$ with high value stands for the image $\mathcal{I}^{(m)}$ where the most expressive changes take place.
+
+%%%%RGN: Seria esta a razão?
+%In order to obtain an expression for the spatio-temporal fluctuations, let consider the following global matrix $\vD$ of order $n \times uv$:
+%%Define the  $n\times p$ matrix
+%\begin{equation}\label{eq_defmatrixD}
+% \vD=\left(
+% \begin{array}{c}
+% vec(\vD^{(1)})^T\\
+% \vdots\\
+% vec(\vD^{(n)})^T\\
+% \end{array}
+% \right),
+%\end{equation}
+%where $vec(\vD^{(m)})$ is a $uv \times 1$ vector representation, using lexicographic ordering, of the wavelet coefficients regarding the instant $m$.
+%%%%RGN: esta parte fica redundante, pois já definimos os elementos envolvidos aqui...
+%% and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$. 
+%\textcolor{red}{(precisa mesmo estar transposto? E se definir como $1 \times uv$?)}
+%
+%
+%%%%%RGN: Acho que a posicao desta citacao esta errada, nao?
+%Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role in this formalization.
+%% here.  
+%We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s$, is such that $s \ll p$.
+This can be represented as the following linear model
+%\begin{equation}\label{E:sparsemodel_d}
+%\vd=\vD\vbeta+\vxi
+%\end{equation}
+%where $\vbeta \in \mathbbm{R}^{uv}$ is sparse, i.e., it has $(uv - s)$ null elements, and $\vxi^{(d)} \in \mathbbm{R}^{n}$ is a random vector.
+\textcolor{red}{(existe esse $(d)$ no $\xi$ mesmo?
+Como $\beta$ e $\xi$ são obtidos?)} % of errors
+%%some $n\times 1$ random vector of errors.
+
+%%%RGN: vamos trocar $s_d$ por $s$? a adicao do d nao faz muita diferenca.
+
+In order to identify spatio-temporal changes, we employ the concept of ultra-high dimensional correlation screening \cite{fan2020statistical} discussed in Section~\ref{section_theroy}.
+For each local squared deviation time series given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, \textcolor{blue}{say $\vD_{kl} = \left\lbrace D_{kl}^{(1)},\ldots,D_{kl}^{(n)} \right\rbrace$}, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
+\begin{equation*}\label{def_Rkld}
+R_{kl}= |\corr \left( \vD_{kl}, \vd \right)|.
+\end{equation*}
+Consequently, rises the matrix $\vR$ from the elements $R_{kl}$ assigned to each $(k,l) \in \mathcal{S}$.
+
+%%%RGN: este trecho acima (...where $\vD_{k,l}=(D_{k,l}(1)) pode ser omitido, pois ja definimos quem é D_kl antes da equacao 2. Ok?
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
+\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
+\label{F:EllipsoidChanges}
+\end{figure}
+
+Let us define a mapping of \textit{important} indices for changes over the image series with respect to $\overline{\mathcal I}$ as $\mathcal{M}^{\star} \subseteq \mathcal{S}$ where \textit{changes in $\{\mathcal{I}^{(m)}\}_{m=1,\ldots,n}$ with respect to $\overline{\mathcal{I}}$ are affected by local changes in the images}.
+%
+%Such mapping, when re-written using lexicographic order as adopted to express the columns of $\vR$ (Eq.~\ref{eq_defmatrixD}), coincides with the non-zero components of $\vbeta$ in Equation~\ref{E:sparsemodel_d}.
+
+%$\mathcal{M}^{\star}=\{(k,l) \in \mathcal{S}: \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
+%This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
+
+An empirical mapping of change locations may be stated by:
+\begin{equation}
+\mathcal{M}_{\tau} = \{(k,l) \in \mathcal{S} : |R_{kl}|>\tau\},
+\label{E:def_Mtaud}
+\end{equation}
+where $\tau \in \mathbbm{R}_{+}^{*}$ is a convenient threshold value \textcolor{blue}{defined as function of $n$}.
+Under \textcolor{red}{some regularity conditions}, the probability $P(\mathcal{M}_{\tau}\supset\mathcal{M}^{\star}) \rightarrow 1$ as $n\rightarrow\infty$ \cite{fan2020statistical}.
+In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ when the number of instants $n$ is large enough.
+
+%\section{Methodology}\label{section_method}
+%
+%Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time.
 These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate.
@@
 Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX^{(m)}$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
-
-We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX^{(m)}\right\}$ as
-\begin{equation} 
-\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
-\end{equation}
-where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
+%
+%We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX^{(m)}\right\}$ as
+%\begin{equation} 
+%\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
+%\end{equation}
+%where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
 The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX^{(m)}$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
@@
 Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX^{(m)}-\vX^{(m-1)}\|^2$, $d(\vX^{(m)},\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
-
-Let $X_{k,l}^{(m)}$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\bar{\mathcal{I}}$, respectively.
+%
+%Let $X_{k,l}^{(m)}$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\bar{\mathcal{I}}$, respectively.
 Computing $D_{k,l}^{(m)}=(X_{k,l}^{(m)}-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD^{(m)} = [D_{k,l}^{(m)}]$.
 We then
-analyze the time series given by
-\begin{equation} 
-\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
-\label{E:eq_defdm}
-\end{equation}
-$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.
-
-The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy.
+%analyze the time series given by
+%\begin{equation} 
+%\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
+%\label{E:eq_defdm}
+%\end{equation}
+%$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.
+%
+%The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy.
 Define the  $n\times p$ matrix
-\begin{equation*}
- \vD=\left(
- \begin{array}{c}
- vec(\vD(1))^T\\
- \vdots\\
- vec(\vD(n))^T\\
- \end{array}
- \right),
-\label{eq_defmatrixD}
-\end{equation*}
-where $vec(\vD^{(m)})$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$.
+%\begin{equation*}
+% \vD=\left(
+% \begin{array}{c}
+% vec(\vD^{(1)})^T\\
+% \vdots\\
+% vec(\vD^{(n)})^T\\
+% \end{array}
+% \right),
+%\label{eq_defmatrixD}
+%\end{equation*}
+%where $vec(\vD^{(m)})$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$.
 Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.
@@
 This can be represented as the following linear model
-\begin{equation} 
-\vd=\vD\vbeta+\vxi
-\label{E:sparsemodel_d}
-\end{equation}
-where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
-
-In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
+%\begin{equation} 
+%\vd=\vD\vbeta+\vxi
+%\label{E:sparsemodel_d}
+%\end{equation}
+%where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
+%
+%In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
 For each local squared mean deviation time series, given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
-\begin{equation*}
-R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
-\label{def_Rkld}
-\end{equation*}
-where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.
-
-\begin{figure}[htb!]
-\centering
-\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
-\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
-\label{F:EllipsoidChanges}
-\end{figure}
-
-We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
-{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
-$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
+%\begin{equation*}
+%R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
+%\label{def_Rkld}
+%\end{equation*}
+%where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.
+%
+%
+%\begin{figure}[htb!]
+%\centering
+%\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
+%\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
+%\label{F:EllipsoidChanges}
+%\end{figure}
+%
+%
+%We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
+%{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
+%$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
 This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
-We build the empirical set of indices flagged as corresponding to change locations by
-\begin{equation}
-\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
-\label{E:def_Mtaud}
-\end{equation}
-where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
-\[
-P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
-\]
-as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.
+%We build the empirical set of indices flagged as corresponding to change locations by
+%\begin{equation}
+%\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
+%\label{E:def_Mtaud}
+%\end{equation}
+%where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
+%\[
+%P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
+%\]
+%as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.
 In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{*}$ when the number of observations $n$ is large.
-
-Further geometrical motivation for our proposal is given as follows.
+%
+%Further geometrical motivation for our proposal is given as follows.
 As defined by (\ref{E:eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values.
@@
 The indices defined by  (\ref{E:def_Mtaud}) are such that
-\[
-\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau,
-\]
-i.e., such that  $\sum_{m=1}^nD_{k,l}^{(m)}\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
-\[
-\left|\sum_{m:\vd(m)>U} D_{k,l}^{(m)}\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}^{(m)}\right|>\Delta,
-\]
-for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$).
+%\[
+%\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau,
+%\]
+%i.e., such that  $\sum_{m=1}^nD_{k,l}^{(m)}\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
+%\[
+%\left|\sum_{m:\vd(m)>U} D_{k,l}^{(m)}\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}^{(m)}\right|>\Delta,
+%\]
+%for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$).
 Thence, when we employ correlation screening we select the two-dimensional wavelet indices which have the closest empirical directions to the vector of image temporal changes.
 Thus we are performing a truly spatio-temporal change detection in a single procedure.
+%
 

### 2021-11-29 tex/paper.wecs/wecs_tgrs_trans26nov21.tex
ajustes semana.22~26nov (29nov21)

--- 2021-11-20 tex/paper.wecs/wecs_tgrs.tex
@@
-%% bare_jrnl.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% see http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% journal paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
+\documentclass[journal]{IEEEtran}
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-\documentclass[journal]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[journal]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
 \ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
 \else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
 
-% *** MATH PACKAGES ***
-%
 \usepackage{amsmath}		
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-%\usepackage{amsaddr}		
-%\usepackage{amsthm}		
-%\usepackage{amsaddr}		
 \usepackage{amsfonts}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
-% Axel Sommerfeldt. This package may be useful when used in conjunction with 
-% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
-% submissions have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.4.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/pkg/endfloat
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
 
@@
 
+\usepackage[export]{adjustbox}
+
 \makeatletter
@@
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+%\hyphenation{op-tical net-works semi-conduc-tor}
 
@@
 % Do not put math or special symbols in the title.
-\title{Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images}
+\title{
+Wavelet Spatio-Temporal Change Detection Method for Multi-Temporal SAR Images
+%Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images
+}
 %
@@
 }
-% <-this % stops a space
-%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
-
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
 
@@
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
 \begin{abstract}
-We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised procedure to detect spatio-temporal change points on multi-temporal SAR images.
+\textcolor{blue}{We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised procedure to detect spatio-temporal change points on multi-temporal SAR images.
 The procedure is based on wavelet approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients.
@@
 We also evaluate the proposed method on a time series of 84 satellite images in a forest region at the border of Brazil and the French Guiana.
-The proposed method displays good results in covering change regions, with the additional benefit of having simple and fast computation.
+The proposed method displays good results in covering change regions, with the additional benefit of having simple and fast computation.}
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Change detection, multi-temporal images, satellite images, wavelets.
+Change detection, Remote Sensing, multi-temporal images, simulated images, wavelets.
 \end{IEEEkeywords}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps.
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by the IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
 
 \IEEEPARstart{C}{hange} detection is an important task performed in remote sensing image that allows researchers and engineers to identify and evaluate modifications on land surfaces captured by multi-temporal satellite images.
-Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping}, for example, are of great importance to study the dynamics of regions sensitive to climate changes and human activity.
+Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping} are of great importance to study the dynamics of regions sensitive to climate changes and human activity.
 Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods to images available at larger and larger from time to time.
@@
 Most methods used for change detection analysis can be classified either as supervised (training data is used to set up the method) or unsupervised (fully data-driven techniques).
-We shall focus in this work on unsupervised approaches, whose examples in the literature include the works of \cite{bruzzone2000automatic}, \cite{celik2010change}, \cite{quin2014mimosa} and \cite{saha2020change}, for example.
+We shall focus in this work on unsupervised approaches, whose examples in the literature include the works of \cite{bruzzone2000automatic,celik2010change,quin2014mimosa,saha2020change,NegriEA2021,NegriFrery2021}.
 Many other proposals of methods vary in their motivations as well as in their applicability.
-Change detection in multi-temporal hyperspectral images is discussed in \cite{bovolo2015time},  \cite{liu2019review}, and \cite{matsunaga2017current}.
+Change detection in multi-temporal hyperspectral images is discussed in \cite{bovolo2015time,liu2019review,matsunaga2017current}.
 \cite{jia2018novel} pursue change detection techniques via non-local means and principal component analysis.
@@
 
-In this paper we analyze Synthetic Aperture Radar (SAR) images through an unsupervised wavelet method.
+%%%RGN: adaptando algumas passagens
+A special attention has been given for multi-temporal change detection using Synthetic Aperture Radar (SAR) images. \textcolor{red}{Include references for change detection with SAR data}. 
+Known to be not affected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
+Conversely, due to its acquisition architecture, the speckle noise typically affects the images obtained by SAR sensors and demands additional pre-processing before its use.
+
+Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
+In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and is allows an efficient computational treatment.
+
+In more details, Wavelet methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multi-scale/resolution information.
+Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. 
+Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical}.
+\textcolor{magenta}{However, the application of feature screening on wavelet coefficients is still a novel idea in change detection literature, and we show it has an interesting potential to provide good results even with simple algorithms.}
+%%%RGN: ainda não falamos de screening...
+
+%In this paper we analyze Synthetic Aperture Radar (SAR) images through an unsupervised wavelet method.
 Known to be not affected by weather, cloud and sunlight conditions, SAR images have been an important source of data for change detection methods in the literature \cite{bovolo2005detail}.
@@
 
-We propose here a novel method for unsupervised spatio-temporal change detection in a time series of SAR images.
-Wavelet Energies Correlation Screening (WECS) is based upon correlation screening for energy apportionment on wavelet approximations.
-The spatial character of the change detection is attained on pixel level, whereas total energy measures for each image in the time series quantify changes happening across time.
-The idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}.
+In the light of presented discussions and motivations, using Wavelet and data screening concepts, we propose the Wavelet Energies Correlation Screening (WECS), a novel unsupervised multi-temporal change detection method for SAR images.
+%
+The main idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}.
 Such method is usually employed in high-dimensional regression models to reduce the problem's dimension by subsetting the available covariates in such a way that true covariates are among the chosen ones with high probability \cite{fan2008sure}.
@@
 
-The rest of the text goes as follows.
+%We propose here a novel method for unsupervised spatio-temporal change detection in a time series of SAR images.
+Wavelet Energies Correlation Screening (WECS) is based upon correlation screening for energy apportionment on wavelet approximations.
+
+%The spatial character of the change detection is attained on pixel level, whereas total energy measures for each image in the time series quantify changes happening across time.
+
+%The idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}.
+Such method is usually employed in high-dimensional regression models to reduce the problem's dimension by subsetting the available covariates in such a way that true covariates are among the chosen ones with high probability \cite{fan2008sure}.
+We show that by applying the feature screening idea on multi-temporal images, we obtain a fast and accurate method to cover change regions with good detection rates.
+
+\textcolor{blue}{This paper is organized as follows:
+% rest of the text goes as follows. 
 Section \ref{section_method} introduces the problem and presents the proposed method.
@@
 In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.
-Section \ref{section_discussion} concludes the paper with a discussion.
+Section \ref{section_discussion} concludes the paper with a discussion.}
 
-\section{Theory Background}\label{section_theroy}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Basic Theory Background}\label{section_theroy}
 
 Wavelet methods have been widely applied to analyze images in signal processing literature, specially for tasks such as signal denoising and compression \cite{mallat1998wavelet}.
-The most common way of describing wavelet representations is as a multiresolution decomposition, where a signal is represented on approximation and detail coefficients, which provide coarse and finer details of the signal, respectively.
+The most common way of describing wavelet representations is as a multi-resolution decomposition, where a signal is represented on approximation and detail coefficients, which provide coarse and finer details of the signal, respectively.
 In practice, the discrete wavelet transform of matrices (image) consists in applying low and high-pass convolution filters to its rows and columns \cite{mallat1989theory}.
@@
 This wavelet smoothing is employed in the next section as a denoising step for image analysis.
+
+\textcolor{red}{Aqui falta uma ligacao entre wavelet-ultradimension-screening. Infelizmente não estou sabendo como ligar esses pontos. Poderia pensar em algo?
+\textit{Um conceito que pode ser empregado para lidar com a alta dimensionalidade dos dados que surge em problemas desta natureza são as técnicas de feature screening, originalmente desenvolvido para...}
+}
 
@@
 
-\section{Methodology}\label{section_method}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Wavelet Energy Correlation Screening}\label{section_method}
+
+Figure~\ref{figDiagram} depicts a high-level conceptualization for the proposed method. The elements included in such representation are formalized in the constructions as follows.
+
+%-------------------------
+\begin{figure}[htb!]
+\centering
+\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
+\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
+\label{figDiagram}
+\end{figure}
 
@@
 
-%%%%RGN: Seria esta a razão?
-%In order to obtain an expression for the spatio-temporal fluctuations, let consider the following global matrix $\vD$ of order $n \times uv$:
-%%Define the  $n\times p$ matrix
-%\begin{equation}\label{eq_defmatrixD}
-% \vD=\left(
-% \begin{array}{c}
-% vec(\vD^{(1)})^T\\
-% \vdots\\
-% vec(\vD^{(n)})^T\\
-% \end{array}
-% \right),
-%\end{equation}
-%where $vec(\vD^{(m)})$ is a $uv \times 1$ vector representation, using lexicographic ordering, of the wavelet coefficients regarding the instant $m$.
-%%%%RGN: esta parte fica redundante, pois já definimos os elementos envolvidos aqui...
-%% and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$. 
-%\textcolor{red}{(precisa mesmo estar transposto? E se definir como $1 \times uv$?)}
-%
-%
-%%%%%RGN: Acho que a posicao desta citacao esta errada, nao?
-%Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role in this formalization.
-%% here.  
-%We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s$, is such that $s \ll p$.
-This can be represented as the following linear model
-%\begin{equation}\label{E:sparsemodel_d}
-%\vd=\vD\vbeta+\vxi
-%\end{equation}
-%where $\vbeta \in \mathbbm{R}^{uv}$ is sparse, i.e., it has $(uv - s)$ null elements, and $\vxi^{(d)} \in \mathbbm{R}^{n}$ is a random vector.
-\textcolor{red}{(existe esse $(d)$ no $\xi$ mesmo?
-Como $\beta$ e $\xi$ são obtidos?)} % of errors
-%%some $n\times 1$ random vector of errors.
-
-%%%RGN: vamos trocar $s_d$ por $s$? a adicao do d nao faz muita diferenca.
-
 In order to identify spatio-temporal changes, we employ the concept of ultra-high dimensional correlation screening \cite{fan2020statistical} discussed in Section~\ref{section_theroy}.
@@
 
-%%%RGN: este trecho acima (...where $\vD_{k,l}=(D_{k,l}(1)) pode ser omitido, pois ja definimos quem é D_kl antes da equacao 2. Ok?
-
-\begin{figure}[htb!]
-\centering
-\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
-\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
-\label{F:EllipsoidChanges}
-\end{figure}
-
 Let us define a mapping of \textit{important} indices for changes over the image series with respect to $\overline{\mathcal I}$ as $\mathcal{M}^{\star} \subseteq \mathcal{S}$ where \textit{changes in $\{\mathcal{I}^{(m)}\}_{m=1,\ldots,n}$ with respect to $\overline{\mathcal{I}}$ are affected by local changes in the images}.
-%
-%Such mapping, when re-written using lexicographic order as adopted to express the columns of $\vR$ (Eq.~\ref{eq_defmatrixD}), coincides with the non-zero components of $\vbeta$ in Equation~\ref{E:sparsemodel_d}.
-
-%$\mathcal{M}^{\star}=\{(k,l) \in \mathcal{S}: \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
-%This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
 
@@
 
-%\section{Methodology}\label{section_method}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}
+
+In order to assess the proposed method, this section presents two distinct studies comprising distinct datasets.
+
+The first study (Sec.~\ref{secExpSimulated}) uses a simulated data set and focus on identify the most appropriated wavelet family and resolution level $J$.
+Specifically, the Haar (haar) \cite{ref}, Daubechies of order 2 and 4 (db2 and db4) \cite{red}, Coiflets of order 4 (coif4) \cite{ref} and Symlets of order 2 and 4 (sym2 and sym4) \cite{ref} wavelet families are analyzed.
+Posterior, once selected the wavelet family, the most suitable resolution level is then pointed out.
+
+Lastly, the performance of the proposed method is measured in terms of True/False Positive Ratios and Receiver Operating Characteristic (ROC) curve. 
+Comparisons with the standard Thresholding of Aggregate Absolute Difference (TAAD) and a non-wavelet version of WECS, herein called Energy Correlation Screening (ECS), são incluídas nos experiments.
+
+In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $A$ where $A_kl = Sum$.
+The ECS stands for the use of I^m instead of $X^m$ when defining $Dkl^m$ at Equation~\ref{E:eq_defdm}.
+
+%%%
+
+The second study (Sec.~\ref{secExpActual}) analyzes the performance of WECS, and respective comparison with TAAD and ECS, in a real-world application with actual SAR image series.
+The appropriate wavelet family and resolution level previously identified are employed.
+ROC curves, F1-Score \cite{ref}, the kappa coefficient, and the variance of kappa \cite{CongaltonGreen2009}, are adopted for this purpose.
+Additionally, the computational run-times are presented and discussed.
+
 %
-%Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time.
-These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate.
-Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
-We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}^{(m)}$.
-Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}.
-After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smooth version of $\mathcal{I}^{(m)}$ with the same dimension.
-The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX^{(m)}$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}^{(m)}$.
-Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX^{(m)}$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
-%
-%We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX^{(m)}\right\}$ as
-%\begin{equation} 
-%\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
-%\end{equation}
-%where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
-The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX^{(m)}$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
-Since each element (pixel) in $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space.
-Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further.
-Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX^{(m)}-\vX^{(m-1)}\|^2$, $d(\vX^{(m)},\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
-%
-%Let $X_{k,l}^{(m)}$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\bar{\mathcal{I}}$, respectively.
-Computing $D_{k,l}^{(m)}=(X_{k,l}^{(m)}-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD^{(m)} = [D_{k,l}^{(m)}]$.
-We then
-%analyze the time series given by
-%\begin{equation} 
-%\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
-%\label{E:eq_defdm}
-%\end{equation}
-%$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.
-%
-%The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy.
-Define the  $n\times p$ matrix
-%\begin{equation*}
-% \vD=\left(
-% \begin{array}{c}
-% vec(\vD^{(1)})^T\\
-% \vdots\\
-% vec(\vD^{(n)})^T\\
-% \end{array}
-% \right),
-%\label{eq_defmatrixD}
-%\end{equation*}
-%where $vec(\vD^{(m)})$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$.
-Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.
-We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d<<p$.
-This can be represented as the following linear model
-%\begin{equation} 
-%\vd=\vD\vbeta+\vxi
-%\label{E:sparsemodel_d}
-%\end{equation}
-%where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
-%
-%In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
-For each local squared mean deviation time series, given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
-%\begin{equation*}
-%R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
-%\label{def_Rkld}
-%\end{equation*}
-%where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.
+%In this section we assess the method formalized in Section~\ref{section_method} with basis in two distinct studies comprising simulated data and an actual SAR multi-temporal image series.
 %
 %
-%\begin{figure}[htb!]
-%\centering
-%\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
-%\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
-%\label{F:EllipsoidChanges}
-%\end{figure}
+% proposed method is are presented two analysis 
 %
 %
-%We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
-%{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
-%$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
-This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
-%We build the empirical set of indices flagged as corresponding to change locations by
-%\begin{equation}
-%\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
-%\label{E:def_Mtaud}
-%\end{equation}
-%where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
-%\[
-%P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
-%\]
-%as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.
-In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{*}$ when the number of observations $n$ is large.
-%
-%Further geometrical motivation for our proposal is given as follows.
-As defined by (\ref{E:eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values.
-This segregates the multi-temporal images, since the former time points identify the images in which significant changes occur, while the latter indices identify time points with no major changes.
-Consider $U>L>0$ such that the $s_d$ highest values of $\vd$ are larger than $U$, and the $n-s_d$ smallest values of $\vd$ are smaller than $L$.
-We also take $\delta=U-L$.
-The indices defined by  (\ref{E:def_Mtaud}) are such that
-%\[
-%\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau,
-%\]
-%i.e., such that  $\sum_{m=1}^nD_{k,l}^{(m)}\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
-%\[
-%\left|\sum_{m:\vd(m)>U} D_{k,l}^{(m)}\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}^{(m)}\right|>\Delta,
-%\]
-%for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$).
-Thence, when we employ correlation screening we select the two-dimensional wavelet indices which have the closest empirical directions to the vector of image temporal changes.
-Thus we are performing a truly spatio-temporal change detection in a single procedure.
-%
-
-\section{Experiments}
-
-In this section we shall evaluate the performance of WECS on two types of data sets.
+%we shall evaluate the performance of WECS on two types of data sets.
 Initially, we generate simulated data where the level of noise and true change regions are known.
@@
 
+%In this section we shall evaluate the performance of WECS on two types of data sets.
+Initially, we generate simulated data where the level of noise and true change regions are known.
+For these images, we analyze different aspects of concerning the application of WECS and compare its performance with a standard change detection method.
+In the following application, real satellite data is employed, where the performance of WECS is evaluated on predefined change regions.
+
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Simulated data analysis}\label{secExpSimulated}
@@
 
-\begin{figure}[htb!]
+%-----------------------
+\begin{figure}[hbt]
 \centering
-%\includegraphics[width=40pc]{EllipsoidChanges/ellipses_all}
-\includegraphics[scale=.11]{../../figs/ellipses_t1}
-\includegraphics[scale=.11]{../../figs/ellipses_t2}
-\includegraphics[scale=.11]{../../figs/ellipses_t3}
-\includegraphics[scale=.11]{../../figs/ellipses_t4}
+
+\mbox{
+\subfigure[cap1]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t1_adjust.png}\label{figSim1}} 
+\subfigure[cap2]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t2_adjust.png}\label{figSim2}}
+}
+\mbox{
+\subfigure[cap3]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t3_adjust.png}\label{figSim3}} 
+\subfigure[cap4]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t4_adjust.png}\label{figSim4}}
+}
+
 \caption{Example of the first four simulated multi-temporal images. Features and changes come as white ellipses and dots.}
@@
 
-\begin{figure}[htp!]
-\includegraphics[scale=.1]{../../figs/total_changes}(a)
-\includegraphics[scale=.1]{../../figs/corr_changes_dm}(b)
-\includegraphics[scale=.1]{../../figs/corr_changes_logratios}(c)
-\includegraphics[scale=.1]{../../figs/corr_changes_dm_nowavelets}(d)
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[Total changes]{\includegraphics[frame,width=0.23\textwidth]{../../figs/total_changes_adjust.png}\label{figResSim1}} 
+\subfigure[dbr WECS d(m), j=2]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_adjust.png}\label{figResSim2}} 
+}
+\mbox{
+\subfigure[Aggregation of absolute differences]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_logratios_adjust.png}\label{figResSim3}} 
+\subfigure[d(m) without wavelets]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_nowavelets_adjust.png}\label{figResSim4}} 
+}
+
 \caption{Simulated images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
@@
 
-\begin{figure}[htp!]
-\includegraphics[scale=.13]{../../figs/methods_comparison}\hspace{-.6cm}(a) 
-\includegraphics[scale=.13]{../../figs/levels_comparison}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{../../figs/families_comparison}\hspace{-.5cm}(c)
-\includegraphics[scale=.13]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
-\caption{ROC curves for detection of changing ellipses in simulated images and different methods.
-(a) The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue).
-(b) db2 $\vd(m)$ with different levels.
-(c) $\vd(m)$ with different wavelet bases and J=2;
-(d) The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line).}
-\label{F:EllipsoidChanges_details}
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue)]{\includegraphics[width=0.23\textwidth]{../../graphs/methodROC.pdf}\label{figROCSim1}}
+\subfigure[db2 $\vd(m)$ with different levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSim2}}
+}
+\mbox{
+\subfigure[$\vd(m)$ with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSim3}} 
+\subfigure[The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line)]{\includegraphics[width=0.23\textwidth]{../../graphs/comparisonROC.pdf}\label{figROCSim4}}
+}
+
+\caption{ROC curves for detection of changing ellipses in simulated images and different methods.}\label{F:EllipsoidChanges_details}
 \end{figure}
@@
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Actual remote sensing application}\label{secExpActual}
@@
 
+%-----------------------
 \begin{figure}[htp!]
@@
 
-\begin{figure}[htp!]
-\includegraphics[scale=.3]{../../figs/forest_changes_time} 
-\caption{Images $\vX^{(m)}$ from times $m$ correspond to top-3 highest values of $\vd(m)$ and the mean image $\bar{\mathcal{I}}$.}
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[$m=25$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_m25.pdf}\label{figImgSentinel2015}} 
+\subfigure[$m=27$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_m27.pdf}\label{figImgSentinel2017}}
+}
+
+\mbox{
+\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_m30.pdf}\label{figImgSentinel2015}} 
+\subfigure[$\verline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_mean.pdf}\label{figImgSentinel2017}}
+}
+
+\caption{Images $\vX^{(m)}$ from distinct instant $m$ and the mean image $\bar{\mathcal{I}}$.}
 \label{F:forest_change_times}
@@
 
-\begin{figure}[htp!]
+%-----------------------
+\begin{figure}[hbt]
 \centering
-\includegraphics[scale=.10]{../../figs/forest_change}\hspace{-.5cm}(a)
-\includegraphics[scale=.15]{../../figs/forest_roc_change}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{../../figs/forest_wecs_abscorr}\hspace{-.5cm}(c)
-\includegraphics[scale=.13]{../../figs/forest_aggreg_ratios}\hspace{-.5cm}(d)\\
-\includegraphics[scale=.10]{../../figs/forest_wecs_change_space}\hspace{-.5cm}(e) 
-\includegraphics[scale=.10]{../../figs/forest_aggreg_change_space}\hspace{-.5cm}(f)\\ 
-\caption{Change points in space of the forest data.
+
+\mbox{
+\subfigure[Ground truth samples]{\includegraphics[width=0.23\textwidth]{../../figs/forest_change}\label{figResForestWecs}} 
+\subfigure[ROC]{\includegraphics[width=0.23\textwidth]{../../figs/forest_roc_change}\label{figResForestAggreg}} 
+}
+
+\caption{
+Change points in space of the forest data. 
 (a) Regions (in white) where changes should be detected.
 (b) ROC curve for detection of changing regions. 
-(c) Matrix of absolute correlations obtained with WECS.
-(d) Change measures from aggregation of absolute differences.
-(e) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension).
-(f) Change regions detected using aggregated absolute differences and Otsu's thresholding.}
-\label{F:forest_roc_wecs_agg}
+}\label{F:forest_roc_wecs_agg___TEMP}
+\end{figure}
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_abscorr}\label{figResForestWecs}} 
+\subfigure[Log.Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_ratios}\label{figResForestAggreg}} 
+}
+
+\mbox{
+\subfigure[Change map -- WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_change_space}\label{figResForestWecs}} 
+\subfigure[Change map -- Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_change_space}\label{figResForestAggreg}} 
+}
+
+\caption{
+Change points in space of the forest data. 
+%(a) Regions (in white) where changes should be detected.
+%(b) ROC curve for detection of changing regions. 
+(a) Matrix of absolute correlations obtained with WECS.
+(b) Change measures from aggregation of absolute differences. 
+(c) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension). 
+(d) Change regions detected using aggregated absolute differences and Otsu's thresholding.}\label{F:forest_roc_wecs_agg}
 \end{figure}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}\label{section_discussion}

### 2021-12-01 tex/paper.wecs/wecs_tgrs_trans26nov21.tex
rev (01dez21)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}
+\section{Experiments}\label{secExperiments}
+
+%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment design}\label{secExpDesign}
 
@@
 The first study (Sec.~\ref{secExpSimulated}) uses a simulated data set and focus on identify the most appropriated wavelet family and resolution level $J$.
-Specifically, the Haar (haar) \cite{ref}, Daubechies of order 2 and 4 (db2 and db4) \cite{red}, Coiflets of order 4 (coif4) \cite{ref} and Symlets of order 2 and 4 (sym2 and sym4) \cite{ref} wavelet families are analyzed.
+Specifically, the Haar (haar), Daubechies of order 2 and 4 (db2 and db4), Coiflets of order 4 (coif4) and Symlets of order 2 and 4 (sym2 and sym4) wavelet families are analyzed \cite{BeylkinEA1991,Daubechies1992}.
 Posterior, once selected the wavelet family, the most suitable resolution level is then pointed out.
@@
 
-In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $A$ where $A_kl = Sum$.
-The ECS stands for the use of I^m instead of $X^m$ when defining $Dkl^m$ at Equation~\ref{E:eq_defdm}.
+In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $\vA$ where $A_{kl} = \sum_{m=1}^{n}\mathcal{I}_{kl}^{(m)}$.
+The ECS stands for the use of $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ when defining $D_{kl}^{(m)}$ at Equation~\ref{E:eq_defdm}.
+
+\textcolor{red}{
+Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}^{(m)} - \mathcal{I}_{k,l}^{(m-1)}|$.
+Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ in the WECS formulation.
+}
 
@@
 The appropriate wavelet family and resolution level previously identified are employed.
-ROC curves, F1-Score \cite{ref}, the kappa coefficient, and the variance of kappa \cite{CongaltonGreen2009}, are adopted for this purpose.
+ROC curves, F1-Score \cite{Rijsbergen1979}, the kappa coefficient, and the variance of kappa \cite{Congalton2019}, are adopted for this purpose.
 Additionally, the computational run-times are presented and discussed.
@@
 
-In this section we apply the change detection methods above on simulated data of multi-temporal images.
+This section applies the WECS, TAAD, and ECS methods to detect changes on a simulated multi-temporal image dataset. 
+Such series comprises 80 multi-temporal images and it is synthesized by repeating a sequence of four images with different types of changes plus a noise. 
+Examples of the first four instants are shown in Figure~\ref{F:EllipsoidChanges}. 
+These images/instants are generated by summing two matrices: (i) a binary signal matrix with ones denoting where the ellipses occur and zero elsewhere; (ii) and a noise matrix with random variables following a standard Gaussian distribution.
+
+The first image, $\mathcal{I}^{(1)}$, presents three elongated ellipses.
+The second image $\mathcal{I}^{(2)}$ has shorter and larger ellipses added.
+Smaller ellipses are then added to form $\mathcal{I}^{(3)}$ and $\mathcal{I}^{(4)}$.
+All the changes made that occur among subsequent images is shown in Figure~\ref{figResSim1}, where white regions (i.e., the ``ones'') correspond to changes along with the instants.
+
+Applying WECS to these images, we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation.
+An example of $\vR$ is presented in Figure~\ref{figResSim2}.
+For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a binary matrix that can be compared with the total true changes shown in Figure \ref{figResSim1}.
+Similarly, Figures~\ref{figResSim3} and \ref{figResSim4} depicts the matrices $\vA$ and $\vR$ provided by TAAD and ECS methods.
+
+%In this section we apply the change detection methods above on simulated data of multi-temporal images.
 The simulated multi-temporal images ($n=80$) is obtained by repeating a sequence of four images with different types of changes plus a noise.
@@
 \mbox{
-\subfigure[cap1]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t1_adjust.png}\label{figSim1}} 
-\subfigure[cap2]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t2_adjust.png}\label{figSim2}}
+\subfigure[$\mathcal{I}^{(1)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t1_adjust.png}\label{figSim1}} 
+\subfigure[$\mathcal{I}^{(2)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t2_adjust.png}\label{figSim2}}
 }
 \mbox{
-\subfigure[cap3]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t3_adjust.png}\label{figSim3}} 
-\subfigure[cap4]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t4_adjust.png}\label{figSim4}}
+\subfigure[$\mathcal{I}^{(3)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t3_adjust.png}\label{figSim3}} 
+\subfigure[$\mathcal{I}^{(4)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t4_adjust.png}\label{figSim4}}
 }
@@
 \subfigure[Total changes]{\includegraphics[frame,width=0.23\textwidth]{../../figs/total_changes_adjust.png}\label{figResSim1}} 
-\subfigure[dbr WECS d(m), j=2]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_adjust.png}\label{figResSim2}} 
+\subfigure[WECS -- db2, $J=2$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_adjust.png}\label{figResSim2}} 
 }
 \mbox{
-\subfigure[Aggregation of absolute differences]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_logratios_adjust.png}\label{figResSim3}} 
-\subfigure[d(m) without wavelets]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_nowavelets_adjust.png}\label{figResSim4}} 
+\subfigure[TAAD -- $\vA$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_logratios_adjust.png}\label{figResSim3}} 
+\subfigure[ECS]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_nowavelets_adjust.png}\label{figResSim4}} 
 }
 
-\caption{Simulated images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
+%\caption{Simulated images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
+\caption{Expected change/non-change regions and typical results provided bu WECS, TAAD and ECS before the thresholding process.}
 \label{F:Changes_methods_images}
@@
 
-Figure \ref{F:Changes_methods_images} illustrates a comparison of different approaches to detect accumulated changes.
+%Figure \ref{F:Changes_methods_images} illustrates a comparison of different approaches to detect accumulated changes.
 Panel (b) shows the result of WECS using Daubechies wavelet with two null moments (db2) and $J=2$.
@@
 
-We compute ROC curves to compare the detection performance of different methods and to verify the influence of some tuning parameters of wavelet smoothing: the resolution level $J$ and the choice of wavelet basis.
+According to the experiment design (Section~\ref{secExpDesign}), ROC curves are employed to compare the performance of each analyzed method and inspect the effects of both $J$ and wavelet basis on the performance of the WECS method.
+In summary, the ROC curves exhibit the performance as a function of true and false positive ratios when distinguishing two events (in this context, the change and non-changed pixels/positions) using distinct thresholds.
+The true and false positive ratios are computed using the ``total change'' image (Fig.~\ref{figResSim1}) as reference.
+Moreover, the tested thresholds embrace all values in $\vR$ (or $\vA$), excluding repetitions and considering it in ascending order.
+
+%We compute ROC curves to compare the detection performance of different methods and to verify the influence of some tuning parameters of wavelet smoothing: the resolution level $J$ and the choice of wavelet basis.
 Each detection method generates a matrix of change detection measures (correlations in the case of WECS).
 The ROC curves present the performance of change detection by applying a threshold on these measures, in the following way:
-\begin{enumerate}
-\item Let $R$ be the matrix of change measures. Compute the range $[r_{\min},r_{\max}]$ of the values in $R$;
-\item Let $(r_{(1)},\ldots,r_{(100)})$ be equally space values between $r_{\min}$ and $r_{\max}$;
-\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$ coincide with the element $(i,j)$ where a change really occurs on the image of total changes.
+%\begin{enumerate}
+%\item Let $R$ be the matrix of change measures. Compute the range $[r_{\min},r_{\max}]$ of the values in $R$;
+%\item Let $(r_{(1)},\ldots,r_{(100)})$ be equally space values between $r_{\min}$ and $r_{\max}$;
+%\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$ coincide with the element $(i,j)$ where a change really occurs on the image of total changes.
 Dividing this number by the total number of changes gives the true positive rate.
-\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$  do not coincide with the element $(i,j)$ where a change really occurs.
+%\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$  do not coincide with the element $(i,j)$ where a change really occurs.
 Dividing this number by the total number of entries where changes do not occur gives the false positive rate.
-\item The ROC curve is the plot of true and false positive rates corresponding to each $k$.
-\end{enumerate}
+%\item The ROC curve is the plot of true and false positive rates corresponding to each $k$.
+%\end{enumerate}
+
+%We compute ROC curves to compare the detection performance of different methods and to verify the influence of some tuning parameters of wavelet smoothing: the resolution level $J$ and the choice of wavelet basis.
+Each detection method generates a matrix of change detection measures (correlations in the case of WECS).
+The ROC curves present the performance of change detection by applying a threshold on these measures, in the following way:
+%\begin{enumerate}
+%\item Let $R$ be the matrix of change measures. Compute the range $[r_{\min},r_{\max}]$ of the values in $R$;
+%\item Let $(r_{(1)},\ldots,r_{(100)})$ be equally space values between $r_{\min}$ and $r_{\max}$;
+%\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$ coincide with the element $(i,j)$ where a change really occurs on the image of total changes.
+Dividing this number by the total number of changes gives the true positive rate.
+%\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$  do not coincide with the element $(i,j)$ where a change really occurs.
+Dividing this number by the total number of entries where changes do not occur gives the false positive rate.
+%\item The ROC curve is the plot of true and false positive rates corresponding to each $k$.
+%\end{enumerate}
 
@@
 \mbox{
-\subfigure[The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue)]{\includegraphics[width=0.23\textwidth]{../../graphs/methodROC.pdf}\label{figROCSim1}}
-\subfigure[db2 $\vd(m)$ with different levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSim2}}
+\subfigure[$\vd(m)$ with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSimBasis}}
+\subfigure[db2 $\vd(m)$ with different levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSimJ}}
 }
+
 \mbox{
-\subfigure[$\vd(m)$ with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSim3}} 
-\subfigure[The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line)]{\includegraphics[width=0.23\textwidth]{../../graphs/comparisonROC.pdf}\label{figROCSim4}}
+\subfigure[The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue)]{\includegraphics[width=0.23\textwidth]{../../graphs/methodROC.pdf}\label{figROCSimRes}}
+%\subfigure[The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line)]{\includegraphics[width=0.23\textwidth]{../../graphs/comparisonROC.pdf}\label{figROCSim4}}
 }
@@
 
-Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the simulated data as follows.
+%basis wavelet
+Figure~\ref{figROCSimBasis} presents the ROC curves considering distinct wavelet basis in the WECS method.
+All instances adopts $J=2$.
+The curves allow concluding that db2 and sym2 deliveries the best trade-off between true and false positive ratios (i.e., high true positive ratios even when the false positive ratios are low).
+As a consequence of this finding, all the following experiments and analyses consider the db2 wavelet basis.
+
+%J-level
+Figure~\ref{figROCSimJ} depicts the ROC curves with $J \in \{1,2,3,4,5\}$ as decomposition levels.
+The profiles exhibited by the ROC curves make evident that $J$ equal to 2 or 3 promotes the best performances, with a slight advantage for $J=2$, since it demands few decomposition levels.
+The overall performance for $J=2$ warrants its use for the rest of the comparisons.
+
+%Compara figA
+Al last, the ROC curves shown in Figure~\ref{figROCSimRes} allow comparing the performances of WECS, TAAD and ECS approaches. 
+Firstly, the low-performance assigned to the ECS method allows concluding that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet concept in the context of the proposed method.
+About the performance of TAAD method, a true positive ratio of 0.8 is guaranteed when tolerating a false positive ratio of approximately 0.4. 
+Reversely, the WECS method provides the same true positive ratio under an almost nill false positive ratio, demonstrating then its superiority compared to the competitors.
+
+%Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the simulated data as follows.
 The effects of wavelet bases, level of decomposition, deep-learning feature extraction are shown on the ROC curves.
@@
 We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS.
-We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
+%We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
 The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}^{(m)}\}$ instead of $\{\vX^{(m)}\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}^{(m)}\}$.
@@
 
-We employed the proposed change detection method on a series of 84 multi-date satellite images.
+This section demonstrates the WECS, TAAD and ECS methods in a real-world change detection application.
+The appropriate wavelet basis and resolution level previously identified in Section~\ref{secExpSimulated} (i.e., db2 and $J=2$) are still considered in this analysis.
+
+In detail, this application considers a large multi-temporal series of 84 images acquired over a forest region at the border of Brazil and the French Guiana, from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
+Each image contains the signal backscatters relative to VV and VH polarizations, a spatial resolution of \SI{10}{\meter} and support of $1538 \times 1556$ pixels wide.
+In order to apply the analyzed methods, the dual-polarized images were combined into a single-band representation considering
+$\mathcal{I}_{kl}^{(m)} = \sqrt{(\mathrm{VV}_{kl}^{(m)})^2+(\mathrm{VH}_{kl}^{(m)})^2}$, with $\mathrm{VV}$ and $\mathrm{VH}$ representing the available polarizations.
+
+%We employed the proposed change detection method on a series of 84 multi-date satellite images.
 The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3.
@@
 
+\textcolor{magenta}{
 A multi-resolution analysis based on a Symlet basis with filter of length 16 (symlet 8) is built.
@@
 Then, we are able to compute the squared mean differences vector $\vd$ and the matrix of absolute correlations $\vR$.
+}
 
-The time change measures computed with $\vd$ are shown in Figure \ref{F:forest_wecs}.
+The instantaneous change values expressed by $\vd$ are shown in Figure~\ref{F:forest_wecs}. 
+The top line represents the median absolute deviation of $\vd$ and allows highlighting the instants 25, 27, and 30; when expressive differences occur in comparison to the other instants.
+The images relative to these instants are in Figure~\ref{F:forest_change_times}.
+
+%The time change measures computed with $\vd$ are shown in Figure \ref{F:forest_wecs}.
 The orange line in Figure \ref{F:forest_wecs}a represents the median absolute deviation of $\vd$, which allow us to notice times that differ expressively from the others.
@@
 \subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_m30.pdf}\label{figImgSentinel2015}} 
-\subfigure[$\verline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_mean.pdf}\label{figImgSentinel2017}}
+\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_mean.pdf}\label{figImgSentinel2017}}
 }
 
-\caption{Images $\vX^{(m)}$ from distinct instant $m$ and the mean image $\bar{\mathcal{I}}$.}
+\caption{Images $\vX^{(m)}$ from distinct instant $m$ and the mean image $\overline{\mathcal{I}}$.}
 \label{F:forest_change_times}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}\label{section_discussion}
+\section{Conclusion}\label{section_discussion}
 

### 2021-12-04 codes/Forest_data_analysis.m
pastes with tif images and csv

@@ -32,8 +32,8 @@ imagesc(sqrt(Y(:,:,1).^2 + Y(:,:,2).^2))
 
 %% Configurations for wavelet decompositions
 
-wname='sym8';
-%wname='db2';
+%wname='sym8';
+wname='db2';
 
 % saving the current boundary handling and then changing it to periodic.
 % The default mode was giving different norms of vC for different
@@ -53,8 +53,9 @@ dwtmode('per');
 %******** analysis for combined chanel ************%
 
 % matrix of mean squared differences of approximation coefficients
-mDifCoefSq = zeros([Nx, Ny, n]);
+mD_wecs = zeros([Nx, Ny, n]);
 
+tic
 % mean image
 imRef = zeros(Nx,Ny);
 for m=1:n
@@ -82,35 +83,43 @@ for m=1:n
     % wavelet approximation for log of VV chanel
     [tmp2,~,~,~] = swt2(tmp(1:2^pwr2,1:2^pwr2),J,wname);
     % sqared mean deviations of level J approx. coeff.
-    mDifCoefSq(:,:,m) = (tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J) - imRef).^2;
+    mD_wecs(:,:,m) = (tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J) - imRef).^2;
     % mean SNR of observed images
     meanSNR = meanSNR + ...
         norm(tmp2(extens+1:extens+Nx,extens+1:extens+Ny,J),'fro')/(norm(data,'fro')*n);
 end
 clear Y t tmp tmp2
 
+% vector with total sum of squared mean deviations for each image
+vd_wecs = reshape(sum(sum(mD_wecs,1),2),1,n);
+
+% computing correlations
+mCorr_wecs = zeros(Nx,Ny);
+for ii=1:Nx
+    for jj=1:Ny
+        tmp = abs(corrcoef(reshape(mD_wecs(ii,jj,:),1,n),vd_wecs));
+        mCorr_wecs(ii,jj) = tmp(1,2);
+    end
+end
+toc
+
 % mean SNR for observed images
 meanSNR
 
-% vector with total sum of squared mean deviations for each image
-vSumDifCoefSq = reshape(sum(sum(mDifCoefSq,1),2),1,n);
-
[…]

### 2021-12-04 codes/SeqEllispseChangeAnalysis_methods.m

@@ -75,7 +75,7 @@ title('db2 WECS d(m), J=2', 'FontSize', 17)
 saveas(mImage,sprintf('../figs/corr_changes_dm.jpg'))
 
 
-%% No wavelets for comparison with mean image
+%% No wavelets for comparison with mean image (ECS)
 
 % matrix of squared mean differences
 mD = zeros(n1,n2,n);

### 2021-12-04 codes/save_tiff_image.m

@@ -1,0 +1,20 @@
+function  save_tiff_image(A,filename)
+% function to save a tiff image from matrix A (single chanel) in a file
+% named filename
+
+% saving a matrix of doubles as tiff image
+t = Tiff(filename, 'w');
+tagstruct.ImageLength = size(A, 1);
+tagstruct.ImageWidth = size(A, 2);
+tagstruct.Compression = Tiff.Compression.None;
+tagstruct.SampleFormat = Tiff.SampleFormat.IEEEFP;
+tagstruct.Photometric = Tiff.Photometric.MinIsBlack;
+tagstruct.BitsPerSample = 64;
+tagstruct.SamplesPerPixel = size(A,3);
+tagstruct.PlanarConfiguration = Tiff.PlanarConfiguration.Chunky;
+t.setTag(tagstruct);
+t.write(A);
+t.close();
+
+end
+

### 2021-12-05 codes/Forest_data_analysis.m
change image with feature screening threshold

@@ -146,11 +146,14 @@ axis off; colorbar
 set(gca,'FontSize',13)
 % saveas(mImage,sprintf('../figs/forest_wecs_abscorr.jpg'))
 
+% Change image using threshold suggested in feature screening literature
 %histogram(mCorr(:))
 mImage = figure;
 cutoff = quantile(mCorr_wecs(:),1-1/log(Nx*Ny));
 imshow(mCorr_wecs>cutoff)
 %saveas(mImage,sprintf('../figs/forest_wecs_change_space.jpg'))
+save_tiff_image(double(mCorr_wecs>cutoff),...
+        sprintf('forest_wecs_abscorr_screethrs.tiff'));
 
 cutoff_KI_wecs = kittler(mCorr_wecs); % Kittler-Illingworth threshold
 cutoff_Otsu_wecs = graythresh(mCorr_wecs); % Otsu's threshold

### 2021-12-05 codes/saving_tif_image.m

@@ -1,16 +1,0 @@
-% saving a matrix of doubles as tiff image
-filename = 'testtifffp.tiff';
-A = mCorr;
-t = Tiff(filename, 'w');
-tagstruct.ImageLength = size(A, 1);
-tagstruct.ImageWidth = size(A, 2);
-tagstruct.Compression = Tiff.Compression.None;
-tagstruct.SampleFormat = Tiff.SampleFormat.IEEEFP;
-tagstruct.Photometric = Tiff.Photometric.MinIsBlack;
-tagstruct.BitsPerSample = 64;
-tagstruct.SamplesPerPixel = size(A,3);
-tagstruct.PlanarConfiguration = Tiff.PlanarConfiguration.Chunky;
-t.setTag(tagstruct);
-t.write(A);
-t.close();
-

### 2021-12-05 tex/paper.wecs/wecs_tgrs.tex
correcting image of some instants

@@
 \IEEEPARstart{C}{hange} detection is an important task performed in remote sensing image that allows researchers and engineers to identify and evaluate modifications on land surfaces captured by multi-temporal satellite images.
-Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping}, for example, are of great importance to study the dynamics of regions sensitive to climate changes and human activity.
+Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping} are of great importance to study the dynamics of regions sensitive to climate changes and human activity.
 Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods to images available at larger and larger from time to time.
@@
 
-In this paper we analyze Synthetic Aperture Radar (SAR) images through an unsupervised wavelet method.
-Known to be not affected by weather, cloud and sunlight conditions, SAR images have been an important source of data for change detection methods in the literature \cite{bovolo2005detail}.
-Unsupervised analysis of such images consists in the steps of data denoising, pixel-level processing and change examination.
-Our idea is to explore the denoising property of wavelet transforms and apply a feature screening idea to identify change regions at pixels represented by wavelet coefficients.
-Wavelet methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multiscale/multiresolution information.
+A special attention has been given for multi-temporal change detection using Synthetic Aperture Radar (SAR) images.
+\textcolor{green}{Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail} }.
+Known to be not affected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
+Conversely, due to its acquisition architecture, the speckle noise typically affects the images obtained by SAR sensors and demands additional pre-processing before its use.
+
+Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
+In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and is allows an efficient computational treatment.
+
+In more details, Wavelet methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multi-scale/resolution information.
 Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. 
 Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical}.
-However, the application of feature screening on wavelet coefficients is still a novel idea in change detection literature, and we show it has an interesting potential to provide good results even with simple algorithms.
-
-We propose here a novel method for unsupervised spatio-temporal change detection in a time series of SAR images.
-Wavelet Energies Correlation Screening (WECS) is based upon correlation screening for energy apportionment on wavelet approximations.
-The spatial character of the change detection is attained on pixel level, whereas total energy measures for each image in the time series quantify changes happening across time.
-The idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}.
+\textcolor{magenta}{However, an application of screening ideas on wavelet coefficients is still a novel approach in change detection literature, and we show it has an interesting potential to provide good results even with simple algorithms.}
+
+In the light of presented discussions and motivations, using Wavelet and data screening concepts, we propose the Wavelet Energies Correlation Screening (WECS), a novel unsupervised multi-temporal change detection method for SAR images.
+%
+The main idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}.
 Such method is usually employed in high-dimensional regression models to reduce the problem's dimension by subsetting the available covariates in such a way that true covariates are among the chosen ones with high probability \cite{fan2008sure}.
@@
 
-The rest of the text goes as follows.
+\textcolor{blue}{This paper is organized as follows:
+% rest of the text goes as follows. 
 Section \ref{section_method} introduces the problem and presents the proposed method.
@@
 In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.
-Section \ref{section_discussion} concludes the paper with a discussion.
+Section \ref{section_discussion} concludes the paper with a discussion.}
 

### 2021-12-05 codes/Forest_data_analysis.m

@@ -122,7 +122,7 @@ saveas(mImage,sprintf('../figs/forest_vSumDifCoefSq.jpg'))
 for m = [25 27 30]
     t = Tiff(strcat('../../Images/timeSeries/ascending/',char(dates{m,:})),'r');
     Y = read(t);
-    data = double(Y(:,:,1).^2 + Y(:,:,2).^2);
+    data = double(Y(:,:,1).^2 + Y(:,:,2).^2)/normconst;
     
     % extending image to be able to use swt2
     pwr2 = ceil(log2(min([Nx Ny])));
@@ -209,7 +209,10 @@ tmp = double(S>cutoff_Otsu_taad);
 save_tiff_image(tmp,...
         sprintf('forest_taad_change_space_otsu.tiff'));
 
-%% No wavelets for comparison with mean image (ECS)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% No wavelets for comparison with mean image (ECS)
 
 % matrix of squared mean differences
 mD_ecs = zeros([Nx, Ny, n]);

### 2021-12-06 tex/paper.wecs/wecs_tgrs.tex
updates in the text

@@
 
-\section{Theory Background}\label{section_theroy}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Basic Theory Background}\label{section_theroy}
 
 Wavelet methods have been widely applied to analyze images in signal processing literature, specially for tasks such as signal denoising and compression \cite{mallat1998wavelet}.
-The most common way of describing wavelet representations is as a multiresolution decomposition, where a signal is represented on approximation and detail coefficients, which provide coarse and finer details of the signal, respectively.
+The most common way of describing wavelet representations is as a multi-resolution decomposition, where a signal is represented on approximation and detail coefficients, which provide coarse and finer details of the signal, respectively.
 In practice, the discrete wavelet transform of matrices (image) consists in applying low and high-pass convolution filters to its rows and columns \cite{mallat1989theory}.
@@
 
-Another idea that shall be employed in this work is the feature screening procedure, a method originally designed for ultra-high dimensional regression models \cite{fan2008sure}.
+\textcolor{green}{
+Using wavelet smoothing on the images consists in an initial step to analyze the data, but the main goal in the analysis is to find changes in space and at instants of time.
+The smoothed images still contain a large number of coefficients that need to be evaluated, what characterizes a high-dimensional problem with multiple time series corresponding to each location in space.
+Since most part of a region is usually not relevant for overall changes, we need to employ a method that retains the most important locations driving these modifications across time.
+Methods like variable selection are largely employed in statistics to identify the most important covariates in a study, and one method particularly efficient in identifying relevant variables is the feature screening technique.
+}
+
+Feature screening procedure is a method originally designed for ultra-high dimensional regression models \cite{fan2008sure}.
 In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, where $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise.
@@
 \end{equation}
-where $\overline{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$ and $\overline{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$; $\| \cdot \|_F$ and $\left\langle \cdot , \cdot \right\rangle_F$ represents the Forbenius norm and inner product, respectively.
-Naturally, both $\overline{\mathcal{I}}$ and $\overline{\vX}$ shares the support $\mathcal{S}$.
+where $\overline{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$ and $\overline{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$; $\| \cdot \|_F$ and $\left\langle \cdot , \cdot \right\rangle_F$ represents the Frobenius norm and inner product, respectively.
+Naturally, both $\overline{\mathcal{I}}$ and $\overline{\vX}$ share the support $\mathcal{S}$.
 
@@
 
-In this context, a instant $m$ in $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$ with high value stands for the image $\mathcal{I}^{(m)}$ where the most expressive changes take place.
-
-%%%%RGN: Seria esta a razão?
-%In order to obtain an expression for the spatio-temporal fluctuations, let consider the following global matrix $\vD$ of order $n \times uv$:
-%%Define the  $n\times p$ matrix
-%\begin{equation}\label{eq_defmatrixD}
-% \vD=\left(
-% \begin{array}{c}
-% vec(\vD^{(1)})^T\\
-% \vdots\\
-% vec(\vD^{(n)})^T\\
-% \end{array}
-% \right),
-%\end{equation}
-%where $vec(\vD^{(m)})$ is a $uv \times 1$ vector representation, using lexicographic ordering, of the wavelet coefficients regarding the instant $m$.
-%%%%RGN: esta parte fica redundante, pois já definimos os elementos envolvidos aqui...
-%% and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$. 
-%\textcolor{red}{(precisa mesmo estar transposto? E se definir como $1 \times uv$?)}
-%
-%
-%%%%%RGN: Acho que a posicao desta citacao esta errada, nao?
-%Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role in this formalization.
-%% here.  
-%We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s$, is such that $s \ll p$.
-This can be represented as the following linear model
-%\begin{equation}\label{E:sparsemodel_d}
-%\vd=\vD\vbeta+\vxi
-%\end{equation}
-%where $\vbeta \in \mathbbm{R}^{uv}$ is sparse, i.e., it has $(uv - s)$ null elements, and $\vxi^{(d)} \in \mathbbm{R}^{n}$ is a random vector.
-\textcolor{red}{(existe esse $(d)$ no $\xi$ mesmo?
-Como $\beta$ e $\xi$ são obtidos?)} % of errors
-%%some $n\times 1$ random vector of errors.
-
-%%%RGN: vamos trocar $s_d$ por $s$? a adicao do d nao faz muita diferenca.
+In this context, an instant $m$ in $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$ with high value stands for the image $\mathcal{I}^{(m)}$ where the most expressive changes take place.
 
 In order to identify spatio-temporal changes, we employ the concept of ultra-high dimensional correlation screening \cite{fan2020statistical} discussed in Section~\ref{section_theroy}.
-For each local squared deviation time series given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, \textcolor{blue}{say $\vD_{kl} = \left\lbrace D_{kl}^{(1)},\ldots,D_{kl}^{(n)} \right\rbrace$}, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
+For each local squared deviation time series given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, \textcolor{blue}{say $\vD_{kl} = \left\lbrace D_{kl}^{(1)},\ldots,D_{kl}^{(n)} \right\rbrace$}, consider the absolute value of its Pearson correlation with the overall squared deviations, given by $\vd$:
 \begin{equation*}\label{def_Rkld}
@@
 \end{equation}
-where $\tau \in \mathbbm{R}_{+}^{*}$ is a convenient threshold value \textcolor{blue}{defined as function of $n$}.
-Under \textcolor{red}{some regularity conditions}, the probability $P(\mathcal{M}_{\tau}\supset\mathcal{M}^{\star}) \rightarrow 1$ as $n\rightarrow\infty$ \cite{fan2020statistical}.
+where $\tau \in \mathbbm{R}_{+}^{*}$ is a convenient threshold value.
+Under \textcolor{green}{some regularity conditions \cite{fan2020statistical}}, the probability $P(\mathcal{M}_{\tau}\supset\mathcal{M}^{\star}) \rightarrow 1$ as $n\rightarrow\infty$.
 In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ when the number of instants $n$ is large enough.
-
-%\section{Methodology}\label{section_method}
-%
-%Let $\mathcal{I}^{(1)},\ldots,\mathcal{I}^{(n)}$ be a set of matrices representing the images of some region of interest and $n$ instants in time.
-These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate.
-Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
-We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}^{(m)}$.
-Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}.
-After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smooth version of $\mathcal{I}^{(m)}$ with the same dimension.
-The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX^{(m)}$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}^{(m)}$.
-Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX^{(m)}$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.
-%
-%We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX^{(m)}\right\}$ as
-%\begin{equation} 
-%\sum_{m=1}^n\|\vX^{(m)}\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX^{(m)}-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
-%\end{equation}
-%where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}^{(m)}$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX^{(m)}$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.
-The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX^{(m)}$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time.
-Since each element (pixel) in $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space.
-Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further.
-Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX^{(m)}-\vX^{(m-1)}\|^2$, $d(\vX^{(m)},\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
-%
-%Let $X_{k,l}^{(m)}$ and $\bar{\mathcal{I}}_{k,l}$ be the entry $(k,l)$ of $\vX^{(m)}$ and $\bar{\mathcal{I}}$, respectively.
-Computing $D_{k,l}^{(m)}=(X_{k,l}^{(m)}-\bar{\mathcal{I}}_{k,l})^2$, we denote the matrix of squared mean differences as $\vD^{(m)} = [D_{k,l}^{(m)}]$.
-We then
-%analyze the time series given by
-%\begin{equation} 
-%\vd(m)= \sum_{k,l}D_{k,l}^{(m)} = \sum_{k,l}(X_{k,l}^{(m)} - \bar{\mathcal{I}}_{k,l})^2,
-%\label{E:eq_defdm}
-%\end{equation}
-%$m=1,\ldots,n$, which measure the temporal overall variation with respect to $\bar{\mathcal{I}}$.
-%
-%The time points with highest values of $\vd(m)$ represent the images for which the most expressive changes take place, where changes here are measured through $\mathbb{L}_2$ energy.
-Define the  $n\times p$ matrix
-%\begin{equation*}
-% \vD=\left(
-% \begin{array}{c}
-% vec(\vD^{(1)})^T\\
-% \vdots\\
-% vec(\vD^{(n)})^T\\
-% \end{array}
-% \right),
-%\label{eq_defmatrixD}
-%\end{equation*}
-%where $vec(\vD^{(m)})$ is the $p\times 1$ vector of wavelet coefficients for time $m$ and $p=\#\{k,l\}$ is total number of locations represented by $\vX^{(m)}$, $m=1,\ldots,n$.
-Sparsity \cite{johnstone2009statistical} on wavelet coefficients plays a special role here.
-We suppose a handful of coefficients drive the changes given by $\vd$, so that the effective dimension of $\vD$ (number of locations where relevant changes occur), say $s_d$, is such that $s_d<<p$.
-This can be represented as the following linear model
-%\begin{equation} 
-%\vd=\vD\vbeta+\vxi
-%\label{E:sparsemodel_d}
-%\end{equation}
-%where $\vbeta$ is sparse, i.e., it has $p-s_d$ null elements, and $\vxi^{(d)}$ is some $n\times 1$ random vector of errors.
-%
-%In order to identify spatio-temporal changes, we employ the idea of ultra-high dimensional correlation screening \cite{fan2020statistical} as follows.
-For each local squared mean deviation time series, given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, consider the absolute value of its Pearson correlation with the overall squared mean deviations, given by $\vd$:
-%\begin{equation*}
-%R_{k,l}= |\corr\left( \vD_{k,l}, \vd\right)|,
-%\label{def_Rkld}
-%\end{equation*}
-%where $\vD_{k,l}=(D_{k,l}(1),\ldots,D_{k,l}(n))^T$ is the time series of squared mean deviations of wavelet coefficients for the two-dimensional index $\{k,l\}$.
-%
-%
-%\begin{figure}[htb!]
-%\centering
-%\includegraphics[scale=.8]{../../drawio/diagram_wecs.drawio_11nov21}
-%\caption{Diagram of steps performed on WECS to compute an absolute correlation of a single point.}
-%\label{F:EllipsoidChanges}
-%\end{figure}
-%
-%
-%We have a matrix $\vR=[R_{k,l}]$ of correlations of ultra-high dimension. Define the set of 
-%{\it important} indices for changes in images with respect to $\bar{\mathcal I}$ as 
-%$\mathcal{M}^{*}=\{(k,l): \textit{changes in } \{\mathcal{I}^{(m)}\}_{m=1}^m \textit{ with respect to}$ $\bar{\mathcal{I}}$ \textit{are affected by changes in location corresponding to index} $(k,l)$ $\textit{in such images}\}$.
-This set coincides with the non-zero vectorized one-dimensional indices for the sparse representation of $\vbeta$ in (\ref{E:sparsemodel_d}).
-%We build the empirical set of indices flagged as corresponding to change locations by
-%\begin{equation}
-%\mathcal{M}_{\tau}=\{(k,l):|R_{k,l}|>\tau\},
-%\label{E:def_Mtaud}
-%\end{equation}
-%where $\tau_d>0$ is a convenient threshold value, function of $n$ and $J$. Under some regularity conditions, the probability that $\mathcal{M}_{\tau}$ contains its target set satisfies
-%\[
-%P(\mathcal{M}_{\tau}\supset\mathcal{M}^{*})\rightarrow 1,
-%\]
-%as $n\rightarrow\infty$ \cite{fan2020statistical}, where $P$ denotes a probability measure.
-In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{*}$ when the number of observations $n$ is large.
-%
-%Further geometrical motivation for our proposal is given as follows.
-As defined by (\ref{E:eq_defdm}), we expect $\vd$ to be a vector with some few high values, say $s_d$, and $n-s_d$ smaller values.
-This segregates the multi-temporal images, since the former time points identify the images in which significant changes occur, while the latter indices identify time points with no major changes.
-Consider $U>L>0$ such that the $s_d$ highest values of $\vd$ are larger than $U$, and the $n-s_d$ smallest values of $\vd$ are smaller than $L$.
-We also take $\delta=U-L$.
-The indices defined by  (\ref{E:def_Mtaud}) are such that
-%\[
-%\frac{<\vD_{k,l},\vd>}{\|\vD_{k,l}\|_2\|\vd\|_2}>\tau,
-%\]
-%i.e., such that  $\sum_{m=1}^nD_{k,l}^{(m)}\vd(m)>\tau\|\vD_{k,l}\|_2\|\vd\|_2$. This can be rewritten as
-%\[
-%\left|\sum_{m:\vd(m)>U} D_{k,l}^{(m)}\right|-\left|\sum_{m:\vd(m)<L} D_{k,l}^{(m)}\right|>\Delta,
-%\]
-%for some arbitrary $\Delta>>0$ (which can be a function of $n$ and $J$).
-Thence, when we employ correlation screening we select the two-dimensional wavelet indices which have the closest empirical directions to the vector of image temporal changes.
-Thus we are performing a truly spatio-temporal change detection in a single procedure.
-%
 

### 2021-12-08 tex/paper.wecs/wecs_tgrs.tex
updates in text

@@
 A special attention has been given for multi-temporal change detection using Synthetic Aperture Radar (SAR) images.
-\textcolor{green}{Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail} }.
+\textcolor{magenta}{Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail} }.
 Known to be not affected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
@@
 
-\textcolor{green}{
+\textcolor{magenta}{
 Using wavelet smoothing on the images consists in an initial step to analyze the data, but the main goal in the analysis is to find changes in space and at instants of time.
-The smoothed images still contain a large number of coefficients that need to be evaluated, what characterizes a high-dimensional problem with multiple time series corresponding to each location in space.
-Since most part of a region is usually not relevant for overall changes, we need to employ a method that retains the most important locations driving these modifications across time.
-Methods like variable selection are largely employed in statistics to identify the most important covariates in a study, and one method particularly efficient in identifying relevant variables is the feature screening technique.
+The smoothed images still contain a large number of coefficients that need to be evaluated simultaneously, what characterizes a high-dimensional problem with multiple time series corresponding to each location in space.
+We need to employ a method that retains the most important locations driving overall modifications across time.
+This type of problem is recurrent in statistics, and particularly in regression analysis it is mainly recognized as a variable selection problem.
+Methods like variable selection are largely employed to identify the most important covariates in a study, and one method particularly efficient in identifying relevant variables when the number of candidates is very large is the feature screening technique.
 }
@@
 where $\tau \in \mathbbm{R}_{+}^{*}$ is a convenient threshold value.
-Under \textcolor{green}{some regularity conditions \cite{fan2020statistical}}, the probability $P(\mathcal{M}_{\tau}\supset\mathcal{M}^{\star}) \rightarrow 1$ as $n\rightarrow\infty$.
-In other words, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ when the number of instants $n$ is large enough.
+\textcolor{magenta}{The idea is that for suitable values of $\tau$, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ \cite{fan2020statistical}.
+This screening idea has been successfully applied in different problems in areas such as graphical models \cite{zheng2020uniform}, time series \cite{yousuf2021targeting} and remote sensing \cite{li2021mapping}.
+We show it can also shed light on problems related to change point detection in time series of images.}
 
-\section{Experiments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}\label{secExperiments}
 
-In this section we shall evaluate the performance of WECS on two types of data sets.
-Initially, we generate simulated data where the level of noise and true change regions are known.
-For these images, we analyze different aspects of concerning the application of WECS and compare its performance with a standard change detection method.
-In the following application, real satellite data is employed, where the performance of WECS is evaluated on predefined change regions.
+%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment design}\label{secExpDesign}
 
+In order to assess the proposed method, this section presents two distinct studies comprising distinct datasets.
+
+The first study (Sec.~\ref{secExpSimulated}) uses a simulated data set and focus on identify the most appropriated wavelet family and resolution level $J$.
+Specifically, the Haar (haar), Daubechies of order 2 and 4 (db2 and db4), Coiflets of order 4 (coif4) and Symlets of order 2 and 4 (sym2 and sym4) wavelet families are analyzed \cite{BeylkinEA1991,Daubechies1992}.
+Posterior, once selected the wavelet family, the most suitable resolution level is then pointed out.
+
+Lastly, the performance of the proposed method is measured in terms of True/False Positive Ratios and Receiver Operating Characteristic (ROC) curve. 
+Comparisons with the standard Thresholding of Aggregate Absolute Difference (TAAD) and a non-wavelet version of WECS, herein called Energy Correlation Screening (ECS), are included in the experiments.
+
+In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $\vA$ where $A_{kl} = \sum_{m=2}^{n}|\mathcal{I}_{kl}^{(m)}-\mathcal{I}_{kl}^{(m-1)}|$.
+The ECS stands for the use of $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ when defining $D_{kl}^{(m)}$ at Equation~\ref{E:eq_defdm}.
+
+%
+%\textcolor{red}{
+%Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}^{(m)} - \mathcal{I}_{k,l}^{(m-1)}|$.
+Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ in the WECS formulation.
+%}
+
+%%%
+
+The second study (Sec.~\ref{secExpActual}) analyzes the performance of WECS, and respective comparison with TAAD and ECS, in a real-world application with actual SAR image series.
+The appropriate wavelet family and resolution level previously identified are employed.
+ROC curves, F1-Score \cite{Rijsbergen1979}, the kappa coefficient, and the variance of kappa \cite{Congalton2019}, are adopted for this purpose.
+Additionally, the computational run-times are presented and discussed.
+
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Simulated data analysis}\label{secExpSimulated}
 
-In this section we apply the change detection methods above on simulated data of multi-temporal images.
-The simulated multi-temporal images ($n=80$) is obtained by repeating a sequence of four images with different types of changes plus a noise.
-Examples of the first four of these images are shown in Figure \ref{F:EllipsoidChanges}.
-These images are generated from the sum of two matrices: a signal matrix with ones on entries where ellipses occur and zero elsewhere, and a noise matrix with random variables following a standard Gaussian distribution.
+This section applies the WECS, TAAD, and ECS methods to detect changes on a simulated multi-temporal image dataset. 
+Such series comprises 80 multi-temporal images and it is synthesized by repeating a sequence of four images with different types of changes plus a noise. 
+Examples of the first four instants are shown in Figure~\ref{F:EllipsoidChanges}. 
+These images/instants are generated by summing two matrices: (i) a binary signal matrix with ones denoting where the ellipses occur and zero elsewhere; (ii) and a noise matrix with random variables following a standard Gaussian distribution.
+
 The first image, $\mathcal{I}^{(1)}$, presents three elongated ellipses.
@@
 Smaller ellipses are then added to form $\mathcal{I}^{(3)}$ and $\mathcal{I}^{(4)}$.
-All the changes made that occur among subsequent images can be seen in Figure \ref{F:Changes_methods_images}(a), which displays a matrix of zeros and ones that correspond, respectively, to locations without and with changes along time.
+All the changes made that occur among subsequent images are shown in Figure~\ref{figResSim1}, where white regions (i.e., the ``ones'') correspond to changes along with the instants.
+
 Applying WECS to these images, we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation.
-An example of $\vR$ is presented in Figure \ref{F:Changes_methods_images}(b).
-For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a matrix of zeros and ones that can be compared with the total true changes displayed in Figure \ref{F:Changes_methods_images}(a).
+An example of $\vR$ is presented in Figure~\ref{figResSim2}.
+For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a binary matrix that can be compared with the total true changes shown in Figure \ref{figResSim1}.
+Similarly, Figures~\ref{figResSim3} and \ref{figResSim4} depicts the matrices $\vA$ and $\vR$ provided by TAAD and ECS methods.
 
-\begin{figure}[htb!]
+%-----------------------
+\begin{figure}[hbt]
 \centering
-%\includegraphics[width=40pc]{EllipsoidChanges/ellipses_all}
-\includegraphics[scale=.11]{../../figs/ellipses_t1}
-\includegraphics[scale=.11]{../../figs/ellipses_t2}
-\includegraphics[scale=.11]{../../figs/ellipses_t3}
-\includegraphics[scale=.11]{../../figs/ellipses_t4}
+
+\mbox{
+\subfigure[$\mathcal{I}^{(1)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t1_adjust.png}\label{figSim1}} 
+\subfigure[$\mathcal{I}^{(2)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t2_adjust.png}\label{figSim2}}
+}
+\mbox{
+\subfigure[$\mathcal{I}^{(3)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t3_adjust.png}\label{figSim3}} 
+\subfigure[$\mathcal{I}^{(4)}$]{\includegraphics[width=0.23\textwidth]{../../figs/ellipses_t4_adjust.png}\label{figSim4}}
+}
+
 \caption{Example of the first four simulated multi-temporal images. Features and changes come as white ellipses and dots.}
@@
 
-\begin{figure}[htp!]
-\includegraphics[scale=.1]{../../figs/total_changes}(a)
-\includegraphics[scale=.1]{../../figs/corr_changes_dm}(b)
-\includegraphics[scale=.1]{../../figs/corr_changes_logratios}(c)
-\includegraphics[scale=.1]{../../figs/corr_changes_dm_nowavelets}(d)
-\caption{Simulated images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[Total changes]{\includegraphics[frame,width=0.23\textwidth]{../../figs/total_changes_adjust.png}\label{figResSim1}} 
+\subfigure[WECS -- db2, $J=2$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_adjust.png}\label{figResSim2}} 
+}
+\mbox{
+\subfigure[TAAD -- $\vA$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_logratios_adjust.png}\label{figResSim3}} 
+\subfigure[ECS]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_nowavelets_adjust.png}\label{figResSim4}} 
+}
+
+%\caption{Simulated images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
+\caption{Expected change/non-change regions and typical results provided bu WECS, TAAD and ECS before the thresholding process.}
 \label{F:Changes_methods_images}
@@
 
-Figure \ref{F:Changes_methods_images} illustrates a comparison of different approaches to detect accumulated changes.
-Panel (b) shows the result of WECS using Daubechies wavelet with two null moments (db2) and $J=2$.
-Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}^{(m)} - \mathcal{I}_{k,l}^{(m-1)}|$.
-Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ in the WECS formulation.
+According to the experiment design (Section~\ref{secExpDesign}), ROC curves are employed to compare the performance of each analyzed method and inspect the effects of both $J$ and wavelet basis on the performance of the WECS method.
+In summary, the ROC curves exhibit the performance as a function of true and false positive ratios when distinguishing two events (in this context, the change and non-changed pixels/positions) using distinct thresholds.
+The true and false positive ratios are computed using the ``total change'' image (Fig.~\ref{figResSim1}) as reference.
+Moreover, the tested thresholds embrace all values in $\vR$ (or $\vA$), excluding repetitions and considering it in ascending order.
 
-We compute ROC curves to compare the detection performance of different methods and to verify the influence of some tuning parameters of wavelet smoothing: the resolution level $J$ and the choice of wavelet basis.
-Each detection method generates a matrix of change detection measures (correlations in the case of WECS).
-The ROC curves present the performance of change detection by applying a threshold on these measures, in the following way:
-\begin{enumerate}
-\item Let $R$ be the matrix of change measures. Compute the range $[r_{\min},r_{\max}]$ of the values in $R$;
-\item Let $(r_{(1)},\ldots,r_{(100)})$ be equally space values between $r_{\min}$ and $r_{\max}$;
-\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$ coincide with the element $(i,j)$ where a change really occurs on the image of total changes.
-Dividing this number by the total number of changes gives the true positive rate.
-\item For each $k=1,\ldots,n$, check how many entries are there such that $R_{i,j}>r_{(k)}$  do not coincide with the element $(i,j)$ where a change really occurs.
-Dividing this number by the total number of entries where changes do not occur gives the false positive rate.
-\item The ROC curve is the plot of true and false positive rates corresponding to each $k$.
-\end{enumerate}
+%-----------------------
+\begin{figure}[hbt]
+\centering
 
-\begin{figure}[htp!]
-\includegraphics[scale=.13]{../../figs/methods_comparison}\hspace{-.6cm}(a) 
-\includegraphics[scale=.13]{../../figs/levels_comparison}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{../../figs/families_comparison}\hspace{-.5cm}(c)
-\includegraphics[scale=.13]{../../figs/dm_comparison_wavelet_deepL}\hspace{-.5cm}(d) \\
-\caption{ROC curves for detection of changing ellipses in simulated images and different methods.
-(a) The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue).
-(b) db2 $\vd(m)$ with different levels.
-(c) $\vd(m)$ with different wavelet bases and J=2;
-(d) The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line).}
-\label{F:EllipsoidChanges_details}
+\mbox{
+\subfigure[$\vd(m)$ with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSimBasis}}
+\subfigure[db2 $\vd(m)$ with different levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSimJ}}
+}
+
+\mbox{
+\subfigure[The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue)]{\includegraphics[width=0.23\textwidth]{../../graphs/methodROC.pdf}\label{figROCSimRes}}
+%\subfigure[The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line)]{\includegraphics[width=0.23\textwidth]{../../graphs/comparisonROC.pdf}\label{figROCSim4}}
+}
+
+\caption{ROC curves for detection of changing ellipses in simulated images and different methods.}\label{F:EllipsoidChanges_details}
 \end{figure}
 
-Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the simulated data as follows.
-The effects of wavelet bases, level of decomposition, deep-learning feature extraction are shown on the ROC curves.
-We employ the following wavelet bases: Haar;  Daubechies db2; Daubechies db4; Coiflets coif4; Symlets sym2 ; and Symlets sym4.
-Panel (c) presents the ROC curves for the proposed method under the aforementioned bases.
-On all instances $J=2$ is employed.
-Comparing the ROC curves of all options, we can notice that Daubechies db2 and Symlets sym2 are the best choices.
-Panel (b) presents the ROC curves for five different levels of decomposition $J=\{1,2,3,4,5\}$ under the Daubechies db2 basis.
-Levels $J=2,3$ have a clear better performance, with a slight advantage to $J=2$, since it is uses less levels on the decomposition.
-The overall performance  of $J=2$ warrants its use for the rest of the comparisons.
-Panel (d) shows how the proposed method performs in comparison to a deep learning feature extraction from a residual learning network \cite{zhang2017beyond}.
-WECS is applied with db2 wavelets and $J=2$, whereas the other method applies WECS's idea, the only difference being that $\vX^{(m)}$ is replaced by another smooth version of $\mathcal{I}^{(m)}$ that employs neural-networks trained to compete with wavelet smoothing.
-We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS.
-We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
-The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}^{(m)}\}$ instead of $\{\vX^{(m)}\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}^{(m)}\}$.
-The ROC curves in Panel (a) show that the proposed WECS outperforms both the other two methods.
+%basis wavelet
+Figure~\ref{figROCSimBasis} presents the ROC curves considering distinct wavelet basis in the WECS method.
+All instances adopts $J=2$.
+The curves allow concluding that db2 and sym2 deliver the best trade-off between true and false positive ratios (i.e., high true positive ratios even when the false positive ratios are low).
+As a consequence of this finding, all the following experiments and analyses consider the db2 wavelet basis.
+
+%J-level
+Figure~\ref{figROCSimJ} depicts the ROC curves with $J \in \{1,2,3,4,5\}$ as decomposition levels.
+The profiles exhibited by the ROC curves make evident that $J$ equal to 2 or 3 promotes the best performances, with a slight advantage for $J=2$, since it demands few decomposition levels.
+The overall performance for $J=2$ warrants its use for the rest of the comparisons.
+
+%Compara figA
+At last, the ROC curves shown in Figure~\ref{figROCSimRes} allow comparing the performances of WECS, TAAD and ECS approaches. 
+Firstly, the low-performance assigned to the ECS method allows concluding that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet smoothing in the context of the proposed method.
+About the performance of TAAD method, a true positive ratio of 0.8 is guaranteed when tolerating a false positive ratio of approximately 0.4. 
+Reversely, the WECS method provides the same true positive ratio under an almost nill false positive ratio, demonstrating then its superiority compared to the competitors.
 

### 2021-12-08 tex/paper.wecs/wecs_tgrs__trans06dez21.tex
atualizacao (08dez21)

--- 2021-12-01 tex/paper.wecs/wecs_tgrs_trans26nov21.tex
@@
 \usepackage[export]{adjustbox}
+
+\usepackage{float}
 
@@
 
-\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Samples.pdf}\label{figImgSentinelROIs}}
+\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.3\textwidth]{../../qgis/maps/SamplesRev.pdf}\label{figImgSentinelROIs}}
 
@@
 %-----------------------
-\begin{figure}[htp!]
+\begin{figure}[htb!]
 \centering
-\includegraphics[scale=.2]{../../figs/forest_vSumDifCoefSq}\hspace{-.5cm}
-\caption{Plot of $\vd(m)$, $m=1,\ldots,85$ with orange horizontal line indicating two times its median absolute deviation.}
+\includegraphics[width=0.485\textwidth]{../../graphs/LineInstants_1-30-59.pdf}
+\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$. The three larger deviations occurs at $m$ equivalent to $25$, $27$ and $30$.
+%with orange horizontal line indicating two times its median absolute deviation.
+}
 \label{F:forest_wecs}
@@
 \mbox{
-\subfigure[$m=25$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_m25.pdf}\label{figImgSentinel2015}} 
-\subfigure[$m=27$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_m27.pdf}\label{figImgSentinel2017}}
+\subfigure[$m=1$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_1.pdf}\label{figImgSentinel2015}} 
+\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_30.pdf}\label{figImgSentinel2017}}
 }
@@
 \mbox{
-\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_m30.pdf}\label{figImgSentinel2015}} 
-\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/forestChange_mean.pdf}\label{figImgSentinel2017}}
+\subfigure[$m=59$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_59.pdf}\label{figImgSentinel2015}} 
+\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_mean.pdf}\label{figImgSentinel2017}}
 }
@@
 
+%%-----------------------
+%\begin{figure}[hbt]
+%\centering
+%
+%\mbox{
+%\subfigure[Ground truth samples]{\includegraphics[width=0.23\textwidth]{../../figs/forest_change}\label{figResForestWecs}} 
+%\subfigure[ROC]{\includegraphics[width=0.23\textwidth]{../../figs/forest_roc_change}\label{figResForestAggreg}} 
+%}
+%
+%\caption{
+%Change points in space of the forest data. 
+%(a) Regions (in white) where changes should be detected.
+%(b) ROC curve for detection of changing regions. 
+%}\label{F:forest_roc_wecs_agg___TEMP}
+%\end{figure}
+
+%%-----------------------
+%\begin{figure}[hbt]
+%\centering
+%
+%\mbox{
+%\subfigure[WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_abscorr}\label{figResForestWecs}} 
+%\subfigure[Log.Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_ratios}\label{figResForestAggreg}} 
+%}
+%
+%\mbox{
+%\subfigure[Change map -- WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_change_space}\label{figResForestWecs}} 
+%\subfigure[Change map -- Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_change_space}\label{figResForestAggreg}} 
+%}
+%
+%\caption{
+%Change points in space of the forest data. 
+%%(a) Regions (in white) where changes should be detected.
+%%(b) ROC curve for detection of changing regions. 
+%(a) Matrix of absolute correlations obtained with WECS.
+%(b) Change measures from aggregation of absolute differences. 
+%(c) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension). 
+%(d) Change regions detected using aggregated absolute differences and Otsu's thresholding.}\label{F:forest_roc_wecs_agg}
+%\end{figure}
+
 %-----------------------
-\begin{figure}[hbt]
+\begin{figure}[h!]
 \centering
@@
 \mbox{
-\subfigure[Ground truth samples]{\includegraphics[width=0.23\textwidth]{../../figs/forest_change}\label{figResForestWecs}} 
-\subfigure[ROC]{\includegraphics[width=0.23\textwidth]{../../figs/forest_roc_change}\label{figResForestAggreg}} 
+\subfigure[WECS-KI]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/WECS-KI.pdf}\label{figResChange1}} 
+\subfigure[WECS-OT]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/WECS-OT.pdf}\label{figResChange2}}
+}
+
+\mbox{
+\subfigure[ECS-OT]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/ECS-OT.pdf}\label{figResChange3}} 
+\subfigure[TAAD-KI]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/TAAD-KI.pdf}\label{figResChange4}}
 }
@@
 \caption{
-Change points in space of the forest data. 
-(a) Regions (in white) where changes should be detected.
-(b) ROC curve for detection of changing regions. 
-}\label{F:forest_roc_wecs_agg___TEMP}
-\end{figure}
-
-%-----------------------
-\begin{figure}[hbt]
-\centering
-
-\mbox{
-\subfigure[WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_abscorr}\label{figResForestWecs}} 
-\subfigure[Log.Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_ratios}\label{figResForestAggreg}} 
-}
-
-\mbox{
-\subfigure[Change map -- WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_change_space}\label{figResForestWecs}} 
-\subfigure[Change map -- Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_change_space}\label{figResForestAggreg}} 
-}
-
-\caption{
-Change points in space of the forest data. 
+\textcolor{red}{Change points in space of the forest data. 
 %(a) Regions (in white) where changes should be detected.
@@
 (c) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension). 
-(d) Change regions detected using aggregated absolute differences and Otsu's thresholding.}\label{F:forest_roc_wecs_agg}
+(d) Change regions detected using aggregated absolute differences and Otsu's thresholding.}}
+%\label{F:forest_roc_wecs_agg}
+\label{figSentinelResults}
 \end{figure}
 
-\begin{table}
-\caption{Accuracy measures $F_1$-score, Precision ($\mathrm{Pr}$) and Recall ($\mathrm{Re}$) computed from detection of changing regions in the forest data using WECS (taking the $[N/\log N]$ largest correlations as corresponding to change) and the aggregation of absolute differences with Otsu and Kittler-Illingworth's (KI) thresholds.}
+%\begin{table}
+%\caption{Accuracy measures $F_1$-score, Precision ($\mathrm{Pr}$) and Recall ($\mathrm{Re}$) computed from detection of changing regions in the forest data using WECS (taking the $[N/\log N]$ largest correlations as corresponding to change) and the aggregation of absolute differences with Otsu and Kittler-Illingworth's (KI) thresholds.}
+%\centering
+%\begin{tabular}{c|ccc}
+%\hline
+%  & Aggregated - Otsu & Aggregated - KI & WECS \\
+%\hline
+%$F_1$-score & 0.2231 & 0.2163 & 0.3253 \\
+%$\mathrm{Pr}$ & 0.1369 & 0.1296 & 0.2390 \\
+%$\mathrm{Re}$ & 0.6022 & 0.6526 & 0.5094 \\
+%\hline
+%\end{tabular}
+%\label{T:forest_accuracy}
+%\end{table}
+
+\begin{table}[H]
+\caption{Inserir...  variances are multiplied by $\times 10^{-5}$}\label{tabAccExpSentinel}
 \centering
-\begin{tabular}{c|ccc}
+
+\begin{tabular}{ccccccc}
+\thickhline
+ & \multicolumn{2}{c}{\textbf{WECS}} & \multicolumn{2}{c}{\textbf{ECS}} & \multicolumn{2}{c}{\textbf{TAAD}}\tabularnewline
 \hline 
-  & Aggregated - Otsu & Aggregated - KI & WECS \\
-\hline
-$F_1$-score & 0.2231 & 0.2163 & 0.3253 \\
-$\mathrm{Pr}$ & 0.1369 & 0.1296 & 0.2390 \\
-$\mathrm{Re}$ & 0.6022 & 0.6526 & 0.5094 \\
-\hline
+Threshold & OT & KI & OT & KI & OT & KI\tabularnewline
+F1-Score & 0.940 & 0.876 & 0.6189 & 0.661 & 0.798 & 0.778\tabularnewline
+kappa & 0.818 & 0.698 & 0.535 & 0.513 & 0.244 & 0.284\tabularnewline
+var. of kappa & 6.90 & 3.43 & 1.91 & 1.72 & 0.76 & 0.91 \tabularnewline
+\thickhline
 \end{tabular}
-\label{T:forest_accuracy}
+
 \end{table}
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+\includegraphics[width=0.485\textwidth]{../../graphs/TPTNs.pdf}
+\caption{Percentages of True/False Positives (TP/FP) and True/False Negatives (TN/FN) relative to change (Positive) and non-change (Negative) reference samples.}\label{figAE}
+\end{figure}
 

### 2021-12-08 tex/paper.wecs/wecs_tgrs__trans07dez21.tex
atualizacao (08dez21)

--- 2021-12-08 tex/paper.wecs/wecs_tgrs__trans06dez21.tex
@@
 \makeatother
+
+\usepackage{scalerel}
+\usepackage{tikz}
+\usetikzlibrary{svg.path}
+
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+\tikzset{orcidlogo/.pic={
+		\fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
+		\fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}	svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
+		svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
+}}
+
+\newcommand\orcidicon[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
+				\begin{tikzpicture}[yscale=-1,transform shape]
+				\pic{orcidlogo};
+				\end{tikzpicture}
+			}{|}}}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\author{Rodney~V.~Fonseca, Rog\'{e}rio~G.~Negri,
+\author{Rodney~V.~Fonseca, 
+		Rog\'{e}rio~G.~Negri \orcidicon{0000-0002-4808-2362},
         Alu\'{i}sio~Pinheiro,
@@
 
-The second study (Sec.~\ref{secExpActual}) analyzes the performance of WECS, and respective comparison with TAAD and ECS, in a real-world application with actual SAR image series.
+The second study (Sec.~\ref{secExpActual}) analyzes the performance of WECS, and respective comparison with ECS and TAAD, in a real-world application with actual SAR image series.
 The appropriate wavelet family and resolution level previously identified are employed.
-ROC curves, F1-Score \cite{Rijsbergen1979}, the kappa coefficient, and the variance of kappa \cite{Congalton2019}, are adopted for this purpose.
+F1-Score \cite{Rijsbergen1979}, the kappa coefficient, and the variance of kappa \cite{Congalton2019}, are adopted for this purpose.
 Additionally, the computational run-times are presented and discussed.
@@
 \mbox{
-\subfigure[The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue)]{\includegraphics[width=0.23\textwidth]{../../graphs/methodROC.pdf}\label{figROCSimRes}}
-%\subfigure[The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line)]{\includegraphics[width=0.23\textwidth]{../../graphs/comparisonROC.pdf}\label{figROCSim4}}
+\subfigure[The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue)]{\includegraphics[width=0.46\textwidth]{../../graphs/methodROC_ext.pdf}\label{figROCSimRes}}
 }
@@
 
-%Figure \ref{F:EllipsoidChanges_details} presents the different ROC curves for change detection methods applied to the simulated data as follows.
-The effects of wavelet bases, level of decomposition, deep-learning feature extraction are shown on the ROC curves.
-We employ the following wavelet bases: Haar;  Daubechies db2; Daubechies db4; Coiflets coif4; Symlets sym2 ; and Symlets sym4.
-Panel (c) presents the ROC curves for the proposed method under the aforementioned bases.
-On all instances $J=2$ is employed.
-Comparing the ROC curves of all options, we can notice that Daubechies db2 and Symlets sym2 are the best choices.
-Panel (b) presents the ROC curves for five different levels of decomposition $J=\{1,2,3,4,5\}$ under the Daubechies db2 basis.
-Levels $J=2,3$ have a clear better performance, with a slight advantage to $J=2$, since it is uses less levels on the decomposition.
-The overall performance  of $J=2$ warrants its use for the rest of the comparisons.
-Panel (d) shows how the proposed method performs in comparison to a deep learning feature extraction from a residual learning network \cite{zhang2017beyond}.
-WECS is applied with db2 wavelets and $J=2$, whereas the other method applies WECS's idea, the only difference being that $\vX^{(m)}$ is replaced by another smooth version of $\mathcal{I}^{(m)}$ that employs neural-networks trained to compete with wavelet smoothing.
-We can see that the ROC curves for images treated with a deep-learning method has a performance worse than that obtained with WECS.
-%We finally have in Panel (a) the proposed WECS with db2 wavelets and $J=2$ compared to two other non-wavelet methods.
-The first involves computing $\vd(m)$ without wavelet smoothing, i.e., the squared deviations are computed using $\{\mathcal{I}^{(m)}\}$ instead of $\{\vX^{(m)}\}$, and the classic method of analyzing aggregated absolute differences of $\{\mathcal{I}^{(m)}\}$.
-The ROC curves in Panel (a) show that the proposed WECS outperforms both the other two methods.
-
 %%%%%%%%%%%%%%%%%%%%
@@
 
-In detail, this application considers a large multi-temporal series of 84 images acquired over a forest region at the border of Brazil and the French Guiana, from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
-Each image contains the signal backscatters relative to VV and VH polarizations, a spatial resolution of \SI{10}{\meter} and support of $1538 \times 1556$ pixels wide.
-In order to apply the analyzed methods, the dual-polarized images were combined into a single-band representation considering
-$\mathcal{I}_{kl}^{(m)} = \sqrt{(\mathrm{VV}_{kl}^{(m)})^2+(\mathrm{VH}_{kl}^{(m)})^2}$, with $\mathrm{VV}$ and $\mathrm{VH}$ representing the available polarizations.
+In detail, this application considers a large multi-temporal series of 84 images acquired over a forest region at the border of Brazil and the French Guiana (Fig.~\ref{figAE}), from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
 
-%We employed the proposed change detection method on a series of 84 multi-date satellite images.
-The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3.
-Each image has two channels (VV and VH) and 1538 by 1556 pixels.
-We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}^{(m)} = ((\mathrm{VV}_{k,l}^{(m)})^2+(\mathrm{VH}_{k,l}^{(m)})^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.
+Each image contains the amplitude signal backscatters relative to VV and VH polarizations, a spatial resolution of \SI{10}{\meter} and support of $1538 \times 1556$ pixels wide.
+Figures~\ref{figImgSentinel2015} and \ref{figImgSentinel2017} depicts the first and last images of the time series, where it s possible to compare and identify some landscape changes.
+
+After careful visual inspection of the backscatter profile of each image in the series, it was possible to identify the regions where the land cover change occurs or does not occur, allowing then collecting reference samples regarding the ``change'' and ``non-change'' conditions.
+Such samples are needed to compute the accuracy measures beforehand mentioned in Section~\ref{secExpDesign}. The spatial distribution of these samples is presented in Figure~\ref{figImgSentinelROIs}.
 
@@
 \centering
-\includegraphics[width=0.485\textwidth]{../../qgis/maps/StudyArea.pdf}
-\caption{Study area location. \textcolor{red}{(preliminar)}}\label{figAE}
+\includegraphics[width=0.485\textwidth]{../../qgis/maps/StudyArea_OSM.pdf}
+\caption{Study area location.}\label{figAE}
 \end{figure}
@@
 \mbox{
-\subfigure[26th December, 2015]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2015.pdf}\label{figImgSentinel2015}} 
-\subfigure[3rd December, 2017]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2017.pdf}\label{figImgSentinel2017}}
+\subfigure[December 26th 2015]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2015.pdf}\label{figImgSentinel2015}} 
+\subfigure[December 3rd 2017]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2017.pdf}\label{figImgSentinel2017}}
 }
 
-\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.3\textwidth]{../../qgis/maps/SamplesRev.pdf}\label{figImgSentinelROIs}}
+\subfigure[Reference change and non-change samples]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/SamplesRev.pdf}\label{figImgSentinelROIs}}
 
-\caption{\textcolor{red}{Inserir caption...}}\label{figImageRef}
+\caption{First and last images in the adopted multitemporal image series, in VV-HV-VV RGB color composition, and reference samples.}\label{figImageRef}
 \end{figure}
 
-\textcolor{magenta}{
-A multi-resolution analysis based on a Symlet basis with filter of length 16 (symlet 8) is built.
+In order to apply the analyzed methods, the dual-polarized images were combined into a single-band representation considering
+$\mathcal{I}_{kl}^{(m)} = \sqrt{(\mathrm{VV}_{kl}^{(m)})^2+(\mathrm{VH}_{kl}^{(m)})^2}$, with $\mathrm{VV}$ and $\mathrm{VH}$ representing the available polarizations.
+
+%%%%RGN: pode verificar como fica a adaptaćão para este trecho?
+%\textcolor{magenta}{
+%A multi-resolution analysis based on a Symlet basis with filter of length 16 (symlet 8) is built.
 In order to have dimensions as power of 2, the matrices $\mathcal{I}_{k,l}$ are extended to a $2048\times2048$ matrix with $\mathcal{I}_{k,l}$ at the center and the remaining parts being completed with mirrored values at the borders.
@@
 Then, we are able to compute the squared mean differences vector $\vd$ and the matrix of absolute correlations $\vR$.
-}
+%}
 
-The instantaneous change values expressed by $\vd$ are shown in Figure~\ref{F:forest_wecs}. 
-The top line represents the median absolute deviation of $\vd$ and allows highlighting the instants 25, 27, and 30; when expressive differences occur in comparison to the other instants.
-The images relative to these instants are in Figure~\ref{F:forest_change_times}.
+Initially, to express the instantaneous change values,
+Figure~\ref{F:forest_wecs} presents the temporal variations given by $\vd$.
+According to the presented profile, it is possible to observe the fluctuations around a central tendency near $12$ as highlight instants with a value similar, higher or lower than this tendency, as pointed out for $m$ equal to 1, 30 and 59, respectively.
+%
+For sake of comparison, the wavelet coefficient representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(10)}$ and $\vX^{(50)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figures~\ref{F:forest_wecs}.
+While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rises in $\vX^{(30)}$.
+Reversely, $\vX^{(59)}$ offers low contribution for identifying change regions.
+
+%The instantaneous change values expressed by $\vd$ are shown in Figure~\ref{F:forest_wecs}. 
+%The top line represents the median absolute deviation of $\vd$ and allows highlighting the instants 25, 27, and 30; when expressive differences occur in comparison to the other instants.
+%The images relative to these instants are in Figure~\ref{F:forest_change_times}.
 
@@
 \includegraphics[width=0.485\textwidth]{../../graphs/LineInstants_1-30-59.pdf}
-\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$. The three larger deviations occurs at $m$ equivalent to $25$, $27$ and $30$.
-%with orange horizontal line indicating two times its median absolute deviation.
-}
-\label{F:forest_wecs}
+\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$.
+Distinct deviation values occurs for $m$ at $1$ (intermediate), $30$ (high) and $59$ (low) in comparison to a central tendency around 12.}\label{F:forest_wecs}
 \end{figure}
@@
 
-The changes in space can be analyzed using the image obtained with $\vR$, which is displayed in Figure \ref{F:forest_wecs}c.
-Denoting the images' dimension as $N=1538\times1556$, if we take the $[N/\log N]$ largest absolute correlations as those corresponding to possible change points, we obtain a matrix of zeros and ones that is presented in Figure \ref{F:forest_wecs}e.
-The white regions in Figure \ref{F:forest_wecs}c (entries with value one) represent the change points, which seem to concentrate mainly on three regions: at the center, to the right of the river; at the top, on the left border of the river; and at the top left.
-Computing aggregated absolute differences to measure changes in space, we obtain Figure \ref{F:forest_wecs}d.
-An image of detected changes corresponding to measures in  Figure \ref{F:forest_wecs}d can be obtained applying a thresholding method for grayscale images.
-Figure \ref{F:forest_wecs}f shows the result of using Otsu's thresholding method \cite{otsu1979threshold} for aggregation of absolute differences.
+After submitting the image series to the methods WECS, ECS and TAAD, are obtained the matrices $\vR$, $\vR_{\star}$ and $\vA$ comprising measure about a tendency of changes in each position of the spatial domain.
+Such matrices are presented in Figure~\ref{figImageMethods}, where high values stand for regions with changes in their land cover over time.
+%
+However, to determine a cut-off value $\tau$ for such matrices, providing then binary maps $\mathcal{M}_{\tau}$ of ``change'' and ``non-change'' areas, the use of thresholding techniques rises as a convenient procedure.
+Among different alternatives in the literature, the Otsu (OT) [figImageMethods ref] and Kittler-Illingworth (KI) \cite{KittlerIllingworth1986} thresholding techniques had been successfully employed for change detection purposes \cite{JohnsonKasischke1998,Nielsen2007,WuEA2014,NegriEA2021}.
 
-The comparison of performance for the two change detection measures considered here can be checked in Figure \ref{F:forest_roc_wecs_agg}, where a ROC curve is computed to check the correct detection of changes.
-WECS attains high true positive rates before the aggregation method, but both methods design competitive results.
-This performance differs from the ROC curve results in the previous section because the real images have a mean signal to noise ratio (SNR) of 3.9327, whereas the simulated images were generated with SNR of 0.2309.
-Hence, WECS is expected to have performance closer to the aggregation method in less noisy applications.
-The reference of correct change regions is shown in Figure \ref{F:forest_roc_wecs_agg}a, which are determined using [XXXXX].
+The accuracy of obtained binary maps $\mathcal{M}_{\tau}$, resulting from the application of OT and KI algorithms on the change images provided by the WECS, ECS and TAAD methods, are measured in terms of F1-Score and kappa coefficient with basis on the reference ground-truth samples (Fig.~\ref{figImgSentinelROIs}).
+Table~\ref{tabAccExpSentinel} presents the computed accuracy measures.
+%Moreover, the percentages of True and False Positives and Negatives are shown in Figure~\ref{figTPTNs}.
 
-We can also compare the change detection methods using the $F_1$-score accuracy measure.
-Denoting $\mathrm{TP}$ as the number of true positives (change pixels correctly detected), $\mathrm{FP}$ as the number of false positives (nonchange pixels flagged as change point) and $\mathrm{FN}$ as the total of false negatives (change points flagged as nonchange point), the $F_1$-score is defined as
-\begin{equation*}
-F_1 = 2\frac{\mathrm{Pr}\cdot\mathrm{Re}}{\mathrm{Pr}+\mathrm{Re}}	,
-\end{equation*}
-where
-\begin{equation*}
-\mathrm{Pr}=\frac{\mathrm{TP}}{\mathrm{TP}+\mathrm{FP}}\quad 
-\text{and} \quad \mathrm{Re} = \frac{\mathrm{TP}}{\mathrm{TP}+\mathrm{FN}}.
-\end{equation*}
-Results of these three accuracy measures are presented in Table \ref{T:forest_accuracy}.
-Results for aggregation of absolutes differences correspond to two types of thresholds: Otsu's method and Kittler-Illingworth \cite{kittler1986minimum}.
-We can observe that WECS presents the highest value of $F_1$-score, which means it has a better performance that the competing method.
+%-----------------------
+\begin{figure}[hbt]
+\centering
 
-%%-----------------------
-%\begin{figure}[hbt]
-%\centering
-%
-%\mbox{
-%\subfigure[Ground truth samples]{\includegraphics[width=0.23\textwidth]{../../figs/forest_change}\label{figResForestWecs}} 
-%\subfigure[ROC]{\includegraphics[width=0.23\textwidth]{../../figs/forest_roc_change}\label{figResForestAggreg}} 
-%}
-%
-%\caption{
-%Change points in space of the forest data. 
-%(a) Regions (in white) where changes should be detected.
-%(b) ROC curve for detection of changing regions. 
-%}\label{F:forest_roc_wecs_agg___TEMP}
-%\end{figure}
+\mbox{
+\subfigure[$\vR$ -- WECS]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageWECS.pdf}\label{figImgSentinel2015}} 
+\subfigure[$\tilde{\vR}$ -- ECS]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageECS.pdf}\label{figImgSentinel2017}}
+}
 
-%%-----------------------
-%\begin{figure}[hbt]
-%\centering
-%
-%\mbox{
-%\subfigure[WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_abscorr}\label{figResForestWecs}} 
-%\subfigure[Log.Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_ratios}\label{figResForestAggreg}} 
-%}
-%
-%\mbox{
-%\subfigure[Change map -- WECS \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_wecs_change_space}\label{figResForestWecs}} 
-%\subfigure[Change map -- Aggreg. \textcolor{red}{(Temp. img)}]{\includegraphics[width=0.23\textwidth]{../../figs/forest_aggreg_change_space}\label{figResForestAggreg}} 
-%}
-%
-%\caption{
-%Change points in space of the forest data. 
-%%(a) Regions (in white) where changes should be detected.
-%%(b) ROC curve for detection of changing regions. 
-%(a) Matrix of absolute correlations obtained with WECS.
-%(b) Change measures from aggregation of absolute differences. 
-%(c) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension). 
-%(d) Change regions detected using aggregated absolute differences and Otsu's thresholding.}\label{F:forest_roc_wecs_agg}
-%\end{figure}
+\subfigure[$\vA$ -- TAAD]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageTAAD.pdf}\label{figImgSentinelROIs}}
+
+\caption{Change tendency matrices $\vR$, $\tilde{\vR}$ and $\vA$ assigned to the WECS, ECS and TAAD methods.}\label{figImageMethods}
+\end{figure}
+
+%%%RGN: ler com calma o trecho abaixo...
+According to the adopted accuracy measures, the WECS method, especially when submitted to the OT algorithm, presents a better performance detecting change locations without increasing/inflating the amount of FP, which is justified by a higher F1-Score level in comparison to the competitors ECS and TAAD.
+Moreover, according to the kappa coefficient, the WECS method also deliveries a more balanced correct classification regarding the change and non-change areas.
+Furthermore, based on the kappa values and respective variances, it is verified that the difference between any pair of results is significant at 1\%.
+
+Although the performance of WECS equipped with the KI algorithm is inferior to the OT algorithm, it is worth observing that the introduced methods still provide more accurate results than ECS and TAAD.
+Regarding the competitors, the TAAD method presents frequent FP errors, leading to lower F1-Score and kappa coefficient than the ECS method.
+
+Figure~\ref{figSentinelResults} depicts the most accurate change/non-change maps according to the measures in Table~\ref{tabAccExpSentinel}.
+It is possible to verify that while the TAAD-KI method assigns the water body as a ``region of change'' and does not detect locations that changed (northwest and southeast portions -- second and fourth quadrants).
+Regarding the ECS method, the provided result has a noise-corrupted aspect due to the frequent FP classification points.
+
+Lastly, as previously observed, the WECS method equipped with the KI algorithm provides a homogeneous mapping over the non-change areas (west portion -- second and third quadrants), accurate detection over the change regions, and low inclusion (FP) and exclusion (FN) error rates.
+In adopting the OT algorithm, the inclusion error increases, resulting in a less-regularized change/non-change map.
+
+The proportions of true/false positive/negative assigned to the analyzed methods are shown in Figure~\ref{figTPTNs} and corroborate the presented discussions.
+
+\begin{table}[H]
+\caption{Accuracy values summary. The variance of kappa are multiplied by $\times 10^{-5}$.}\label{tabAccExpSentinel}
+\centering
+\begin{tabular}{ccccccc}
+\thickhline
+ & \multicolumn{2}{c}{\textbf{WECS}} & \multicolumn{2}{c}{\textbf{ECS}} & \multicolumn{2}{c}{\textbf{TAAD}}\tabularnewline
+Threshold & OT & KI & OT & KI & OT & KI\tabularnewline
+\hline 
+F1-Score & 0.940 & 0.876 & 0.6189 & 0.661 & 0.798 & 0.778\tabularnewline
+Kappa & 0.818 & 0.698 & 0.535 & 0.513 & 0.244 & 0.284\tabularnewline
+Var. of kappa & 6.90 & 3.43 & 1.91 & 1.72 & 0.76 & 0.91 \tabularnewline
+\thickhline
+\end{tabular}
+\end{table}
 
@@
 
-\caption{
-\textcolor{red}{Change points in space of the forest data. 
-%(a) Regions (in white) where changes should be detected.
-%(b) ROC curve for detection of changing regions. 
-(a) Matrix of absolute correlations obtained with WECS.
-(b) Change measures from aggregation of absolute differences. 
-(c) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension). 
-(d) Change regions detected using aggregated absolute differences and Otsu's thresholding.}}
-%\label{F:forest_roc_wecs_agg}
-\label{figSentinelResults}
+\caption{Resulting maps from the analyzed methods.}\label{figSentinelResults}
 \end{figure}
-
-%\begin{table}
-%\caption{Accuracy measures $F_1$-score, Precision ($\mathrm{Pr}$) and Recall ($\mathrm{Re}$) computed from detection of changing regions in the forest data using WECS (taking the $[N/\log N]$ largest correlations as corresponding to change) and the aggregation of absolute differences with Otsu and Kittler-Illingworth's (KI) thresholds.}
-%\centering
-%\begin{tabular}{c|ccc}
-%\hline
-%  & Aggregated - Otsu & Aggregated - KI & WECS \\
-%\hline
-%$F_1$-score & 0.2231 & 0.2163 & 0.3253 \\
-%$\mathrm{Pr}$ & 0.1369 & 0.1296 & 0.2390 \\
-%$\mathrm{Re}$ & 0.6022 & 0.6526 & 0.5094 \\
-%\hline
-%\end{tabular}
-%\label{T:forest_accuracy}
-%\end{table}
-
-\begin{table}[H]
-\caption{Inserir...  variances are multiplied by $\times 10^{-5}$}\label{tabAccExpSentinel}
-\centering
-
-\begin{tabular}{ccccccc}
-\thickhline
- & \multicolumn{2}{c}{\textbf{WECS}} & \multicolumn{2}{c}{\textbf{ECS}} & \multicolumn{2}{c}{\textbf{TAAD}}\tabularnewline
-\hline 
-Threshold & OT & KI & OT & KI & OT & KI\tabularnewline
-F1-Score & 0.940 & 0.876 & 0.6189 & 0.661 & 0.798 & 0.778\tabularnewline
-kappa & 0.818 & 0.698 & 0.535 & 0.513 & 0.244 & 0.284\tabularnewline
-var. of kappa & 6.90 & 3.43 & 1.91 & 1.72 & 0.76 & 0.91 \tabularnewline
-\thickhline
-\end{tabular}
-
-\end{table}
 
@@
 \includegraphics[width=0.485\textwidth]{../../graphs/TPTNs.pdf}
-\caption{Percentages of True/False Positives (TP/FP) and True/False Negatives (TN/FN) relative to change (Positive) and non-change (Negative) reference samples.}\label{figAE}
+\caption{Percentages of True/False Positives (TP/FP) and True/False Negatives (TN/FN) relative to change (Positive) and non-change (Negative) reference samples.}\label{figTPTNs}
 \end{figure}

### 2021-12-09 tex/paper.wecs/wecs_tgrs.tex
updates no texto

@@
 
+\usepackage{float}
+
 \makeatletter
@@
 
-\author{Rodney~V.~Fonseca, Rog\'{e}rio~G.~Negri,
-        Alu\'{i}sio~Pinheiro,
-        and~Abdourrahmane~Atto,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
+\author{Rodney~V.~Fonseca \orcidicon{0000-0003-3948-3145}, 
+		Rog\'{e}rio~G.~Negri \orcidicon{0000-0002-4808-2362},
+        Alu\'{i}sio~Pinheiro \orcidicon{0000-0002-6132-7731},
+        and~Abdourrahmane~Atto \orcidicon{0000-0003-1753-4917},~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
 \thanks{This work was supported by FAPESP grants 2016/24469-6 and 2018/04654-9 and CNPq grants 309230/2017-9 and 310991/2020-0.}
@@
 
-In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $\vA$ where $A_{kl} = \sum_{m=2}^{n}|\mathcal{I}_{kl}^{(m)}-\mathcal{I}_{kl}^{(m-1)}|$.
+In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $\vA$ where \textcolor{magenta}{$A_{kl} = \sum_{m=2}^{n}|\mathcal{I}_{kl}^{(m)}-\mathcal{I}_{kl}^{(m-1)}|$}.
 The ECS stands for the use of $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ when defining $D_{kl}^{(m)}$ at Equation~\ref{E:eq_defdm}.
@@
 
+%%%RGN: adaptar...
+\textcolor{magenta}{
+We used a computer with an Intel Intel i-7 processor (\SI{8}{core}, \SI{3.5}{\giga\hertz}), and \SI{16}{\giga\byte} of RAM running the Ubuntu Linux version~20.04 operating system.
+The code of the proposed method is freely available at \url{https://github.com/rodneyfv/WECS}.
+The times to run WECS, TAAD and ECS were, respectively, 143.45s, 7.89s and 110.08s.
+Despite taking a longer time to run, we shall see that WECS provides more accurate results than the other two methods (c.f.
+Table \ref{tabAccExpSentinel}).
+}
+
 %%%%%%%%%%%%%%%%%%%%
@@
 For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a binary matrix that can be compared with the total true changes shown in Figure \ref{figResSim1}.
-Similarly, Figures~\ref{figResSim3} and \ref{figResSim4} depicts the matrices $\vA$ and $\vR$ provided by TAAD and ECS methods.
+Similarly, Figures~\ref{figResSim3} and \ref{figResSim4} depict the matrices $\vA$ and $\vR$ provided by TAAD and ECS methods.
 
@@
 %\caption{Simulated images with changing ellipses. (a) Image composed by the total changes over time. (b) Proposed db2 WECS $\vd(m)$ with $J=2$; (c) Standard approach. (d) $\vd(m)$ without wavelets.}
-\caption{Expected change/non-change regions and typical results provided bu WECS, TAAD and ECS before the thresholding process.}
+\caption{Expected change/non-change regions and typical results provided by WECS, TAAD and ECS before the thresholding process.}
 \label{F:Changes_methods_images}
@@
 \mbox{
-\subfigure[$\vd(m)$ with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSimBasis}}
-\subfigure[db2 $\vd(m)$ with different levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSimJ}}
-}
-
-\mbox{
-\subfigure[The proposed methods in black (db2 WECS $\vd(m)$) vs two non-wavelet methods: standard log-ratio aggregation (red stars); and $\vd(m)$ (blue)]{\includegraphics[width=0.23\textwidth]{../../graphs/methodROC.pdf}\label{figROCSimRes}}
-%\subfigure[The proposed db2 WECS $\vd(m)$ (red circles) and $\vd(m)$ with deep-learning feature extraction and without wavelets (yellow line)]{\includegraphics[width=0.23\textwidth]{../../graphs/comparisonROC.pdf}\label{figROCSim4}}
+\subfigure[WECS with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSimBasis}}
+\subfigure[db2 WECS with different decomposition levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSimJ}}
+}
+
+\mbox{
+\subfigure[The proposed methods in red (db2 WECS) vs two non-wavelet methods: standard thresholding of aggregate absolute difference (TAAD - blue); and energy correlation screening (ECS - green)]{\includegraphics[width=0.46\textwidth]{../../graphs/methodROC_ext.pdf}\label{figROCSimRes}}
 }
@@
 Figure~\ref{figROCSimBasis} presents the ROC curves considering distinct wavelet basis in the WECS method.
-All instances adopts $J=2$.
+All instances adopt $J=2$.
 The curves allow concluding that db2 and sym2 deliver the best trade-off between true and false positive ratios (i.e., high true positive ratios even when the false positive ratios are low).
@@
 Figure~\ref{figROCSimJ} depicts the ROC curves with $J \in \{1,2,3,4,5\}$ as decomposition levels.
-The profiles exhibited by the ROC curves make evident that $J$ equal to 2 or 3 promotes the best performances, with a slight advantage for $J=2$, since it demands few decomposition levels.
+The profiles exhibited by the ROC curves make evident that $J$ equal to 2 or 3 promotes the best performances, with a slight advantage for $J=2$, since it demands fewer decomposition levels.
 The overall performance for $J=2$ warrants its use for the rest of the comparisons.
@@
 At last, the ROC curves shown in Figure~\ref{figROCSimRes} allow comparing the performances of WECS, TAAD and ECS approaches. 
-Firstly, the low-performance assigned to the ECS method allows concluding that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet smoothing in the context of the proposed method.
+Firstly, the low-performance assigned to the ECS method allows concluding that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet \textcolor{magenta}{smoothing} in the context of the proposed method.
 About the performance of TAAD method, a true positive ratio of 0.8 is guaranteed when tolerating a false positive ratio of approximately 0.4. 
@@
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Actual remote sensing application}\label{secExpActual}
 
-We employed the proposed change detection method on a series of 84 multi-date satellite images.
-The images were taken on a forest region at the border of Brazil and the French Guiana from 2015-12-26 to 2017-12-3.
-Each image has two channels (VV and VH) and 1538 by 1556 pixels.
-We perform a change detection wavelet analyses on the combined image by considering each observed entry as $\mathcal{I}_{k,l}^{(m)} = ((\mathrm{VV}_{k,l}^{(m)})^2+(\mathrm{VH}_{k,l}^{(m)})^2)^{1/2}$, where $\mathrm{VV}$ and $\mathrm{VH}$ represent the matrices of observations from VV and VH channels, respectively.
+This section demonstrates the WECS, TAAD and ECS methods in a real-world change detection application.
+The appropriate wavelet basis and resolution level previously identified in Section~\ref{secExpSimulated} (i.e., db2 and $J=2$) are still considered in this analysis.
+
+In detail, this application considers a large multi-temporal series of 84 images acquired over a forest region at the border of Brazil and the French Guiana (Fig.~\ref{figAE}), from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
+
+Each image contains the amplitude signal backscatters relative to VV and VH polarizations, a spatial resolution of \SI{10}{\meter} and support of $1538 \times 1556$ pixels wide.
+Figures~\ref{figImgSentinel2015} and \ref{figImgSentinel2017} depict the first and last images of the time series, where it s possible to compare and identify some landscape changes.
+
+After careful visual inspection of the backscatter profile of each image in the series, it was possible to identify the regions where the land cover change occurs or does not occur, allowing then collecting reference samples regarding the ``change'' and ``non-change'' conditions.
+Such samples are needed to compute the accuracy measures beforehand mentioned in Section~\ref{secExpDesign}. The spatial distribution of these samples is presented in Figure~\ref{figImgSentinelROIs}.
 
@@
 \mbox{
-\subfigure[26th December, 2015]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2015.pdf}\label{figImgSentinel2015}} 
-\subfigure[3rd December, 2017]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2017.pdf}\label{figImgSentinel2017}}
-}
-
-\subfigure[Ground truth change/non-change samples]{\includegraphics[width=0.23\textwidth]{../../figs/coming-soon.png}\label{figImgSentinelROIs}}
-
-\caption{\textcolor{red}{Inserir caption...}}\label{figImageRef}
+\subfigure[December 26th 2015]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2015.pdf}\label{figImgSentinel2015}} 
+\subfigure[December 3rd 2017]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/Sentinel_2017.pdf}\label{figImgSentinel2017}}
+}
+
+\subfigure[Reference change and non-change samples]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/SamplesRev.pdf}\label{figImgSentinelROIs}}
+
+\caption{First and last images in the adopted multitemporal image series, in VV-HV-VV RGB color composition, and reference samples.}\label{figImageRef}
 \end{figure}
 
-A multi-resolution analysis based on a Symlet basis with filter of length 16 (symlet 8) is built.
+In order to apply the analyzed methods, the dual-polarized images were combined into a single-band representation considering
+$\mathcal{I}_{kl}^{(m)} = \sqrt{(\mathrm{VV}_{kl}^{(m)})^2+(\mathrm{VH}_{kl}^{(m)})^2}$, with $\mathrm{VV}$ and $\mathrm{VH}$ representing the available polarizations.
+
+%%%%RGN: pode verificar como fica a adaptaćão para este trecho?
+%\textcolor{magenta}{
+%A multi-resolution analysis based on a Symlet basis with filter of length 16 (symlet 8) is built.
 In order to have dimensions as power of 2, the matrices $\mathcal{I}_{k,l}$ are extended to a $2048\times2048$ matrix with $\mathcal{I}_{k,l}$ at the center and the remaining parts being completed with mirrored values at the borders.
@@
 Then, we are able to compute the squared mean differences vector $\vd$ and the matrix of absolute correlations $\vR$.
-
-The time change measures computed with $\vd$ are shown in Figure \ref{F:forest_wecs}.
-The orange line in Figure \ref{F:forest_wecs}a represents the median absolute deviation of $\vd$, which allow us to notice times that differ expressively from the others.
-We can notice that times $m=25,27,30$ are highlighted as having the most expressive changes.
-The images  corresponding to these times can be seen in Figure \ref{F:forest_change_times}.
-
-\begin{figure}[htp!]
-\centering
-\includegraphics[scale=.2]{../../figs/forest_vSumDifCoefSq}\hspace{-.5cm}
-\caption{Plot of $\vd(m)$, $m=1,\ldots,85$ with orange horizontal line indicating two times its median absolute deviation.}
-\label{F:forest_wecs}
+%}
+
+Initially, to express the instantaneous change values,
+Figure~\ref{F:forest_wecs} presents the temporal variations given by $\vd$.
+According to the presented profile, it is possible to observe the fluctuations around a central tendency near $12$ as highlight instants with a value similar, higher or lower than this tendency, as pointed out for $m$ equal to 1, 30 and 59, respectively.
+%
+For sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(10)}$ and $\vX^{(50)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_wecs}.
+While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rises in $\vX^{(30)}$.
+Reversely, $\vX^{(59)}$ offers low contribution for identifying change regions.
+
+%-----------------------
+\begin{figure}[htb!]
+\centering
+\includegraphics[width=0.485\textwidth]{../../graphs/LineInstants_1-30-59.pdf}
+\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$.
+Distinct deviation values occurs for $m$ at $1$ (intermediate), $30$ (high) and $59$ (low) in comparison to a central tendency around 12.}\label{F:forest_wecs}
 \end{figure}
 
-\begin{figure}[htp!]
-\includegraphics[scale=.3]{../../figs/forest_changes_time} 
-\caption{Images $\vX^{(m)}$ from times $m$ correspond to top-3 highest values of $\vd(m)$ and the mean image $\bar{\mathcal{I}}$.}
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[$m=1$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_1.pdf}\label{figImgSentinel2015}} 
+\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_30.pdf}\label{figImgSentinel2017}}
+}
+
+\mbox{
+\subfigure[$m=59$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_59.pdf}\label{figImgSentinel2015}} 
+\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_mean.pdf}\label{figImgSentinel2017}}
+}
+
+\caption{Images $\vX^{(m)}$ from distinct instant $m$ and the mean image $\overline{\mathcal{I}}$.}
 \label{F:forest_change_times}
@@
 
-The changes in space can be analyzed using the image obtained with $\vR$, which is displayed in Figure \ref{F:forest_wecs}c.
-Denoting the images' dimension as $N=1538\times1556$, if we take the $[N/\log N]$ largest absolute correlations as those corresponding to possible change points, we obtain a matrix of zeros and ones that is presented in Figure \ref{F:forest_wecs}e.
-The white regions in Figure \ref{F:forest_wecs}c (entries with value one) represent the change points, which seem to concentrate mainly on three regions: at the center, to the right of the river; at the top, on the left border of the river; and at the top left.
-Computing aggregated absolute differences to measure changes in space, we obtain Figure \ref{F:forest_wecs}d.
-An image of detected changes corresponding to measures in  Figure \ref{F:forest_wecs}d can be obtained applying a thresholding method for grayscale images.
-Figure \ref{F:forest_wecs}f shows the result of using Otsu's thresholding method \cite{otsu1979threshold} for aggregation of absolute differences.
-
-The comparison of performance for the two change detection measures considered here can be checked in Figure \ref{F:forest_roc_wecs_agg}, where a ROC curve is computed to check the correct detection of changes.
-WECS attains high true positive rates before the aggregation method, but both methods design competitive results.
-This performance differs from the ROC curve results in the previous section because the real images have a mean signal to noise ratio (SNR) of 3.9327, whereas the simulated images were generated with SNR of 0.2309.
-Hence, WECS is expected to have performance closer to the aggregation method in less noisy applications.
-The reference of correct change regions is shown in Figure \ref{F:forest_roc_wecs_agg}a, which are determined using [XXXXX].
-
-We can also compare the change detection methods using the $F_1$-score accuracy measure.
-Denoting $\mathrm{TP}$ as the number of true positives (change pixels correctly detected), $\mathrm{FP}$ as the number of false positives (nonchange pixels flagged as change point) and $\mathrm{FN}$ as the total of false negatives (change points flagged as nonchange point), the $F_1$-score is defined as
-\begin{equation*}
-F_1 = 2\frac{\mathrm{Pr}\cdot\mathrm{Re}}{\mathrm{Pr}+\mathrm{Re}}	,
-\end{equation*}
-where
-\begin{equation*}
-\mathrm{Pr}=\frac{\mathrm{TP}}{\mathrm{TP}+\mathrm{FP}}\quad 
-\text{and} \quad \mathrm{Re} = \frac{\mathrm{TP}}{\mathrm{TP}+\mathrm{FN}}.
-\end{equation*}
-Results of these three accuracy measures are presented in Table \ref{T:forest_accuracy}.
-Results for aggregation of absolutes differences correspond to two types of thresholds: Otsu's method and Kittler-Illingworth \cite{kittler1986minimum}.
-We can observe that WECS presents the highest value of $F_1$-score, which means it has a better performance that the competing method.
-
-\begin{figure}[htp!]
-\centering
-\includegraphics[scale=.10]{../../figs/forest_change}\hspace{-.5cm}(a)
-\includegraphics[scale=.15]{../../figs/forest_roc_change}\hspace{-.5cm}(b)\\
-\includegraphics[scale=.13]{../../figs/forest_wecs_abscorr}\hspace{-.5cm}(c)
-\includegraphics[scale=.13]{../../figs/forest_aggreg_ratios}\hspace{-.5cm}(d)\\
-\includegraphics[scale=.10]{../../figs/forest_wecs_change_space}\hspace{-.5cm}(e) 
-\includegraphics[scale=.10]{../../figs/forest_aggreg_change_space}\hspace{-.5cm}(f)\\ 
-\caption{Change points in space of the forest data.
-(a) Regions (in white) where changes should be detected.
-(b) ROC curve for detection of changing regions.
-(c) Matrix of absolute correlations obtained with WECS.
-(d) Change measures from aggregation of absolute differences.
-(e) Change regions detected by highlighting the $[N/\log N]$ largest correlations in $\vR$ ($N$ denoting the image's dimension).
-(f) Change regions detected using aggregated absolute differences and Otsu's thresholding.}
-\label{F:forest_roc_wecs_agg}
+After submitting the image series to the methods WECS, ECS and TAAD, are obtained the matrices $\vR$, $\vR_{\star}$ and $\vA$, respectively, comprising measure about a tendency of changes in each position of the spatial domain.
+Such matrices are presented in Figure~\ref{figImageMethods}, where high values stand for regions with changes in their land cover over time.
+%
+However, to determine a cut-off value $\tau$ for such matrices, providing then binary maps $\mathcal{M}_{\tau}$ of ``change'' and ``non-change'' areas, the use of thresholding techniques rises as a convenient procedure.
+Among different alternatives in the literature, the Otsu (OT) [figImageMethods ref] and Kittler-Illingworth (KI) \cite{KittlerIllingworth1986} thresholding techniques had been successfully employed for change detection purposes \cite{JohnsonKasischke1998,Nielsen2007,WuEA2014,NegriEA2021}.
+
+The accuracy of obtained binary maps $\mathcal{M}_{\tau}$, resulting from the application of OT and KI algorithms on the change images provided by the WECS, ECS and TAAD methods, are measured in terms of F1-Score and kappa coefficient with basis on the reference ground-truth samples (Fig.~\ref{figImgSentinelROIs}).
+Table~\ref{tabAccExpSentinel} presents the computed accuracy measures.
+%Moreover, the percentages of True and False Positives and Negatives are shown in Figure~\ref{figTPTNs}.
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+
+\mbox{
+\subfigure[$\vR$ -- WECS]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageWECS.pdf}\label{figImgSentinel2015}} 
+\subfigure[$\tilde{\vR}$ -- ECS]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageECS.pdf}\label{figImgSentinel2017}}
+}
+
+\subfigure[$\vA$ -- TAAD]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageTAAD.pdf}\label{figImgSentinelROIs}}
+
+\caption{Change tendency matrices $\vR$, $\tilde{\vR}$ and $\vA$ assigned to the WECS, ECS and TAAD methods.}\label{figImageMethods}
 \end{figure}
 
-\begin{table}
-\caption{Accuracy measures $F_1$-score, Precision ($\mathrm{Pr}$) and Recall ($\mathrm{Re}$) computed from detection of changing regions in the forest data using WECS (taking the $[N/\log N]$ largest correlations as corresponding to change) and the aggregation of absolute differences with Otsu and Kittler-Illingworth's (KI) thresholds.}
-\centering
-\begin{tabular}{c|ccc}
+%%%RGN: ler com calma o trecho abaixo...
+According to the adopted accuracy measures, the WECS method, especially when submitted to the OT algorithm, presents a better performance detecting change locations without increasing/inflating the amount of false positives (FP), which is justified by a higher F1-Score level in comparison to the competitors ECS and TAAD.
+Moreover, according to the kappa coefficient, the WECS method also deliveries a more balanced correct classification regarding the change and non-change areas.
+Furthermore, based on the kappa values and respective variances, it is verified that the difference between any pair of results is significant at 1\%.
+
+Although the performance of WECS equipped with the KI algorithm is inferior to the OT algorithm, it is worth observing that the introduced methods still provide more accurate results than ECS and TAAD.
+Regarding the competitors, the TAAD method presents frequent FP errors, leading to lower F1-Score and kappa coefficient than the ECS method.
+
+Figure~\ref{figSentinelResults} depicts the most accurate change/non-change maps according to the measures in Table~\ref{tabAccExpSentinel}.
+It is possible to verify that while the TAAD-KI method assigns the water body as a ``region of change'' and does not detect locations that changed (northwest and southeast portions -- second and fourth quadrants).
+Regarding the ECS method, the provided result has a noise-corrupted aspect due to the frequent FP classification points.
+
+Lastly, as previously observed, the WECS method equipped with the KI algorithm provides a homogeneous mapping over the non-change areas (west portion -- second and third quadrants), accurate detection over the change regions, and low inclusion (FP) and exclusion (FN) error rates.
+In adopting the OT algorithm, the inclusion error increases, resulting in a less-regularized change/non-change map.
+
+The proportions of true/false positive/negative assigned to the analyzed methods are shown in Figure~\ref{figTPTNs} and corroborate the presented discussions.
+
+\begin{table}[H]
+\caption{Accuracy values summary. The variance of kappa are multiplied by $\times 10^{-5}$.}\label{tabAccExpSentinel}
+\centering
+\begin{tabular}{ccccccc}
+\thickhline
+ & \multicolumn{2}{c}{\textbf{WECS}} & \multicolumn{2}{c}{\textbf{ECS}} & \multicolumn{2}{c}{\textbf{TAAD}}\tabularnewline
+Threshold & OT & KI & OT & KI & OT & KI\tabularnewline
 \hline 
-  & Aggregated - Otsu & Aggregated - KI & WECS \\
-\hline
-$F_1$-score & 0.2231 & 0.2163 & 0.3253 \\
-$\mathrm{Pr}$ & 0.1369 & 0.1296 & 0.2390 \\
-$\mathrm{Re}$ & 0.6022 & 0.6526 & 0.5094 \\
-\hline
+F1-Score & 0.940 & 0.876 & 0.6189 & 0.661 & 0.798 & 0.778\tabularnewline
+Kappa & 0.818 & 0.698 & 0.535 & 0.513 & 0.244 & 0.284\tabularnewline
+Var. of kappa & 6.90 & 3.43 & 1.91 & 1.72 & 0.76 & 0.91 \tabularnewline
+\thickhline
 \end{tabular}
-\label{T:forest_accuracy}
 \end{table}
 
-\section{Discussion}\label{section_discussion}
+%-----------------------
+\begin{figure}[h!]
+\centering
+
+\mbox{
+\subfigure[WECS-KI]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/WECS-KI.pdf}\label{figResChange1}} 
+\subfigure[WECS-OT]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/WECS-OT.pdf}\label{figResChange2}}
+}
+
+\mbox{
+\subfigure[ECS-OT]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/ECS-OT.pdf}\label{figResChange3}} 
+\subfigure[TAAD-KI]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/TAAD-KI.pdf}\label{figResChange4}}
+}
+
+\caption{Resulting maps from the analyzed methods.}\label{figSentinelResults}
+\end{figure}
+
+%-----------------------
+\begin{figure}[hbt]
+\centering
+\includegraphics[width=0.485\textwidth]{../../graphs/TPTNs.pdf}
+\caption{Percentages of True/False Positives (TP/FP) and True/False Negatives (TN/FN) relative to change (Positive) and non-change (Negative) reference samples.}\label{figTPTNs}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}\label{section_discussion}
 

### 2021-12-10 tex/paper.wecs/wecs_tgrs.tex
quase... (10dez21)

@@
         and~Abdourrahmane~Atto \orcidicon{0000-0003-1753-4917},~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
-\thanks{This work was supported by FAPESP grants 2016/24469-6 and 2018/04654-9 and CNPq grants 309230/2017-9 and 310991/2020-0.}
+\thanks{This work was supported by FAPESP (grants 2016/24469-6, 2018/04654-9 and 2021/01305-6) and CNPq (grants 309230/2017-9 and 310991/2020-0).}
 \thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
@@
 A special attention has been given for multi-temporal change detection using Synthetic Aperture Radar (SAR) images. 
-\textcolor{magenta}{Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail} }.
+Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail}.
 Known to be not affected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
@@
 
-In more details, Wavelet methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multi-scale/resolution information.
+In more details, wavelet-based methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multi-scale/resolution information.
 Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. 
-Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical}.
-\textcolor{magenta}{However, an application of screening ideas on wavelet coefficients is still a novel approach in change detection literature, and we show it has an interesting potential to provide good results even with simple algorithms.}
+Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate,bouhlel2015multivariate,celik2009multiscale,cui2012statistical}. 
+%%%RGN: não foi definido/motivado o que é screening
+\textcolor{orange}{However, an application of screening ideas on wavelet coefficients is still a novel approach in change detection literature, and we show it has an interesting potential to provide good results even with simple algorithms.}
 
@@
 
-\textcolor{blue}{This paper is organized as follows:
-% rest of the text goes as follows. 
+This paper is organized as follows:
+A basic concepts and definitions are presented in Section~\ref{section_theroy}.
 Section \ref{section_method} introduces the problem and presents the proposed method.
-We show WECS performance on simulated multi-temporal image data in Section \ref{section_validation}.
-In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.
-Section \ref{section_discussion} concludes the paper with a discussion.}
+We show WECS performance on simulated multi-temporal image data in Section \ref{secExpSimulated}.
+In Section \ref{secExpActual} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from December 26th 2015 to December 3rd 2017.
+Section \ref{section_discussion} concludes this study.% paper with a discussion.
 
@@
 The number $J$ is also called resolution level, a tuning parameter for wavelet smoothing.
-This wavelet smoothing is employed in the next section as a denoising step for image analysis.
-
-\textcolor{magenta}{
-Using wavelet smoothing on the images consists in an initial step to analyze the data, but the main goal in the analysis is to find changes in space and at instants of time.
+%This wavelet smoothing is employed in the next section as a denoising step for image analysis.
+
+Although the wavelet smoothing on images embraces an initial step to analyze the data, the main goal is to find the spatial changes over time. 
+The smoothed images still contain many coefficients that need to be evaluated simultaneously, which characterizes a high-dimensional problem with multiple time series corresponding to each location in space.
+Consequently, retaining only the most essential locations driving overall modifications across time is needed. 
+This issue is recurrent in regression analysis when selecting the most relevant variables.
+The feature screening technique is a particularly efficient method to identify relevant variables, especially when the number of candidates is large.
+
+%Using wavelet smoothing on images consists in an initial step to analyze the data, but the main goal in the analysis is to find changes in space and at instants of time.
 The smoothed images still contain a large number of coefficients that need to be evaluated simultaneously, what characterizes a high-dimensional problem with multiple time series corresponding to each location in space.
@@
 Methods like variable selection are largely employed to identify the most important covariates in a study, and one method particularly efficient in identifying relevant variables when the number of candidates is very large is the feature screening technique.
-}
-
-Feature screening procedure is a method originally designed for ultra-high dimensional regression models \cite{fan2008sure}.
+
+In details, the feature screening is a method originally designed for ultra-high dimensional regression models \cite{fan2008sure}.
 In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, where $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise.
@@
 %
-\textcolor{blue}{These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate}.
+Such images are relative to one SAR channel or a combination of channels; this will be specified when appropriate. 
 Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
@@
 Among several approaches able to detect change events, the Pearson correlation coefficient rises as an convenient alternative.
-Furthermore, such measure \textcolor{blue}{shares connections with the idea of feature screening employed in high-dimensional regression, as explained earlier.}
+Furthermore, such measure shares connections with the idea of feature screening employed in high-dimensional regression, as explained earlier.
 
@@
 %
-The matrix $\vD^{(m)}$ embraces the squared differences between $\vX^{(m)}$ and $\overline{\mathcal{I}}$, where $D_{kl}^{(m)} = \left( \vX_{kl}^{m} - \overline{\mathcal{I}}_{kl} \right)^2$.
+The matrix $\vD^{(m)}$ embraces the squared differences between $\vX^{(m)}$ and $\overline{\mathcal{I}}$, where $D_{kl}^{(m)} = \left( \vX_{kl}^{(m)} - \overline{\mathcal{I}}_{kl} \right)^2$.
 %
@@
 In order to identify spatio-temporal changes, we employ the concept of ultra-high dimensional correlation screening \cite{fan2020statistical} discussed in Section~\ref{section_theroy}.
-For each local squared deviation time series given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, \textcolor{blue}{say $\vD_{kl} = \left\lbrace D_{kl}^{(1)},\ldots,D_{kl}^{(n)} \right\rbrace$}, consider the absolute value of its Pearson correlation with the overall squared deviations, given by $\vd$:
-\begin{equation*}\label{def_Rkld}
-R_{kl}= |\corr \left( \vD_{kl}, \vd \right)|.
-\end{equation*}
+For each local squared deviation time series given by individual elements of $\vD^{(m)}$ across $m=1,\ldots,n$, say $\vD_{kl} = \left\lbrace D_{kl}^{(1)},\ldots,D_{kl}^{(n)} \right\rbrace$, consider the absolute value of its Pearson correlation with the overall squared deviations $\vd$:
+\begin{equation}\label{def_Rkld}
+R_{kl}= |\corr \left( \vD_{kl}, \vd \right)|
+\end{equation}
 Consequently, rises the matrix $\vR$ from the elements $R_{kl}$ assigned to each $(k,l) \in \mathcal{S}$.
 
-Let us define a mapping of \textit{important} indices for changes over the image series with respect to $\overline{\mathcal I}$ as $\mathcal{M}^{\star} \subseteq \mathcal{S}$ where \textit{changes in $\{\mathcal{I}^{(m)}\}_{m=1,\ldots,n}$ with respect to $\overline{\mathcal{I}}$ are affected by local changes in the images}.
-
-An empirical mapping of change locations may be stated by:
+Let us define a mapping of \textit{relevant} indices for changes over the image series with respect to $\overline{\mathcal I}$ as $\mathcal{M}^{\star} \subseteq \mathcal{S}$ where \textit{changes in $\{\mathcal{I}^{(m)}\}_{m=1,\ldots,n}$ with respect to $\overline{\mathcal{I}}$ are affected by local changes in the images}.
+
+An empirical mapping of change locations may be stated as:
 \begin{equation}
@@
 where $\tau \in \mathbbm{R}_{+}^{*}$ is a convenient threshold value.
-\textcolor{magenta}{The idea is that for suitable values of $\tau$, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ \cite{fan2020statistical}.
+The idea is that for suitable values of $\tau$, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ \cite{fan2020statistical}.
 This screening idea has been successfully applied in different problems in areas such as graphical models \cite{zheng2020uniform}, time series \cite{yousuf2021targeting} and remote sensing \cite{li2021mapping}.
-We show it can also shed light on problems related to change point detection in time series of images.}
+We show it can also shed light on problems related to change detection in image time series.
 
@@
 
-In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $\vA$ where \textcolor{magenta}{$A_{kl} = \sum_{m=2}^{n}|\mathcal{I}_{kl}^{(m)}-\mathcal{I}_{kl}^{(m-1)}|$}.
-The ECS stands for the use of $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ when defining $D_{kl}^{(m)}$ at Equation~\ref{E:eq_defdm}.
-
-%
-%\textcolor{red}{
-%Panel (c) presents the result of using aggregated absolute differences, a standard approach where the accumulation of changes are measured by a matrix $\vS=\{\vS_{k,l}\}$ with $S_{kl} = \sum_{m=2}^n	|\mathcal{I}_{k,l}^{(m)} - \mathcal{I}_{k,l}^{(m-1)}|$.
-Finally, in Panel (d) we can see the result if $\vd(m)$ is performed purely on the spatial domain, using $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ in the WECS formulation.
-%}
-
-%%%
-
-The second study (Sec.~\ref{secExpActual}) analyzes the performance of WECS, and respective comparison with TAAD and ECS, in a real-world application with actual SAR image series.
+In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $\vA$ where $A_{kl} = \sum_{m=2}^{n}|\mathcal{I}_{kl}^{(m)}-\mathcal{I}_{kl}^{(m-1)}|$.
+The ECS stands for the use of $\mathcal{I}^{(m)}$ instead of $\vX^{(m)}$ when defining $D_{kl}^{(m)}$ at Equation~\ref{E:eq_defdm}, and consequently Equation~\ref{def_Rkld}.
+The resulting ``Pearson correlation'' matrix from ECS approach is denoted by $\widetilde{\vR}$.
+
+The second study (Sec.~\ref{secExpActual}) presents an analysis on the performance of WECS, and respective comparison with TAAD and ECS, in a real-world application with actual SAR image series.
 The appropriate wavelet family and resolution level previously identified are employed.
-ROC curves, F1-Score \cite{Rijsbergen1979}, the kappa coefficient, and the variance of kappa \cite{Congalton2019}, are adopted for this purpose.
+F1-Score \cite{Rijsbergen1979}, True/False Positive/Negative (TP, TN, FP and FN) rates, the kappa coefficient, and the variance of kappa \cite{Congalton2019}, are adopted for this purpose.
 Additionally, the computational run-times are presented and discussed.
 
-%%%RGN: adaptar...
-\textcolor{magenta}{
 We used a computer with an Intel Intel i-7 processor (\SI{8}{core}, \SI{3.5}{\giga\hertz}), and \SI{16}{\giga\byte} of RAM running the Ubuntu Linux version~20.04 operating system.
-The code of the proposed method is freely available at \url{https://github.com/rodneyfv/WECS}.
-The times to run WECS, TAAD and ECS were, respectively, 143.45s, 7.89s and 110.08s.
-Despite taking a longer time to run, we shall see that WECS provides more accurate results than the other two methods (c.f.
-Table \ref{tabAccExpSentinel}).
-}
+The code of the proposed method is freely available at \url{https://github.com/rodneyfv/wecs}.
 
@@
 For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a binary matrix that can be compared with the total true changes shown in Figure \ref{figResSim1}.
-Similarly, Figures~\ref{figResSim3} and \ref{figResSim4} depict the matrices $\vA$ and $\vR$ provided by TAAD and ECS methods.
+Similarly, Figures~\ref{figResSim3} and \ref{figResSim4} depict the matrices $\vA$ and $\widetilde{\vR}$ provided by TAAD and ECS methods.
 
@@
 \subfigure[Total changes]{\includegraphics[frame,width=0.23\textwidth]{../../figs/total_changes_adjust.png}\label{figResSim1}} 
-\subfigure[WECS -- db2, $J=2$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_adjust.png}\label{figResSim2}} 
-}
-\mbox{
-\subfigure[TAAD -- $\vA$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_logratios_adjust.png}\label{figResSim3}} 
-\subfigure[ECS]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_nowavelets_adjust.png}\label{figResSim4}} 
+\subfigure[WECS (db2, $J=2$) -- $\vR$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_adjust.png}\label{figResSim2}} 
+}
+\mbox{
+\subfigure[TAAD -- $\vA$]{\includegraphics[width=0.23\textwidth]{../../figs/corr_changes_logratios_adjust.png}\label{figResSim3}} 
+\subfigure[ECS -- $\widetilde{\vR}$]{\includegraphics[frame,width=0.23\textwidth]{../../figs/corr_changes_dm_nowavelets_adjust.png}\label{figResSim4}} 
 }
@@
 \mbox{
-\subfigure[WECS with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSimBasis}}
-\subfigure[db2 WECS with different decomposition levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSimJ}}
-}
-
-\mbox{
-\subfigure[The proposed methods in red (db2 WECS) vs two non-wavelet methods: standard thresholding of aggregate absolute difference (TAAD - blue); and energy correlation screening (ECS - green)]{\includegraphics[width=0.46\textwidth]{../../graphs/methodROC_ext.pdf}\label{figROCSimRes}}
-}
-
-\caption{ROC curves for detection of changing ellipses in simulated images and different methods.}\label{F:EllipsoidChanges_details}
+\subfigure[WECS with different wavelet bases and J=2]{\includegraphics[width=0.23\textwidth]{../../graphs/basisROC.pdf}\label{figROCSimBasis}} \quad
+\subfigure[WECS with db2 basis and different decomposition levels]{\includegraphics[width=0.23\textwidth]{../../graphs/levelsROC.pdf}\label{figROCSimJ}}
+}
+
+\mbox{
+\subfigure[ROC curve comparison between WECS (db2 and $J=2$), TAAD and ECS methods]{\includegraphics[width=0.46\textwidth]{../../graphs/methodROC_ext.pdf}\label{figROCSimRes}}
+}
+
+\caption{ROC curves related to the experiment with simulated datasey.}\label{F:EllipsoidChanges_details}
 \end{figure}
@@
 At last, the ROC curves shown in Figure~\ref{figROCSimRes} allow comparing the performances of WECS, TAAD and ECS approaches. 
-Firstly, the low-performance assigned to the ECS method allows concluding that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet \textcolor{magenta}{smoothing} in the context of the proposed method.
-About the performance of TAAD method, a true positive ratio of 0.8 is guaranteed when tolerating a false positive ratio of approximately 0.4. 
-Reversely, the WECS method provides the same true positive ratio under an almost nill false positive ratio, demonstrating then its superiority compared to the competitors.
+Firstly, the low-performance assigned to the ECS method allows concluding that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet smoothing in the context of the proposed method.
+About the performance of TAAD method, a True Positive Ratio of 0.8 is guaranteed when tolerating a False Positive Ratio of approximately 0.4. 
+Reversely, the WECS method provides the same True Positive Ratio under an almost nil false positive ratio, demonstrating then its superiority compared to the competitors.
 
@@
 \includegraphics[width=0.485\textwidth]{../../qgis/maps/StudyArea_OSM.pdf}
-\caption{Study area location.}\label{figAE}
+\caption{The study area location.}\label{figAE}
 \end{figure}
@@
 
-Initially, to express the instantaneous change values,
+Initially, to express the instantaneous change values with respect to the WECS method,
 Figure~\ref{F:forest_wecs} presents the temporal variations given by $\vd$.
@@
 %
-For sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(10)}$ and $\vX^{(50)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_wecs}.
+For sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(30)}$ and $\vX^{(59)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_wecs}.
 While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rises in $\vX^{(30)}$.
@@
 \mbox{
-\subfigure[$m=1$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_1.pdf}\label{figImgSentinel2015}} 
-\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_30.pdf}\label{figImgSentinel2017}}
-}
-
-\mbox{
-\subfigure[$m=59$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_59.pdf}\label{figImgSentinel2015}} 
-\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_mean.pdf}\label{figImgSentinel2017}}
+\subfigure[$m=1$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_1.png}\label{figImgSentinel2015}} 
+\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_30.png}\label{figImgSentinel2017}}
+}
+
+\mbox{
+\subfigure[$m=59$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_59.png}\label{figImgSentinel2015}} 
+\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_mean.png}\label{figImgSentinel2017}}
 }
@@
 
-After submitting the image series to the methods WECS, ECS and TAAD, are obtained the matrices $\vR$, $\vR_{\star}$ and $\vA$, respectively, comprising measure about a tendency of changes in each position of the spatial domain.
+After submitting the image series to the methods WECS, ECS and TAAD, are obtained the matrices $\vR$, $\widetilde{\vR}$ and $\vA$, respectively, comprising measure about a tendency of changes in each position of the spatial domain.
 Such matrices are presented in Figure~\ref{figImageMethods}, where high values stand for regions with changes in their land cover over time.
 %
-However, to determine a cut-off value $\tau$ for such matrices, providing then binary maps $\mathcal{M}_{\tau}$ of ``change'' and ``non-change'' areas, the use of thresholding techniques rises as a convenient procedure.
-Among different alternatives in the literature, the Otsu (OT) [figImageMethods ref] and Kittler-Illingworth (KI) \cite{KittlerIllingworth1986} thresholding techniques had been successfully employed for change detection purposes \cite{JohnsonKasischke1998,Nielsen2007,WuEA2014,NegriEA2021}.
+However, to determine a cut-off value $\tau$ for such matrices, providing then binary maps $\mathcal{M}_{\tau}$ of ``change'' and ``non-change'' areas (Equation~\ref{E:def_Mtaud}), the use of thresholding techniques rises as a convenient procedure.
+Among different alternatives in the literature, the Otsu (OT) \cite{otsu1979threshold} and Kittler-Illingworth (KI) \cite{KittlerIllingworth1986} thresholding techniques had been successfully employed for change detection purposes \cite{JohnsonKasischke1998,Nielsen2007,WuEA2014,NegriEA2021}.
 
@@
 \mbox{
-\subfigure[$\vR$ -- WECS]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageWECS.pdf}\label{figImgSentinel2015}} 
-\subfigure[$\tilde{\vR}$ -- ECS]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageECS.pdf}\label{figImgSentinel2017}}
-}
-
-\subfigure[$\vA$ -- TAAD]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageTAAD.pdf}\label{figImgSentinelROIs}}
-
-\caption{Change tendency matrices $\vR$, $\tilde{\vR}$ and $\vA$ assigned to the WECS, ECS and TAAD methods.}\label{figImageMethods}
+\subfigure[WECS -- $\vR$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageWECS.png}\label{figImgSentinel2015}} 
+\subfigure[ECS -- $\widetilde{\vR}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageECS.png}\label{figImgSentinel2017}}
+}
+
+\subfigure[TAAD -- $\vA$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageTAAD.png}\label{figImgSentinelROIs}}
+
+\caption{The ``tendency of change'' matrices $\vR$, $\widetilde{\vR}$ and $\vA$  assigned to the WECS, ECS and TAAD methods, respectively. The values are scaled to $[0,1]$.}\label{figImageMethods}
 \end{figure}
@@
 
-Lastly, as previously observed, the WECS method equipped with the KI algorithm provides a homogeneous mapping over the non-change areas (west portion -- second and third quadrants), accurate detection over the change regions, and low inclusion (FP) and exclusion (FN) error rates.
+As previously observed, the WECS method equipped with the KI algorithm provides a homogeneous mapping over the non-change areas (west portion -- second and third quadrants), accurate detection over the change regions, and low inclusion (FP) and exclusion (FN) error rates.
 In adopting the OT algorithm, the inclusion error increases, resulting in a less-regularized change/non-change map.
 
-The proportions of true/false positive/negative assigned to the analyzed methods are shown in Figure~\ref{figTPTNs} and corroborate the presented discussions.
+The proportions of True/False Positive/Negative assigned to the analyzed methods are shown in Figure~\ref{figTPTNs} and corroborate the presented discussions.
 
@@
 
+Lastly, the computational run-time of WECS, TAAD and ECS were, respectively, \SI{143.45}{\seconds}, \SI{7.89}{\seconds} and \SI{110.08}{\seconds}.
+Despite expending a longer execution time, its worth mention that WECS' running time is not excessive in the context of remote sensing image processing, beyond provides more accurate results in comparison to the competitors.
+
+%see that WECS provides more accurate results than the other two methods (c.f. Table \ref{tabAccExpSentinel}).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%RGN: Esquecemos dessa parte... :(
+\textcolor{orange}{
 We present a novel way of detecting changes in multi-temporal satellite images, WECS.
@@
 The method is employed to analyze a time series of 84 images of a forest.
+}
 

### 2021-12-11 tex/paper.wecs/wecs_tgrs.tex
updates no tex

@@
 %%%RGN: não foi definido/motivado o que é screening
-\textcolor{orange}{However, an application of screening ideas on wavelet coefficients is still a novel approach in change detection literature, and we show it has an interesting potential to provide good results even with simple algorithms.}
+\textcolor{orange}{However, there are variable selection methods that have been successfully applied in high-dimensional statistical models, but that are still novel ideas on wavelet coefficients and in the change detection literature.
+We show such approach has an interesting potential to provide good results even with simple algorithms.}
 
@@
 %%%RGN: Esquecemos dessa parte... :(
-\textcolor{orange}{
-We present a novel way of detecting changes in multi-temporal satellite images, WECS.
+
+We present a novel way of detecting changes in multi-temporal satellite images called WECS.
 The procedure is based on wavelet energies from both the estimated individual coefficients as well as the whole mean image.
-It makes use of correlation screening for ultra-high dimensional data.
-Thereupon, WECS is expected to provide a sample of points in space in a way that such set, contains real change points with high probability.
-The proposed method's performance is shown using both simulated and real data.
-The proposed method is useful to detect spatio-temporal change points, which is illustrated on data analyses.
-The method is employed to analyze a time series of 84 images of a forest.
-}
+It makes use of correlation screening for ultra-high dimensional data to identify which locations (pixels) are the most related to an overall change measure of the image time series.
+Thereupon, WECS is expected to provide a sample of points in space in a way that such set contains real change points with high probability.
+
+The performance of WECS was evaluated in studies involving simulated and real data.
+In both experiments, WECS is compared with two standard approaches for analyzing multi-temporal images.
+One drawback of WECS compared to standard methods is that it takes longer time to process the images.
+However, its performance is shown to be superior on both real and simulated image, and the processing time is still not prohibitive.
+
+The current paper warrants future research in different directions, for example: adapting the idea of WECS to different types of images (multispectral, polarimetric SAR, etc); studying energy correlation screening method with for distinct smoothing techniques; deducing change detection rates theoretically for a statistical models.
 

### 2021-12-12 tex/paper.wecs/wecs_tgrs.tex
versao para revisao

@@
 %%%RGN: não foi definido/motivado o que é screening
-\textcolor{orange}{However, there are variable selection methods that have been successfully applied in high-dimensional statistical models, but that are still novel ideas on wavelet coefficients and in the change detection literature.
-We show such approach has an interesting potential to provide good results even with simple algorithms.}
+However, there are variable selection methods that have been successfully applied in high-dimensional statistical models but that are still novel ideas in the wavelet and change detection literature.
+We show that one of these methods has an interesting potential to provide good results even with simple algorithms.
 
@@
 
-Lastly, the computational run-time of WECS, TAAD and ECS were, respectively, \SI{143.45}{\seconds}, \SI{7.89}{\seconds} and \SI{110.08}{\seconds}.
+Lastly, the computational run-time of WECS, TAAD and ECS were, respectively, \SI{143.45}{\second}, \SI{7.89}{\second} and \SI{110.08}{\second}.
 Despite expending a longer execution time, its worth mention that WECS' running time is not excessive in the context of remote sensing image processing, beyond provides more accurate results in comparison to the competitors.
-
-%see that WECS provides more accurate results than the other two methods (c.f. Table \ref{tabAccExpSentinel}).
 

### 2021-12-13 tex/paper.wecs/corrections_Aluisio/mudancas_texto_DEC2021.tex
Aluisio's corrections

PAGE 1 COLUMN 1

LINE 24 "to images available at larger and larger from time to time" ==> "to large images available over long periods".

PAGE 1 COLUMN 2

LINE 7 "not affected" ==> "unaffected"

LINE 18 "is" XXXXX

LINE 20 "In more details," XXXXX

LINES 21-22 "not only ... problems" XXXXXX

LINES 33-34 "In the light ... concepts" ==> "Given the aforementioned discussion and motivation"

LINE -2 "A basic" ==> "basic"

PAGE 2 COLUMN 1

PARAGRAPH 3

LINE 8 'needed" ==> "desirable".

PARAGRAPH 4

LINE 1 "In details," XXXXX

LINE 2 "for ultra-high" ==> "for high and ultra-high"

LINE 7 ", where " XXXXX

LINE 9 "is that" ==> "is such that"

LINE 9 "what" ==> "which"

LINE 10 "unfeasible, but only" ==> "unfeasible. Moreover, only"

LINE -3 "could" ==> "should"

PAGE 2 COLUMN 2

PARAGRAPH 2

LINE 9 "along" ==> "over"

PARAGRAPH 3

LINE 1 "After" ==> "The"

LINE 2 ", one of its byproducts is" ==> "results in"

LINE 3 ", of so" ==> "of the so"

PARAGRAPH 4

LINE 1 "can be" ==> "may be"

LINE 2 "analysis" ==> "analyses"

LINES 2-3 "different types of wavelet transform and basis" ==> "different wavelet bases"

LINE 3 "as well the use of" ==> "as well as the"

LINE 4 "for detail" ==> "of detail"

PAGE 3 COLUMN 1

PARAGRAPH 2

LINE 1 "The most right-hand term in " ==> "The last term on the right-hand side of "

LINE 3 "regarding" ==> "and"

LINE 5 "the time" ==> "time"

LINE 8 "the spatial" ==> "spatial"

LINES 9-10 "the Pearson correlation coefficient rises as an convenient alternative" ==> 
"the Pearson correlation coefficient is an interesting alternative".

LINE -0 ", as explained earlier" ==> "[26]"

FORMULAS (2) e (3) Faltam pontos finais.

PARAGRAPH 6

LINES 1-2  "Consequen ... R ... assigned" ==> "Let R=[R_{kl}], where R_{kl} is assigned"

PARAGRAPH 7

LINE 1 "Let us define" ==> "Define"

LINE -2 "that for " ==> "that, for

ACRESCENTAR AQUI A IDEIA DE REGRESSAO ... ******Está no final do arquivo

PAGE 3 COLUMN 2

PARAGRAPH 2

LINE 2 "distinct studies" ==> "studies"

PARAGRAPH 3

LINE 2 "identify" ==> "identifying"

PARAGRAPH 3

LINES -1-(-2) "Posterior ... out" XXXXX

PARAGRAPH 4

LINE 1 "Lastly, the" ==> "The"

 PAGE 4 COLUMN 1

 PARAGRAPH 1

 LINES 1-2 "All the changes made that occur among subsequent" ==> "All changes between subsequent"

 LINE 3 "the "ones")" ==> " "ones")"

 LINE 4 "along ...instants" XXXXXXX

 PARAGRAPH 2

 LINE 1 "images,"  ==> "images"

  PARAGRAPH 2

 LINE 3 "The curves allow concluding that" ==> "We conclude that"

 PAGE 4 COLUMN 2

 PARAGRAPH 2

 LINE -1 "to the competitors" ==> "both competitors"

 PARAGRAPH 3

 LINE 1 "This sections demonstrates the WECS" ==> "This sections compares the performance of WECS"

LINE 2 "methods" XXXXXXX

LINE 3 "In detail, this application considers" ==> "We consider here

PAGE 5 COLUMN 1

 PARAGRAPH 2

 LINES 3-4 "Figures 9(a)-9(b)" ==> "Figures 6(a)-6(b)"

 PARAGRAPH 3

 LINE 2 "it was" ==> "it is"

 PARAGRAPH 4

 LINE -2 "Figure 7" ==> "Figure 8"

 PAGE 5 COLUMN 2

 PARAGRAPH 2

 LINES 1-3 "After submitting the image series to the methods WECS, ECS, and TAAD, are obtained ... domain" ==> "Through the application WECS, ECS and TAAD to the image series, we obtain matrices R, \tilde{R} and A. They represent the spatially-localized change measures given by WECS, ECS and TAAD, respectively".

PAGE 6 COLUMN 1

 PARAGRAPH 2

 LINE 4 "kappa coefficient with basis" ==> "kappa coefficient based"

PARAGRAPH 3

 LINES 1-2 "the WECS method" ==> "WECS"

 PAGE 6 COLUMN 2

 PARAGRAPH 1

 LINE 5 "the WECS method" ==> "WECS"

 PARAGRAPH 2

 LINE 4 "than ECS and TAAD" ==> "than either ECS or TAAD".

 LINES 4-5 "the TAAD method" ==> "TAAD"

 PARAGRAPH 3

 LINE 1 "the WECS method" ==> "WECS"

PAGE 7 COLUMN 2

 PARAGRAPH 3

 LINE 3-6 "studying energy correlation screening method with for distinct smoothing techniques; deducing change detection rates theoretically for a statistical models" ==> "extending energy correlation screening for distinct smoothing techniques; deducing sharp theoretical change detection rates for appropriate statistical models"

************************************ IDEIA DE REGRESSAO *****************

Na página 3 coluna 1 o antepenúltimo parágrafo se lê " Define a mapping of {\it relevant} indices ... as ${\mathcal M}_* ...$ where ... in the images".

Aqui, acrescenta-se

"If we apply to Equation (2) the index dicothomy defined by ${\mathcal M}_*$, we have
\begin{equation}
d(m)=\sum_{(k,l)\in\mathcal M_*}\beta_{k,l}D_{k,l}^{(m)}+\varepilon(m), \label{Regressao}
\end{equation}
where $\beta_{k,l}$ are non-null regression coefficients for $(k,l)\in\mathcal M_*$, and $\varepilon(m)$ are stochastic error terms.

The error terms allow for both the apportionment of spurious correlation for indices $notin \mathcal M_*$ as well as for the energies not represented by the wavelet smoothing.

It is a well-known property of discrete wavelet transforms that it statistically decorrelates the original data [19,]-[20]. For instance, this motivates the use of WECS instead of ECS or TAAD, since wavelets will result in sparser representations for $\mathcal M_*$ against either method. Moreover, the sure screening theoretical results for independent data motivates our conjecture that the regression set-up given by Equation (\ref{Regressao}) should have a good performance. A rigorous proof for dependent data sets such as multitemporal series of satellite images is beyond the scope of this manuscript, but the numerical results provide us with some solid evidences."

### 2021-12-14 tex/paper.wecs/corrections_Aluisio/mudancas_texto_DEC2021.tex
some corrections from Aluisio

@@
 
-LINE 24 "to images available at larger and larger from time to time" ==> "to large images available over long periods".
+* LINE 24 "to images available at larger and larger from time to time" ==> "to large images available over long periods".
 
@@
 
-LINE 7 "not affected" ==> "unaffected"
+* LINE 7 "not affected" ==> "unaffected"
 
-LINE 18 "is" XXXXX
+* LINE 18 "is" XXXXX
 
-LINE 20 "In more details," XXXXX
+* LINE 20 "In more details," XXXXX
 
-LINES 21-22 "not only ... problems" XXXXXX
+* LINES 21-22 "not only ... problems" XXXXXX
 
-LINES 33-34 "In the light ... concepts" ==> "Given the aforementioned discussion and motivation"
+* LINES 33-34 "In the light ... concepts" ==> "Given the aforementioned discussion and motivation"
 
-LINE -2 "A basic" ==> "basic"
+* LINE -2 "A basic" ==> "basic"
 
@@
 
-LINE 8 'needed" ==> "desirable".
+* LINE 8 'needed" ==> "desirable".
 
@@
 
-LINE 1 "In details," XXXXX
+* LINE 1 "In details," XXXXX
 
-LINE 2 "for ultra-high" ==> "for high and ultra-high"
+* LINE 2 "for ultra-high" ==> "for high and ultra-high"
 
-LINE 7 ", where " XXXXX
+* LINE 7 ", where " XXXXX
 
-LINE 9 "is that" ==> "is such that"
+* LINE 9 "is that" ==> "is such that"
 
-LINE 9 "what" ==> "which"
+* LINE 9 "what" ==> "which"
 
-LINE 10 "unfeasible, but only" ==> "unfeasible. Moreover, only"
+* LINE 10 "unfeasible, but only" ==> "unfeasible. Moreover, only"
 
-LINE -3 "could" ==> "should"
+* LINE -3 "could" ==> "should"
 

### 2021-12-14 tex/paper.wecs/wecs_tgrs.tex
some corrections from Aluisio

@@
 Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping} are of great importance to study the dynamics of regions sensitive to climate changes and human activity.
-Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods to images available at larger and larger from time to time.
+Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods to large images available over long periods.
 A review for change detection in multi-temporal remote sensing is given by \cite{ban2016change}.
@@
 Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail}.
-Known to be not affected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
+Known to be unaffected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
 Conversely, due to its acquisition architecture, the speckle noise typically affects the images obtained by SAR sensors and demands additional pre-processing before its use.
@@
 Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
-In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and is allows an efficient computational treatment.
-
-In more details, wavelet-based methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multi-scale/resolution information.
+In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and allows an efficient computational treatment.
+
+Wavelet-based methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} thanks to wavelet capabilities in capturing multi-scale/resolution information.
 Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. 
@@
 
-In the light of presented discussions and motivations, using Wavelet and data screening concepts, we propose the Wavelet Energies Correlation Screening (WECS), a novel unsupervised multi-temporal change detection method for SAR images.
+Given the aforementioned discussion and motivation, we propose the Wavelet Energies Correlation Screening (WECS), a novel unsupervised multi-temporal change detection method for SAR images.
 %
@@
 
-This paper is organized as follows:
-A basic concepts and definitions are presented in Section~\ref{section_theroy}.
+This paper is organized as follows: basic concepts and definitions are presented in Section~\ref{section_theroy}.
 Section \ref{section_method} introduces the problem and presents the proposed method.
@@
 The number $J$ is also called resolution level, a tuning parameter for wavelet smoothing.
-%This wavelet smoothing is employed in the next section as a denoising step for image analysis.
 
@@
 The smoothed images still contain many coefficients that need to be evaluated simultaneously, which characterizes a high-dimensional problem with multiple time series corresponding to each location in space.
-Consequently, retaining only the most essential locations driving overall modifications across time is needed. 
+Consequently, retaining only the most essential locations driving overall modifications across time is desirable. 
 This issue is recurrent in regression analysis when selecting the most relevant variables.
@@
 
-%Using wavelet smoothing on images consists in an initial step to analyze the data, but the main goal in the analysis is to find changes in space and at instants of time.
-The smoothed images still contain a large number of coefficients that need to be evaluated simultaneously, what characterizes a high-dimensional problem with multiple time series corresponding to each location in space.
-We need to employ a method that retains the most important locations driving overall modifications across time.
-This type of problem is recurrent in statistics, and particularly in regression analysis it is mainly recognized as a variable selection problem.
-Methods like variable selection are largely employed to identify the most important covariates in a study, and one method particularly efficient in identifying relevant variables when the number of candidates is very large is the feature screening technique.
-
-In details, the feature screening is a method originally designed for ultra-high dimensional regression models \cite{fan2008sure}.
-In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, where $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise.
-The problem setup is that $p$ is much larger than $n$, what makes standard regression methods unfeasible, but only a handful of the available covariates are relevant for the model, i.e., have a nonzero corresponding parameter $\beta_i$.
+The feature screening is a method originally designed for high and ultra-high dimensional regression models \cite{fan2008sure}.
+In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise.
+The problem setup is such that $p$ is much larger than $n$, which makes standard regression methods unfeasible.
+Moreover, only a handful of the available covariates are relevant for the model, i.e., have a nonzero corresponding parameter $\beta_i$.
 The feature screening idea consists in computing the sample correlation $\mathrm{corr}(\vy,\vW_i)$ among response and explanatory variables, and then selecting those covariates whose correlation are among the highest values.
 Under suitable conditions, such method is known to select a set containing all true covariates with high probability.
-In the image change detection problem, our idea is that a similar approach could be used to detect change locations by taking an overall change measure as response variable and local (pixel) measures as potential covariates.
+In the image change detection problem, our idea is that a similar approach should be used to detect change locations by taking an overall change measure as response variable and local (pixel) measures as potential covariates.
 

### 2021-12-15 tex/paper.wecs/corrections_Aluisio/mudancas_texto_DEC2021.tex
corrections of Aluisio

@@
 
-LINE 9 "along" ==> "over"
+* LINE 9 "along" ==> "over"
 
@@
 
-LINE 1 "After" ==> "The"
+* LINE 1 "After" ==> "The"
 
-LINE 2 ", one of its byproducts is" ==> "results in"
+* LINE 2 ", one of its byproducts is" ==> "results in"
 
-LINE 3 ", of so" ==> "of the so"
+* LINE 3 ", of so" ==> "of the so"
 
@@
 
-LINE 1 "can be" ==> "may be"
+* LINE 1 "can be" ==> "may be"
 
-LINE 2 "analysis" ==> "analyses"
+* LINE 2 "analysis" ==> "analyses"
 
-LINES 2-3 "different types of wavelet transform and basis" ==> "different wavelet bases"
+* LINES 2-3 "different types of wavelet transform and basis" ==> "different wavelet bases"
 
-LINE 3 "as well the use of" ==> "as well as the"
+* LINE 3 "as well the use of" ==> "as well as the"
 
-LINE 4 "for detail" ==> "of detail"
+* LINE 4 "for detail" ==> "of detail"
 
@@
 
-LINE 1 "The most right-hand term in " ==> "The last term on the right-hand side of "
+* LINE 1 "The most right-hand term in " ==> "The last term on the right-hand side of "
 
-LINE 3 "regarding" ==> "and"
+* LINE 3 "regarding" ==> "and"
 
-LINE 5 "the time" ==> "time"
+* LINE 5 "the time" ==> "time"
 
-LINE 8 "the spatial" ==> "spatial"
+* LINE 8 "the spatial" ==> "spatial"
 
-LINES 9-10 "the Pearson correlation coefficient rises as an convenient alternative" ==> 
+* LINES 9-10 "the Pearson correlation coefficient rises as an convenient alternative" ==> 
 "the Pearson correlation coefficient is an interesting alternative".
 
-LINE -0 ", as explained earlier" ==> "[26]"
+* LINE -0 ", as explained earlier" ==> "[26]"
 
-FORMULAS (2) e (3) Faltam pontos finais.
+* FORMULAS (2) e (3) Faltam pontos finais.
 
@@
 
-LINES 1-2  "Consequen ... R ... assigned" ==> "Let R=[R_{kl}], where R_{kl} is assigned"
+* LINES 1-2  "Consequen ... R ... assigned" ==> "Let R=[R_{kl}], where R_{kl} is assigned"
 
@@
 
-LINE 1 "Let us define" ==> "Define"
+* LINE 1 "Let us define" ==> "Define"
 
-LINE -2 "that for " ==> "that, for
+* LINE -2 "that for " ==> "that, for
 
-ACRESCENTAR AQUI A IDEIA DE REGRESSAO ... ******Está no final do arquivo
+* ACRESCENTAR AQUI A IDEIA DE REGRESSAO ... ******Está no final do arquivo
 
@@
 
-LINE 2 "distinct studies" ==> "studies"
+* LINE 2 "distinct studies" ==> "studies"
 
@@
 
-LINE 2 "identify" ==> "identifying"
+* LINE 2 "identify" ==> "identifying"
 
@@
 
-LINES -1-(-2) "Posterior ... out" XXXXX
+* LINES -1-(-2) "Posterior ... out" XXXXX
 
@@
 
-LINE 1 "Lastly, the" ==> "The"
+* LINE 1 "Lastly, the" ==> "The"
 
@@
 
- LINES 1-2 "All the changes made that occur among subsequent" ==> "All changes between subsequent"
+* LINES 1-2 "All the changes made that occur among subsequent" ==> "All changes between subsequent"
 
- LINE 3 "the "ones")" ==> " "ones")"
+* LINE 3 "the "ones")" ==> " "ones")"
 
- LINE 4 "along ...instants" XXXXXXX
+* LINE 4 "along ...instants" XXXXXXX
 
@@
 
- LINE 1 "images,"  ==> "images"
+* LINE 1 "images,"  ==> "images"
 
@@
 
- LINE 3 "The curves allow concluding that" ==> "We conclude that"
+* LINE 3 "The curves allow concluding that" ==> "We conclude that"
 
@@
 
- LINE -1 "to the competitors" ==> "both competitors"
+* LINE -1 "to the competitors" ==> "both competitors"
 
@@
 
- LINE 1 "This sections demonstrates the WECS" ==> "This sections compares the performance of WECS"
+* LINE 1 "This sections demonstrates the WECS" ==> "This sections compares the performance of WECS"
 
-LINE 2 "methods" XXXXXXX
+* LINE 2 "methods" XXXXXXX
 
-LINE 3 "In detail, this application considers" ==> "We consider here
+* LINE 3 "In detail, this application considers" ==> "We consider here
 
@@
 
- LINES 3-4 "Figures 9(a)-9(b)" ==> "Figures 6(a)-6(b)"
+* LINES 3-4 "Figures 9(a)-9(b)" ==> "Figures 6(a)-6(b)"
 
@@
 
- LINE 2 "it was" ==> "it is"
+* LINE 2 "it was" ==> "it is"
 
@@
 
- LINE -2 "Figure 7" ==> "Figure 8"
+* LINE -2 "Figure 7" ==> "Figure 8"
 
@@
 
- LINES 1-3 "After submitting the image series to the methods WECS, ECS, and TAAD, are obtained ...
+* LINES 1-3 "After submitting the image series to the methods WECS, ECS, and TAAD, are obtained ...
 domain" ==> "Through the application WECS, ECS and TAAD to the image series, we obtain matrices R, \tilde{R} and A.
@@
 
- LINE 4 "kappa coefficient with basis" ==> "kappa coefficient based"
+* LINE 4 "kappa coefficient with basis" ==> "kappa coefficient based"
 
@@
 
- LINES 1-2 "the WECS method" ==> "WECS"
+* LINES 1-2 "the WECS method" ==> "WECS"
 
@@
 
- LINE 5 "the WECS method" ==> "WECS"
+* LINE 5 "the WECS method" ==> "WECS"
 
@@
 
- LINE 4 "than ECS and TAAD" ==> "than either ECS or TAAD".
+* LINE 4 "than ECS and TAAD" ==> "than either ECS or TAAD".
 
- LINES 4-5 "the TAAD method" ==> "TAAD"
+* LINES 4-5 "the TAAD method" ==> "TAAD"
 
@@
 
- LINE 1 "the WECS method" ==> "WECS"
+* LINE 1 "the WECS method" ==> "WECS"
 
@@
 
- LINE 3-6 "studying energy correlation screening method with for distinct smoothing techniques; deducing change detection rates theoretically for a statistical models" ==> "extending energy correlation screening for distinct smoothing techniques; deducing sharp theoretical change detection rates for appropriate statistical models"
+* LINE 3-6 "studying energy correlation screening method with for distinct smoothing techniques; deducing change detection rates theoretically for a statistical models" ==> "extending energy correlation screening for distinct smoothing techniques; deducing sharp theoretical change detection rates for appropriate statistical models"
 

### 2021-12-15 tex/paper.wecs/wecs_tgrs.tex
corrections of Aluisio

@@
         and~Abdourrahmane~Atto \orcidicon{0000-0003-1753-4917},~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
-\thanks{This work was supported by FAPESP (grants 2016/24469-6, 2018/04654-9 and 2021/01305-6) and CNPq (grants 309230/2017-9 and 310991/2020-0).}
+\thanks{This work was supported by FAPESP (grants 2016/24469-6, 2018/04654-9, 2021/01305-6 and 2021/04513-9) and CNPq (grants 309230/2017-9 and 310991/2020-0).}
 \thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
@@
 Such images are relative to one SAR channel or a combination of channels; this will be specified when appropriate. 
-Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time.
+Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}^{(m)}$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes over time.
 We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}^{(m)}$.
@@
 
-After application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX^{(m)}$, a smoothed version of $\mathcal{I}^{(m)}$
+The application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$ results in a matrix of the so called approximation wavelet coefficients $\vX^{(m)}$, a smoothed version of $\mathcal{I}^{(m)}$
 defined on the same support $\mathcal{S}$. The higher $J\in\{1,\ldots,\log_2( \min{u,v} )\}$ is, the smoother $\vX^{(m)}$ gets.
 
-Beyond many other aspects that can be involved in wavelet analysis of images, which may include different types of wavelet transform and basis as well the use of thresholding for detail coefficients, the current construction is focused on $\vX^{(m)}$ to provide a simple wavelet smoothing.
+Beyond many other aspects that may be involved in wavelet analyses of images, which may include different wavelet bases as well as the use of thresholding of detail coefficients, the current construction is focused on $\vX^{(m)}$ to provide a simple wavelet smoothing.
 Nevertheless, extensions based on distinct aspects are straightforward.
@@
 
-The most right-hand term in Equation~(\ref{E:eq_ANOVAlogimage}) measures the deviations 
-between the wavelet coefficients at distinct instants ($\vX^{(m)}; \ m=1,\ldots,n$) regarding the \textit{average image} (i.e., $\overline{\mathcal{I}}$).
-Such deviations may favor detecting relevant changes over the time.
-Since each element (i.e., a pixel/position $(k,l)\in \mathcal{S}$) of $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the local deviation to the overall measure
-may also allow detecting the spatial changes.
-Among several approaches able to detect change events, the Pearson correlation coefficient rises as an convenient alternative.
-Furthermore, such measure shares connections with the idea of feature screening employed in high-dimensional regression, as explained earlier.
+The last term on the right-hand side of Equation~(\ref{E:eq_ANOVAlogimage}) measures the deviations between the wavelet coefficients at distinct instants ($\vX^{(m)}; \ m=1,\ldots,n$) and the \textit{average image} (i.e., $\overline{\mathcal{I}}$).
+Such deviations may favor detecting relevant changes over time.
+Since each element (i.e., a pixel/position $(k,l)\in \mathcal{S}$) of $\vX^{(m)}$ also has a corresponding sequence of deviations in time, the local deviation to the overall measure may also allow detecting spatial changes.
+Among several approaches able to detect change events, the Pearson correlation coefficient is an interesting alternative.
+Furthermore, such measure shares connections with the idea of feature screening employed in high-dimensional regression \cite{fan2008sure}.
 
@@
 \begin{equation}\label{E:eq_defdm}
-\vd(m)= \sum_{k=1}^{u} \sum_{l=1}^{v} D_{kl}^{(m)} %; \ m=1,\ldots,n
+\vd^{(m)} = \sum_{k=1}^{u} \sum_{l=1}^{v} D_{kl}^{(m)}.
 \end{equation}
@@
 \begin{equation}\label{def_Rkld}
-R_{kl}= |\corr \left( \vD_{kl}, \vd \right)|
+R_{kl}= |\corr \left( \vD_{kl}, \vd \right)|.
 \end{equation}
-Consequently, rises the matrix $\vR$ from the elements $R_{kl}$ assigned to each $(k,l) \in \mathcal{S}$.
-
-Let us define a mapping of \textit{relevant} indices for changes over the image series with respect to $\overline{\mathcal I}$ as $\mathcal{M}^{\star} \subseteq \mathcal{S}$ where \textit{changes in $\{\mathcal{I}^{(m)}\}_{m=1,\ldots,n}$ with respect to $\overline{\mathcal{I}}$ are affected by local changes in the images}.
+Let $\vR = \{R_{kl}\}$ be the matrix of absolute correlations, where $R_{kl}$ is assigned to each $(k,l) \in \mathcal{S}$.
+
+Define a mapping of \textit{relevant} indices for changes over the image series with respect to $\overline{\mathcal I}$ as $\mathcal{M}^{\star} \subseteq \mathcal{S}$ where \textit{changes in $\{\mathcal{I}^{(m)}\}_{m=1,\ldots,n}$ with respect to $\overline{\mathcal{I}}$ are affected by local changes in the images}.
+
+\textcolor{magenta}{
+If we apply to Equation (\ref{E:eq_defdm}) the index dichotomy defined by $\mathcal{M}_*$, we have
+\begin{align}
+\vd^{(m)}=\sum_{(k,l)\in\mathcal{M}_*}\beta_{k,l}D_{k,l}^{(m)}+\varepsilon^{(m)}, \label{Regressao}
+\end{align}
+where $\beta_{k,l}$ are non-null regression coefficients for $(k,l)\in\mathcal{M}_*$, and $\varepsilon^{(m)}$ are stochastic error terms. 
+The error terms allow for both the apportionment of spurious correlation for indices not in $\mathcal{M}_*$ as well as for the energies not represented by the wavelet smoothing.
+}
+
+\textcolor{magenta}{
+It is a well-known property of discrete wavelet transforms that it statistically decorrelates the original data \cite{vidakovic1999statistical,morettin2017wavelets}.
+For instance, this motivates the use of WECS instead of a non-wavelet version of energy correlation, since wavelets will result in sparser representations for $\mathcal{M}_*$.
+Moreover, the sure screening theoretical results for independent data motivates our conjecture that the regression set-up given by Equation (\ref{Regressao}) should have a good performance.
+A rigorous proof for dependent data sets such as multitemporal series of satellite images is beyond the scope of this manuscript, but the numerical results provide us with some solid evidences.
+}
 
@@
 where $\tau \in \mathbbm{R}_{+}^{*}$ is a convenient threshold value.
-The idea is that for suitable values of $\tau$, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ \cite{fan2020statistical}.
+The idea is that, for suitable values of $\tau$, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ \cite{fan2020statistical}.
 This screening idea has been successfully applied in different problems in areas such as graphical models \cite{zheng2020uniform}, time series \cite{yousuf2021targeting} and remote sensing \cite{li2021mapping}.
@@
 
-In order to assess the proposed method, this section presents two distinct studies comprising distinct datasets.
-
-The first study (Sec.~\ref{secExpSimulated}) uses a simulated data set and focus on identify the most appropriated wavelet family and resolution level $J$.
+In order to assess the proposed method, this section presents two studies comprising distinct datasets.
+
+The first study (Sec.~\ref{secExpSimulated}) uses a simulated data set and focus on identifying the most appropriated wavelet family and resolution level $J$.
 Specifically, the Haar (haar), Daubechies of order 2 and 4 (db2 and db4), Coiflets of order 4 (coif4) and Symlets of order 2 and 4 (sym2 and sym4) wavelet families are analyzed \cite{BeylkinEA1991,Daubechies1992}.
-Posterior, once selected the wavelet family, the most suitable resolution level is then pointed out.
-
-Lastly, the performance of the proposed method is measured in terms of True/False Positive Ratios and Receiver Operating Characteristic (ROC) curve. 
+%Posterior, once selected the wavelet family, the most suitable resolution level is then pointed out.
+
+The performance of the proposed method is measured in terms of True/False Positive Ratios and Receiver Operating Characteristic (ROC) curve. 
 Comparisons with the standard Thresholding of Aggregate Absolute Difference (TAAD) and a non-wavelet version of WECS, herein called Energy Correlation Screening (ECS), are included in the experiments.
@@
 Smaller ellipses are then added to form $\mathcal{I}^{(3)}$ and $\mathcal{I}^{(4)}$.
-All the changes made that occur among subsequent images are shown in Figure~\ref{figResSim1}, where white regions (i.e., the ``ones'') correspond to changes along with the instants.
-
-Applying WECS to these images, we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation.
+All changes between subsequent images are shown in Figure~\ref{figResSim1}, where white regions (i.e., ``ones'') correspond to changes.
+
+Applying WECS to these images we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation.
 An example of $\vR$ is presented in Figure~\ref{figResSim2}.
@@
 All instances adopt $J=2$.
-The curves allow concluding that db2 and sym2 deliver the best trade-off between true and false positive ratios (i.e., high true positive ratios even when the false positive ratios are low).
+We conclude that db2 and sym2 deliver the best trade-off between true and false positive ratios (i.e., high true positive ratios even when the false positive ratios are low).
 As a consequence of this finding, all the following experiments and analyses consider the db2 wavelet basis.
@@
 About the performance of TAAD method, a True Positive Ratio of 0.8 is guaranteed when tolerating a False Positive Ratio of approximately 0.4. 
-Reversely, the WECS method provides the same True Positive Ratio under an almost nil false positive ratio, demonstrating then its superiority compared to the competitors.
+Reversely, the WECS method provides the same True Positive Ratio under an almost nil false positive ratio, demonstrating then its superiority compared to both competitors.
 
@@
 
-This section demonstrates the WECS, TAAD and ECS methods in a real-world change detection application.
+This section compares the performance of WECS, TAAD and ECS in a real-world change detection application.
 The appropriate wavelet basis and resolution level previously identified in Section~\ref{secExpSimulated} (i.e., db2 and $J=2$) are still considered in this analysis.
 
-In detail, this application considers a large multi-temporal series of 84 images acquired over a forest region at the border of Brazil and the French Guiana (Fig.~\ref{figAE}), from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
+We consider here a large multi-temporal series of 84 images acquired over a forest region at the border of Brazil and the French Guiana (Fig.~\ref{figAE}), from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
 
@@
 
-After careful visual inspection of the backscatter profile of each image in the series, it was possible to identify the regions where the land cover change occurs or does not occur, allowing then collecting reference samples regarding the ``change'' and ``non-change'' conditions.
+After careful visual inspection of the backscatter profile of each image in the series, it is possible to identify the regions where the land cover change occurs or does not occur, allowing then collecting reference samples regarding the ``change'' and ``non-change'' conditions.
 Such samples are needed to compute the accuracy measures beforehand mentioned in Section~\ref{secExpDesign}. The spatial distribution of these samples is presented in Figure~\ref{figImgSentinelROIs}.
@@
 %
-For sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(30)}$ and $\vX^{(59)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_wecs}.
+For sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(30)}$ and $\vX^{(59)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_change_times}.
 While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rises in $\vX^{(30)}$.
@@
 \mbox{
-\subfigure[$m=1$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_1.png}\label{figImgSentinel2015}} 
-\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_30.png}\label{figImgSentinel2017}}
-}
-
-\mbox{
-\subfigure[$m=59$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_59.png}\label{figImgSentinel2015}} 
-\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_mean.png}\label{figImgSentinel2017}}
+\subfigure[$m=1$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_1.png}\label{maps_X_1}} 
+\subfigure[$m=30$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_30.png}\label{maps_X_30}}
+}
+
+\mbox{
+\subfigure[$m=59$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_59.png}\label{maps_X_59}} 
+\subfigure[$\overline{\mathcal{I}}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/X_mean.png}\label{maps_X_mean}}
 }
@@
 
-After submitting the image series to the methods WECS, ECS and TAAD, are obtained the matrices $\vR$, $\widetilde{\vR}$ and $\vA$, respectively, comprising measure about a tendency of changes in each position of the spatial domain.
+%After submitting the image series to the methods WECS, ECS and TAAD, are obtained the matrices $\vR$, $\widetilde{\vR}$ and $\vA$, respectively, comprising measure about a tendency of changes in each position of the spatial domain.
+Through the application WECS, ECS and TAAD to the image series, we obtain matrices $\vR$, $\widetilde{\vR}$ and $\vA$.
+They represent the spatially-localized change measures given by WECS, ECS and TAAD, respectively.
 Such matrices are presented in Figure~\ref{figImageMethods}, where high values stand for regions with changes in their land cover over time.
@@
 
-The accuracy of obtained binary maps $\mathcal{M}_{\tau}$, resulting from the application of OT and KI algorithms on the change images provided by the WECS, ECS and TAAD methods, are measured in terms of F1-Score and kappa coefficient with basis on the reference ground-truth samples (Fig.~\ref{figImgSentinelROIs}).
+The accuracy of obtained binary maps $\mathcal{M}_{\tau}$, resulting from the application of OT and KI algorithms on the change images provided by the WECS, ECS and TAAD methods, are measured in terms of F1-Score and kappa coefficient based on the reference ground-truth samples (Fig.~\ref{figImgSentinelROIs}).
 Table~\ref{tabAccExpSentinel} presents the computed accuracy measures.
@@
 \mbox{
-\subfigure[WECS -- $\vR$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageWECS.png}\label{figImgSentinel2015}} 
-\subfigure[ECS -- $\widetilde{\vR}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageECS.png}\label{figImgSentinel2017}}
-}
-
-\subfigure[TAAD -- $\vA$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageTAAD.png}\label{figImgSentinelROIs}}
+\subfigure[WECS -- $\vR$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageWECS.png}\label{imageWECS}} 
+\subfigure[ECS -- $\widetilde{\vR}$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageECS.png}\label{imageECS}}
+}
+
+\subfigure[TAAD -- $\vA$]{\includegraphics[width=0.23\textwidth]{../../qgis/maps/imageTAAD.png}\label{imageTAAD}}
 
@@
 %%%RGN: ler com calma o trecho abaixo...
-According to the adopted accuracy measures, the WECS method, especially when submitted to the OT algorithm, presents a better performance detecting change locations without increasing/inflating the amount of false positives (FP), which is justified by a higher F1-Score level in comparison to the competitors ECS and TAAD.
-Moreover, according to the kappa coefficient, the WECS method also deliveries a more balanced correct classification regarding the change and non-change areas.
+According to the adopted accuracy measures, WECS, especially when submitted to the OT algorithm, presents a better performance detecting change locations without increasing/inflating the amount of false positives (FP), which is justified by a higher F1-Score level in comparison to the competitors ECS and TAAD.
+Moreover, according to the kappa coefficient, WECS also deliveries a more balanced correct classification regarding the change and non-change areas.
 Furthermore, based on the kappa values and respective variances, it is verified that the difference between any pair of results is significant at 1\%.
 
-Although the performance of WECS equipped with the KI algorithm is inferior to the OT algorithm, it is worth observing that the introduced methods still provide more accurate results than ECS and TAAD.
-Regarding the competitors, the TAAD method presents frequent FP errors, leading to lower F1-Score and kappa coefficient than the ECS method.
+Although the performance of WECS equipped with the KI algorithm is inferior to the OT algorithm, it is worth observing that the introduced methods still provide more accurate results than either ECS or TAAD.
+Regarding the competitors, TAAD presents frequent FP errors, leading to lower F1-Score and kappa coefficient than ECS.
 
@@
 
-%%%RGN: Esquecemos dessa parte... :(
-
 We present a novel way of detecting changes in multi-temporal satellite images called WECS.
@@
 
-The current paper warrants future research in different directions, for example: adapting the idea of WECS to different types of images (multispectral, polarimetric SAR, etc); studying energy correlation screening method with for distinct smoothing techniques; deducing change detection rates theoretically for a statistical models.
+The current paper warrants future research in different directions, for example: adapting the idea of WECS to different types of images (multispectral, polarimetric SAR, etc); extending energy correlation screening for distinct smoothing techniques; deducing sharp theoretical change detection rates for appropriate statistical models.
 

### 2021-12-16 tex/paper.wecs/wecs_tgrs.tex
pequenos ajustes (16dez21)

@@
 			}{|}}}}
+
+\usepackage{mathtools}
+\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
+\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We show WECS performance on simulated multi-temporal image data.
-We also evaluate the proposed method on a time series of 84 satellite images in a forest region at the border of Brazil and the French Guiana.
+We also evaluate the proposed method on a time series of 85 satellite images in a forest region at the border of Brazil and the French Guiana.
 The proposed method displays good results in covering change regions, with the additional benefit of having simple and fast computation.
@@
 The application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$ results in a matrix of the so called approximation wavelet coefficients $\vX^{(m)}$, a smoothed version of $\mathcal{I}^{(m)}$
-defined on the same support $\mathcal{S}$. The higher $J\in\{1,\ldots,\log_2( \min{u,v} )\}$ is, the smoother $\vX^{(m)}$ gets.
+defined on the same support $\mathcal{S}$. The higher $J\in\{1,2,\ldots, \floor{ \log_2( \min\{u,v\} ) } \}$ is, the smoother $\vX^{(m)}$ gets.
 
@@
 \end{equation}
-Let $\vR = \{R_{kl}\}$ be the matrix of absolute correlations, where $R_{kl}$ is assigned to each $(k,l) \in \mathcal{S}$.
+
+Consequently, it is defined $\vR = \{R_{kl}\}$ as the matrix of absolute correlations, where $R_{kl}$ is assigned to each $(k,l) \in \mathcal{S}$.
+%Let $\vR = \{R_{kl}\}$ be the matrix of absolute correlations, where $R_{kl}$ is assigned to each $(k,l) \in \mathcal{S}$.
 
@@
 
-\textcolor{magenta}{
-If we apply to Equation (\ref{E:eq_defdm}) the index dichotomy defined by $\mathcal{M}_*$, we have
+If we apply to Equation (\ref{E:eq_defdm}) the index dichotomy defined by $\mathcal{M}^\star$, we have
 \begin{align}
-\vd^{(m)}=\sum_{(k,l)\in\mathcal{M}_*}\beta_{k,l}D_{k,l}^{(m)}+\varepsilon^{(m)}, \label{Regressao}
+\vd^{(m)}=\sum_{(k,l)}\beta_{kl}D_{kl}^{(m)}+\varepsilon^{(m)}, \label{Regressao}
 \end{align}
-where $\beta_{k,l}$ are non-null regression coefficients for $(k,l)\in\mathcal{M}_*$, and $\varepsilon^{(m)}$ are stochastic error terms. 
-The error terms allow for both the apportionment of spurious correlation for indices not in $\mathcal{M}_*$ as well as for the energies not represented by the wavelet smoothing.
-}
-
-\textcolor{magenta}{
+where $\beta_{kl}$ are non-null regression coefficients for $(k,l)\in\mathcal{M}^\star$, and $\varepsilon^{(m)}$ are stochastic error terms. 
+The error terms allow for both the apportionment of spurious correlation for indices not in $\mathcal{M}^\star$ as well as for the energies not represented by the wavelet smoothing.
+
 It is a well-known property of discrete wavelet transforms that it statistically decorrelates the original data \cite{vidakovic1999statistical,morettin2017wavelets}.
-For instance, this motivates the use of WECS instead of a non-wavelet version of energy correlation, since wavelets will result in sparser representations for $\mathcal{M}_*$.
+For instance, this motivates the use of WECS instead of a non-wavelet version of energy correlation, since wavelets will result in sparser representations for $\mathcal{M}^\star$.
 Moreover, the sure screening theoretical results for independent data motivates our conjecture that the regression set-up given by Equation (\ref{Regressao}) should have a good performance.
 A rigorous proof for dependent data sets such as multitemporal series of satellite images is beyond the scope of this manuscript, but the numerical results provide us with some solid evidences.
-}
-
-An empirical mapping of change locations may be stated as:
+
+Finally, an empirical mapping of change locations may be stated as:
 \begin{equation}
@@
 %Posterior, once selected the wavelet family, the most suitable resolution level is then pointed out.
-
 The performance of the proposed method is measured in terms of True/False Positive Ratios and Receiver Operating Characteristic (ROC) curve.
+
 Comparisons with the standard Thresholding of Aggregate Absolute Difference (TAAD) and a non-wavelet version of WECS, herein called Energy Correlation Screening (ECS), are included in the experiments.
-
 In summary, the TAAD comprises the application of a thresholding algorithm on the accumulated change representation image $\vA$ where $A_{kl} = \sum_{m=2}^{n}|\mathcal{I}_{kl}^{(m)}-\mathcal{I}_{kl}^{(m-1)}|$.
@@
 
-We consider here a large multi-temporal series of 84 images acquired over a forest region at the border of Brazil and the French Guiana (Fig.~\ref{figAE}), from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
+We consider here a large multi-temporal series of 85 images acquired over a forest region at the border of Brazil and the French Guiana (Fig.~\ref{figAE}), from December 26th 2015 to December 3rd 2017, by the SAR sensor onboard the Sentinel-1 satellite.
 
@@
 \begin{table}[H]
-\caption{Accuracy values summary. The variance of kappa are multiplied by $\times 10^{-5}$.}\label{tabAccExpSentinel}
+\caption{Accuracy values summary. The variance of kappa are multiplied by $\times 10^{5}$.}\label{tabAccExpSentinel}
 \centering

### 2021-12-16 tex/paper.wecs/corrections_Aluisio/mudancas_texto_DEC2021.tex
Aluisio's corrections

@@
 
+PAGE 3 COLUMN 2
+
+* A sugestão é suprimir "This screening idea has been successfully applied in different problems in areas such as graphical models [31], time series [32] and remote sensing [33].
+We show it can also shed light on problems related to change detection in image time series" na página 3.
+Conseqüentemente, retiram-se as referências [31]-[33] da página 8.
+
+PAGE 1 COLUMN 2
+
+* Trocar "Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
+In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and allows an efficient computational
+treatment." por "Change detection process using SAR images is a challenging task, due to the the high-dimensionality of multi-temporal datasets associated with the speckle noise.
+In this context, the use of Wavelets rises as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms."
+
+PAGE 2 COLUMN 1
+
+* Trocar "The feature screening" por "Feature screening".
+
+PAGE 2 COLUMN 2
+
+* Trocar "that may be involved in wavelet" por "that are involved in wavelet".
+
+Página 1 Coluna 1
+
+Par2 Introduction
+
+* "Many other proposals of methods" => "Many other proposals"
+
+Página 1 Coluna 2
+
+Par2 l5
+
+* ", [18]" ==> XXXXXXX
+
+Par3
+
+* Trocar "Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
+In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and allows an efficient computational
+treatment." por "Change detection process using SAR images is a challenging task., due to the the high-dimensionality of multi-temporal datasets associated with the speckle noise.
+In this context, the use of Wavelets rises as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms."
+
+Par4
+
+* "Analysis of SAR ... methods, such as [21]-[24]. However," ==>  "SAR ... methods [21]-[24]. On the other hand,"
+
+Página 2 Coluna 1
+
+Par3
+
+* Trocar "The feature screening" por "Feature screening".
+
+* "In order to explain how it works, consider" ==> "Consider"
+
+Página 2 Coluna 2
+
+Par3
+* Trocar "that may be involved in wavelet" por "that are involved in wavelet".
+
+Página 3 Coluna 1
+
+Par2
+
+* "Therefore, it is defined the temporal overall variation sequence" ==> "Therefore, the temporal overall variation sequence is defined as"
+
+Página 3 Coluna 2
+
+Par2
+
+* A sugestão é suprimir "This screening idea has been successfully applied in different problems in areas such as graphical models [31], time series [32] and remote sensing [33].
+We show it can also shed light on problems related to change detection in image time series" na página 3.
+Conseqüentemente, retiram-se as referências [31]-[33] da página 8.
+
+Section IV.A
+
+* "A. Experiment design" ==> "A. Experimental design"
+
+Par3
+
+* "proposed method," ==> "proposed method's perfomance,"
+* "most appropriated" ==> "most appropriate"
+
+Página 4 Coluna 1
+
+Par1
+
+* "This section applies the WECS, TAAD, and ECS methods to detect changes" ==> "We discuss here the application of  WECS, TAAD, and ECS"
+
+* "plus a noise" ==> "plus noise".
+
+* "first four instants" ==> "first four time instants"
+
+Par2
+
+* "All changes" ==> "Change patterns"
+
+Par3
+
+* "An example of" ==> "A typical example of"
+
+Par4
+
+* "According to the experiment design" ==> "According to the experimental design"
+
+Página 4 Coluna 2
+
+Par1
+
+* "the change and non-changed" ==> "the change and no-change"
+
+Par2
+
+* "considering distint wavelet basis" ==> "considering different wavelet bases"
+
+Par3
+
+* "curves make evident that $J$ equal to 2 or 3 promotes the best performances, with a slight advantage to $J=2$," ==> "curves provides evidence that $J$ equal to 2 or 3 lead to the best performances.
+$J=2$ has a slight advantage,
+
+Par4
+
+* "At last, the ROC curves ... 4(c), allow comparing the ... and ECS approaches." ==> "The ROC curves ... 4(c), present the ... and ECS approaches."
+
+* "Firstly, the low-performance ... concluding that the simple swap" ==> "ECS's low performance shows that the simple swap"
+
+* "Reversely, the WECS method provides" ==> "Since WECS provides"
+
+Página 5 Coluna 1
+
+* Fig.4 "ROC curves ... datasey" ==> "ROC curves ... dataset"
+
+Par1
+
+* ", demonstrating then its superiority compared to both competitors" ==> ", its superior performance is clearly established"
+
+Par2
+
+* "are still considered in this analysis" ==> "are employed"
+
+Par4
+
+* "it s possible" ==> "it is possible"
+
+Página 5 Coluna 2
+
+Par2
+
+* "the WECS method" ==> "WECS"
+
+* "central tendency near 12" ==> "central value near 12"
+
+* "than this tendency," ==> "than this baseline,"
+
+* "For sake" ==> "For the sake"
+
+Página 6 Coluna 1
+
+* Fig.7 "Plot ... central tendency around 12" ==> "Plot ... baseline around 12"
+
+* Fig.8 "from distinct instant $m$" ==> ", $m=1, 30, 59$,"
+
+Página 6 Coluna 2
+
+Par2
+
+* "The accuracy of obtained binary maps" ==> "The accuracy of the binary maps"
+
+* "The WECS, ECS and TAAD methods" ==> "WECS, ECS and TAAD"
+
+* "F1-Score and kappa" ==> "the F1-Score and the kappa"
+
+* Fig.9 "assigned to the WECS, ECS and TAAD methods," ==> "associated to WECS, ECS and TAAD,"
+
+Par4
+
+* "Although the performance of WECS equipped" ==> "Although WECS's performance when equipped"
+
+Página 7 Coluna 1
+
+Par1
+
+* "verify that while the TAAD-K1 method assigns" ==> "verify that, while TAAD-K1 assigns"
+
+* " and does not" ==> ", it does not"
+
+* "Regarding the ECS method, the provided result has a noise-corrupt aspect due to the frequent" ==> 
+"Regarding ECS, its result has a noise-corrupt character, with the frequent"
+
+Par2
+
+* "the WECS method" ==> "WECS"
+
+Par3
+
+* "corroborate the presented discussions" ==> "corroborate the aforementioned discussion"
+
+* Table 1 "Accuracy ... The Variance of Kappa are multiplied ..." ==> "Accuracy ... Variance of Kappa  multiplied ..."
+
+* Fig10 "Resulting maps" ==> "Resulting change maps"
+
+Página 7 Coluna 2
+
+Par1
+
+* "processing beyond provides more accurate results in comparison to the competitors" ==> 
+"processing, beyond providing systematically more accurate results"
+
+Página 8 Coluna 2
+
+* Retirar referências [31]-[33] Zheng at al., Yousuf and Feng, Li et al. .
+
 ************************************ IDEIA DE REGRESSAO *****************

### 2021-12-16 tex/paper.wecs/wecs_tgrs.tex
Aluisio's corrections

@@
 We shall focus in this work on unsupervised approaches, whose examples in the literature include the works of \cite{bruzzone2000automatic,celik2010change,quin2014mimosa,saha2020change,NegriEA2021,NegriFrery2021}.
-Many other proposals of methods vary in their motivations as well as in their applicability.
+Many other proposals vary in their motivations as well as in their applicability.
 Change detection in multi-temporal hyperspectral images is discussed in \cite{bovolo2015time,liu2019review,matsunaga2017current}.
@@
 A special attention has been given for multi-temporal change detection using Synthetic Aperture Radar (SAR) images. 
-Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail}.
+Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}.
 Known to be unaffected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
@@
 
-Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
-In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and allows an efficient computational treatment.
+Change detection process using SAR images is a challenging task due to the the high-dimensionality of multi-temporal datasets associated with the speckle noise.
+In this context, the use of Wavelets rises as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms.
 
@@
 Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. 
-Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate,bouhlel2015multivariate,celik2009multiscale,cui2012statistical}. 
-%%%RGN: não foi definido/motivado o que é screening
-However, there are variable selection methods that have been successfully applied in high-dimensional statistical models but that are still novel ideas in the wavelet and change detection literature.
+SAR images have been investigated under different approaches using wavelet methods \cite{atto2012multidate,bouhlel2015multivariate,celik2009multiscale,cui2012statistical}. 
+On the other hand, there are variable selection methods that have been successfully applied in high-dimensional statistical models but that are still novel ideas in the wavelet and change detection literature.
 We show that one of these methods has an interesting potential to provide good results even with simple algorithms.
@@
 
-The feature screening is a method originally designed for high and ultra-high dimensional regression models \cite{fan2008sure}.
-In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise.
+Feature screening is a method originally designed for high and ultra-high dimensional regression models \cite{fan2008sure}.
+Consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise.
 The problem setup is such that $p$ is much larger than $n$, which makes standard regression methods unfeasible.
@@
 
-Beyond many other aspects that may be involved in wavelet analyses of images, which may include different wavelet bases as well as the use of thresholding of detail coefficients, the current construction is focused on $\vX^{(m)}$ to provide a simple wavelet smoothing.
+Beyond many other aspects that are involved in wavelet analyses of images, which may include different wavelet bases as well as the use of thresholding of detail coefficients, the current construction is focused on $\vX^{(m)}$ to provide a simple wavelet smoothing.
 Nevertheless, extensions based on distinct aspects are straightforward.
@@
 %
-Therefore, it is defined the temporal overall variation sequence $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$, whose elements are given by:
+Therefore, the temporal overall variation sequence is defined as $\left\lbrace \vd^{(m)}  \right\rbrace_{m=1,\ldots,n}$, whose elements are given by:
 \begin{equation}\label{E:eq_defdm}
@@
 The idea is that, for suitable values of $\tau$, the empirical set $\mathcal{M}_{\tau}$ has high probability of detecting the correct change locations in $\mathcal{M}^{\star}$ \cite{fan2020statistical}.
-This screening idea has been successfully applied in different problems in areas such as graphical models \cite{zheng2020uniform}, time series \cite{yousuf2021targeting} and remote sensing \cite{li2021mapping}.
-We show it can also shed light on problems related to change detection in image time series.
 
@@
 %%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment design}\label{secExpDesign}
-
-In order to assess the proposed method, this section presents two studies comprising distinct datasets.
-
-The first study (Sec.~\ref{secExpSimulated}) uses a simulated data set and focus on identifying the most appropriated wavelet family and resolution level $J$.
+\subsection{Experimental design}\label{secExpDesign}
+
+In order to assess the proposed method's performance, this section presents two studies comprising distinct datasets.
+
+The first study (Sec.~\ref{secExpSimulated}) uses a simulated data set and focus on identifying the most appropriate wavelet family and resolution level $J$.
 Specifically, the Haar (haar), Daubechies of order 2 and 4 (db2 and db4), Coiflets of order 4 (coif4) and Symlets of order 2 and 4 (sym2 and sym4) wavelet families are analyzed \cite{BeylkinEA1991,Daubechies1992}.
-%Posterior, once selected the wavelet family, the most suitable resolution level is then pointed out.
 The performance of the proposed method is measured in terms of True/False Positive Ratios and Receiver Operating Characteristic (ROC) curve.
@@
 
-This section applies the WECS, TAAD, and ECS methods to detect changes on a simulated multi-temporal image dataset. 
-Such series comprises 80 multi-temporal images and it is synthesized by repeating a sequence of four images with different types of changes plus a noise. 
-Examples of the first four instants are shown in Figure~\ref{F:EllipsoidChanges}. 
+We discuss here the application of  WECS, TAAD, and ECS on a simulated multi-temporal image dataset. 
+Such series comprises 80 multi-temporal images and it is synthesized by repeating a sequence of four images with different types of changes plus noise. 
+Examples of the first four time instants are shown in Figure~\ref{F:EllipsoidChanges}. 
 These images/instants are generated by summing two matrices: (i) a binary signal matrix with ones denoting where the ellipses occur and zero elsewhere; (ii) and a noise matrix with random variables following a standard Gaussian distribution.
@@
 Smaller ellipses are then added to form $\mathcal{I}^{(3)}$ and $\mathcal{I}^{(4)}$.
-All changes between subsequent images are shown in Figure~\ref{figResSim1}, where white regions (i.e., ``ones'') correspond to changes.
+Changes patterns between subsequent images are shown in Figure~\ref{figResSim1}, where white regions (i.e., ``ones'') correspond to changes.
 
 Applying WECS to these images we obtain a matrix $\vR$ of correlations between deviations of each $\mathcal{I}$ entry with the total squared mean deviation.
-An example of $\vR$ is presented in Figure~\ref{figResSim2}.
+A typical example of $\vR$ is presented in Figure~\ref{figResSim2}.
 For some choice of threshold $\tau$ on absolute values of $\vR$, we obtain a binary matrix that can be compared with the total true changes shown in Figure \ref{figResSim1}.
@@
 
-According to the experiment design (Section~\ref{secExpDesign}), ROC curves are employed to compare the performance of each analyzed method and inspect the effects of both $J$ and wavelet basis on the performance of the WECS method.
-In summary, the ROC curves exhibit the performance as a function of true and false positive ratios when distinguishing two events (in this context, the change and non-changed pixels/positions) using distinct thresholds.
+According to the experimental design (Section~\ref{secExpDesign}), ROC curves are employed to compare the performance of each analyzed method and inspect the effects of both $J$ and wavelet basis on the performance of the WECS method.
+In summary, the ROC curves exhibit the performance as a function of true and false positive ratios when distinguishing two events (in this context, the change and no-change pixels/positions) using distinct thresholds.
 The true and false positive ratios are computed using the ``total change'' image (Fig.~\ref{figResSim1}) as reference.
@@
 
-\caption{ROC curves related to the experiment with simulated datasey.}\label{F:EllipsoidChanges_details}
+\caption{ROC curves related to the experiment with simulated dataset.}\label{F:EllipsoidChanges_details}
 \end{figure}
@@
 %basis wavelet
-Figure~\ref{figROCSimBasis} presents the ROC curves considering distinct wavelet basis in the WECS method.
+Figure~\ref{figROCSimBasis} presents the ROC curves considering different wavelet bases in the WECS method.
 All instances adopt $J=2$.
@@
 Figure~\ref{figROCSimJ} depicts the ROC curves with $J \in \{1,2,3,4,5\}$ as decomposition levels.
-The profiles exhibited by the ROC curves make evident that $J$ equal to 2 or 3 promotes the best performances, with a slight advantage for $J=2$, since it demands fewer decomposition levels.
+The profiles exhibited by the ROC curves provides evidence that $J$ equal to 2 or 3 lead to the best performances.
+$J=2$ has a slight advantage, since it demands fewer decomposition levels.
 The overall performance for $J=2$ warrants its use for the rest of the comparisons.
@@
 %Compara figA
-At last, the ROC curves shown in Figure~\ref{figROCSimRes} allow comparing the performances of WECS, TAAD and ECS approaches. 
-Firstly, the low-performance assigned to the ECS method allows concluding that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet smoothing in the context of the proposed method.
+The ROC curves shown in Figure~\ref{figROCSimRes} present the performances of WECS, TAAD and ECS approaches. 
+ECS's low performance shows that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet smoothing in the context of the proposed method.
 About the performance of TAAD method, a True Positive Ratio of 0.8 is guaranteed when tolerating a False Positive Ratio of approximately 0.4. 
-Reversely, the WECS method provides the same True Positive Ratio under an almost nil false positive ratio, demonstrating then its superiority compared to both competitors.
+Since WECS provides the same True Positive Ratio under an almost nil false positive ratio, its superior performance is clearly established.
 
@@
 This section compares the performance of WECS, TAAD and ECS in a real-world change detection application.
-The appropriate wavelet basis and resolution level previously identified in Section~\ref{secExpSimulated} (i.e., db2 and $J=2$) are still considered in this analysis.
+The appropriate wavelet basis and resolution level previously identified in Section~\ref{secExpSimulated} (i.e., db2 and $J=2$) are employed.
 
@@
 Each image contains the amplitude signal backscatters relative to VV and VH polarizations, a spatial resolution of \SI{10}{\meter} and support of $1538 \times 1556$ pixels wide.
-Figures~\ref{figImgSentinel2015} and \ref{figImgSentinel2017} depict the first and last images of the time series, where it s possible to compare and identify some landscape changes.
+Figures~\ref{figImgSentinel2015} and \ref{figImgSentinel2017} depict the first and last images of the time series, where it is possible to compare and identify some landscape changes.
 
@@
 
-%%%%RGN: pode verificar como fica a adaptaćão para este trecho?
-%\textcolor{magenta}{
-%A multi-resolution analysis based on a Symlet basis with filter of length 16 (symlet 8) is built.
-In order to have dimensions as power of 2, the matrices $\mathcal{I}_{k,l}$ are extended to a $2048\times2048$ matrix with $\mathcal{I}_{k,l}$ at the center and the remaining parts being completed with mirrored values at the borders.
-The wavelet transform at resolution level $J=2$ is applied to these matrices and only the portion corresponding to the $1538 \times 1556$ is kept for further processing.
-Then, we are able to compute the squared mean differences vector $\vd$ and the matrix of absolute correlations $\vR$.
-%}
-
-Initially, to express the instantaneous change values with respect to the WECS method,
+Initially, to express the instantaneous change values with respect to WECS,
 Figure~\ref{F:forest_wecs} presents the temporal variations given by $\vd$.
-According to the presented profile, it is possible to observe the fluctuations around a central tendency near $12$ as highlight instants with a value similar, higher or lower than this tendency, as pointed out for $m$ equal to 1, 30 and 59, respectively.
-%
-For sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(30)}$ and $\vX^{(59)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_change_times}.
+According to the presented profile, it is possible to observe the fluctuations around a central value near $12$ as highlight instants with a value similar, higher or lower than this baseline, as pointed out for $m$ equal to 1, 30 and 59, respectively.
+%
+For the sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(30)}$ and $\vX^{(59)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_change_times}.
 While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rises in $\vX^{(30)}$.
@@
 \includegraphics[width=0.485\textwidth]{../../graphs/LineInstants_1-30-59.pdf}
-\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$.
-Distinct deviation values occurs for $m$ at $1$ (intermediate), $30$ (high) and $59$ (low) in comparison to a central tendency around 12.}\label{F:forest_wecs}
+\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$. Distinct deviation values occurs for $m$ at $1$ (intermediate), $30$ (high) and $59$ (low) in comparison to a baseline around 12.}\label{F:forest_wecs}
 \end{figure}
@@
 
-\caption{Images $\vX^{(m)}$ from distinct instant $m$ and the mean image $\overline{\mathcal{I}}$.}
+\caption{Images $\vX^{(m)}$ from distinct instant $m=1,30,59$ and the mean image $\overline{\mathcal{I}}$.}
 \label{F:forest_change_times}
@@
 
-The accuracy of obtained binary maps $\mathcal{M}_{\tau}$, resulting from the application of OT and KI algorithms on the change images provided by the WECS, ECS and TAAD methods, are measured in terms of F1-Score and kappa coefficient based on the reference ground-truth samples (Fig.~\ref{figImgSentinelROIs}).
+The accuracy of the binary maps $\mathcal{M}_{\tau}$, resulting from the application of OT and KI algorithms on the change images provided by WECS, ECS and TAAD, are measured in terms of the F1-Score and the kappa coefficient based on the reference ground-truth samples (Fig.~\ref{figImgSentinelROIs}).
 Table~\ref{tabAccExpSentinel} presents the computed accuracy measures.
@@
 
-\caption{The ``tendency of change'' matrices $\vR$, $\widetilde{\vR}$ and $\vA$  assigned to the WECS, ECS and TAAD methods, respectively. The values are scaled to $[0,1]$.}\label{figImageMethods}
+\caption{The ``tendency of change'' matrices $\vR$, $\widetilde{\vR}$ and $\vA$  assigned to WECS, ECS and TAAD, respectively. The values are scaled to $[0,1]$.}\label{figImageMethods}
 \end{figure}
@@
 
-Although the performance of WECS equipped with the KI algorithm is inferior to the OT algorithm, it is worth observing that the introduced methods still provide more accurate results than either ECS or TAAD.
+Although WECS's performance when equipped with the KI algorithm is inferior to the OT algorithm, it is worth observing that the introduced methods still provide more accurate results than either ECS or TAAD.
 Regarding the competitors, TAAD presents frequent FP errors, leading to lower F1-Score and kappa coefficient than ECS.
@@
 Figure~\ref{figSentinelResults} depicts the most accurate change/non-change maps according to the measures in Table~\ref{tabAccExpSentinel}.
-It is possible to verify that while the TAAD-KI method assigns the water body as a ``region of change'' and does not detect locations that changed (northwest and southeast portions -- second and fourth quadrants).
-Regarding the ECS method, the provided result has a noise-corrupted aspect due to the frequent FP classification points.
-
-As previously observed, the WECS method equipped with the KI algorithm provides a homogeneous mapping over the non-change areas (west portion -- second and third quadrants), accurate detection over the change regions, and low inclusion (FP) and exclusion (FN) error rates.
+It is possible to verify that, while TAAD-K1 assigns the water body as a ``region of change'', it does not detect locations that changed (northwest and southeast portions -- second and fourth quadrants).
+Regarding ECS, its result has a noise-corrupt character, with frequent FP classification points.
+
+As previously observed, WECS equipped with the KI algorithm provides a homogeneous mapping over the non-change areas (west portion -- second and third quadrants), accurate detection over the change regions, and low inclusion (FP) and exclusion (FN) error rates.
 In adopting the OT algorithm, the inclusion error increases, resulting in a less-regularized change/non-change map.
 
-The proportions of True/False Positive/Negative assigned to the analyzed methods are shown in Figure~\ref{figTPTNs} and corroborate the presented discussions.
+The proportions of True/False Positive/Negative assigned to the analyzed methods are shown in Figure~\ref{figTPTNs} and corroborate the aforementioned discussion.
 
 \begin{table}[H]
-\caption{Accuracy values summary. The variance of kappa are multiplied by $\times 10^{5}$.}\label{tabAccExpSentinel}
+\caption{Accuracy values summary. Variance of kappa multiplied by $\times 10^{5}$.}\label{tabAccExpSentinel}
 \centering
@@
 
-\caption{Resulting maps from the analyzed methods.}\label{figSentinelResults}
+\caption{Resulting change maps from the analyzed methods.}\label{figSentinelResults}
 \end{figure}
@@
 Lastly, the computational run-time of WECS, TAAD and ECS were, respectively, \SI{143.45}{\second}, \SI{7.89}{\second} and \SI{110.08}{\second}.
-Despite expending a longer execution time, its worth mention that WECS' running time is not excessive in the context of remote sensing image processing, beyond provides more accurate results in comparison to the competitors.
+Despite expending a longer execution time, its worth mention that WECS' running time is not excessive in the context of remote sensing image processing, beyond providing systematically more accurate results.
 

### 2021-12-17 tex/RebutalLetter/RebutalLetter.tex
modelo rebuttal (17dez21)

%\documentclass[smallextended]{svjour3}
\documentclass[11pt]{report}
%\pagestyle{empty}

%\usepackage[latin1]{inputenc}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage[many,breakable]{tcolorbox}
\usepackage{natbib}

%\usepackage{xr}
%\externaldocument{NegriEA_AssemblyUCD_R1}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\newtcolorbox{mybox}[1]{
colback = white!5!white,
colframe = white!75!black,
fonttitle=\bfseries,
title=#1,breakable
}

\usepackage{graphicx}
\usepackage{subfigure}

%\newtcolorbox{mybox}[1]{
%colback = gray!5!white,
%colframe = gray!75!black,
%fonttitle=\bfseries,
%title=#1
%}

%%%% determina o tamanho da folha
\usepackage[pdftex]{geometry}
\geometry{a4paper,left=2.8cm,right=2.8cm,top=1.5cm,bottom=1.5cm,twoside} % tamanho da pagina

%%% --------------------------------

%-----------------------------------------------------------------------------------
\begin{document}

\begin{center}
\large{\textbf{Response Letter}}

\vglue 0.3cm

\huge{A thresholding process on the dissimilarities between probability models for change detection on Remote Sensing data}
\end{center}

%\author
\begin{center}
\textbf{Luiz Gustavo Rodrigues Godoy, Rog\'{e}rio~G.~Negri, Diogo de Jesus Amore}
\end{center}

\date{\today}

\vspace{2cm}
\noindent Dear Associate Editor,
\bigskip

\textcolor{black}{We would like to thank the reviewers for the valuable comments and the time spent on checking our manuscript. 
We have addressed their comments towards further improving the paper contents and presentation. 
Below we detail the changes and updates incorporated into the manuscript in this round of review.}

\medskip
\noindent Yours sincerely,

\begin{flushright}
\noindent The authors.
\end{flushright}
%-----------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%
\noindent-----------------------------------------------------------------------------------------
\section*{Editor's decision}

%-------------------------------------
\textit{Only a few issues need to be solved as the two reviewers suggested. Please revise your manuscript.}

\medskip

\textbf{R:} We have done our best to answer all comments and included the corresponding modifications in the new version of the manuscript. 
To facilitate the identification of new content added in the revised manuscript, we highlight in red the text inclusions/changes.
%-------------------------------------

%%%%%%%%%%%%%%%%%%%
\noindent-----------------------------------------------------------------------------------------
\section*{Reviewer \#1}

%-------------------------------------
\textit{This paper presents an application of unsupervised algorithms for change detection analysis based on stochastic distance and threshold. And the author compared it with CVA method to prove it robust and high performance. However, there are several major comments in the manuscript that must be addressed before considering its publication.}

\medskip

\textbf{R:} We thank the reviewer for all his/her contributions, technical insights, and the time spent on reviewing our manuscript.
%-------------------------------------

\medskip

\medskip
%-------------------------------------
\begin{mybox}{Comment 1}
\textit{The method applied three different types of stochastic distances, while the difference of these distances is not explained, and the reason for different performances when they are applied to remote sensing images is not analyzed.}

\medskip

\textbf{R:} As a consequence of its particular formulation, each stochastic distance plays a distinct way of measuring the similarity between two probability distributions. Therefore, while two objects (i.e., regions/neighborhoods) may share high similarities according to a given stochastic distance, such objects may also demonstrate elevated divergence levels concerning other stochastic distances.
Consequently, we should understand that each stochastic distance may provide a distinct representation of the spatio-temporal dynamics.
In the context of the proposed change detection framework, it is essential to assess different stochastic distance alternatives and quantify the respective performances on mapping the landscape changes.

\medskip

The presented response was adapted and then inserted in the Section 2.2, as showed bellow:
\begin{quotation}

As a consequence of its particular formulation, each stochastic distance plays a distinct way of measuring the similarity between two probability distributions.
Therefore, while two distributions may present a certain level of similarity according to a given stochastic distance, these distributions may also demonstrate divergences concerning other stochastic distances.
Consequently, admitting that distinct stochastic distances may lead to particular conclusions, the identification of the most convenient distance for a specific application is needed.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 2}
\textit{Fig.1 is the key block diagram illustrating the proposed approach. Maybe I missed some key points, but I want to clarify some questions. Each pixel in the neighborhood is the same important for V? And the probability density functions fu and fv are the multivariate Gaussian distributions in experiment?}

\medskip

\textbf{R:} The reviewer's interpretation is right. All the pixels/positions $s$ into the neighborhood $\mathcal{V}_{\rho}(s)$ has the same importance when modeling the probability distribution functions $f_U$ and $f_V$. Such modeling is made by estimating the parameters $\mu$ and $\Sigma$ from $\mathcal{V}_{\rho}(s)$ but considering different instants, resulting in the models $f_U$ and $f_V$. In the experiments, $f_U$ and $f_V$ are both multivariate Gaussian distributions.

\medskip

The following text was added in Section~3:
\begin{quotation}

Equivalently, it may be understood that the information from $V(\mathcal{I}_1,(i,j),\rho)$ and $V(\mathcal{I}_2,(i,j),\rho)$ are employed to estimate the parameters of $f_{\mathbf{U}}$ and $f_{\mathbf{V}}$. 
In this process, the radiometric values concerning the elements/coordinates into the mentioned neighborhoods, regardless of their distance from the reference position $(i,j)$, equal importance on the distributions parameters estimation.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 3}
\textit{The author chose the reference samples manually. And I want to ask that what is the rule and way when you take the samples because I notice that the distribution of most samples at different times are located in similar areas.}

\medskip

\textbf{R:} The ground-reference samples were collected through a careful visual inspection process regarding pairs of instants. On certain occasions, when selecting a ``non-change'' sample that does not change over time, such sample tends to be reconsidered in distinct pairs of instants. In the same sense, targets/regions with frequent spectral change also may be re-employed.

\medskip

An additional explanation concerning the reference samples was inserted in the experiments section:
\begin{quotation}

Under certain conditions, places of ``non-change'' samples that do not change over time tend to be reconsidered in distinct pairs of instants. In the same sense, regions with frequent spectral transition may be re-employed as a ``change'' sample.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 4}
\textit{The author listed some literature in the introduction. I think the authors might have missed some latest and relevant papers in this area. Please comment on the following recent papers in the revised paper. [a] Change detection using deep learning approach with object-based image analysis. [b] An unsupervised domain adaptation approach for change detection and its application to deforestation mapping in tropical biomes. [c] A deep translation (gan) based change detection network for optical and sar remote sensing images}

\medskip

\textbf{R:} Thanks for the suggested papers. We included the respective citations in the introduction section.

\medskip

\begin{quotation}

Recently, diverse supervised approaches have been proposed in the literature to deal with change detection on Remote Sensing data. 
Ref.~\citenum{LiuEA2021} combines object-based image analysis and convolutional neural network concepts to perform %a supervised 
change detection. %approach. 
Ref.~\citenum{LiEA2021} employs generative adversarial nets to propose a 
%supervised 
change detection method that simultaneously analyzes optical and radar data. 
Ref.~\citenum{VegaEA2021} exploits the idea of ``Domain Adaptation'' to train a neural network using information from one source for posterior generic use on the change detection over distinct imagery datasets.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 5}
\textit{The experiments need to be further analyzed. For instance, the intermediate result of Id should be displayed and explained, because it make the experiment more convincing.}

\medskip
\textbf{R:} Thanks for this suggestion.

The distance ($\mathcal{I}_{d}$ -- SDTA) and amplitude ($\mathcal{I}_{a}$ -- CVA) images, related to the change detection maps presented in Section 4.3, were incorporated in the manuscript. Additional discussions regarding such images were also added to the text.

\medskip

\begin{quotation}

As a complementary analysis, Figures~10 to 13 show the distance ($\mathcal{I}_d$) and amplitude ($\mathcal{I}_a$) images referent to the change maps presented in Figures~6 to 9, respectively. Once the values may change as a function of the input data, distance type, and neighborhood size, the depicted maps use a quantile-based color scale.
%
Concerning both methods, it may observe that the Kittler-Illingworth technique tends to threshold the distance/amplitude values around the second and third quantile.
%
In consonance with the change maps (Figs.6-9), the SDTA method expresses the changes through homogeneous regions. The same is not observed for the CVA method, where isolated points of high amplitude values are spread along the study area.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\newpage

\vspace{0.25cm}

%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#2}

%-------------------------------------
\begin{mybox}{General Comment}

\textit{The paper proposes a change detection method for Landsat 5/8 images, referred to as Stochastic Distance Thresholding Analysis (SDTA). Please consider the following issues:}

\medskip

\textbf{R:}  We thank the reviewer for all his/her contributions, technical insights and the time spent on reviewing our manuscript.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 1}
\textit{Line 149. Known densities. Please add details, by considering both the sensor/band/resolution characteristics and the neighborhood size.}

\medskip

\textbf{R:} Further details about the densities/neighborhoods were incorporated in Section~3, as pointed below. Details concerning the adopted sensor and neighborhood size ($\rho$) are found in Sections~4.1 and 4.2.

\begin{quotation}
Equivalently, it may be understood that the information from $V(\mathcal{I}_1,(i,j),\rho)$ and $V(\mathcal{I}_2,(i,j),\rho)$ are employed to estimate the parameters of $f_{\mathbf{U}}$ and $f_{\mathbf{V}}$. 
In this process, the radiometric values concerning the elements/coordinates into the mentioned neighborhoods, regardless of their distance from the reference position $(i,j)$, have equal importance on the distributions parameters estimation.
\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 2}
\textit{Was pansharpening applied? If it was not, add results on pansharpened MS bands.}

\medskip

\textbf{R:} The pansharpening was not applied. One of the main reasons is the absence of a panchromatic band for images acquired in 1999, 2003, and 2008 by the TM sensor. 
However, according to the reviewer's comment, the conclusion section mentions the importance of future studies and analysis with higher resolution images.

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 3}
\textit{Line 208: GitHub link is not working.}

\medskip
\textbf{R:} 
In fact, the repository has been created, but we have not yet provided the code. The codes will be made available immediately after the manuscript acceptance/publishing. The main reason to make it unavailable is due to possible adjusts that may be necessary through the revision process.

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 4 and 5}
\textit{Please better explain the anomalous behavior in Fig.5b.
\\
Fig.10b shows poor performance for low FPR with respect to CVA. Justify that from a methodological point of view or with considerations on the specificity of the 2003-2008 scenario.
}

\medskip

\textbf{R:} For the sake of simplicity, we discussed comments 4 and 5 together. We included the following text in the ``results and discussion'' section.

\medskip

\begin{quotation}

The behavior observed in Figure~\cite{5b} is a consequence of the spectral behavior of the ``2003 image'' (Fig.~\ref{2003}). External factors (like atmospheric, climatic, or even seasonality) may impair the target distinction and the measured stochastic distance/amplitude values.
%
It is essential to attend that the ``2003 image'' affects both analyzed methods (SDTA and CVA) concerning 1999--2003 and 2003--2008 comparisons in terms of results accuracy, with more influence on this latter pair.
%
Furthermore, according to the ROC curves (Fig.14(a) and (b)), it is possible to identify that the ``2003 image'' demands higher False Positive Rate values (approximately 0.195 and 0.3 concerning the pairs 1999--2003 and 2003--2008, respectively) in order to deliver a True Positive Rate above 0.8.
%
In summary, we may consider that the interference mentioned is the cause of large ``blobs of changes'' that overlaps the non-change reference samples and consequently drop the F1-Score values.
%
On the other hand, since the CVA does not consider the neighborhood information, avoiding the insurgence of large ``blobs of changes'' yet producing very noisy results, it is still less susceptible to False Positive errors in this case.
%
However, from a qualitative point of view, we should stress that the change maps delivered by the SDTA method are superior to those from CVA.

\end{quotation}

\end{mybox}
%-------------------------------------

%\vspace{0.3cm}
%
%\medskip
%%-------------------------------------
%\begin{mybox}{Comment 5}
%\textit{Fig.10b shows poor performance for low FPR with respect to CVA. Justify that from a methodological point of view or with considerations on the specificity of the 2003-2008 scenario.}
%
%\medskip
%\textbf{R:} \textcolor{red}{Que explicaćão temos para este caso? (que faz o CVA ganhar da proposta). O que foi mesmo??}
%\end{mybox}
%%-------------------------------------

\vspace{0.3cm}

\bibliographystyle{apalike}
\bibliography{../JARS_DA_review/refsartigo}

\end{document}

### 2021-12-17 tex/RebutalLetter/RebutalLetter.tex
Corrections from Aluisio and rebutal letter

@@
 
-\huge{A thresholding process on the dissimilarities between probability models for change detection on Remote Sensing data}
+\huge{Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images}
 \end{center}
@@
 \begin{center}
-\textbf{Luiz Gustavo Rodrigues Godoy, Rog\'{e}rio~G.~Negri, Diogo de Jesus Amore}
+\textbf{Rodney~V.~Fonseca, Rog\'{e}rio~G.~Negri, Alu\'{i}sio~Pinheiro and Abdourrahmane~Atto}
 \end{center}
@@
 %%%%%%%%%%%%%%%%%%%
-\noindent-----------------------------------------------------------------------------------------
+\noindent---------------------------------------------------------------------------------------
 \section*{Editor's decision}
@@
 %-------------------------------------
-\textit{Only a few issues need to be solved as the two reviewers suggested. Please revise your manuscript.}
+\textit{
+Dear Dr. Fonseca:\\
+I am sorry to inform you that your manuscript:\\
+Wavelet Spatio-Temporal Change Detection on multi-temporal POLSAR images (TGRS-2021-01131)\\
+has been carefully reviewed by the Editorial Board of the IEEE Transactions on Geoscience and Remote Sensing
+(TGRS) and has not been recommended for publication at this time. It is felt that, while there may be sufficient original
+content in the manuscript to warrant publication, it still requires a more substantial amount of work than is consistent with
+a ‘Major Revision’ decision. For this reason, we encourage you to consider revising the manuscript in response to the
+comments given below and submitting it as a new submission to TGRS. If you do decide to resubmit the paper, please
+include a point-by-point response to the comments of the reviewers along with the new paper in order to expedite its
+review. Use the "Submit a Resubmission" link in your author center when you submit the new manuscript.\\
+Your resubmission deadline is 20-Jan-2022.\\
+Below are summary comments from the TGRS Editorial Board for your information.\\
+Thank you for submitting your manuscript to TGRS.\\
+Sincerely,\\
+Dr. Simon Yueh\\
+Editor, IEEE Trans. on Geoscience and Remote Sensing
+}
 
@@
 
-\textbf{R:} We have done our best to answer all comments and included the corresponding modifications in the new version of the manuscript. 
-To facilitate the identification of new content added in the revised manuscript, we highlight in red the text inclusions/changes.
+\textbf{R:} .
 %-------------------------------------
@@
 %%%%%%%%%%%%%%%%%%%
-\noindent-----------------------------------------------------------------------------------------
+\noindent---------------------------------------------------------------------------------------
+\section*{Associate Editor}
+
+%-------------------------------------
+\textit{Dear Authors,\\
+I regret to inform you that I cannot recommend publication of your paper in the IEEE Transactions on Geoscience and
+Remote Sensing journal. This rejection is based upon the enclosed comments of the reviewers.
+In case you decide to resubmit the paper to this journal, I suggest that you take into account all the reviewers’ comments
+to improve your manuscript.
+Thank you for considering TGARS for the publication of your manuscript. We welcome the future submission of your
+work to the journal.\\
+Sincerely,}
+
+\medskip
+
+\textbf{R:} .
+%-------------------------------------
+
+\medskip
+
+\newpage
+
+\vspace{0.25cm}
+
+%%%%%%%%%%%%%%%%%%%
+\noindent---------------------------------------------------------------------------------------
 \section*{Reviewer \#1}
@@
 %-------------------------------------
-\textit{This paper presents an application of unsupervised algorithms for change detection analysis based on stochastic distance and threshold.
-And the author compared it with CVA method to prove it robust and high performance.
-However, there are several major comments in the manuscript that must be addressed before considering its publication.}
+\textit{This paper presents a wavelet decomposition based change detection method for POLSAR images.
+The method seems original but the paper is poorly written.
+The detailed comments are listed as follows.}
 
@@
 
-\textbf{R:} We thank the reviewer for all his/her contributions, technical insights, and the time spent on reviewing our manuscript.
+\textbf{R:} .
 %-------------------------------------
@@
 \begin{mybox}{Comment 1}
-\textit{The method applied three different types of stochastic distances, while the difference of these distances is not explained, and the reason for different performances when they are applied to remote sensing images is not analyzed.}
+\textit{This paper lacks a thorough review of POLSAR image change detection methods. What are the challenges of
+POLSAR image change detection? What are the advantages and defects of existing methods? What’s the superiority of
+the proposed method over existing ones? What are the contributions of this paper to remote sensing community? There
+are many works about change detection in POLSAR images. This paper only addresses the method without necessary
+comparison.}
 
@@
 
-\textbf{R:} As a consequence of its particular formulation, each stochastic distance plays a distinct way of measuring the similarity between two probability distributions.
-Therefore, while two objects (i.e., regions/neighborhoods) may share high similarities according to a given stochastic distance, such objects may also demonstrate elevated divergence levels concerning other stochastic distances.
-Consequently, we should understand that each stochastic distance may provide a distinct representation of the spatio-temporal dynamics.
-In the context of the proposed change detection framework, it is essential to assess different stochastic distance alternatives and quantify the respective performances on mapping the landscape changes.
+\textbf{R:} .
 
 \medskip
-
-The presented response was adapted and then inserted in the Section 2.2, as showed bellow:
-\begin{quotation}
-
-As a consequence of its particular formulation, each stochastic distance plays a distinct way of measuring the similarity between two probability distributions.
-Therefore, while two distributions may present a certain level of similarity according to a given stochastic distance, these distributions may also demonstrate divergences concerning other stochastic distances.
-Consequently, admitting that distinct stochastic distances may lead to particular conclusions, the identification of the most convenient distance for a specific application is needed.
-
-\end{quotation}
 
@@
 \begin{mybox}{Comment 2}
-\textit{Fig.1 is the key block diagram illustrating the proposed approach.
-Maybe I missed some key points, but I want to clarify some questions.
-Each pixel in the neighborhood is the same important for V?
-And the probability density functions fu and fv are the multivariate Gaussian distributions in experiment?}
+\textit{The authors claim that “it is sparse, very fast and scalable. Finally, it is easily adapted to be updatable”. However, the
+change detection in POLSAR images has been significantly developed and there are many excellent methods. Many of
+them are also fast and scalable. Without any necessary comparison, how to demonstrate that “it is very fast and scalable”? What’s the novelty of the proposed method?}
 
@@
 
-\textbf{R:} The reviewer's interpretation is right.
-All the pixels/positions $s$ into the neighborhood $\mathcal{V}_{\rho}(s)$ has the same importance when modeling the probability distribution functions $f_U$ and $f_V$.
-Such modeling is made by estimating the parameters $\mu$ and $\Sigma$ from $\mathcal{V}_{\rho}(s)$ but considering different instants, resulting in the models $f_U$ and $f_V$.
-In the experiments, $f_U$ and $f_V$ are both multivariate Gaussian distributions.
+\textbf{R:} .
 
 \medskip
-
-The following text was added in Section~3:
-\begin{quotation}
-
-Equivalently, it may be understood that the information from $V(\mathcal{I}_1,(i,j),\rho)$ and $V(\mathcal{I}_2,(i,j),\rho)$ are employed to estimate the parameters of $f_{\mathbf{U}}$ and $f_{\mathbf{V}}$.
-In this process, the radiometric values concerning the elements/coordinates into the mentioned neighborhoods, regardless of their distance from the reference position $(i,j)$, equal importance on the distributions parameters estimation.
-
-\end{quotation}
 
@@
 \begin{mybox}{Comment 3}
-\textit{The author chose the reference samples manually.
-And I want to ask that what is the rule and way when you take the samples because I notice that the distribution of most samples at different times are located in similar areas.}
+\textit{What’s the relationship between this paper and [i]? [i] also uses wavelet transformation for change detection.
+[i] Fonseca, R. , Ludwig, G. , Montoril, M. , Pinheiro, A. . (2020). Nonparametric methods for detecting change in Multitemporal SAR-PolSAR Satellite Data. arXiv: 2001.05764.}
 
@@
 
-\textbf{R:} The ground-reference samples were collected through a careful visual inspection process regarding pairs of instants.
-On certain occasions, when selecting a ``non-change'' sample that does not change over time, such sample tends to be reconsidered in distinct pairs of instants.
-In the same sense, targets/regions with frequent spectral change also may be re-employed.
+\textbf{R:} .
 
 \medskip
-
-An additional explanation concerning the reference samples was inserted in the experiments section:
-\begin{quotation}
-
-Under certain conditions, places of ``non-change'' samples that do not change over time tend to be reconsidered in distinct pairs of instants.
-In the same sense, regions with frequent spectral transition may be re-employed as a ``change'' sample.
-
-\end{quotation}
 
@@
 \begin{mybox}{Comment 4}
-\textit{The author listed some literature in the introduction.
-I think the authors might have missed some latest and relevant papers in this area.
-Please comment on the following recent papers in the revised paper.
-[a] Change detection using deep learning approach with object-based image analysis.
-[b] An unsupervised domain adaptation approach for change detection and its application to deforestation mapping in tropical biomes.
-[c] A deep translation (gan) based change detection network for optical and sar remote sensing images}
+\textit{Methodology describes the proposed method in detail, but what’s new in this paper? All the operators are existing or
+intuitive.}
 
@@
 
-\textbf{R:} Thanks for the suggested papers. We included the respective citations in the introduction section.
+\textbf{R:} .
 
 \medskip
-
-\begin{quotation}
-
-Recently, diverse supervised approaches have been proposed in the literature to deal with change detection on Remote Sensing data. 
-Ref.~\citenum{LiuEA2021} combines object-based image analysis and convolutional neural network concepts to perform %a supervised 
-change detection. %approach. 
-Ref.~\citenum{LiEA2021} employs generative adversarial nets to propose a 
-%supervised 
-change detection method that simultaneously analyzes optical and radar data. 
-Ref.~\citenum{VegaEA2021} exploits the idea of ``Domain Adaptation'' to train a neural network using information from one source for posterior generic use on the change detection over distinct imagery datasets.
-
-\end{quotation}
 
@@
 \begin{mybox}{Comment 5}
-\textit{The experiments need to be further analyzed. For instance, the intermediate result of Id should be displayed and explained, because it make the experiment more convincing.}
+\textit{The method is not clearly described. For example, what’s log-image? What’s the epsilon in Eq. (1). Eq. (1) seems
+problematic. It denotes wavelet decomposition?}
 
 \medskip
-\textbf{R:} Thanks for this suggestion.
-
-The distance ($\mathcal{I}_{d}$ -- SDTA) and amplitude ($\mathcal{I}_{a}$ -- CVA) images, related to the change detection maps presented in Section 4.3, were incorporated in the manuscript.
-Additional discussions regarding such images were also added to the text.
+\textbf{R:} .
 
 \medskip
-
-\begin{quotation}
-
-As a complementary analysis, Figures~10 to 13 show the distance ($\mathcal{I}_d$) and amplitude ($\mathcal{I}_a$) images referent to the change maps presented in Figures~6 to 9, respectively.
-Once the values may change as a function of the input data, distance type, and neighborhood size, the depicted maps use a quantile-based color scale.
-%
-Concerning both methods, it may observe that the Kittler-Illingworth technique tends to threshold the distance/amplitude values around the second and third quantile.
-%
-In consonance with the change maps (Figs.6-9), the SDTA method expresses the changes through homogeneous regions.
-The same is not observed for the CVA method, where isolated points of high amplitude values are spread along the study area.
-
-\end{quotation}
 
@@
 \vspace{0.3cm}
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 6}
+\textit{Without comparison with recent change detection methods in experiments, how to demonstrate the advantages of the
+proposed method? For example, if there is a method that is faster and performs better than existing ones, why I have to
+use the proposed one?}
+
+\medskip
+\textbf{R:} .
+
+\medskip
+
+\end{mybox}
+%-------------------------------------
+
+\vspace{0.3cm}
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 7}
+\textit{The authors claims that “Its performance is similar to a deep learning feature extraction method”. However, the deep
+learning feature extraction method is proposed for image denoising and it is not specially designed for POLSAR images.
+It is obviously unfair to compare the proposed method with it. Why not compare the proposed method with recent
+POLSAR image change detection methods? The proposed one cannot outperform them?}
+
+\medskip
+\textbf{R:} .
+
+\medskip
+
+\end{mybox}
+%-------------------------------------
+
+\vspace{0.3cm}
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 8}
+\textit{For real data, there are only self comparisons without any comparison with existing excellent methods. Therefore, the
+conclusions may be not correct.}
+
+\medskip
+\textbf{R:} .
+
+\medskip
+
+\end{mybox}
+%-------------------------------------
 
@@
 
-\textit{The paper proposes a change detection method for Landsat 5/8 images, referred to as Stochastic Distance Thresholding Analysis (SDTA). Please consider the following issues:}
+\textit{Fonseca et al. present a wavelet-based approach to detect spatiotemporal change points on multi-temporal Polarimetric
+SAR. The approach is potentially interesting and might be of use for a broader community. However, it suffered from
+several shortcomings, which prevents me from recommending the publication of the manuscript in its current form. In the
+following o elaborate my concerns, and I hope authors find them useful.}
 
@@
 
-\textbf{R:}  We thank the reviewer for all his/her contributions, technical insights and the time spent on reviewing our manuscript.
+\textbf{R:}  .
 \end{mybox}
@@
 \begin{mybox}{Comment 1}
-\textit{Line 149. Known densities. Please add details, by considering both the sensor/band/resolution characteristics and the neighborhood size.}
+\textit{the method section is unclear and partly incomplete. There are equations whose terms are not defined (e.g., Eq 1 and
+that between 7 and 8 without a number). Furthermore, the procedure for Wavelet multiresolution analysis is not
+described. Although, provided in the cited literature, it is crucial to carefully layout the procedure used to create the input
+to the change detection algorithm. Also, statements like "J-th approximation level" are unclear. I suspect the author
+meant to say "J-th decomposition level."}
+
+\medskip
+\textbf{R:} .
+\end{mybox}
+%-------------------------------------
 
@@
 
-\textbf{R:} Further details about the densities/neighborhoods were incorporated in Section~3, as pointed below.
-Details concerning the adopted sensor and neighborhood size ($\rho$) are found in Sections~4.1 and 4.2.
+%-------------------------------------
 
-\begin{quotation}
-Equivalently, it may be understood that the information from $V(\mathcal{I}_1,(i,j),\rho)$ and $V(\mathcal{I}_2,(i,j),\rho)$ are employed to estimate the parameters of $f_{\mathbf{U}}$ and $f_{\mathbf{V}}$.
-In this process, the radiometric values concerning the elements/coordinates into the mentioned neighborhoods, regardless of their distance from the reference position $(i,j)$, have equal importance on the distributions parameters estimation.
-\end{quotation}
+\begin{mybox}{Comment 2}
+\textit{synthetic test is confusing. Figure 1 shows a sequence of 4 synthetic images with changes appearing consecutively.
+The expectation from the presented method and promises made in the abstract and introduction is that 4 corresponding
+change maps can be obtained characterizing the changes made in the synthetic datasets. However, Figure 2 shows a
+total change and some other maps, which is unclear how they were obtained or what is the relevance to the presented
+method and promises of the paper.}
 
+\medskip
+
+\textbf{R:} .
 \end{mybox}
+%-------------------------------------
+
 %-------------------------------------
@@
 %-------------------------------------
-\begin{mybox}{Comment 2}
-\textit{Was pansharpening applied? If it was not, add results on pansharpened MS bands.}
+
+\begin{mybox}{Comment 3}
+\textit{the real data set is not introduced carefully, and it is not clear what datasets from what sensor at what resolution and
+quality are used. How the real data is proceeded and prepared for this analysis, etc. In Figures 7-9 most of the panels
+(e.g., c-j) show many overlapping curves that color-coded without providing a colorbar. These figures carry very little
+useful information and I encourage authors to rethink ways to present their results. I also suggest authors to provide
+examples of images and show the wavelet decomposition results. Also, show the average approximate coefficient
+matrix. Last but not least, quantify a change map and provide a validation test such as an independent optical image
+taken from the same area...}
 
@@
 
-\textbf{R:} The pansharpening was not applied. One of the main reasons is the absence of a panchromatic band for images acquired in 1999, 2003, and 2008 by the TM sensor. 
-However, according to the reviewer's comment, the conclusion section mentions the importance of future studies and analysis with higher resolution images.
-
+\textbf{R:} .
 \end{mybox}
@@
 
-\vspace{0.3cm}
-
-\medskip
 %-------------------------------------
-\begin{mybox}{Comment 3}
-\textit{Line 208: GitHub link is not working.}
-
-\medskip
-\textbf{R:} 
-In fact, the repository has been created, but we have not yet provided the code.
-The codes will be made available immediately after the manuscript acceptance/publishing.
-The main reason to make it unavailable is due to possible adjusts that may be necessary through the revision process.
-
-\end{mybox}
-%-------------------------------------
-
-\vspace{0.3cm}
-
-\medskip
-%-------------------------------------
-\begin{mybox}{Comment 4 and 5}
-\textit{Please better explain the anomalous behavior in Fig.5b.
-\\
-Fig.10b shows poor performance for low FPR with respect to CVA. Justify that from a methodological point of view or with considerations on the specificity of the 2003-2008 scenario.
-}
-
-\medskip
-
-\textbf{R:} For the sake of simplicity, we discussed comments 4 and 5 together. We included the following text in the ``results and discussion'' section.
-
-\medskip
-
-\begin{quotation}
-
-The behavior observed in Figure~\cite{5b} is a consequence of the spectral behavior of the ``2003 image'' (Fig.~\ref{2003}).
-External factors (like atmospheric, climatic, or even seasonality) may impair the target distinction and the measured stochastic distance/amplitude values.
-%
-It is essential to attend that the ``2003 image'' affects both analyzed methods (SDTA and CVA) concerning 1999--2003 and 2003--2008 comparisons in terms of results accuracy, with more influence on this latter pair.
-%
-Furthermore, according to the ROC curves (Fig.14(a) and (b)), it is possible to identify that the ``2003 image'' demands higher False Positive Rate values (approximately 0.195 and 0.3 concerning the pairs 1999--2003 and 2003--2008, respectively) in order to deliver a True Positive Rate above 0.8.
-%
-In summary, we may consider that the interference mentioned is the cause of large ``blobs of changes'' that overlaps the non-change reference samples and consequently drop the F1-Score values.
-%
-On the other hand, since the CVA does not consider the neighborhood information, avoiding the insurgence of large ``blobs of changes'' yet producing very noisy results, it is still less susceptible to False Positive errors in this case.
-%
-However, from a qualitative point of view, we should stress that the change maps delivered by the SDTA method are superior to those from CVA.
-
-\end{quotation}
-
-\end{mybox}
-%-------------------------------------
-
-%\vspace{0.3cm}
-%
-%\medskip
-%%-------------------------------------
-%\begin{mybox}{Comment 5}
-%\textit{Fig.10b shows poor performance for low FPR with respect to CVA. Justify that from a methodological point of view or with considerations on the specificity of the 2003-2008 scenario.}
-%
-%\medskip
-%\textbf{R:} \textcolor{red}{Que explicaćão temos para este caso? (que faz o CVA ganhar da proposta). O que foi mesmo??}
-%\end{mybox}
-%%-------------------------------------
 

### 2021-12-17 tex/RebutalLetter/RebutalLetter_modelo.tex
Corrections from Aluisio and rebutal letter

%\documentclass[smallextended]{svjour3}
\documentclass[11pt]{report}
%\pagestyle{empty}

%\usepackage[latin1]{inputenc}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage[many,breakable]{tcolorbox}
\usepackage{natbib}

%\usepackage{xr}
%\externaldocument{NegriEA_AssemblyUCD_R1}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\newtcolorbox{mybox}[1]{
colback = white!5!white,
colframe = white!75!black,
fonttitle=\bfseries,
title=#1,breakable
}

\usepackage{graphicx}
\usepackage{subfigure}

%\newtcolorbox{mybox}[1]{
%colback = gray!5!white,
%colframe = gray!75!black,
%fonttitle=\bfseries,
%title=#1
%}

%%%% determina o tamanho da folha
\usepackage[pdftex]{geometry}
\geometry{a4paper,left=2.8cm,right=2.8cm,top=1.5cm,bottom=1.5cm,twoside} % tamanho da pagina

%%% --------------------------------

%-----------------------------------------------------------------------------------
\begin{document}

\begin{center}
\large{\textbf{Response Letter}}

\vglue 0.3cm

\huge{A thresholding process on the dissimilarities between probability models for change detection on Remote Sensing data}
\end{center}

%\author
\begin{center}
\textbf{Luiz Gustavo Rodrigues Godoy, Rog\'{e}rio~G.~Negri, Diogo de Jesus Amore}
\end{center}

\date{\today}

\vspace{2cm}
\noindent Dear Associate Editor,
\bigskip

\textcolor{black}{We would like to thank the reviewers for the valuable comments and the time spent on checking our manuscript. 
We have addressed their comments towards further improving the paper contents and presentation. 
Below we detail the changes and updates incorporated into the manuscript in this round of review.}

\medskip
\noindent Yours sincerely,

\begin{flushright}
\noindent The authors.
\end{flushright}
%-----------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%
\noindent-----------------------------------------------------------------------------------------
\section*{Editor's decision}

%-------------------------------------
\textit{Only a few issues need to be solved as the two reviewers suggested. Please revise your manuscript.}

\medskip

\textbf{R:} We have done our best to answer all comments and included the corresponding modifications in the new version of the manuscript. 
To facilitate the identification of new content added in the revised manuscript, we highlight in red the text inclusions/changes.
%-------------------------------------

%%%%%%%%%%%%%%%%%%%
\noindent-----------------------------------------------------------------------------------------
\section*{Reviewer \#1}

%-------------------------------------
\textit{This paper presents an application of unsupervised algorithms for change detection analysis based on stochastic distance and threshold. And the author compared it with CVA method to prove it robust and high performance. However, there are several major comments in the manuscript that must be addressed before considering its publication.}

\medskip

\textbf{R:} We thank the reviewer for all his/her contributions, technical insights, and the time spent on reviewing our manuscript.
%-------------------------------------

\medskip

\medskip
%-------------------------------------
\begin{mybox}{Comment 1}
\textit{The method applied three different types of stochastic distances, while the difference of these distances is not explained, and the reason for different performances when they are applied to remote sensing images is not analyzed.}

\medskip

\textbf{R:} As a consequence of its particular formulation, each stochastic distance plays a distinct way of measuring the similarity between two probability distributions. Therefore, while two objects (i.e., regions/neighborhoods) may share high similarities according to a given stochastic distance, such objects may also demonstrate elevated divergence levels concerning other stochastic distances.
Consequently, we should understand that each stochastic distance may provide a distinct representation of the spatio-temporal dynamics.
In the context of the proposed change detection framework, it is essential to assess different stochastic distance alternatives and quantify the respective performances on mapping the landscape changes.

\medskip

The presented response was adapted and then inserted in the Section 2.2, as showed bellow:
\begin{quotation}

As a consequence of its particular formulation, each stochastic distance plays a distinct way of measuring the similarity between two probability distributions.
Therefore, while two distributions may present a certain level of similarity according to a given stochastic distance, these distributions may also demonstrate divergences concerning other stochastic distances.
Consequently, admitting that distinct stochastic distances may lead to particular conclusions, the identification of the most convenient distance for a specific application is needed.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 2}
\textit{Fig.1 is the key block diagram illustrating the proposed approach. Maybe I missed some key points, but I want to clarify some questions. Each pixel in the neighborhood is the same important for V? And the probability density functions fu and fv are the multivariate Gaussian distributions in experiment?}

\medskip

\textbf{R:} The reviewer's interpretation is right. All the pixels/positions $s$ into the neighborhood $\mathcal{V}_{\rho}(s)$ has the same importance when modeling the probability distribution functions $f_U$ and $f_V$. Such modeling is made by estimating the parameters $\mu$ and $\Sigma$ from $\mathcal{V}_{\rho}(s)$ but considering different instants, resulting in the models $f_U$ and $f_V$. In the experiments, $f_U$ and $f_V$ are both multivariate Gaussian distributions.

\medskip

The following text was added in Section~3:
\begin{quotation}

Equivalently, it may be understood that the information from $V(\mathcal{I}_1,(i,j),\rho)$ and $V(\mathcal{I}_2,(i,j),\rho)$ are employed to estimate the parameters of $f_{\mathbf{U}}$ and $f_{\mathbf{V}}$. 
In this process, the radiometric values concerning the elements/coordinates into the mentioned neighborhoods, regardless of their distance from the reference position $(i,j)$, equal importance on the distributions parameters estimation.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 3}
\textit{The author chose the reference samples manually. And I want to ask that what is the rule and way when you take the samples because I notice that the distribution of most samples at different times are located in similar areas.}

\medskip

\textbf{R:} The ground-reference samples were collected through a careful visual inspection process regarding pairs of instants. On certain occasions, when selecting a ``non-change'' sample that does not change over time, such sample tends to be reconsidered in distinct pairs of instants. In the same sense, targets/regions with frequent spectral change also may be re-employed.

\medskip

An additional explanation concerning the reference samples was inserted in the experiments section:
\begin{quotation}

Under certain conditions, places of ``non-change'' samples that do not change over time tend to be reconsidered in distinct pairs of instants. In the same sense, regions with frequent spectral transition may be re-employed as a ``change'' sample.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 4}
\textit{The author listed some literature in the introduction. I think the authors might have missed some latest and relevant papers in this area. Please comment on the following recent papers in the revised paper. [a] Change detection using deep learning approach with object-based image analysis. [b] An unsupervised domain adaptation approach for change detection and its application to deforestation mapping in tropical biomes. [c] A deep translation (gan) based change detection network for optical and sar remote sensing images}

\medskip

\textbf{R:} Thanks for the suggested papers. We included the respective citations in the introduction section.

\medskip

\begin{quotation}

Recently, diverse supervised approaches have been proposed in the literature to deal with change detection on Remote Sensing data. 
Ref.~\citenum{LiuEA2021} combines object-based image analysis and convolutional neural network concepts to perform %a supervised 
change detection. %approach. 
Ref.~\citenum{LiEA2021} employs generative adversarial nets to propose a 
%supervised 
change detection method that simultaneously analyzes optical and radar data. 
Ref.~\citenum{VegaEA2021} exploits the idea of ``Domain Adaptation'' to train a neural network using information from one source for posterior generic use on the change detection over distinct imagery datasets.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 5}
\textit{The experiments need to be further analyzed. For instance, the intermediate result of Id should be displayed and explained, because it make the experiment more convincing.}

\medskip
\textbf{R:} Thanks for this suggestion.

The distance ($\mathcal{I}_{d}$ -- SDTA) and amplitude ($\mathcal{I}_{a}$ -- CVA) images, related to the change detection maps presented in Section 4.3, were incorporated in the manuscript. Additional discussions regarding such images were also added to the text.

\medskip

\begin{quotation}

As a complementary analysis, Figures~10 to 13 show the distance ($\mathcal{I}_d$) and amplitude ($\mathcal{I}_a$) images referent to the change maps presented in Figures~6 to 9, respectively. Once the values may change as a function of the input data, distance type, and neighborhood size, the depicted maps use a quantile-based color scale.
%
Concerning both methods, it may observe that the Kittler-Illingworth technique tends to threshold the distance/amplitude values around the second and third quantile.
%
In consonance with the change maps (Figs.6-9), the SDTA method expresses the changes through homogeneous regions. The same is not observed for the CVA method, where isolated points of high amplitude values are spread along the study area.

\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\newpage

\vspace{0.25cm}

%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#2}

%-------------------------------------
\begin{mybox}{General Comment}

\textit{The paper proposes a change detection method for Landsat 5/8 images, referred to as Stochastic Distance Thresholding Analysis (SDTA). Please consider the following issues:}

\medskip

\textbf{R:}  We thank the reviewer for all his/her contributions, technical insights and the time spent on reviewing our manuscript.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 1}
\textit{Line 149. Known densities. Please add details, by considering both the sensor/band/resolution characteristics and the neighborhood size.}

\medskip

\textbf{R:} Further details about the densities/neighborhoods were incorporated in Section~3, as pointed below. Details concerning the adopted sensor and neighborhood size ($\rho$) are found in Sections~4.1 and 4.2.

\begin{quotation}
Equivalently, it may be understood that the information from $V(\mathcal{I}_1,(i,j),\rho)$ and $V(\mathcal{I}_2,(i,j),\rho)$ are employed to estimate the parameters of $f_{\mathbf{U}}$ and $f_{\mathbf{V}}$. 
In this process, the radiometric values concerning the elements/coordinates into the mentioned neighborhoods, regardless of their distance from the reference position $(i,j)$, have equal importance on the distributions parameters estimation.
\end{quotation}

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 2}
\textit{Was pansharpening applied? If it was not, add results on pansharpened MS bands.}

\medskip

\textbf{R:} The pansharpening was not applied. One of the main reasons is the absence of a panchromatic band for images acquired in 1999, 2003, and 2008 by the TM sensor. 
However, according to the reviewer's comment, the conclusion section mentions the importance of future studies and analysis with higher resolution images.

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 3}
\textit{Line 208: GitHub link is not working.}

\medskip
\textbf{R:} 
In fact, the repository has been created, but we have not yet provided the code. The codes will be made available immediately after the manuscript acceptance/publishing. The main reason to make it unavailable is due to possible adjusts that may be necessary through the revision process.

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 4 and 5}
\textit{Please better explain the anomalous behavior in Fig.5b.
\\
Fig.10b shows poor performance for low FPR with respect to CVA. Justify that from a methodological point of view or with considerations on the specificity of the 2003-2008 scenario.
}

\medskip

\textbf{R:} For the sake of simplicity, we discussed comments 4 and 5 together. We included the following text in the ``results and discussion'' section.

\medskip

\begin{quotation}

The behavior observed in Figure~\cite{5b} is a consequence of the spectral behavior of the ``2003 image'' (Fig.~\ref{2003}). External factors (like atmospheric, climatic, or even seasonality) may impair the target distinction and the measured stochastic distance/amplitude values.
%
It is essential to attend that the ``2003 image'' affects both analyzed methods (SDTA and CVA) concerning 1999--2003 and 2003--2008 comparisons in terms of results accuracy, with more influence on this latter pair.
%
Furthermore, according to the ROC curves (Fig.14(a) and (b)), it is possible to identify that the ``2003 image'' demands higher False Positive Rate values (approximately 0.195 and 0.3 concerning the pairs 1999--2003 and 2003--2008, respectively) in order to deliver a True Positive Rate above 0.8.
%
In summary, we may consider that the interference mentioned is the cause of large ``blobs of changes'' that overlaps the non-change reference samples and consequently drop the F1-Score values.
%
On the other hand, since the CVA does not consider the neighborhood information, avoiding the insurgence of large ``blobs of changes'' yet producing very noisy results, it is still less susceptible to False Positive errors in this case.
%
However, from a qualitative point of view, we should stress that the change maps delivered by the SDTA method are superior to those from CVA.

\end{quotation}

\end{mybox}
%-------------------------------------

%\vspace{0.3cm}
%
%\medskip
%%-------------------------------------
%\begin{mybox}{Comment 5}
%\textit{Fig.10b shows poor performance for low FPR with respect to CVA. Justify that from a methodological point of view or with considerations on the specificity of the 2003-2008 scenario.}
%
%\medskip
%\textbf{R:} \textcolor{red}{Que explicaćão temos para este caso? (que faz o CVA ganhar da proposta). O que foi mesmo??}
%\end{mybox}
%%-------------------------------------

\vspace{0.3cm}

\bibliographystyle{apalike}
\bibliography{../JARS_DA_review/refsartigo}

\end{document}

### 2021-12-17 tex/paper.wecs/corrections_Aluisio/mudancas_texto_DEC2021.tex
Corrections from Aluisio and rebutal letter

@@
 
+Página 4 Col. 1
+
+* Par2 "Changes patterns" ==> "Patterns of changes"
+
+* Par3 "provides evidence" ==> "provide evidences"
+
+* Par4 "WECS, TAAD and ECS approaches." ==> "WECS, TAAD and ECS."
+
+* Par4 "About the performance of TAAD method," ==> "Regarding TAAD's performance,"
+
+Página 6 Col. 1
+* Fig.7 "Distinct deviation values occurs for m at 1" ==> "Distinct deviations occur at m=1"
+
+* Fig.7 "and 59 (low) in comparison to a baseline around 12." ==> "and 59 (low). Baseline  around 12."
+
+* Fig.8 "from distinct instant m = 1, 30, 59 and the mean image" ==>  "for m = 1, 30, 59 and mean image"
+
 ************************************ IDEIA DE REGRESSAO *****************

### 2021-12-17 tex/paper.wecs/wecs_tgrs.tex
Corrections from Aluisio and rebutal letter

@@
 Smaller ellipses are then added to form $\mathcal{I}^{(3)}$ and $\mathcal{I}^{(4)}$.
-Changes patterns between subsequent images are shown in Figure~\ref{figResSim1}, where white regions (i.e., ``ones'') correspond to changes.
+Patterns of changes between subsequent images are shown in Figure~\ref{figResSim1}, where white regions (i.e., ``ones'') correspond to changes.
 
@@
 Figure~\ref{figROCSimJ} depicts the ROC curves with $J \in \{1,2,3,4,5\}$ as decomposition levels.
-The profiles exhibited by the ROC curves provides evidence that $J$ equal to 2 or 3 lead to the best performances.
+The profiles exhibited by the ROC curves provide evidences that $J$ equal to 2 or 3 lead to the best performances.
 $J=2$ has a slight advantage, since it demands fewer decomposition levels.
@@
 %Compara figA
-The ROC curves shown in Figure~\ref{figROCSimRes} present the performances of WECS, TAAD and ECS approaches. 
+The ROC curves shown in Figure~\ref{figROCSimRes} present the performances of WECS, TAAD and ECS. 
 ECS's low performance shows that the simple swap of the wavelet transform by a ``deviation image'' into the proposed correlation screening pipeline is an inconvenient choice and reinforces the importance of the wavelet smoothing in the context of the proposed method.
-About the performance of TAAD method, a True Positive Ratio of 0.8 is guaranteed when tolerating a False Positive Ratio of approximately 0.4. 
+Regarding TAAD's performance, a True Positive Ratio of 0.8 is guaranteed when tolerating a False Positive Ratio of approximately 0.4. 
 Since WECS provides the same True Positive Ratio under an almost nil false positive ratio, its superior performance is clearly established.
@@
 \includegraphics[width=0.485\textwidth]{../../graphs/LineInstants_1-30-59.pdf}
-\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$. Distinct deviation values occurs for $m$ at $1$ (intermediate), $30$ (high) and $59$ (low) in comparison to a baseline around 12.}\label{F:forest_wecs}
+\caption{Plot of $\vd(m)$ for $m=1,\ldots,85$. Distinct deviations occur at $m=1$ (intermediate), $30$ (high) and $59$ (low). Baseline around 12.}\label{F:forest_wecs}
 \end{figure}
@@
 
-\caption{Images $\vX^{(m)}$ from distinct instant $m=1,30,59$ and the mean image $\overline{\mathcal{I}}$.}
+\caption{Images $\vX^{(m)}$ for $m=1,30,59$ and mean image $\overline{\mathcal{I}}$.}
 \label{F:forest_change_times}

### 2021-12-18 tex/RebutalLetter/RebutalLetter.tex
response to reviewers

@@
 \vspace{2cm}
-\noindent Dear Associate Editor,
+\noindent Dear Editors,
 \bigskip
@@
 We have addressed their comments towards further improving the paper contents and presentation. 
-Below we detail the changes and updates incorporated into the manuscript in this round of review.}
+Below we detail the changes and updates incorporated into the manuscript in this new submission.}
 
@@
 
-\textbf{R:} .
+\textbf{R:} We would like to thank the editor for his comments and for giving us the opportunity to improve the manuscript and submit it again to TGRS.
+Another coauthor has joined our group and we have been able to improve considerably the manuscript's readability and presentation of our contribution.
+We hope the editors and reviewers find the current version significantly more appropriate for the journal.
 %-------------------------------------
@@
 
-\textbf{R:} .
+\textbf{R:} We would like to thank the associate editor for his consideration with our manuscript and for allowing us to re-submit our work.
+The manuscript was carefully reviewed and a thorough analysis was carried out to improve its content.
+The reviewers' comments have been addressed one by one and below you can find a response to each one of them.
 %-------------------------------------
@@
 
-\textbf{R:} .
+\textbf{R:} We thank the reviewer for all his/her contributions, technical insights, and the time spent on reviewing our manuscript.
+The manuscript was re-written and the its content was carefully improve.
 %-------------------------------------
@@
 
-\textbf{R:} .
+\textbf{R:} The reviewer is totally right.
+Even though we wrote PolSAR in the title, the first manuscript did not address the characteristics of this type of image or mentioned related methods from the literature.
+The term PolSAR was removed from the manuscript and now we let it clear that our method is suitable for single channel images.
+In the application with real data, we focus on SAR images, what is mentioned in title in the new manuscript.
 
@@
 
-\textbf{R:} .
+\textbf{R:} These remarks about the proposed method were removed from the manuscript. Additionally, the current version does not refers to PolSAR images anymore.
 
@@
 
-\textbf{R:} .
+\textbf{R:} In the cited paper [i], the authors estimate probability density functions for each instant using the images' wavelet coefficients.
+Next, a time series of curves is estimated using such density functions, which are later used to compute a matrix of differences for pairs of instants.
+The method in the present manuscript is significantly simpler, basically using discrete wavelet transform for images and Pearson correlation, whereas [i] employs functional data analysis tools.
 
@@
 
-\textbf{R:} .
+\textbf{R:} The idea of using correlation screening to detect spatio/temporal changes has not been proposed in the literature yet.
+Works in feature screening basically employ existing and well known correlation measures, but the application of these techniques in the statistical learning literature is an active research topic regarding high-dimensional data.
+We show this idea can also offer useful results for satellite image data as well.
 
@@
 \medskip
-\textbf{R:} .
+\textbf{R:} Log-image referred to the logarithm of a matrix representing a image, but such terminology is not longer used in the current manuscript.
+The epsilon mentioned referred to noise present in an observed image and it did not involve wavelet decomposition.
+We realized such problems with the presentation of the methodology and re-wrote this section.
+We hope it is more clear now.
 
@@
 \medskip
-\textbf{R:} .
+\textbf{R:} We have not been able to find another method that address simultaneously spatial and temporal change detection for a full time series of satellite images.
+Most methods deal only with pairs of images.
+Therefore, we decided to compare our results with the standard approaches cited in the manuscript.
 
@@
 \medskip
-\textbf{R:} .
+\textbf{R:} The comparison with a deep learning method was removed, as well as any mention to PolSAR images.
 
@@
 \medskip
-\textbf{R:} .
+\textbf{R:} The real data application was improved and now we offer much more details about the region analyzed.
+The analyses are once again compared with standard approaches for the type of analyses we make.
 
@@
 
-\textbf{R:}  .
+\textbf{R:} We thank the reviewer for all his/her contributions, technical insights and the time
+spent on reviewing our manuscript. The manuscript was carefully improved and we hope the recent version is suitable to be considered for publication.
 \end{mybox}
@@
 \begin{mybox}{Comment 1}
-\textit{the method section is unclear and partly incomplete. There are equations whose terms are not defined (e.g., Eq 1 and
+\textit{1 - the method section is unclear and partly incomplete. There are equations whose terms are not defined (e.g., Eq 1 and
 that between 7 and 8 without a number). Furthermore, the procedure for Wavelet multiresolution analysis is not
@@
 \medskip
-\textbf{R:} .
+\textbf{R:} Presentation of the methodology was carefully revised and now all terms and methods are introduced before being referred to.
+We present a brief introduction of wavelet decomposition and feature screening that might be enough to understand the following parts of the manuscript.
+Concerning the approximation level, it is indeed one of the decomposition levels, but we meant that only coefficients related to coarse details are used.
+This is explained with more details in the current version.
 \end{mybox}
@@
 \begin{mybox}{Comment 2}
-\textit{synthetic test is confusing. Figure 1 shows a sequence of 4 synthetic images with changes appearing consecutively.
+\textit{2 - synthetic test is confusing. Figure 1 shows a sequence of 4 synthetic images with changes appearing consecutively.
 The expectation from the presented method and promises made in the abstract and introduction is that 4 corresponding
@@
 
-\textbf{R:} .
+\textbf{R:} The four images are now presented separately in Figure 2 of the recent version.
+The total changes are showed in Figure 3 of this version.
+The idea is that, if only the first and last images were used (images 1 and 4, respectively), changes in instants between these times would be missed.
+The proposed method is suited to detect this type of change.
+Explanations in the section about the simulated images were improved and we added more details on how they were generated.
 \end{mybox}
@@
 \begin{mybox}{Comment 3}
-\textit{the real data set is not introduced carefully, and it is not clear what datasets from what sensor at what resolution and
+\textit{3 - the real data set is not introduced carefully, and it is not clear what datasets from what sensor at what resolution and
 quality are used. How the real data is proceeded and prepared for this analysis, etc. In Figures 7-9 most of the panels
@@
 
-\textbf{R:} .
+\textbf{R:} Details about the images were included in the text: type of satellite image; type of sensor; resolution; and description of pre-processing steps used before applying the proposed method.
+We also removed similar figures that were not adding new information to the analysis.
+The Figures with overlapping curves are no longer used, and different details about the region studied are presented now, including change maps in Figure 6c and Figure 10.
+In particular, Figure 6c is a validation change map that is used to compute accuracy methods presented in Table 1 of the recent manuscript version.
 \end{mybox}
@@
 
+%-------------------------------------
+%%%%%%%%%%%%%%%%%%%
+\section*{Reviewer \#3}
+
+%-------------------------------------
+\begin{mybox}{General Comment}
+
+\textit{The manuscript presents a methodology to detect changes in the wavelet domain using multitemporal SAR
+data. Although all the mathematical formalism looks sound and the results on simulated data are promising,
+the manuscript is still lacking of some important parts. Most importantly, a proper analysis of results using
+real data should be done. Real data are indeed used, but their analysis has little to do with validating change
+detection and it is more about showing how the new observables look like. More information should also be
+added to the simulations. After the addition of these information and a proper validation the paper should be
+publishable in TGRS.}
+\medskip
+
+\textbf{R:} We thank the reviewer for all his/her contributions, technical insights and the time
+spent on reviewing our manuscript.
+The analysis of real data was improved and now we offer more details about the images used and present a validation change image to measure the accuracy of the proposed method.
+We also provide now a better description of how the simulated images were generated and analyzed.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 1}
+\textit{Data used: In the entire paper there is no mention of the type of data used. Are this Sentinel-1 GRD images?
+\textcolor{red}{Also please consider acknowledging ESA for the use of data.}}
+
+\medskip
+\textbf{R:} The data was obatined from the SAR sensor onboard the Sentinel-1 satellite.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 2}
+\textit{Simulations: Some more details should be provided to describe how the simulations were performed.
+I assume these are Monte Carlo simulations where an Additive Gaussian Noise is added to the image (representing
+the log intensity). This is not ideal, since it opens the question on the noise level you simulate (i.e. the energy
+of the noise). Why did you selected that energy for the noise? Ideally, you would simulate the complex pixel
+as a Circular Complex Gaussian random variable, then take the intensity (maybe use multi-look) and then the
+log. Please justify the value of noise energy you selected or please repeat the simulations starting from complex
+pixels.}
+
+\medskip
+\textbf{R:} The simulated images were indeed generated with additive Gaussian noise.
+We included more details about how the simulated images were generated and what level of noise was used.
+The signal to noise to ration considered was chosen in order to not let the true changes too apparent in the observed images (especially image 4), as can be seen in Figure 2.
+Such level also offers a harder scenario than what is analyzed in the application with real data.
+Such explanations are presented in the manuscript now.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 3}
+\textit{Figure 2: Is (a) a validation mask where each pixel that changes at any time is flagged as 1? Please add
+more details on the content of subplot (a).}
+
+\medskip
+\textbf{R:} Yes, ones (white) correspond to locations were change happened at any instant, whereas zero (black) correspond to places where no change takes place.
+This information is given now in the recent version and the figure now is on Figure 3a.
+We also present more details about the content of other images as well.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 4}
+\textit{Timeseries vs pairwise: At the moment, it is not clear what is the advantage of using the full time series
+compared to apply a pairwise approach using first and last image. In the simulation, is this a better choice in
+detecting changes that applying a single pairwise change detector between image 1 and image 4? So not do the
+accumulation, but just use first and last change detection for comparison.}
+
+\medskip
+\textbf{R:} The simulated images illustrate the point in considering the full time series: if only images 1 and 4 were considered in a pairwise analysis, the changes that happened between these instants would be missing.
+The proposed method is able to detect this type of change, as can be seen in Figure 3b.
+Additionally, we also look for instants when a relevant change might have happened.
+The method proposed can be used to detect this type of change as well through analyzing the the measures $\mathbf{d}^{(1)},\ldots,\mathbf{d}^{(n)}$.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 5}
+\textit{Real data info and preprocessing: There is absolutely no information about what satellite data are used. I
+assume this are Sentinel-1 GRD, but please provide info on that. Also there is no info on what pre-processing
+you use. \textcolor{red}{Did you do calibration, terrain correction, noise removal, geocoding?}}
+
+\medskip
+\textbf{R:} We included information about the images used in the real data application. They are indeed Sentinel-1 images.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 6}
+\textit{Real data validation: The test with real data is very poor and should be improved. For instance, there is
+no information on what changed in the scene. Also there is no detection mask obtained from the analysis and
+no quantitative results (e.g. a ROC curve). These informations must be provided in order to make this paper
+publishable.}
+
+\medskip
+\textbf{R:} A validation image with change locations is used in the recent version and can be seen in Figure 6c.
+This image is used to compute accuracy measures of the proposed method and standard approaches.
+A change map is also provided using two different types of thresholding (Otsu and Kittler-Illingworth).
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 7}
+\textit{Figure 4 to 6: These are temporal trends, but there is no information on what the reader is supposed to see
+in these trends. If you take the backscattering (or any other measure, e.g. kurtosis) it will always show a trend...
+why are the trends coming from your method better than any other trend one could get?\\
+Each plot I assume is referred to a region. What is the region you are looking at (forest, river, agriculture)?
+How big is the region? Are we supposed to have changes in these regions? If so, which changes and when they
+happen?}
+
+\medskip
+\textbf{R:} These figures are not presented in the recent version, and only an overall change measure is presented in Figure 7.
+An increasing trend in this plot means that the corresponding images are changing with respect to the mean image.
+A decreasing trend indicates that images are becoming more similar to the mean image.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 8}
+\textit{Figure 7 to 9: What do all these colour lines in the plots refer too? Are different different regions?}
+
+\medskip
+\textbf{R:} Yes, each color correspond to a different location (pixel).
+The goal was to show that locations flagged as change point have squared deviations that behave similarly as the overall measure.
+Since these Figures were somewhat confusing, we decided not to show them in the recent version.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 9}
+\textit{Figure 10 and 11: Do you have information on the area that changed? If you want to use R (d) for detection,
+and low correlation are flagged as changing areas, doe s this mean that the river (permanent water body) is the
+most changing part?}
+
+\medskip
+\textbf{R:} We present now a change map that is used to evaluate the methods.
+When using $\mathbf{R}$, high correlations are the ones flagged as corresponding to changing areas.
+This is because the correlation is computed between a local change measure and an overall change measure.
+Therefore, the river is not detected as changing region by the proposed method.
+\end{mybox}
+%-------------------------------------
+
+\medskip
+%-------------------------------------
+\begin{mybox}{Comment 10}
+\textit{P.2, c.1, l.31: Please change "practive" into "practice".}
+
+\medskip
+\textbf{R:} The word was corrected in the recent version.
+\end{mybox}
 %-------------------------------------

### 2021-12-18 tex/paper.wecs/wecs_tgrs.tex
response to reviewers

@@
 
-Change detection process using SAR images is a challenging task due to the the high-dimensionality of multi-temporal datasets associated with the speckle noise.
-In this context, the use of Wavelets rises as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms.
+Change detection process using SAR images is a challenging task due to the high-dimensionality of multi-temporal datasets associated with the speckle noise.
+In this context, the use of Wavelets comes as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms.
 
@@
 
-In order to apply the analyzed methods, the dual-polarized images were combined into a single-band representation considering
+In order to apply the analyzed methods, the dual-polarized images were combined into a single-band representation considering the so-called "span" image
 $\mathcal{I}_{kl}^{(m)} = \sqrt{(\mathrm{VV}_{kl}^{(m)})^2+(\mathrm{VH}_{kl}^{(m)})^2}$, with $\mathrm{VV}$ and $\mathrm{VH}$ representing the available polarizations.
@@
 For the sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(30)}$ and $\vX^{(59)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_change_times}.
-While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rises in $\vX^{(30)}$.
+While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rise in $\vX^{(30)}$.
 Reversely, $\vX^{(59)}$ offers low contribution for identifying change regions.
@@
 Lastly, the computational run-time of WECS, TAAD and ECS were, respectively, \SI{143.45}{\second}, \SI{7.89}{\second} and \SI{110.08}{\second}.
-Despite expending a longer execution time, its worth mention that WECS' running time is not excessive in the context of remote sensing image processing, beyond providing systematically more accurate results.
+Despite expending a longer execution time, it is worth mention that WECS' running time is not excessive in the context of remote sensing image processing, beyond providing systematically more accurate results.
 

### 2021-12-19 tex/RebutalLetter/RebutalLetter.tex
updates in the rebutal letter

@@
 
-\textcolor{black}{We would like to thank the reviewers for the valuable comments and the time spent on checking our manuscript. 
+\textcolor{black}{We would like to thank you and the reviewers for the valuable comments and the time spent on checking our manuscript. 
 We have addressed their comments towards further improving the paper contents and presentation. 
@@
 \textbf{R:} We would like to thank the editor for his comments and for giving us the opportunity to improve the manuscript and submit it again to TGRS.
-Another coauthor has joined our group and we have been able to improve considerably the manuscript's readability and presentation of our contribution.
+Another coauthor has joined our group and we have been able to improve the manuscript's readability.
+We now provide more details in the analyses, show different figure with more informative content and the presentation of our contribution has improved considerably.
 We hope the editors and reviewers find the current version significantly more appropriate for the journal.
@@
 
-\textbf{R:} We would like to thank the associate editor for his consideration with our manuscript and for allowing us to re-submit our work.
+\textbf{R:} We would like to thank the associate editor for his consideration with our manuscript and for allowing us to resubmit our work.
 The manuscript was carefully reviewed and a thorough analysis was carried out to improve its content.
-The reviewers' comments have been addressed one by one and below you can find a response to each one of them.
+The reviewers' comments have been addressed and below you can find an explanation to each one of them.
 %-------------------------------------
@@
 
-\textbf{R:} We thank the reviewer for all his/her contributions, technical insights, and the time spent on reviewing our manuscript.
-The manuscript was re-written and the its content was carefully improve.
+\textbf{R:} We thank the reviewer for all his/her contributions, technical insights, and the time spent on reviewing our manuscript. It was re-written and its content was carefully improved.
 %-------------------------------------
@@
 The term PolSAR was removed from the manuscript and now we let it clear that our method is suitable for single channel images.
-In the application with real data, we focus on SAR images, what is mentioned in title in the new manuscript.
+In the application with real data we focus on SAR images, which is the only type of image mentioned in title now.
 
@@
 Works in feature screening basically employ existing and well known correlation measures, but the application of these techniques in the statistical learning literature is an active research topic regarding high-dimensional data.
-We show this idea can also offer useful results for satellite image data as well.
+We show that this idea can offer useful results for satellite image data analysis as well.
 
@@
 \medskip
-\textbf{R:} Log-image referred to the logarithm of a matrix representing a image, but such terminology is not longer used in the current manuscript.
+\textbf{R:} Log-image referred to the logarithm of a matrix representing an image, but such terminology is no longer used in the current manuscript.
 The epsilon mentioned referred to noise present in an observed image and it did not involve wavelet decomposition.
-We realized such problems with the presentation of the methodology and re-wrote this section.
+We realized that such problems were undermining the presentation of our methodology.
+Therefore, we decided to rewrite this section.
 We hope it is more clear now.
@@
 \textbf{R:} Presentation of the methodology was carefully revised and now all terms and methods are introduced before being referred to.
-We present a brief introduction of wavelet decomposition and feature screening that might be enough to understand the following parts of the manuscript.
+We present a brief introduction of wavelet decomposition and feature screening that might be enough to understand further parts of the manuscript.
 Concerning the approximation level, it is indeed one of the decomposition levels, but we meant that only coefficients related to coarse details are used.
@@
 The idea is that, if only the first and last images were used (images 1 and 4, respectively), changes in instants between these times would be missed.
-The proposed method is suited to detect this type of change.
-Explanations in the section about the simulated images were improved and we added more details on how they were generated.
+The proposed method is suited to detect this type of change only observed when considering the full time series of images.
+Explanations were improved in the section about the simulated images and we added more details on how they were generated.
 \end{mybox}
@@
 
-\textbf{R:} Details about the images were included in the text: type of satellite image; type of sensor; resolution; and description of pre-processing steps used before applying the proposed method.
+\textbf{R:} Details about the images were included in the text: type of satellite image; type of sensor; resolution; \textcolor{red}{and description of pre-processing steps used before applying the proposed method}.
 We also removed similar figures that were not adding new information to the analysis.
-The Figures with overlapping curves are no longer used, and different details about the region studied are presented now, including change maps in Figure 6c and Figure 10.
-In particular, Figure 6c is a validation change map that is used to compute accuracy methods presented in Table 1 of the recent manuscript version.
+The figures with overlapping curves are no longer used, and different details about the region studied are presented now, including change maps in Figure 6c and Figure 10.
+In particular, Figure 6c is a validation change map that is used to compute accuracy measure of the methods considered.
+These measures are given in Table 1 of the recent manuscript version.
 \end{mybox}
@@
 
+\newpage
 %-------------------------------------
@@
 The signal to noise to ration considered was chosen in order to not let the true changes too apparent in the observed images (especially image 4), as can be seen in Figure 2.
-Such level also offers a harder scenario than what is analyzed in the application with real data.
+Such level of noise is also more challenging to analyze than what is observed in the application with real data.
 Such explanations are presented in the manuscript now.
@@
 \medskip
-\textbf{R:} Yes, ones (white) correspond to locations were change happened at any instant, whereas zero (black) correspond to places where no change takes place.
+\textbf{R:} Yes, ones (white) correspond to locations were change happened at any instant, whereas zeros (black) correspond to places where no change takes place.
 This information is given now in the recent version and the figure now is on Figure 3a.
-We also present more details about the content of other images as well.
+We also present more details about the content of the remaining images as well.
 \end{mybox}
@@
 \medskip
-\textbf{R:} We included information about the images used in the real data application. They are indeed Sentinel-1 images.
+\textbf{R:} We included more detailed information about the images used in the real data application. They are indeed Sentinel-1 images.
 \end{mybox}
@@
 \medskip
-\textbf{R:} Yes, each color correspond to a different location (pixel).
-The goal was to show that locations flagged as change point have squared deviations that behave similarly as the overall measure.
-Since these Figures were somewhat confusing, we decided not to show them in the recent version.
+\textbf{R:} Yes, each color corresponds to a different location (pixel).
+The goal was to show that locations flagged as change point have squared deviations that behave similarly as the overall change measure.
+Since these figures were somewhat confusing, we decided not to show them in the recent version.
 \end{mybox}

### 2021-12-19 tex/paper.wecs/wecs_tgrs.tex
updates in the rebutal letter

@@
 Most methods used for change detection analysis can be classified either as supervised (training data is used to set up the method) or unsupervised (fully data-driven techniques).
-We shall focus in this work on unsupervised approaches, whose examples in the literature include the works of \cite{bruzzone2000automatic,celik2010change,quin2014mimosa,saha2020change,NegriEA2021,NegriFrery2021}.
+We focus in this work on unsupervised approaches, whose examples in the literature include the works of \cite{bruzzone2000automatic,celik2010change,quin2014mimosa,saha2020change,NegriEA2021,NegriFrery2021}.
 Many other proposals vary in their motivations as well as in their applicability.
@@
 Compressed projection and image fusion are employed by \cite{hou2014unsupervised}.
-Deep learning by slow feature analysis for change detection is the subject of \cite{du2019unsupervised}.
+Deep slow feature analysis for change detection is the subject of \cite{du2019unsupervised}.
 \cite{chen2020change} proposes a change detection method driven by adaptive parameter estimation.
@@
 Change detection process using SAR images is a challenging task due to the high-dimensionality of multi-temporal datasets associated with the speckle noise.
-In this context, the use of Wavelets comes as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms.
+In this context, the use of Wavelets rises as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms.
 
@@
 
-The application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$ results in a matrix of the so called approximation wavelet coefficients $\vX^{(m)}$, a smoothed version of $\mathcal{I}^{(m)}$
+The application of this wavelet transform to $\mathcal{I}^{(m)}$ at some appropriate resolution level $J\geq 1$ results in a matrix of the so called approximation wavelet coefficients $\vX^{(m)}$ of $\mathcal{I}^{(m)}$
 defined on the same support $\mathcal{S}$. The higher $J\in\{1,2,\ldots, \floor{ \log_2( \min\{u,v\} ) } \}$ is, the smoother $\vX^{(m)}$ gets.
@@
 For the sake of comparison, the wavelet representation at each mentioned instant (i.e., $\vX^{(1)}$, $\vX^{(30)}$ and $\vX^{(59)}$) and the mean image (i.e., $\overline{\mathcal{I}}$) are exhibited in Figure~\ref{F:forest_change_times}.
-While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes rise in $\vX^{(30)}$.
+While $\vX^{(1)}$ shares similarities with $\overline{\mathcal{I}}$, evident changes increase in $\vX^{(30)}$.
 Reversely, $\vX^{(59)}$ offers low contribution for identifying change regions.

### 2021-12-20 tex/RebutalLetter/RebutalLetter.tex
ajustes menores na rebuttal (20dez21)

@@
 
-\newpage
+\medskip
+
+\medskip
+
+%\newpage
 
@@
 
-\newpage
+\medskip
+\medskip
+\medskip
+
+%\newpage
 
@@
 \begin{mybox}{Comment 1}
-\textit{1 - the method section is unclear and partly incomplete. There are equations whose terms are not defined (e.g., Eq 1 and
+\textit{The method section is unclear and partly incomplete. There are equations whose terms are not defined (e.g., Eq 1 and
 that between 7 and 8 without a number). Furthermore, the procedure for Wavelet multiresolution analysis is not
@@
 \begin{mybox}{Comment 2}
-\textit{2 - synthetic test is confusing. Figure 1 shows a sequence of 4 synthetic images with changes appearing consecutively.
+\textit{Synthetic test is confusing. Figure 1 shows a sequence of 4 synthetic images with changes appearing consecutively.
 The expectation from the presented method and promises made in the abstract and introduction is that 4 corresponding
@@
 \begin{mybox}{Comment 3}
-\textit{3 - the real data set is not introduced carefully, and it is not clear what datasets from what sensor at what resolution and
+\textit{The real data set is not introduced carefully, and it is not clear what datasets from what sensor at what resolution and
 quality are used. How the real data is proceeded and prepared for this analysis, etc. In Figures 7-9 most of the panels
@@
 
-\textbf{R:} Details about the images were included in the text: type of satellite image; type of sensor; resolution; \textcolor{red}{and description of pre-processing steps used before applying the proposed method}.
+\textbf{R:} Details about the images were included in the text:
+the acquisition satellite/sensor; spatial resolution; and the image size.
+% type of satellite image; type of sensor; resolution; \textcolor{red}{and description of pre-processing steps used before applying the proposed method}. 
  We also removed similar figures that were not adding new information to the analysis.
-The figures with overlapping curves are no longer used, and different details about the region studied are presented now, including change maps in Figure 6c and Figure 10.
-In particular, Figure 6c is a validation change map that is used to compute accuracy measure of the methods considered.
-These measures are given in Table 1 of the recent manuscript version.
+The figures with overlapping curves are no longer used, and different details about the region studied are presented now, including change maps in Figures 6(c) and 10.
+In particular, Figure 6(c) is a validation/reference change map that is used to compute accuracy measure of the methods considered.
+These measures are given in Table 1 of the revised manuscript.% recent manuscript version.
 \end{mybox}
@@
 
-\newpage
+%\newpage
+\medskip
+\medskip
+\medskip
+
 %-------------------------------------
@@
 \textbf{R:} The data was obatined from the SAR sensor onboard the Sentinel-1 satellite.
+\textcolor{green}{Verdade... vamos citar a fonte, apenas foi dito o sensor.}
 \end{mybox}
@@
 \textbf{R:} We included more detailed information about the images used in the real data application. They are indeed Sentinel-1 images.
+
+\textcolor{green}{Não houve pre-processamento, mas é interessante incluir isso no texto. Eu farei isso.}
 \end{mybox}
@@
 \medskip
-\textbf{R:} A validation image with change locations is used in the recent version and can be seen in Figure 6c.
+\textbf{R:} A validation image with change locations is used in the recent version and can be seen in Figure 6(c).
 This image is used to compute accuracy measures of the proposed method and standard approaches.
-A change map is also provided using two different types of thresholding (Otsu and Kittler-Illingworth).
+A change map is also provided using two different types of thresholding algorithms (Otsu and Kittler-Illingworth).
 \end{mybox}

### 2021-12-20 tex/RebutalLetter/RebutalLetter.tex
ajustes++ (20dez21)

@@
 \begin{mybox}{Comment 1}
-\textit{Data used: In the entire paper there is no mention of the type of data used. Are this Sentinel-1 GRD images?
-\textcolor{red}{Also please consider acknowledging ESA for the use of data.}}
+\textit{Data used: In the entire paper there is no mention of the type of data used. Are this Sentinel-1 GRD images? Also please consider acknowledging ESA for the use of data.}
 
 \medskip
-\textbf{R:} The data was obatined from the SAR sensor onboard the Sentinel-1 satellite.
-\textcolor{green}{Verdade... vamos citar a fonte, apenas foi dito o sensor.}
+\textbf{R:} The reviewer is correct, the Sentinel-1 GRD images were employed in this research.
+The revised text includes a description of this sensor/dataset.
+Acknowledgments to ESA were also included at the end of the manuscript.
+Thank you.
 \end{mybox}
@@
 assume this are Sentinel-1 GRD, but please provide info on that. Also there is no info on what pre-processing
-you use. \textcolor{red}{Did you do calibration, terrain correction, noise removal, geocoding?}}
+you use. Did you do calibration, terrain correction, noise removal, geocoding?}
 
 \medskip
-\textbf{R:} We included more detailed information about the images used in the real data application. They are indeed Sentinel-1 images.
+\textbf{R:} As mentioned (in comment \#1), we added more details about the adopted images in the experiment with actual remote sensing data.
 
-\textcolor{green}{Não houve pre-processamento, mas é interessante incluir isso no texto. Eu farei isso.}
+% used in the real data application. They are indeed Sentinel-1 images.
+
+%\textcolor{green}{Não houve pre-processamento, mas é interessante incluir isso no texto. Eu farei isso.}
 \end{mybox}

### 2021-12-20 tex/paper.wecs/wecs_tgrs.tex
ajustes++ (20dez21)

@@
 
-Each image contains the amplitude signal backscatters relative to VV and VH polarizations, a spatial resolution of \SI{10}{\meter} and support of $1538 \times 1556$ pixels wide.
+Each image contains the amplitude signal backscatters relative to VV and VH polarizations, a spatial resolution of \SI{10}{\meter} and a support of $1538 \times 1556$ pixels wide in ground range projection.
+Specifically, these images where obtained from the \textit{Sentinel-1 SAR GRD} dataset maintained by the European Space Agency \cite{ESA2021}.
+%
 Figures~\ref{figImgSentinel2015} and \ref{figImgSentinel2017} depict the first and last images of the time series, where it is possible to compare and identify some landscape changes.
@@
 % use section* for acknowledgment
-%\section*{Acknowledgment}
+\section*{Acknowledgment}
+
+The authors would like to thank the European Space Agency (ESA) for providing the Sentinel-1 data used in this research.
 

### 2021-12-21 tex/RebutalLetter/ResponseLetter.tex
Change rebuttal for response

%\documentclass[smallextended]{svjour3}
\documentclass[11pt]{report}
%\pagestyle{empty}

%\usepackage[latin1]{inputenc}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage[many,breakable]{tcolorbox}
\usepackage{natbib}

%\usepackage{xr}
%\externaldocument{NegriEA_AssemblyUCD_R1}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\newtcolorbox{mybox}[1]{
colback = white!5!white,
colframe = white!75!black,
fonttitle=\bfseries,
title=#1,breakable
}

\usepackage{graphicx}
\usepackage{subfigure}

%\newtcolorbox{mybox}[1]{
%colback = gray!5!white,
%colframe = gray!75!black,
%fonttitle=\bfseries,
%title=#1
%}

%%%% determina o tamanho da folha
\usepackage[pdftex]{geometry}
\geometry{a4paper,left=2.8cm,right=2.8cm,top=1.5cm,bottom=1.5cm,twoside} % tamanho da pagina

%%% --------------------------------

%-----------------------------------------------------------------------------------
\begin{document}

\begin{center}
\large{\textbf{Response Letter}}

\vglue 0.3cm

\huge{Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images}
\end{center}

%\author
\begin{center}
\textbf{Rodney~V.~Fonseca, Rog\'{e}rio~G.~Negri, Alu\'{i}sio~Pinheiro and Abdourrahmane~Atto}
\end{center}

\date{\today}

\vspace{2cm}
\noindent Dear Editors,
\bigskip

\textcolor{black}{We would like to thank you and the reviewers for the valuable comments and the time spent on checking our manuscript. 
We have addressed their comments towards further improving the paper contents and presentation. 
Below we detail the changes and updates incorporated into the manuscript in this new submission.}

\medskip
\noindent Yours sincerely,

\begin{flushright}
\noindent The authors.
\end{flushright}
%-----------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%
\noindent---------------------------------------------------------------------------------------
\section*{Editor's decision}

%-------------------------------------
\textit{
Dear Dr. Fonseca:\\
I am sorry to inform you that your manuscript:\\
Wavelet Spatio-Temporal Change Detection on multi-temporal POLSAR images (TGRS-2021-01131)\\
has been carefully reviewed by the Editorial Board of the IEEE Transactions on Geoscience and Remote Sensing
(TGRS) and has not been recommended for publication at this time. It is felt that, while there may be sufficient original
content in the manuscript to warrant publication, it still requires a more substantial amount of work than is consistent with
a ‘Major Revision’ decision. For this reason, we encourage you to consider revising the manuscript in response to the
comments given below and submitting it as a new submission to TGRS. If you do decide to resubmit the paper, please
include a point-by-point response to the comments of the reviewers along with the new paper in order to expedite its
review. Use the "Submit a Resubmission" link in your author center when you submit the new manuscript.\\
Your resubmission deadline is 20-Jan-2022.\\
Below are summary comments from the TGRS Editorial Board for your information.\\
Thank you for submitting your manuscript to TGRS.\\
Sincerely,\\
Dr. Simon Yueh\\
Editor, IEEE Trans. on Geoscience and Remote Sensing
}

\medskip

\textbf{R:} We would like to thank the editor for his comments and for giving us the opportunity to improve the manuscript and submit it again to TGRS. Another coauthor has joined our group and we have been able to improve the manuscript's readability. We now provide more details in the analyses, show different figures with more informative content and the presentation of our contribution has improved considerably. We hope the editors and reviewers find the current version more appropriate for the journal.
%-------------------------------------

%%%%%%%%%%%%%%%%%%%
\noindent---------------------------------------------------------------------------------------
\section*{Associate Editor}

%-------------------------------------
\textit{Dear Authors,\\
I regret to inform you that I cannot recommend publication of your paper in the IEEE Transactions on Geoscience and
Remote Sensing journal. This rejection is based upon the enclosed comments of the reviewers.
In case you decide to resubmit the paper to this journal, I suggest that you take into account all the reviewers’ comments
to improve your manuscript.
Thank you for considering TGARS for the publication of your manuscript. We welcome the future submission of your
work to the journal.\\
Sincerely,}

\medskip

\textbf{R:} We would like to thank the associate editor for his consideration with our manuscript and for allowing us to resubmit our work. The manuscript was carefully reviewed and a thorough analysis was carried out to improve its content. The reviewers' comments have been addressed and below you can find an explanation to each one of them.
%-------------------------------------

\medskip

\medskip

\medskip

%\newpage

\vspace{0.25cm}

%%%%%%%%%%%%%%%%%%%
\noindent---------------------------------------------------------------------------------------
\section*{Reviewer \#1}

%-------------------------------------
\textit{This paper presents a wavelet decomposition based change detection method for POLSAR images. The method seems original but the paper is poorly written. The detailed comments are listed as follows.}

\medskip

\textbf{R:} We thank the reviewer for all his/her contributions, technical insights, and the time spent on reviewing our manuscript. It was rewritten and its content was carefully improved.
%-------------------------------------

\medskip

\medskip
%-------------------------------------
\begin{mybox}{Comment 1}
\textit{This paper lacks a thorough review of POLSAR image change detection methods. What are the challenges of
POLSAR image change detection? What are the advantages and defects of existing methods? What’s the superiority of
the proposed method over existing ones? What are the contributions of this paper to remote sensing community? There
are many works about change detection in POLSAR images. This paper only addresses the method without necessary
comparison.}

\medskip

\textbf{R:} The reviewer is totally right. Even though we wrote PolSAR in the title, the first manuscript did not address the characteristics of this type of image or mentioned related methods from the literature. The term PolSAR was removed from the manuscript and now we let it clear that our method is suitable for single channel images. We focus on SAR images in the application with real data. This is the only type of image mentioned in title now.

\medskip

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 2}
\textit{The authors claim that “it is sparse, very fast and scalable. Finally, it is easily adapted to be updatable”. However, the
change detection in POLSAR images has been significantly developed and there are many excellent methods. Many of
them are also fast and scalable. Without any necessary comparison, how to demonstrate that “it is very fast and scalable”? What’s the novelty of the proposed method?}

\medskip

\textbf{R:} These remarks about the proposed method were removed from the manuscript. Additionally, the current version does not refer to PolSAR images anymore.

\medskip

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 3}
\textit{What’s the relationship between this paper and [i]? [i] also uses wavelet transformation for change detection.
[i] Fonseca, R. , Ludwig, G. , Montoril, M. , Pinheiro, A. . (2020). Nonparametric methods for detecting change in Multitemporal SAR-PolSAR Satellite Data. arXiv: 2001.05764.}

\medskip

\textbf{R:} In the cited paper [i], the authors estimate probability density functions for each instant using the images' wavelet coefficients. Next, a time series of curves is estimated using such density functions, which are later used to compute a matrix of differences for pairs of instants. The method in the present manuscript is significantly simpler, basically using discrete wavelet transform for images and Pearson correlation, whereas [i] employs functional data analysis tools.

\medskip

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 4}
\textit{Methodology describes the proposed method in detail, but what’s new in this paper? All the operators are existing or
intuitive.}

\medskip

\textbf{R:} The idea of using correlation screening to detect spatio/temporal changes has not been proposed in the literature yet. Works in feature screening basically employ existing and well known correlation measures, but the application of these techniques in the statistical learning literature is an active research topic regarding high-dimensional data. We show that this idea can offer useful results for satellite image data analysis as well.

\medskip

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 5}
\textit{The method is not clearly described. For example, what’s log-image? What’s the epsilon in Eq. (1). Eq. (1) seems
problematic. It denotes wavelet decomposition?}

\medskip
\textbf{R:} Log-image referred to the logarithm of a matrix representing an image, but such terminology is no longer used in the current manuscript. The epsilon mentioned referred to noise present in an observed image and it did not involve wavelet decomposition. We realized that such problems were undermining the presentation of our methodology. Therefore, we decided to rewrite this section and we hope it is more clear now.

\medskip

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 6}
\textit{Without comparison with recent change detection methods in experiments, how to demonstrate the advantages of the
proposed method? For example, if there is a method that is faster and performs better than existing ones, why I have to
use the proposed one?}

\medskip
\textbf{R:} We have not been able to find another method that address simultaneously spatial and temporal change detection for a full time series of satellite images. Most methods deal only with pairs of images. Therefore, we decided to compare our results with the standard approaches cited in the manuscript.

\medskip

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 7}
\textit{The authors claims that “Its performance is similar to a deep learning feature extraction method”. However, the deep
learning feature extraction method is proposed for image denoising and it is not specially designed for POLSAR images.
It is obviously unfair to compare the proposed method with it. Why not compare the proposed method with recent
POLSAR image change detection methods? The proposed one cannot outperform them?}

\medskip
\textbf{R:} The comparison with a deep learning method was removed, as well as any mention to PolSAR images.

\medskip

\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------
\begin{mybox}{Comment 8}
\textit{For real data, there are only self comparisons without any comparison with existing excellent methods. Therefore, the
conclusions may be not correct.}

\medskip
\textbf{R:} The real data application was improved and now we offer much more details about the region analyzed. The analyses are once again compared with standard approaches for the type of analyses we make.

\medskip

\end{mybox}
%-------------------------------------

\medskip
\medskip
\medskip

%\newpage

\vspace{0.25cm}

%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#2}

%-------------------------------------
\begin{mybox}{General Comment}

\textit{Fonseca et al. present a wavelet-based approach to detect spatiotemporal change points on multi-temporal Polarimetric
SAR. The approach is potentially interesting and might be of use for a broader community. However, it suffered from
several shortcomings, which prevents me from recommending the publication of the manuscript in its current form. In the
following o elaborate my concerns, and I hope authors find them useful.}

\medskip

\textbf{R:} We thank the reviewer for all his/her contributions, technical insights and the time
spent on reviewing our manuscript. The manuscript was carefully improved and we hope the recent version is suitable to be considered for publication.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 1}
\textit{The method section is unclear and partly incomplete. There are equations whose terms are not defined (e.g., Eq 1 and
that between 7 and 8 without a number). Furthermore, the procedure for Wavelet multiresolution analysis is not
described. Although, provided in the cited literature, it is crucial to carefully layout the procedure used to create the input
to the change detection algorithm. Also, statements like "J-th approximation level" are unclear. I suspect the author
meant to say "J-th decomposition level."}

\medskip
\textbf{R:} Presentation of the methodology was carefully revised and now all terms and methods are introduced before being referred to. We present a brief introduction of wavelet decomposition and feature screening that might be enough to understand further parts of the manuscript. Concerning the approximation level, it is indeed one of the decomposition levels, but we meant that only coefficients related to coarse details are used. This is explained with more details in the current version.
\end{mybox}
%-------------------------------------

\medskip

%-------------------------------------

\begin{mybox}{Comment 2}
\textit{Synthetic test is confusing. Figure 1 shows a sequence of 4 synthetic images with changes appearing consecutively.
The expectation from the presented method and promises made in the abstract and introduction is that 4 corresponding
change maps can be obtained characterizing the changes made in the synthetic datasets. However, Figure 2 shows a
total change and some other maps, which is unclear how they were obtained or what is the relevance to the presented
method and promises of the paper.}

\medskip

\textbf{R:} The four images are now presented separately in Figure 2 of the recent version. The total changes are showed in Figure 3 of this version. The idea is that, if only the first and last images were used (images 1 and 4, respectively), changes in instants between these times would be missed. The proposed method is suited to detect this type of change only observed when considering the full time series of images. Explanations were improved in the section about the simulated images and we added more details on how they were generated.
\end{mybox}
%-------------------------------------

%-------------------------------------

\vspace{0.3cm}

\medskip
%-------------------------------------

\begin{mybox}{Comment 3}
\textit{The real data set is not introduced carefully, and it is not clear what datasets from what sensor at what resolution and
quality are used. How the real data is proceeded and prepared for this analysis, etc. In Figures 7-9 most of the panels
(e.g., c-j) show many overlapping curves that color-coded without providing a colorbar. These figures carry very little
useful information and I encourage authors to rethink ways to present their results. I also suggest authors to provide
examples of images and show the wavelet decomposition results. Also, show the average approximate coefficient
matrix. Last but not least, quantify a change map and provide a validation test such as an independent optical image
taken from the same area...}

\medskip

\textbf{R:} Details about the images were included in the text:
the acquisition satellite/sensor; spatial resolution; and the image size.
% type of satellite image; type of sensor; resolution; \textcolor{red}{and description of pre-processing steps used before applying the proposed method}. 
 We also removed similar figures that were not adding new information to the analysis. The figures with overlapping curves are no longer used, and different details about the region studied are presented now, including change maps in Figures 6(c) and 10. In particular, Figure 6(c) is a validation/reference change map that is used to compute accuracy measure of the methods considered. These measures are given in Table 1 of the revised manuscript.% recent manuscript version.
\end{mybox}
%-------------------------------------

%\newpage
\medskip
\medskip
\medskip

%-------------------------------------
%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#3}

%-------------------------------------
\begin{mybox}{General Comment}

\textit{The manuscript presents a methodology to detect changes in the wavelet domain using multitemporal SAR
data. Although all the mathematical formalism looks sound and the results on simulated data are promising,
the manuscript is still lacking of some important parts. Most importantly, a proper analysis of results using
real data should be done. Real data are indeed used, but their analysis has little to do with validating change
detection and it is more about showing how the new observables look like. More information should also be
added to the simulations. After the addition of these information and a proper validation the paper should be
publishable in TGRS.}
\medskip

\textbf{R:} We thank the reviewer for all his/her contributions, technical insights and the time
spent on reviewing our manuscript. The analysis of real data was improved and now we offer more details about the images used and present a validation change image to measure the accuracy of the proposed method. We also provide now a better description of how the simulated images were generated and analyzed.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 1}
\textit{Data used: In the entire paper there is no mention of the type of data used. Are this Sentinel-1 GRD images? Also please consider acknowledging ESA for the use of data.}

\medskip
\textbf{R:} The reviewer is correct, the Sentinel-1 GRD images were employed in this research. The revised text includes a description of this sensor/dataset. Acknowledgments to ESA were also included at the end of the manuscript. Thank you.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 2}
\textit{Simulations: Some more details should be provided to describe how the simulations were performed. I assume these are Monte Carlo simulations where an Additive Gaussian Noise is added to the image (representing
the log intensity). This is not ideal, since it opens the question on the noise level you simulate (i.e. the energy
of the noise). Why did you selected that energy for the noise? Ideally, you would simulate the complex pixel
as a Circular Complex Gaussian random variable, then take the intensity (maybe use multi-look) and then the
log. Please justify the value of noise energy you selected or please repeat the simulations starting from complex
pixels.}

\medskip
\textbf{R:} The simulated images were indeed generated with additive Gaussian noise. We included more details about how the simulated images were generated and what level of noise was used. The signal to noise to ratio considered was chosen in order to not let the true changes too apparent in the observed images (especially image 4), as can be seen in Figure 2. Such level of noise is also more challenging to analyze than what is observed in the application with real data. Such explanations are presented in the manuscript now.   
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 3}
\textit{Figure 2: Is (a) a validation mask where each pixel that changes at any time is flagged as 1? Please add
more details on the content of subplot (a).}

\medskip
\textbf{R:} Yes, ones (white) correspond to locations were change happened at any instant, whereas zeros (black) correspond to places where no change takes place.  This information is given now in the recent version and the figure now is on Figure 3a. We also present more details about the content of the remaining images as well.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 4}
\textit{Timeseries vs pairwise: At the moment, it is not clear what is the advantage of using the full time series
compared to apply a pairwise approach using first and last image. In the simulation, is this a better choice in
detecting changes that applying a single pairwise change detector between image 1 and image 4? So not do the
accumulation, but just use first and last change detection for comparison.}

\medskip
\textbf{R:} The simulated images illustrate the point in considering the full time series: if only images 1 and 4 were considered in a pairwise analysis, the changes that happened between these instants would be missing. The proposed method is able to detect this type of change, as can be seen in Figure 3b. Additionally, we also look for instants when a relevant change might have happened. The method proposed can be used to detect this type of change as well through analyzing the the measures $\mathbf{d}^{(1)},\ldots,\mathbf{d}^{(n)}$.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 5}
\textit{Real data info and preprocessing: There is absolutely no information about what satellite data are used. I
assume this are Sentinel-1 GRD, but please provide info on that. Also there is no info on what pre-processing
you use. Did you do calibration, terrain correction, noise removal, geocoding?}

\medskip
\textbf{R:} As mentioned (in comment \#1), we added more details about the adopted images in the experiment with actual remote sensing data.

% used in the real data application. They are indeed Sentinel-1 images.

%\textcolor{green}{Não houve pre-processamento, mas é interessante incluir isso no texto. Eu farei isso.}
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 6}
\textit{Real data validation: The test with real data is very poor and should be improved. For instance, there is
no information on what changed in the scene. Also there is no detection mask obtained from the analysis and
no quantitative results (e.g. a ROC curve). These informations must be provided in order to make this paper
publishable.}

\medskip
\textbf{R:} A validation image with change locations is used in the recent version and can be seen in Figure 6(c). This image is used to compute accuracy measures of the proposed method and standard approaches. A change map is also provided using two different types of thresholding algorithms (Otsu and Kittler-Illingworth).
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 7}
\textit{Figure 4 to 6: These are temporal trends, but there is no information on what the reader is supposed to see
in these trends. If you take the backscattering (or any other measure, e.g. kurtosis) it will always show a trend...
why are the trends coming from your method better than any other trend one could get?\\
Each plot I assume is referred to a region. What is the region you are looking at (forest, river, agriculture)?
How big is the region? Are we supposed to have changes in these regions? If so, which changes and when they
happen?}

\medskip
\textbf{R:} These figures are not presented in the recent version, and only an overall change measure is presented in Figure 7. An increasing trend in this plot means that the corresponding images are changing with respect to the mean image. A decreasing trend indicates that images are becoming more similar to the mean image.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 8}
\textit{Figure 7 to 9: What do all these colour lines in the plots refer too? Are different different regions?}

\medskip
\textbf{R:} Yes, each color corresponds to a different location (pixel). The goal was to show that locations flagged as change point have squared deviations that behave similarly as the overall change measure. Since these figures were somewhat confusing, we decided not to show them in the recent version.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 9}
\textit{Figure 10 and 11: Do you have information on the area that changed? If you want to use R (d) for detection,
and low correlation are flagged as changing areas, doe s this mean that the river (permanent water body) is the
most changing part?}

\medskip
\textbf{R:} We present now a change map that is used to evaluate the methods. When using $\mathbf{R}$, high correlations are the ones flagged as corresponding to changing areas. This is because the correlation is computed between a local change measure and an overall change measure. Therefore, the river is not detected as changing region by the proposed method.
\end{mybox}
%-------------------------------------

\medskip
%-------------------------------------
\begin{mybox}{Comment 10}
\textit{P.2, c.1, l.31: Please change "practive" into "practice".}

\medskip
\textbf{R:} The word was corrected in the recent version.
\end{mybox}
%-------------------------------------

\vspace{0.3cm}

\bibliographystyle{apalike}
\bibliography{../JARS_DA_review/refsartigo}

\end{document}
